### 2017-02-15 paper.tex
all

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% This is a (brief) model paper using the achemso class
%% The document class accepts keyval options, which should include
%% the target journal and optionally the manuscript type.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[journal=jprobs,manuscript=article]{achemso}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional packages needed here.  Only include packages
%% which are essential, to avoid problems later. Do NOT use any
%% packages which require e-TeX (for example etoolbox): the e-TeX
%% extensions are not currently available on the ACS conversion
%% servers.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% If issues arise when submitting your manuscript, you may want to
%% un-comment the next line.  This provides information on the
%% version of every file you have used.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%\listfiles

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Place any additional macros here.  Please use \newcommand* where
%% possible, and avoid layout-changing macros (which are not used
%% when typesetting).
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcommand*\mycommand[1]{\texttt{\emph{#1}}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Meta-data block
%% ---------------
%% Each author should be given as a separate \author command.
%%
%% Corresponding authors should have an e-mail given after the author
%% name as an \email command. Phone and fax numbers can be given
%% using \phone and \fax, respectively; this information is optional.
%%
%% The affiliation of authors is given after the authors; each
%% \affiliation command applies to all preceding authors not already
%% assigned an affiliation.
%%
%% The affiliation takes an option argument for the short name.  This
%% will typically be something like "University of Somewhere".
%%
%% The \altaffiliation macro should be used for new address, etc.
%% On the other hand, \alsoaffiliation is used on a per author basis
%% when authors are associated with multiple institutions.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\author{Andrew N. Other}
\altaffiliation{A shared footnote}
\author{Fred T. Secondauthor}
\altaffiliation{Current address: Some other place, Othert\"own,
Germany}
\author{I. Ken Groupleader}
\altaffiliation{A shared footnote}
\email{i.k.groupleader@unknown.uu}
\phone{+123 (0)123 4445556}
\fax{+123 (0)123 4445557}
\affiliation[Unknown University]
{Department of Chemistry, Unknown University, Unknown Town}
\alsoaffiliation[Second University]
{Department of Chemistry, Second University, Nearby Town}
\author{Susanne K. Laborator}
\email{s.k.laborator@bigpharma.co}
\affiliation[BigPharma]
{Lead Discovery, BigPharma, Big Town, USA}
\author{Kay T. Finally}
\affiliation[Unknown University]
{Department of Chemistry, Unknown University, Unknown Town}
\alsoaffiliation[Second University]
{Department of Chemistry, Second University, Nearby Town}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[An \textsf{achemso} demo]
  {Rapid and Accurate Global PTM Discovery (GPTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{IR,NMR,UV}
\keywords{Proteomics}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The manuscript does not need to include \maketitle, which is
%% executed automatically.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "tocentry" environment can be used to create an entry for the
%% graphical table of contents. It is given here as some journals
%% require that it is printed as part of the abstract page. It will
%% be automatically moved as appropriate.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{tocentry}

Some journals require a graphical entry for the Table of Contents.
This should be laid out ``print ready'' so that the sizing of the
text is correct.

Inside the \texttt{tocentry} environment, the font used is Helvetica
8\,pt, as required by \emph{Journal of the American Chemical
Society}.

The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
permitted for  \emph{Journal of the American Chemical Society}
graphical table of content entries. The box will not resize if the
content is too big: instead it will overflow the edge of the box.

This box and the associated title will always be printed on a
separate page at the end of the document.

\end{tocentry}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The abstract environment will automatically gobble the contents
%% if an abstract is not used by the target journal.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
%  This is an example document for the \textsf{achemso} document
%  class, intended for submissions to the American Chemical Society
%  for publication. The class is based on the standard \LaTeXe\
%  \textsf{report} file, and does not seek to reproduce the appearance
%  of a published paper.
%
%  This is an abstract for the \textsf{achemso} document class
%  demonstration document.  An abstract is only allowed for certain
%  manuscript types.  The selection of \texttt{journal} and
%  \texttt{manuscript} will determine if an abstract is valid.  If
%  not, the class will issue an appropriate error.
Posttranslational-modifications (PTMs) influence many aspects of protein function in biological processes, and correctly identifying the various protein modifications in biological samples is crucial for understanding proteins. Ways of identifying and localizing PTMs are limited, but emerging techniques in the field of mass spectrometry are becoming available. GPTM-D [Journal of Very Important Results, 1, 1 (2016)] is a recently developed tool for global identification of PTMs using a single pass database search that is promising. Spectra file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of new PTM identification. We describe the calibration tool developed, and present numerical results that validate the proposed enhancement.

Correctly identifying protein posttranslational-modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes. Methods to identify and localize PTMs from complex biological samples have been limited, but a few techniques are emerging. G-PTM-D\cite{doi:10.1021/acs.jproteome.6b00034} is a recently developed tool for global identification of PTMs. Spectra file calibration prior to applying the tool, We enhanced this method by using non-standard mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations. The number of identified modifications increased by 20\% and the search time decreased by an order of magnitude. We also show the advantages of using mass windows in a standard database search, when PTM discovery is not the primary goal.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
%This is a paragraph of text to fill the introduction of the
%demonstration file.  The demonstration file attempts to show the
%modifications of the standard \LaTeX\ macros that are implemented by
%the \textsf{achemso} class.  These are mainly concerned with content,
%as opposed to appearance.
Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available. Information about a modification can include the chemical or isotopic composition, average mass, monoisotopic mass, specificity to certain residues, and possible restriction of placement to certain peptide or protein termini. While this data is almost comprehensive, information about the localization of these modification to certain residues in specific proteins is scarce. 
We have recently described a bioinformatics tool, GPTM-D, for the discovery and localization of new PTMs from “bottom-up” tandem mass spectrometric datasets1. The GPTM-D workflow consists of three stages: 1) An open mass database search2-3 that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).  2) For those peptides for which the mass difference corresponds to a known PTM, a database augmentation step adds plausible localized PTMs to the peptides in the search database. 3) A final standard narrow mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR). 
We propose an extension to GPTM-D that addresses the long search times of the open mass search, and the limitations of the standard final narrow mass search. We propose a change to the standard GPTM-d workflow using a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications. This change improves the search time by orders of magnitude, and improves the specificity as well. 
Along with algorithmic improvements, we recommend using spectral calibration prior to running the GPTM-d procedure. A critical parameter for peptide identification and PTM localization is mass accuracy4. Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.  Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.  Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories:  External calibration prior to the MS experiment (e.g. standard instrument calibration protocols); internal calibration during the MS experiment (e.g. real-time calibration using a lock mass standard5); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).  We use a post-acquisition calibration procedure, that builds upon the software lock mass concept6 recently reported by the Mann group.  In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.  The increased mass accuracy of the recalibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.  
The new search strategy implementing these changes was tested in the analysis of 2 deep proteomic datasets, where we saw a\% increase in overall peptide identification, and a \% increase in the identification of post-translationally modified peptides.
Finally, we propose a carefully drafted heuristic for identifying and localizing modifications not currently present in any database.

\section{Experimental Procedures}
We developed a modified version of the Morpheus software for bottom-up spectral database searching that we call MetaMorpheus, which integrates a spectral calibration procedure with the G-PTM-D workflow. Two multi-fraction mammalian datasets with deep proteome coverage were used to evaluate the performance of the methods. Uniprot XML protein databases containing only reviewed human/mouse proteins were employed. A 10ppm precursor mass tolerance was used for the initial calibration step, and then reduced to 5ppm for subsequent searches in defined narrow mass windows (notch searches). For the Global PTM Discovery (G-PTM-D) process, we allowed 120 modifications including PTMs, adducts, and chemical modifications. For the discovery of other modifications, we used a novel wide mass-window search (interval search).
The database search software employed is a modified version of Morpheus7, MetaMorpheus. We incorporated the calibration procedure, and the GPTM-D workflow in this modified version for convenience.
Database search parameters are 10ppm precursor mass tolerance for uncalibrated and 5ppm precursor tolerance for calibrated spectra. We use a 0.01 Dalton product mass tolerance for uncalibrated and 0.006 Dalton for calibrated spectra. On-the-fly decoy database generation was used everywhere.
For both mouse and human datasets we use XML databases from uniprot acquired on 7/28/16 and containing only reviewed proteins. 
The datasets tested are described in detail in 9-10. You can find them at http://www.LloydSmithDatasets.com.

\section{Results and Discussion}
By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PTMs increased by 8\%. The overall search time dropped significantly, from 30 hours to 3.5 hours for all datasets. Using an expanded modification list that includes chemical modifications such as adducts and large mass PTMs such as glycosylations, the overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%. We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Daltons. Algorithmic improvements in the G-PTM-D procedure itself, such as searching for combinations of modifications and for modifications with neutral losses make it an excellent tool for PTM discovery.
Previously described wide mass search strategies for discovery and localization of unknown modifications benefit from using alternative, carefully designed search modes based on interval and notch searches. We discovered that limiting mass shifts to a lower bound of -187 Da (corresponding to the largest mass difference that could be attributed to loss of a single residue, Tryptophan), and no upper bound, is an important step in eliminating spurious PSMs. Furthermore, treating highly suspect mass shifts corresponding to residue additions/substitutions/deletions in a separate notch search with individualized false discovery rate estimates is beneficial. An automated tool built into MetaMorpheus allows confidently identifying novel modifications based on these search results. These strategies for increasing confidence in identifications with unusual mass shifts can be used in any database search.
optional Novel Modification Discovery step. The input into the workflow is a standard bottom-up tandem-MS dataset obtained from a tryptically digested protein sample along with a corresponding protein sequence database, and the output is a comprehensive set of peptide and PTM identifications.  Before describing the workflow in detail, it is useful to review and define the different database search modes that are employed. These are illustrated in Figure 1B, which shows four variations: standard search, notch search, comb search, and open search. These differ with respect to whether the entire mass space is searched, as opposed to defined subsets of the mass space; and with respect to the narrowness of the mass window employed for each element in the search space.  The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum). In the following sections we describe the various steps in the flowchart.
a)	b)

Figure 1: a) The blue components mark the enhanced GPTM-D workflow for identifying and localizing PTMs. The green components are the steps in the novel Modification discovery heuristic. b) The four different search strategies
CALIBRATION
The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g. 10 ppm) database search is performed on the dataset. This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1%; based upon the target:decoy strategy (ref)).  Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed. The calibration algorithm is described in detail in Supplementary Material 1.
A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value. We withheld 30% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states. The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset. The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively. Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
a)	b)

Figure 2: a) b)

ENHANCED GPTM-D
The purpose of the Notch Search is to identify peptides that have a mass that is different from the selected ion mass by a mass corresponding to a known modification. This search is expected to produce some peptide spectrum matches with a significantly different mass than the corresponding extracted mass from the MS spectrum. This difference is hypothesized to arise from the presence of PTMs, adducts such as oxidation or metal contamination, or combinations thereof. Even though adducts arising from experimental artefacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS. Using the notch search capability, we can perform this search an order of magnitude faster than the equivalent open mass search in GPTM-D. A curated set of notches and the corresponding modifications is given in Appendix A. Using an off-the shelf database of modification is not advisable, due to shortcomings in each examined database, see next section.
The database augmentation step is like the procedure described in the original GPTM-d work: an option to have the modification with the appropriate mass is added to the protein database, so that a subsequent search may include it as an option. An important difference is the ability to deal with combinations of modifications.
Since the XML database is annotated with PTMs from the UniProt PTM list, it is a natural first idea to use it in the augmentation step as well. It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many things are missing, including adducts and glycosylations.
The final search with the augmented database is conducted with notches: the final notch search is specific to each dataset: we only recommend including the common missed monoisotopic mass values in the final search: see the following section for justification. Newly identified localized PTMs are now readily identified
The effects of the increased mass accuracy were demonstrated in the numerical results for step 2, so to make the comparison fair we use the same precursor and product mass tolerances for both the standard and the notch mass search. The search time is understandably slightly higher for the Notch search, but it is not a big deal since the search times for both are not long, and orders of magnitude shorter than the searches in step 3. 
	Standard Search-Calibrated	Notch Search-Calibrated
Time (jurkat)	22.1 hrs	2.9 hrs
Time (Mouse)	13.5 hrs	1.2 hrs

DISCOVERY OF MODIFICATIONS NOT IN DATABASE
The proposed notch search strategy can be used for discovery of novel modifications not present in a database. We describe the heuristic approach for augmenting the modification database and the final notch list. 
The discovery heuristic starts with conducting a modified comb interval search of a calibrated file. The highly non-uniform shape of the allowable mass differences to search is motivated by the Open Search Case Study in the following section. This allows us to pre-emptively restrict the database search to only look for peptides in intervals that satisfy the condition of being within the specified search windows. This alternative to the open search improves search times and increases specificity.
After the conclusion of the search, the PSMs within 1% FDR are selected, and the corresponding mass differences are grouped. Then the peaks (mass difference values with multiple identifications) in this histogram are identified. We then attempt to categorize these peaks into one of the following categories:
1.	No mass error. Exact match (within a certain noise tolerance) between the peptide corresponding to the MS2 fragmentation and the reported precursor monoisotopic mass.
2.	Monoisotopic error. Peaks around values such as 1.003, 2.006, 3.009 (multiples of the difference in mass between C12 and C13). These correspond to having misidentified the monoisotopic peak in the preprocessing deconvolution step. In this case, the identified peptide is still correct. 
3.	Peaks corresponding to PTMs or adducts, e.g. 42.010 for Acetylation or 21.981 for Sodium adduct. In this analysis, there is no conceptual difference between PTMs and adducts: both are observed simply as mass differences between the identified peptide and the precursor mass.  These peaks are further classified into
a.	Localizeable – e.g. Methylation. There is evidence for the modification in the MS2 spectrum: the corresponding peaks are shifted by an appropriate amount. 
b.	Labile – e.g. Sulfonation. There is no evidence of the modification in the fragmentation spectrum. 
c.	Either Localizeable or Labile, e.g. Phosphorylation. Sometimes there is evidence of it in the fragmentation spectrum, and sometimes there is not. 
4.	Amino acid removals, additions or substitutions.
5.	Modification dependent mass shifts – these peaks occur only in presence of certain modifications in the identified peptide. 	E.g. -15.995 in presence of an identified oxidation, or -79.966 in presence on an identified phosphorylation. 
6.	Combinations of any of the above, e.g. 1.987 for combination of a monoisotopic error and a deamidation. Some of these combinations occur frequently, and it is crucial to account for them. 
An automated script analyzes each identified peak, and provides clues to the nature of the peak. For every peak, a profile is automatically generated. It includes the total number of unique peptides associated with the mass shift, the fraction of decoys, mass match with any known entry in the Unimod or UniProt database, mass match to an amino acid addition/removal combination, mass match to a combination of higher frequency peaks, fraction of localizable targets, localization residues and/or termini, and presence of any modifications in the matched peptides. We follow the following procedure to classify mass shift peaks. 
1.	A z score is computed for the fraction of decoys in the peak, comparing with 1%FDR. If it corresponds to p>0.05, we do not include this peak in the analysis, because there is no sufficient confidence it produces legitimate results.
2.

An attempted localization of the mass difference is automatically performed on every peptide-spectrum match. We carefully examine the mass differences with a localization fraction greater than 0.2, and attempt to deduce the chemical formula, specificity sites and positions within a peptide. Once a determination has been made, we add the new modification type to the modifications list.
An example output of this step is:

The improvements of using a Comb search instead of an Open search are two-fold. We summarize the differences in Table 2.
	Open Search	Comb Search
Search Time	54.36 hrs	29.66 hrs

The improvement in the discernibility of PTMs is apparent when running a comb search on calibrated vs uncalibrated spectra files. Specifically, the discernibility of PTMs with similar mass errors (ones that are only different because of the mass defect). This was not really possible previously. In the current run, Sulfation and Phosphorylation are readily distinguishable. We present the numbers for the example below in the step 5 numerical validation section.
OPEN SEARCH CASE STUDY
Introduction of the comb search instead of an open search is motivated by a careful review of the Unimod database. Known modifications with mass difference within 200 Daltons have values that are within [-0.1, 0.2] of every integer. This interval choice stems from the mass defect in the primary isotope of the common elements. PTM combinations also have this property.
We tested multiple search strategies in a search for one that gives the smallest number of false positives. 
KNOWN MODIFICATION CONFIRMATION CASE STUDY
Confirmation of the properties of Sodium Adducts, Phosphorylation and Sulfonation.
GLYCOPROTEINS CASE STUDY
Proteins that contain modifications such as Hexose, HexNAC, and similar ones are interesting. The uniprot database only lists some possible sites for such modifications, but does not specify the actual type – thus creating ambiguity regarding the type of the modification. The new modification discovery workflow is well suited to identify such proteins. 
NOVEL MODIFICATION CASE STUDY

%
%\subsection{References}
%
%The class makes various changes to the way that references are
%handled.  The class loads \textsf{natbib}, and also the
%appropriate bibliography style.  References can be made using
%the normal method; the citation should be placed before any
%punctuation, as the class will move it if using a superscript
%citation style
%\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
%The use of \textsf{natbib} allows the use of the various citation
%commands of that package: \citeauthor{Abernethy2003} have shown
%something, in \citeyear{Cotton1999}, or as given by
%Ref.~\citenum{Mena2000}.  Long lists of authors will be
%automatically truncated in most article formats, but not in
%supplementary information or reviews \cite{Pople2003}. If you
%encounter problems with the citation macros, please check that
%your copy of \textsf{natbib} is up to date. The demonstration
%database file \texttt{achemso-demo.bib} shows how to complete
%entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
%using the \texttt{\textbackslash latin} command.
%
%Multiple citations to be combined into a list can be given as
%a single citation.  This uses the \textsf{mciteplus} package
%\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
%Citations other than the first of the list should be indicated
%with a star. If the \textsf{mciteplus} package is not installed,
%the standard bibliography tools will still work but starred
%references will be ignored. Individual references can be referred
%to using \texttt{\textbackslash mciteSubRef}:
%``ref.~\mciteSubRef{Eisenstein2005}''.
%
%The class also handles notes to be added to the bibliography.  These
%should be given in place in the document \bibnote{This is a note.
%The text will be moved the the references section.  The title of the
%section will change to ``Notes and References''.}.  As with
%citations, the text should be placed before punctuation.  A note is
%also generated if a citation has an optional note.  This assumes that
%the whole work has already been cited: odd numbering will result if
%this is not the case \cite[p.~1]{Cotton1999}.

%\newpage
%
%\subsection{Floats}
%
%New float types are automatically set up by the class file.  The
%means graphics are included as follows (Scheme~\ref{sch:example}).  As
%illustrated, the float is ``here'' if possible.
%\begin{scheme}
%  Your scheme graphic would go here: \texttt{.eps} format\\
%  for \LaTeX\, or \texttt{.pdf} (or \texttt{.png}) for pdf\LaTeX\\
%  \textsc{ChemDraw} files are best saved as \texttt{.eps} files:\\
%  these can be scaled without loss of quality, and can be\\
%  converted to \texttt{.pdf} files easily using \texttt{eps2pdf}.\\
%  %\includegraphics{graphic}
%  \caption{An example scheme}
%  \label{sch:example}
%\end{scheme}peptideIndex.Count: 10465024
%
%
%\begin{figure}
%  As well as the standard float types \texttt{table}\\
%  and \texttt{figure}, the class also recognises\\
%  \texttt{scheme}, \texttt{chart} and \texttt{graph}.
%  \caption{An example figure}
%  \label{fgr:example}
%\end{figure}
%
%Charts, figures and schemes do not necessarily have to be labelled or
%captioned.  However, tables should always have a title. It is
%possible to include a number and label for a graphic without any
%title, using an empty argument to the \texttt{\textbackslash caption}
%macro.
%
%The use of the different floating environments is not required, but
%it is intended to make document preparation easier for authors. In
%general, you should place your graphics where they make logical
%sense; the production process will move them if needed.
%
%\subsection{Math(s)}
%
%The \textsf{achemso} class does not load any particular additional
%support for mathematics.  If packages such as \textsf{amsmath} are
%required, they should be loaded in the preamble.  However,
%the basic \LaTeX\ math(s) input should work correctly without
%this.  Some inline material \( y = mx + c \) or $ 1 + 1 = 2 $
%followed by some display. \[ A = \pi r^2 \]
%
%It is possible to label equations in the usual way (Eq.~\ref{eqn:example}).
%\begin{equation}
%  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
%\end{equation}
%This can also be used to have equations containing graphical
%content. To align the equation number with the middle of the graphic,
%rather than the bottom, a minipage may be used.
%\begin{equation}
%  \begin{minipage}[c]{0.80\linewidth}
%    \centering
%    As illustrated here, the width of \\
%    the minipage needs to allow some  \\
%    space for the number to fit in to.
%    %\includegraphics{graphic}
%  \end{minipage}
%  \label{eqn:graphic}
%\end{equation}
%
%\section{Experimental}
%
%The usual experimental details should appear here.  This could
%include a table, which can be referenced as Table~\ref{tbl:example}.
%Notice that the caption is positioned at the top of the table.
%\begin{table}
%  \caption{An example table}
%  \label{tbl:example}
%  \begin{tabular}{ll}
%    \hline
%    Header one  & Header two  \\
%    \hline
%    Entry one   & Entry two   \\
%    Entry three & Entry four  \\
%    Entry five  & Entry five  \\
%    Entry seven & Entry eight \\
%    \hline
%  \end{tabular}
%\end{table}
%
%Adding notes to tables can be complicated.  Perhaps the easiest
%method is to generate these using the basic
%\texttt{\textbackslash textsuperscript} and
%\texttt{\textbackslash emph} macros, as illustrated (Table~\ref{tbl:notes}).
%\begin{table}
%  \caption{A table with notes}
%  \label{tbl:notes}
%  \begin{tabular}{ll}
%    \hline
%    Header one                            & Header two \\
%    \hline
%    Entry one\textsuperscript{\emph{a}}   & Entry two  \\
%    Entry three\textsuperscript{\emph{b}} & Entry four \\
%    \hline
%  \end{tabular}
%
%  \textsuperscript{\emph{a}} Some text;
%  \textsuperscript{\emph{b}} Some more text.
%\end{table}
%
%The example file also loads the optional \textsf{mhchem} package, so
%that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
%gives \ce{H2SO4}.  See the use in the bibliography file (when using
%titles in the references section).
%
%The use of new commands should be limited to simple things which will
%not interfere with the production process.  For example,
%\texttt{\textbackslash mycommand} has been defined in this example,
%to give italic, mono-spaced text: \mycommand{some text}.

%\section{Extra information when writing JACS Communications}
%
%When producing communications for \emph{J.~Am.\ Chem.\ Soc.}, the
%class will automatically lay the text \begin{flushleft}

%\end{flushleft}out in the style of the
%journal. This gives a guide to the length of text that can be
%accommodated in such a publication. There are some points to bear in
%mind when preparing a JACS Communication in this way.  The layout
%produced here is a \emph{model} for the published result, and the
%outcome should be taken as a \emph{guide} to the final length. The
%spacing and sizing of graphical content is an area where there is
%some flexibility in the process.  You should not worry about the
%space before and after graphics, which is set to give a guide to the
%published size. This is very dependant on the final published layout.
%
%You should be able to use the same source to produce a JACS
%Communication and a normal article.  For example, this demonstration
%file will work with both \texttt{type=article} and
%\texttt{type=communication}. Sections and any abstract are
%automatically ignored, although you will get warnings to this effect.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The "Acknowledgement" section can be given in all manuscript
%% classes.  This should be given within the "acknowledgement"
%% environment, which will make the correct section or running title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{acknowledgement}

The authors thank \ldots
\end{acknowledgement}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The same is true for Supporting Information, which should use the
%% suppinfo environment.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{suppinfo}

A listing of the contents of each file supplied as Supporting Information
should be included. For instructions on what should be included in the
Supporting Information as well as how to prepare this material for
publications, refer to the journal's Instructions for Authors.

The following files are available free of charge.
\begin{itemize}
  \item Filename: brief description
  \item Filename: brief description
\end{itemize}

\end{suppinfo}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The appropriate \bibliography command should be placed here.
%% Notice that the class file automatically sets \bibliographystyle
%% and also names the section correctly.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\bibliography{citations}

\end{document}

### 2017-02-16 paper.tex
chages

@@
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% This is a (brief) model paper using the achemso class
-%% The document class accepts keyval options, which should include
-%% the target journal and optionally the manuscript type.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \documentclass[journal=jprobs,manuscript=article]{achemso}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Place any additional packages needed here.  Only include packages
-%% which are essential, to avoid problems later. Do NOT use any
-%% packages which require e-TeX (for example etoolbox): the e-TeX
-%% extensions are not currently available on the ACS conversion
-%% servers.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% If issues arise when submitting your manuscript, you may want to
-%% un-comment the next line.  This provides information on the
-%% version of every file you have used.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%\listfiles
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Place any additional macros here.  Please use \newcommand* where
-%% possible, and avoid layout-changing macros (which are not used
-%% when typesetting).
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \newcommand*\mycommand[1]{\texttt{\emph{#1}}}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Meta-data block
-%% ---------------
-%% Each author should be given as a separate \author command.
-%%
-%% Corresponding authors should have an e-mail given after the author
-%% name as an \email command. Phone and fax numbers can be given
-%% using \phone and \fax, respectively; this information is optional.
-%%
-%% The affiliation of authors is given after the authors; each
-%% \affiliation command applies to all preceding authors not already
-%% assigned an affiliation.
-%%
-%% The affiliation takes an option argument for the short name.  This
-%% will typically be something like "University of Somewhere".
-%%
-%% The \altaffiliation macro should be used for new address, etc.
-%% On the other hand, \alsoaffiliation is used on a per author basis
-%% when authors are associated with multiple institutions.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\author{Andrew N. Other}
-\altaffiliation{A shared footnote}
-\author{Fred T. Secondauthor}
-\altaffiliation{Current address: Some other place, Othert\"own,
-Germany}
-\author{I. Ken Groupleader}
-\altaffiliation{A shared footnote}
-\email{i.k.groupleader@unknown.uu}
-\phone{+123 (0)123 4445556}
-\fax{+123 (0)123 4445557}
-\affiliation[Unknown University]
-{Department of Chemistry, Unknown University, Unknown Town}
-\alsoaffiliation[Second University]
-{Department of Chemistry, Second University, Nearby Town}
-\author{Susanne K. Laborator}
-\email{s.k.laborator@bigpharma.co}
-\affiliation[BigPharma]
-{Lead Discovery, BigPharma, Big Town, USA}
-\author{Kay T. Finally}
-\affiliation[Unknown University]
-{Department of Chemistry, Unknown University, Unknown Town}
-\alsoaffiliation[Second University]
-{Department of Chemistry, Second University, Nearby Town}
+\author{Stefan K. Solntsev}
+\email{solntsev@wisc.edu}
+\author{Michael R. Shortreed}
+\author{Brian L. Frey}
+\author{Lloyd M. Smith}
+\affiliation[UwMadison]
+{University of Wisconsin-Madison}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The document title should be given as usual. Some journals require
-%% a running title from the author: this should be supplied as an
-%% optional argument to \title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \title[An \textsf{achemso} demo]
-  {Rapid and Accurate Global PTM Discovery (GPTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows}
+  {Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Some journals require a list of abbreviations or keywords to be
-%% supplied. These should be set up here, and will be printed after
-%% the title and author information, if needed.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\abbreviations{IR,NMR,UV}
-\keywords{Proteomics}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The manuscript does not need to include \maketitle, which is
-%% executed automatically.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{document}
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "tocentry" environment can be used to create an entry for the
-%% graphical table of contents. It is given here as some journals
-%% require that it is printed as part of the abstract page. It will
-%% be automatically moved as appropriate.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-\begin{tocentry}
-
-Some journals require a graphical entry for the Table of Contents.
-This should be laid out ``print ready'' so that the sizing of the
-text is correct.
-
-Inside the \texttt{tocentry} environment, the font used is Helvetica
-8\,pt, as required by \emph{Journal of the American Chemical
-Society}.
-
-The surrounding frame is 9\,cm by 3.5\,cm, which is the maximum
-permitted for  \emph{Journal of the American Chemical Society}
-graphical table of content entries. The box will not resize if the
-content is too big: instead it will overflow the edge of the box.
-
-This box and the associated title will always be printed on a
-separate page at the end of the document.
-
-\end{tocentry}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The abstract environment will automatically gobble the contents
-%% if an abstract is not used by the target journal.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{abstract}
-%  This is an example document for the \textsf{achemso} document
-%  class, intended for submissions to the American Chemical Society
-%  for publication. The class is based on the standard \LaTeXe\
-%  \textsf{report} file, and does not seek to reproduce the appearance
-%  of a published paper.
-%
-%  This is an abstract for the \textsf{achemso} document class
-%  demonstration document.  An abstract is only allowed for certain
-%  manuscript types.  The selection of \texttt{journal} and
-%  \texttt{manuscript} will determine if an abstract is valid.  If
-%  not, the class will issue an appropriate error.
-Posttranslational-modifications (PTMs) influence many aspects of protein function in biological processes, and correctly identifying the various protein modifications in biological samples is crucial for understanding proteins.
-Ways of identifying and localizing PTMs are limited, but emerging techniques in the field of mass spectrometry are becoming available.
-GPTM-D [Journal of Very Important Results, 1, 1 (2016)] is a recently developed tool for global identification of PTMs using a single pass database search that is promising.
-Spectra file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of new PTM identification.
-We describe the calibration tool developed, and present numerical results that validate the proposed enhancement.
 
@@
 G-PTM-D\cite{doi:10.1021/acs.jproteome.6b00034} is a recently developed tool for global identification of PTMs.
-Spectra file calibration prior to applying the tool, We enhanced this method by using non-standard mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
+Spectra file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of new PTM identification.
+We enhanced this method by using non-standard mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
 The number of identified modifications increased by 20\% and the search time decreased by an order of magnitude.
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% Start the main part of the manuscript here.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \section{Introduction}
-%This is a paragraph of text to fill the introduction of the
-%demonstration file.  The demonstration file attempts to show the
-%modifications of the standard \LaTeX\ macros that are implemented by
-%the \textsf{achemso} class.  These are mainly concerned with content,
-%as opposed to appearance.
+
 Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available.
@@
 While this data is almost comprehensive, information about the localization of these modification to certain residues in specific proteins is scarce.
-We have recently described a bioinformatics tool, GPTM-D, for the discovery and localization of new PTMs from “bottom-up” tandem mass spectrometric datasets1.
-The GPTM-D workflow consists of three stages: 1) An open mass database search2-3 that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
+
+We have recently described a bioinformatics tool, G-PTM-D, for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets\cite{doi:10.1021/acs.jproteome.6b00034}.
+The GPTM-D workflow consists of three stages: 1) An wide-mass database search2-3 that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
 2) For those peptides for which the mass difference corresponds to a known PTM, a database augmentation step adds plausible localized PTMs to the peptides in the search database.
@@
 1\% FDR).
+
 We propose an extension to GPTM-D that addresses the long search times of the open mass search, and the limitations of the standard final narrow mass search.
@@
 This change improves the search time by orders of magnitude, and improves the specificity as well.
+We propose using a notch search in the final phase.
+
+The mass windows concept was extended to other procedures, beyond the G-PTM-D workflow. We propose using mass windows as an alternative to any wide or open-mass search.
+
 Along with algorithmic improvements, we recommend using spectral calibration prior to running the GPTM-d procedure.
@@
 The increased mass accuracy of the recalibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
+
 The new search strategy implementing these changes was tested in the analysis of 2 deep proteomic datasets, where we saw a\% increase in overall peptide identification, and a \% increase in the identification of post-translationally modified peptides.

### 2017-02-28 paper.tex
many

@@
 \usepackage[T1]{fontenc}       % Use modern font encodings
+\usepackage{multirow}
 
@@
 Methods to identify and localize PTMs from complex biological samples have been limited, but a few techniques are emerging.
-G-PTM-D\cite{doi:10.1021/acs.jproteome.6b00034} is a recently developed tool for global identification of PTMs.
+G-PTM-D\cite{Li_2016} is a recently developed tool for global identification of PTMs.
 Spectra file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of new PTM identification.
-We enhanced this method by using non-standard mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
+We enhanced G-PTM-D by using non-standard mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
 The number of identified modifications increased by 20\% and the search time decreased by an order of magnitude.
-We also show the advantages of using mass windows in a standard database search, when PTM discovery is not the primary goal.
+We also show the advantages of using mass-window searches in other contexts, when PTM discovery is not the primary goal.
 \end{abstract}
@@
 Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available.
-Information about a modification can include the chemical or isotopic composition, average mass, monoisotopic mass, specificity to certain residues, and possible restriction of placement to certain peptide or protein termini.
+Information about a modification can include the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
 While this data is almost comprehensive, information about the localization of these modification to certain residues in specific proteins is scarce.
 
-We have recently described a bioinformatics tool, G-PTM-D, for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets\cite{doi:10.1021/acs.jproteome.6b00034}.
-The GPTM-D workflow consists of three stages: 1) An wide-mass database search2-3 that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
-2) For those peptides for which the mass difference corresponds to a known PTM, a database augmentation step adds plausible localized PTMs to the peptides in the search database.
-3) A final standard narrow mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g.
+We have recently described a bioinformatics tool, G-PTM-D, for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets\cite{Li_2016}.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015,Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
+2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database.
+3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g.
 1\% FDR).
 
-We propose an extension to GPTM-D that addresses the long search times of the open mass search, and the limitations of the standard final narrow mass search.
-We propose a change to the standard GPTM-d workflow using a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
-This change improves the search time by orders of magnitude, and improves the specificity as well.
-We propose using a notch search in the final phase.
+Before describing the proposed changes to G-PTM-D, we introduce the database search mode definitions we use throughout.
+A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e.
+5ppm, or 2.1 Da).
+An \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e.
+$\pm 500$ Da) between the observed precursor mass and the peptide mass.
+An even less restrictive \textit{open-mass} search, matches peptides from a database to  fragmentation spectra without taking the precursor mass into account.
+We propose alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
+The notch search is an extension of the narrow-mass search: it allows exact matches (within a tolerance) to specific mass differences between the theoretical peptide and observed precursor masses.
+The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
+The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum).
+Figure~\ref{fgr:differentSearchModes} provides a convenient way to visualize the various search modes.
 
-The mass windows concept was extended to other procedures, beyond the G-PTM-D workflow. We propose using mass windows as an alternative to any wide or open-mass search.
+We propose an extension to G-PTM-D (Enhanced G-PTM-D) that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
+The wide-mass search is replaced by a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
+This change improves the specificity and significantly decreases search.
+The final narrow-mass search is also replaced by a notch search, further increasing the number of confidently identified peptides.
 
-Along with algorithmic improvements, we recommend using spectral calibration prior to running the GPTM-d procedure.
-A critical parameter for peptide identification and PTM localization is mass accuracy4.
+We recommend using the mass-window search in other procedures as well, beyond the G-PTM-D workflow.
+Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on a wide-mass or an open-mass search.
+We demonstrate that using a combination of specialized notch and interval searches produces superior identification and discovery results.
+
+Along with algorithmic improvements, we recommend using spectral calibration prior to running the G-PTM-D procedure.
+A critical parameter for peptide identification and PTM localization is mass accuracy\cite{Scherl_2008}.
 Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
@@
 standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
-real-time calibration using a lock mass standard5); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).
-We use a post-acquisition calibration procedure, that builds upon the software lock mass concept6 recently reported by the Mann group.
+real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).
+We use a post-acquisition calibration procedure, that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
@@
 
-The new search strategy implementing these changes was tested in the analysis of 2 deep proteomic datasets, where we saw a\% increase in overall peptide identification, and a \% increase in the identification of post-translationally modified peptides.
-Finally, we propose a carefully drafted heuristic for identifying and localizing modifications not currently present in any database.
+\begin{figure}
+  \includegraphics{figureDifferentSearchModes.eps}
+  \caption{Search Modes}
+  \label{fgr:differentSearchModes}
+\end{figure}
 
 \section{Experimental Procedures}
-We developed a modified version of the Morpheus software for bottom-up spectral database searching that we call MetaMorpheus, which integrates a spectral calibration procedure with the G-PTM-D workflow.
-Two multi-fraction mammalian datasets with deep proteome coverage were used to evaluate the performance of the methods.
-Uniprot XML protein databases containing only reviewed human/mouse proteins were employed.
-A 10ppm precursor mass tolerance was used for the initial calibration step, and then reduced to 5ppm for subsequent searches in defined narrow mass windows (notch searches).
-For the Global PTM Discovery (G-PTM-D) process, we allowed 120 modifications including PTMs, adducts, and chemical modifications.
-For the discovery of other modifications, we used a novel wide mass-window search (interval search).
-The database search software employed is a modified version of Morpheus7, MetaMorpheus. We incorporated the calibration procedure, and the GPTM-D workflow in this modified version for convenience.
-Database search parameters are 10ppm precursor mass tolerance for uncalibrated and 5ppm precursor tolerance for calibrated spectra.
-We use a 0.01 Dalton product mass tolerance for uncalibrated and 0.006 Dalton for calibrated spectra.
-On-the-fly decoy database generation was used everywhere.
-For both mouse and human datasets we use XML databases from uniprot acquired on 7/28/16 and containing only reviewed proteins. 
-The datasets tested are described in detail in 9-10. You can find them at http://www.LloydSmithDatasets.com.
+We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the Enhanced G-PTM-D workflow.
+Two multi-fraction mammalian datasets (\textit{Mouse} and \textit{Jurkat}) with deep proteome coverage were used to evaluate the performance of the methods: these datasets are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
+Uniprot XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
+
+A 10ppm precursor mass tolerance was used for the initial calibration step, and then reduced to 5ppm for subsequent notch searches.
+For the database augmentation step, we allowed 120 modifications including PTMs, adducts, and chemical modifications.
+Final notch search used notches centered at \{0, 1.003, 2.005, 3.008\} in order to capture missed identifications due to missed monoisotopic peaks.
+
+For the discovery of novel modifications, we used a combination of two searches: A notch search with notches at all combinations of up to 2 residue mass additions or removals, and an interval search with lower bound of -187 Daltons, and with exclusions at masses corresponding to the notch search.
 
 \section{Results and Discussion}
+
+\subsection{Enhanced G-PTM-D}
+
+The input into the workflow is a standard bottom-up tandem-MS dataset obtained from a tryptically digested protein sample along with a corresponding protein sequence database, and the output is a comprehensive set of peptide and PTM identifications.
+
+Even though adducts arising from experimental artefacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS.
+A curated set of notches and the corresponding modifications is given in Appendix A.
+Using an off-the shelf database of modification is not advisable.
+The database augmentation step is like the procedure described in the original G-PTM-D work: modifications with appropriate mass are added to the protein database, so that subsequent searches may include them as variable localized modifications.
+Since the XML database is annotated with PTMs from the UniProt PTM list, it is a natural first idea to use it in the augmentation step as well.
+It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many PTMs are missing, including adducts and glycosylations.
+The final search with the augmented database is conducted with notches: the final notch search is specific to each dataset: we only recommend including the common missed monoisotopic mass values in the final search: see the following section for justification.
+Newly identified localized PTMs are now readily identified.
+
 By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PTMs increased by 8\%.
-The overall search time dropped significantly, from 30 hours to 3.5 hours for all datasets.
+The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-label}.
 Using an expanded modification list that includes chemical modifications such as adducts and large mass PTMs such as glycosylations, the overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%.
@@
 Algorithmic improvements in the G-PTM-D procedure itself, such as searching for combinations of modifications and for modifications with neutral losses make it an excellent tool for PTM discovery.
+
+\begin{table}[]
+\centering
+\caption{Comparison of Wide-Mass and Notch Searches}
+\label{my-label}
+\begin{tabular}{ll|l|l|l}
+                      &        & Wide-Mass Search & Notch Search & Notch Search With Expanded Mod List \\
+\hline
+\multirow{2}{*}{Time} & Jurkat & 22.1 hrs         & 2.9 hrs   &  4.5 hrs  \\
+                      & Mouse  & 13.5 hrs         & 1.2 hrs   &  1.8 hrs \\
+\hline
+\multirow{2}{*}{PSMs} & Jurkat & 203237           & 210566   &  220574  \\
+                      & Mouse  & 153674           & 162473   &   170743
+\end{tabular}
+\end{table}
+
+\subsection{Mass Windows in an Open Search}
+
+Restricting the allowed precursor masses is beneficial in searches that are intended to discover mass shift peaks.
+We ran a completely open precursor mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
+The plots in Figure~\ref{fgr:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
+It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.
+
+This effect can be explained by the fact that it's easier to match long sequences to short peptide fragmentation profiles than it is to match short sequences to fragmentation pattern of a long peptide.
+
+\begin{figure}
+\caption{Upper and Lower bound cutoffs in Mouse Data}
+\label{fig:figure2-upperlowerbounds}
+\includegraphics[scale=0.5]{figure_2-upperlowerbounds}
+\end{figure}
+
+MetaMorpheus includes a capability to identify peaks in a mass shift histogram obtained from a search, using a recently described peak finding algorithm\cite{Rodriguez_2014}.
+We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
+Some histogram peaks are illustrated in Figure~\ref{fig:figure3}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
+
+\begin{figure}
+\caption{Some Histogram Peaks For Mouse Data}
+\label{fig:figure3}
+\includegraphics[scale=0.6]{figure_3peaks}
+\end{figure}
+
+We identified the property of some specific mass shifts that correspond to anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
+
+Mass windows provide an additional mechanism in improving confident identification rates in a wide mass search.
+
+\subsection{Discovery of Modifications not in Database}
+
 Previously described wide mass search strategies for discovery and localization of unknown modifications benefit from using alternative, carefully designed search modes based on interval and notch searches.
@@
 An automated tool built into MetaMorpheus allows confidently identifying novel modifications based on these search results.
-These strategies for increasing confidence in identifications with unusual mass shifts can be used in any database search.
-optional Novel Modification Discovery step.
-The input into the workflow is a standard bottom-up tandem-MS dataset obtained from a tryptically digested protein sample along with a corresponding protein sequence database, and the output is a comprehensive set of peptide and PTM identifications.
-Before describing the workflow in detail, it is useful to review and define the different database search modes that are employed.
-These are illustrated in Figure 1B, which shows four variations: standard search, notch search, comb search, and open search.
-These differ with respect to whether the entire mass space is searched, as opposed to defined subsets of the mass space; and with respect to the narrowness of the mass window employed for each element in the search space.
-The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum).
-In the following sections we describe the various steps in the flowchart.
-a)	b)
 
-Figure 1: a) The blue components mark the enhanced GPTM-D workflow for identifying and localizing PTMs.
-The green components are the steps in the novel Modification discovery heuristic.
-b) The four different search strategies
-CALIBRATION
-The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g.
-10 ppm) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1%; based upon the target:decoy strategy (ref)).
-Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
-The calibration algorithm is described in detail in Supplementary Material 1.
-A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value.
-We withheld 30% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states.
-The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
-The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively.
-Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
-a)	b)
-
-Figure 2: a) b)
-
-ENHANCED GPTM-D
-The purpose of the Notch Search is to identify peptides that have a mass that is different from the selected ion mass by a mass corresponding to a known modification.
-This search is expected to produce some peptide spectrum matches with a significantly different mass than the corresponding extracted mass from the MS spectrum.
-This difference is hypothesized to arise from the presence of PTMs, adducts such as oxidation or metal contamination, or combinations thereof.
-Even though adducts arising from experimental artefacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS.
-Using the notch search capability, we can perform this search an order of magnitude faster than the equivalent open mass search in GPTM-D.
-A curated set of notches and the corresponding modifications is given in Appendix A.
-Using an off-the shelf database of modification is not advisable, due to shortcomings in each examined database, see next section.
-The database augmentation step is like the procedure described in the original GPTM-d work: an option to have the modification with the appropriate mass is added to the protein database, so that a subsequent search may include it as an option.
-An important difference is the ability to deal with combinations of modifications.
-Since the XML database is annotated with PTMs from the UniProt PTM list, it is a natural first idea to use it in the augmentation step as well.
-It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many things are missing, including adducts and glycosylations.
-The final search with the augmented database is conducted with notches: the final notch search is specific to each dataset: we only recommend including the common missed monoisotopic mass values in the final search: see the following section for justification.
-Newly identified localized PTMs are now readily identified
-The effects of the increased mass accuracy were demonstrated in the numerical results for step 2, so to make the comparison fair we use the same precursor and product mass tolerances for both the standard and the notch mass search.
-The search time is understandably slightly higher for the Notch search, but it is not a big deal since the search times for both are not long, and orders of magnitude shorter than the searches in step 3.
-	Standard Search-Calibrated	Notch Search-Calibrated
-Time (jurkat)	22.1 hrs	2.9 hrs
-Time (Mouse)	13.5 hrs	1.2 hrs
-
-DISCOVERY OF MODIFICATIONS NOT IN DATABASE
 The proposed notch search strategy can be used for discovery of novel modifications not present in a database.
@@
 This alternative to the open search improves search times and increases specificity.
-After the conclusion of the search, the PSMs within 1% FDR are selected, and the corresponding mass differences are grouped.
+After the conclusion of the search, the PSMs within 1\% FDR are selected, and the corresponding mass differences are grouped.
 Then the peaks (mass difference values with multiple identifications) in this histogram are identified.
 We then attempt to categorize these peaks into one of the following categories:
-1.	No mass error. Exact match (within a certain noise tolerance) between the peptide corresponding to the MS2 fragmentation and the reported precursor monoisotopic mass.
-2.	Monoisotopic error.
-Peaks around values such as 1.003, 2.006, 3.009 (multiples of the difference in mass between C12 and C13).
+\begin{itemize}
+\item \textit{No mass error} Exact match (within a certain noise tolerance) between the peptide corresponding to the MS2 fragmentation and the reported precursor monoisotopic mass.
+\item \textit{Monoisotopic error} Peaks around values such as 1.003, 2.006, 3.009 (multiples of the difference in mass between C12 and C13).
 These correspond to having misidentified the monoisotopic peak in the preprocessing deconvolution step.
 In this case, the identified peptide is still correct.
-3.	Peaks corresponding to PTMs or adducts, e.g.
+\item \textit{Exact PTM or Adduct Masses}e.g.
 42.010 for Acetylation or 21.981 for Sodium adduct.
@@
 These peaks are further classified into
-a.	Localizeable – e.g. Methylation. There is evidence for the modification in the MS2 spectrum: the corresponding peaks are shifted by an appropriate amount. 
-b.	Labile – e.g. Sulfonation. There is no evidence of the modification in the fragmentation spectrum. 
-c.	Either Localizeable or Labile, e.g. Phosphorylation. Sometimes there is evidence of it in the fragmentation spectrum, and sometimes there is not. 
-4.	Amino acid removals, additions or substitutions.
-5.	Modification dependent mass shifts – these peaks occur only in presence of certain modifications in the identified peptide.
+\begin{itemize}
+\item Localizeable - e.g. Methylation. There is evidence for the modification in the MS2 spectrum: the corresponding peaks are shifted by an appropriate amount. 
+\item Labile - e.g. Sulfonation. There is no evidence of the modification in the fragmentation spectrum. 
+\item Either Localizeable or Labile, e.g. Phosphorylation. Sometimes there is evidence of it in the fragmentation spectrum, and sometimes there is not. 
+\end{itemize}
+\item Amino acid removals, additions or substitutions.
+\item Modification dependent mass shifts - these peaks occur only in presence of certain modifications in the identified peptide.
 	E.g.
 -15.995 in presence of an identified oxidation, or -79.966 in presence on an identified phosphorylation.
-6.	Combinations of any of the above, e.g. 1.987 for combination of a monoisotopic error and a deamidation. Some of these combinations occur frequently, and it is crucial to account for them. 
+\item Combinations of any of the above, e.g. 1.987 for combination of a monoisotopic error and a deamidation. Some of these combinations occur frequently, and it is crucial to account for them. 
+\end{itemize}
+
 An automated script analyzes each identified peak, and provides clues to the nature of the peak.
@@
 We follow the following procedure to classify mass shift peaks.
-1.	A z score is computed for the fraction of decoys in the peak, comparing with 1%FDR.
+1.	A z score is computed for the fraction of decoys in the peak, comparing with 1\%FDR.
 If it corresponds to p>0.05, we do not include this peak in the analysis, because there is no sufficient confidence it produces legitimate results.
@@
 We tested multiple search strategies in a search for one that gives the smallest number of false positives. 
-KNOWN MODIFICATION CONFIRMATION CASE STUDY
+\subsubsection{Known Modification Confirmation Case Study}
+
 Confirmation of the properties of Sodium Adducts, Phosphorylation and Sulfonation.
-GLYCOPROTEINS CASE STUDY
+\subsubsection{Glycans Case Study}
 Proteins that contain modifications such as Hexose, HexNAC, and similar ones are interesting.
-The uniprot database only lists some possible sites for such modifications, but does not specify the actual type – thus creating ambiguity regarding the type of the modification.
+The uniprot database only lists some possible sites for such modifications, but does not specify the actual type, thus creating ambiguity regarding the type of the modification.
 The new modification discovery workflow is well suited to identify such proteins.
-NOVEL MODIFICATION CASE STUDY
+Table~\ref{tbl:glyco} lists the automatically identified glyco modification masses and identities, sorted by frequency of occurence.
+
+\begin{table}[]
+\centering
+\caption{Glyco}
+\label{tbl:glyco}
+\begin{tabular}{lll}
+MassShift & Count & UnimodID                        \\
+\hline
+203.079   & 30    & HexNAc                          \\
+162.052   & 18    & Hex                             \\
+876.322   & 10    & Hex(2)HexNAc(2)dHex(1)          \\
+656.228   & 8     & Hex(1)HexNAc(1)NeuAc(1)         \\
+365.132   & 8     & Hex(1)HexNAc(1)                 \\
+1216.422  & 8     & Hex(5)HexNAc(2)                 \\
+730.265   & 6     & Hex(2)HexNAc(2)                 \\
+1378.475  & 4     & Hex(6)HexNAc(2)                 \\
+947.323   & 4     & Hex(1)HexNAc(1)NeuAc(2)         \\
+963.317   & 3     & Hex(1)HexNAc(1)NeuAc(1)NeuGc(1) \\
+178.047   & 3     & Galactosyl                      \\
+1038.375  & 3     & dHex(1)Hex(3)HexNAc(2)          \\
+146.058   & 2     & dHex                            \\
+340.1     & 2     & Glucosylgalactosyl              \\
+406.158   & 2     & HexNAc(2)                       \\
+697.256   & 2     & HexNAc(2)NeuAc(1)               \\
+672.223   & 2     & Hex(1)HexNAc(1)NeuGc(1)         \\
+860.326   & 2     & Hex(1)HexNAc(2)dHex(2)         
+\end{tabular}
+\end{table}
+
+\subsubsection{Novel Modification Case Study}
+
+\subsection{Calibration}
+
+The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g.
+10 ppm) database search is performed on the dataset.
+This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy (ref)).
+Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
+The calibration algorithm is described in detail in Supplementary Material 1.
+A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value.
+We withheld 30\% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states.
+The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
+The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively.
+Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
+
+A 10 ppm narrow tolerance search produces matches with mass differences in ppm given in~\ref{fgr:figure1}.
+The average and standard deviations along with 1, 2, and 3 standard deviation widths are given in Table~\ref{tbl:calib}.
+
+\begin{table}
+  \caption{Calibration Results: ppm Mass Differences}
+  \label{tbl:calib}
+  \begin{tabular}{llll}
+    \hline
+    & prior& after smooth & after rf  \\
+    \hline
+    stdev&	1.553	&1.039&	0.898 \\
+	avg&	-3.148&	0.204&0.215\\
+	68.3\% within	& [-4.701,-1.595 ]&	[-0.834,1.243	]&	[-0.683,1.114]\\
+	95.4\% within&[	-6.254,	-0.042]&[	-1.873,2.282]&[	-1.581,	2.012]\\
+	99.7\% within&	[-7.807,	1.511]&[	-2.912,3.321]&[	-2.479,	2.910]\\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}
+  \includegraphics[scale=0.6]{figure_1.eps}
+  \caption{An example figure}
+  \label{fgr:figure1}
+\end{figure}
 
@@
 %
-%New float types are automatically set up by the class file.  The
-%means graphics are included as follows (Scheme~\ref{sch:example}).  As
-%illustrated, the float is ``here'' if possible.
-%\begin{scheme}
-%  Your scheme graphic would go here: \texttt{.eps} format\\
-%  for \LaTeX\, or \texttt{.pdf} (or \texttt{.png}) for pdf\LaTeX\\
-%  \textsc{ChemDraw} files are best saved as \texttt{.eps} files:\\
-%  these can be scaled without loss of quality, and can be\\
-%  converted to \texttt{.pdf} files easily using \texttt{eps2pdf}.\\
-%  %\includegraphics{graphic}
-%  \caption{An example scheme}
-%  \label{sch:example}
-%\end{scheme}peptideIndex.Count: 10465024
-%
-%
-%\begin{figure}
-%  As well as the standard float types \texttt{table}\\
-%  and \texttt{figure}, the class also recognises\\
-%  \texttt{scheme}, \texttt{chart} and \texttt{graph}.
-%  \caption{An example figure}
-%  \label{fgr:example}
-%\end{figure}
-%
-%Charts, figures and schemes do not necessarily have to be labelled or
-%captioned.  However, tables should always have a title. It is
-%possible to include a number and label for a graphic without any
-%title, using an empty argument to the \texttt{\textbackslash caption}
-%macro.
-%
-%The use of the different floating environments is not required, but
-%it is intended to make document preparation easier for authors. In
-%general, you should place your graphics where they make logical
-%sense; the production process will move them if needed.
-%
-%\subsection{Math(s)}
-%
-%The \textsf{achemso} class does not load any particular additional
-%support for mathematics.  If packages such as \textsf{amsmath} are
-%required, they should be loaded in the preamble.  However,
-%the basic \LaTeX\ math(s) input should work correctly without
-%this.  Some inline material \( y = mx + c \) or $ 1 + 1 = 2 $
-%followed by some display. \[ A = \pi r^2 \]
-%
-%It is possible to label equations in the usual way (Eq.~\ref{eqn:example}).
-%\begin{equation}
-%  \frac{\mathrm{d}}{\mathrm{d}x} \, r^2 = 2r \label{eqn:example}
-%\end{equation}
-%This can also be used to have equations containing graphical
-%content. To align the equation number with the middle of the graphic,
-%rather than the bottom, a minipage may be used.
-%\begin{equation}
-%  \begin{minipage}[c]{0.80\linewidth}
-%    \centering
-%    As illustrated here, the width of \\
-%    the minipage needs to allow some  \\
-%    space for the number to fit in to.
-%    %\includegraphics{graphic}
-%  \end{minipage}
-%  \label{eqn:graphic}
-%\end{equation}
-%
-%\section{Experimental}
-%
-%The usual experimental details should appear here.  This could
-%include a table, which can be referenced as Table~\ref{tbl:example}.
-%Notice that the caption is positioned at the top of the table.
-%\begin{table}
-%  \caption{An example table}
-%  \label{tbl:example}
-%  \begin{tabular}{ll}
-%    \hline
-%    Header one  & Header two  \\
-%    \hline
-%    Entry one   & Entry two   \\
-%    Entry three & Entry four  \\
-%    Entry five  & Entry five  \\
-%    Entry seven & Entry eight \\
-%    \hline
-%  \end{tabular}
-%\end{table}
-%
-%Adding notes to tables can be complicated.  Perhaps the easiest
-%method is to generate these using the basic
-%\texttt{\textbackslash textsuperscript} and
-%\texttt{\textbackslash emph} macros, as illustrated (Table~\ref{tbl:notes}).
-%\begin{table}
-%  \caption{A table with notes}
-%  \label{tbl:notes}
-%  \begin{tabular}{ll}
-%    \hline
-%    Header one                            & Header two \\
-%    \hline
-%    Entry one\textsuperscript{\emph{a}}   & Entry two  \\
-%    Entry three\textsuperscript{\emph{b}} & Entry four \\
-%    \hline
-%  \end{tabular}
-%
-%  \textsuperscript{\emph{a}} Some text;
-%  \textsuperscript{\emph{b}} Some more text.
-%\end{table}
-%
 %The example file also loads the optional \textsf{mhchem} package, so
@@
 %
-%The use of new commands should be limited to simple things which will
-%not interfere with the production process.  For example,
-%\texttt{\textbackslash mycommand} has been defined in this example,
-%to give italic, mono-spaced text: \mycommand{some text}.
-
-%\section{Extra information when writing JACS Communications}
-%
-%When producing communications for \emph{J.~Am.\ Chem.\ Soc.}, the
-%class will automatically lay the text \begin{flushleft}
-
-%\end{flushleft}out in the style of the
-%journal. This gives a guide to the length of text that can be
-%accommodated in such a publication. There are some points to bear in
-%mind when preparing a JACS Communication in this way.  The layout
-%produced here is a \emph{model} for the published result, and the
-%outcome should be taken as a \emph{guide} to the final length. The
-%spacing and sizing of graphical content is an area where there is
-%some flexibility in the process.  You should not worry about the
-%space before and after graphics, which is set to give a guide to the
-%published size. This is very dependant on the final published layout.
-%
-%You should be able to use the same source to produce a JACS
-%Communication and a normal article.  For example, this demonstration
-%file will work with both \texttt{type=article} and
-%\texttt{type=communication}. Sections and any abstract are
-%automatically ignored, although you will get warnings to this effect.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The "Acknowledgement" section can be given in all manuscript
-%% classes.  This should be given within the "acknowledgement"
-%% environment, which will make the correct section or running title.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{acknowledgement}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The appropriate \bibliography command should be placed here.
-%% Notice that the class file automatically sets \bibliographystyle
-%% and also names the section correctly.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+\newpage
+
 \bibliography{citations}

### 2017-02-28 GraphGeneration/fig1-SearchTypesDiagram.py

@@ -1,0 +1,50 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+plt.figure(1)
+
+ylabels = ['Reject', 'Accept']
+yTicks = [1,2]
+
+plt.subplot(411)
+plt.plot([0],[2], 'go')
+plt.hlines(1,-100,-2, colors = 'r')
+plt.hlines(1,2,100, colors = 'r')
+plt.ylim([0,3])
+plt.yticks(yTicks, ylabels)
+plt.title("Narrow-Mass Search")
+
+plt.subplot(412)
+plt.plot([0],[2], 'go')
+plt.plot([16],[2], 'go')
+plt.plot([-17],[2], 'go')
+plt.plot([42],[2], 'go')
+plt.hlines(1,-100,-19, colors = 'r')
+plt.hlines(1,-15,-2, colors = 'r')
+plt.hlines(1,2,14, colors = 'r')
+plt.hlines(1,18,40, colors = 'r')
+plt.hlines(1,44,100, colors = 'r')
+plt.ylim([0,3])
+plt.yticks(yTicks, ylabels)
+plt.title("Notch Search")
+
+plt.subplot(413)
+plt.hlines(1,-100,-57, colors = 'r')
+plt.plot([-16],[1], 'ro')
+plt.plot([57],[1], 'ro')
+plt.hlines(2,-57,-18, colors = 'g')
+plt.hlines(2,-14,55, colors = 'g')
+plt.hlines(2,59,100, colors = 'g')
+plt.ylim([0,3])
+plt.yticks(yTicks, ylabels)
+plt.title("Interval Search")
+
+plt.subplot(414)
+plt.hlines(2,-100,100, colors = 'g')
+plt.ylim([0,3])
+plt.yticks(yTicks, ylabels)
+plt.title("Wide-Mass Search")
+
+plt.tight_layout()
+
+plt.show()

### 2017-02-28 GraphGeneration/fig2-LimitsOnOpenSearch.py

@@ -1,0 +1,94 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+offset = 200
+
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateOpenSearch.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateOpenSearch.psmtsv", sep='\t')
+
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+
+A=data_sorted['MassDiff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+print(A[0])
+print(A[-1])
+
+AhighConf = A[C<0.01]
+BhighConf = B[C<0.01]
+
+print(AhighConf[0])
+print(AhighConf[-1])
+
+print(BhighConf[0])
+print(BhighConf[1])
+print(BhighConf[2])
+print(BhighConf[3])
+print(BhighConf[4])
+
+totalCount = len(AhighConf)
+   
+forward = [0]
+forwardDiv = []
+forwardInd= []
+for i in range(totalCount):
+    forward.append(forward[-1]+1 if BhighConf[i]=='D' else forward[-1])
+    forwardDiv.append(forward[-1]/(i+1))
+    forwardInd.append(AhighConf[i])
+
+print(forwardDiv[0])
+print(forwardDiv[1])
+print(forwardDiv[2])
+print(forwardDiv[3])
+print(forwardDiv[4])
+print(forwardDiv[5])
+print(forwardDiv[6])
+print(forwardDiv[7])
+print(forwardDiv[8])
+
+ok = plt.subplot(121)
+
+plt.scatter(forwardInd[offset:totalCount-offset:100], forwardDiv[offset:totalCount-offset:100])
+plt.xscale('symlog')
+plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
+plt.ylabel("FDR")
+plt.xlabel("Mass Diff")
+plt.title("Limit From Below")
+
+
+backward = [0]
+backwardDiv= []
+backwardInd= []
+for i in range(totalCount-1,-1,-1):
[…]

### 2017-02-28 GraphGeneration/fig3-Histograms.py

@@ -1,0 +1,63 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+
+massDiffs=data_sorted['MassDiff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(122)
+
+halfWidth = 0.01
+min1 = 113.08406-halfWidth
+max1 = 113.08406+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+print(massDiffsHighConf1)
+
+
+a = plt.hist(massDiffsHighConf1, bins  = 25, range=[min1, max1], label = "Targets")
+b = plt.hist(massDiffsHighConf1decoy, bins  =25, range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+plt.xlabel("Mass Diff")
+plt.ylabel("Count")
+plt.title("PSMs around 113.084")
+plt.annotate('Leucine Mass', xy=(113.08406, 50), xytext=(113.074, 40),
+            arrowprops=dict(facecolor='black', shrink=0.05))
+ 
+ok = plt.subplot(121)
+
+halfWidth = 0.01
+min1 = 79.961573-halfWidth
+max1 = 79.961573+halfWidth
+
+massDiffsHighConf2 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-02-28 GraphGeneration/test.py

@@ -1,0 +1,74 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+fileName = r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-13-26-00\02-14-17_Cl-1_rep1beforesc1.ms1dptsv"
+
+nice = np.loadtxt(fileName, skiprows=1, unpack = True)
+
+print(sum(nice[-1])/len(nice[-1]))
+
+plt.figure(1)
+plt.subplot(231)
+plt.plot(nice[0],nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(nice[0]), max(nice[0]), zorder = 10)
+plt.xlabel("mz")
+plt.subplot(232)
+plt.plot(nice[1],nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(nice[1]), max(nice[1]), zorder = 10)
+plt.xlabel("rt")
+plt.subplot(233)
+plt.plot(np.log(nice[2]),nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(np.log(nice[2])), max(np.log(nice[2])), zorder = 10)
+plt.xlabel("log(intensity)")
+plt.subplot(234)
+plt.plot(np.log(nice[3]),nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(np.log(nice[3])), max(np.log(nice[3])), zorder = 10)
+plt.xlabel("log(TIC)")
+plt.subplot(235)
+plt.plot(np.log(nice[4]),nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(np.log(nice[4])), max(np.log(nice[4])), zorder = 10)
+plt.xlabel("log(injection time)")
+
+if len(nice)==7:
+    plt.subplot(236)
+    plt.plot(nice[5],nice[-1], 'bo',markersize = .3)
+    plt.hlines(0, min(nice[5]), max(nice[5]), zorder = 10)
+    plt.xlabel("isolation mz")
+
+
+plt.figure(2)
+
+fileName = r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-13-26-00\02-14-17_Cl-1_rep1afterfc2after.ms1dptsv"
+
+nice = np.loadtxt(fileName, skiprows=1, unpack = True)
+
+print(sum(nice[-1])/len(nice[-1]))
+
+plt.subplot(231)
[…]

### 2017-02-28 GraphGeneration/test2.py

@@ -1,0 +1,27 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+fileName = r"C:\Users\stepa\Source\CalibrationPaper\GraphGeneration\ppm.txt"
+
+nice = np.genfromtxt(fileName, delimiter='\t', unpack = True)
+plt.figure(1)
+
+plt.subplot(131)
+A = nice[0]
+plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
+plt.ylim([0,2000])
+plt.title("prior")
+
+plt.subplot(132)
+A = nice[1]
+plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
+plt.ylim([0,2000])
+plt.title("after smooth")
+
+plt.subplot(133)
+A = nice[2]
+plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
+plt.ylim([0,2000])
+plt.title("after rf")
+
+plt.show()

### 2017-02-28 GraphGeneration/test3.py

@@ -1,0 +1,74 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+maxi = 15000
+
+#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
+#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
+
+#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
+#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
+
+#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Br
+#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Br
+
+#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-01-16\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Cl
+ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-27-09-26-26\Task2Search\aggregate20ppmAroundZero.psmtsv", sep='\t')  #Cl
+
+A=ok[r"MassDiff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue_notch"]
+
+plt.figure(1)
+
+ok = plt.subplot(131)
+plt.hist(A[C<0.01], bins  = 101, range=[-20.2,20.2])
+plt.ylim([1,maxi])
+plt.vlines(0,0,maxi,zorder = 10)
[…]

### 2017-02-28 paper.tex
fixes

@@
 5ppm, or 2.1 Da).
-An \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e.
+A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e.
 $\pm 500$ Da) between the observed precursor mass and the peptide mass.
@@
 The wide-mass search is replaced by a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
-This change improves the specificity and significantly decreases search.
+This change improves the specificity and significantly decreases search time.
 The final narrow-mass search is also replaced by a notch search, further increasing the number of confidently identified peptides.
@@
 Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on a wide-mass or an open-mass search.
-We demonstrate that using a combination of specialized notch and interval searches produces superior identification and discovery results.
+We demonstrate that a combination of specialized notch and interval searches produces superior identification and discovery results.
 
@@
 
-\subsection{Enhanced G-PTM-D}
+\subsection{Enhanced G~-PTM~-D}
 
@@
 Using an off-the shelf database of modification is not advisable.
-The database augmentation step is like the procedure described in the original G-PTM-D work: modifications with appropriate mass are added to the protein database, so that subsequent searches may include them as variable localized modifications.
+The database augmentation step is like the procedure described in the original G~-PTM~-D work: modifications with appropriate mass are added to the protein database, so that subsequent searches may include them as variable localized modifications.
 Since the XML database is annotated with PTMs from the UniProt PTM list, it is a natural first idea to use it in the augmentation step as well.
@@
 
-By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PTMs increased by 8\%.
+By employing the novel notch search strategy instead of the wide-mass search in the G~-PTM~-D process, using only the modifications listed in Uniprot, the number of confidently identified PTMs increased by 8\%.
 The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-label}.
@@
 We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Daltons.
-Algorithmic improvements in the G-PTM-D procedure itself, such as searching for combinations of modifications and for modifications with neutral losses make it an excellent tool for PTM discovery.
+Algorithmic improvements in the G~-PTM~-D procedure itself, such as searching for combinations of modifications and for modifications with neutral losses make it an excellent tool for PTM discovery.
 

### 2017-02-28 paper.tex
ok

@@
 
-\subsection{Enhanced G~-PTM~-D}
-
 The input into the workflow is a standard bottom-up tandem-MS dataset obtained from a tryptically digested protein sample along with a corresponding protein sequence database, and the output is a comprehensive set of peptide and PTM identifications.
 
-Even though adducts arising from experimental artefacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS.
-A curated set of notches and the corresponding modifications is given in Appendix A.
-Using an off-the shelf database of modification is not advisable.
-The database augmentation step is like the procedure described in the original G~-PTM~-D work: modifications with appropriate mass are added to the protein database, so that subsequent searches may include them as variable localized modifications.
-Since the XML database is annotated with PTMs from the UniProt PTM list, it is a natural first idea to use it in the augmentation step as well.
-It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many PTMs are missing, including adducts and glycosylations.
-The final search with the augmented database is conducted with notches: the final notch search is specific to each dataset: we only recommend including the common missed monoisotopic mass values in the final search: see the following section for justification.
-Newly identified localized PTMs are now readily identified.
+\subsection{Enhanced G-PTM-D}
 
-By employing the novel notch search strategy instead of the wide-mass search in the G~-PTM~-D process, using only the modifications listed in Uniprot, the number of confidently identified PTMs increased by 8\%.
+By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PSMs increased by 6\%.
 The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-label}.
-Using an expanded modification list that includes chemical modifications such as adducts and large mass PTMs such as glycosylations, the overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%.
-We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Daltons.
-Algorithmic improvements in the G~-PTM~-D procedure itself, such as searching for combinations of modifications and for modifications with neutral losses make it an excellent tool for PTM discovery.
 
@@
 \centering
-\caption{Comparison of Wide-Mass and Notch Searches}
+\caption{Effects of Replacing Initial Wide-Mass Search by a Notch Search}
 \label{my-label}
+\begin{tabular}{ll|l|l}
+                      &        & Wide-Mass Search & Notch Search\\
+\hline
+\multirow{2}{*}{Time} & Jurkat & 22.1 hrs         & 2.9 hrs    \\
+                      & Mouse  & 13.5 hrs         & 1.2 hrs   \\
+\hline
+\multirow{2}{*}{PSMs} & Jurkat & 203237           & 210566    \\
+                      & Mouse  & 153674           & 162473   
+\end{tabular}
+\end{table}
+
+The final search with the augmented database is conducted with notches: the final notch search is specific to each dataset: we only recommend including the common missed monoisotopic mass values in the final search, the effect of using these notches is shown in Table~\ref{tab:table2}.
+
+\begin{table}[]
+\centering
+\caption{Effects of Replacing Final Narrow-Mass Search by a Notch Search}
+\label{tab:table2}
+\begin{tabular}{ll|l|l}
+                      &        & Narrow-Mass Search & Notch Search\\
+\hline
+\multirow{2}{*}{PSMs} & Jurkat  & 210566   &  220574  \\
+                      & Mouse    & 162473   &   170743
+\end{tabular}
+\end{table}
+
+Since the XML database is annotated with PTMs from the UniProt PTM list, it is natural to use it in the augmentation step as well.
+It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many PTMs are missing, including adducts and glycosylations.
+Notch searches allow using expanded modification lists that include chemical modifications such as adducts and large mass PTMs such as glycosylations.
+Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS.
+A curated set of notches and the corresponding modifications is given in Appendix A.
+The overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%.
+We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Daltons.
+
+\begin{table}[]
+\centering
+\caption{Effects of Using Different Modification Lists}
+\label{tab:table3}
 \begin{tabular}{ll|l|l|l}
-                      &        & Wide-Mass Search & Notch Search & Notch Search With Expanded Mod List \\
+                      &        & Only Uniprot & Uniprot+Glyco & Uniprot+Glyco+Adducts\\
 \hline
-\multirow{2}{*}{Time} & Jurkat & 22.1 hrs         & 2.9 hrs   &  4.5 hrs  \\
-                      & Mouse  & 13.5 hrs         & 1.2 hrs   &  1.8 hrs \\
-\hline
-\multirow{2}{*}{PSMs} & Jurkat & 203237           & 210566   &  220574  \\
-                      & Mouse  & 153674           & 162473   &   170743
+\multirow{2}{*}{PSMs} & Jurkat  & 220574   &  222985 & 223578\\
+                      & Mouse    & 170743   &   171432& 174783 
 \end{tabular}
@@
 
-Mass windows provide an additional mechanism in improving confident identification rates in a wide mass search.
+Mass windows provide an mechanism to exclude these problematic mass shifts from a wide-mass search by employing a interval search.
 
@@
 
-The proposed notch search strategy can be used for discovery of novel modifications not present in a database.
 We describe the heuristic approach for augmenting the modification database and the final notch list.
-The discovery heuristic starts with conducting a modified comb interval search of a calibrated file.
-The highly non-uniform shape of the allowable mass differences to search is motivated by the Open Search Case Study in the following section.
-This allows us to pre-emptively restrict the database search to only look for peptides in intervals that satisfy the condition of being within the specified search windows.
-This alternative to the open search improves search times and increases specificity.
-After the conclusion of the search, the PSMs within 1\% FDR are selected, and the corresponding mass differences are grouped.
-Then the peaks (mass difference values with multiple identifications) in this histogram are identified.
-We then attempt to categorize these peaks into one of the following categories:
+After the conclusion of the search, the PSMs within 1\% FDR are selected, and the corresponding mass differences are grouped into peaks, which are categorized into one of the following categories:
 \begin{itemize}
 \item \textit{No mass error} Exact match (within a certain noise tolerance) between the peptide corresponding to the MS2 fragmentation and the reported precursor monoisotopic mass.
-\item \textit{Monoisotopic error} Peaks around values such as 1.003, 2.006, 3.009 (multiples of the difference in mass between C12 and C13).
+\item \textit{Monoisotopic error} Peaks around values such as 1.003, 2.006, 3.009 (average differences in mass between the monoisotopic peaks and the first few isotope peks).
 These correspond to having misidentified the monoisotopic peak in the preprocessing deconvolution step.

### 2017-02-28 GraphGeneration/test3.py

@@ -15,6 +15,7 @@ maxi = 15000
 #ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-01-16\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Cl
 ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-27-09-26-26\Task2Search\aggregate20ppmAroundZero.psmtsv", sep='\t')  #Cl
 
+
 A=ok[r"MassDiff (ppm)"]
 B=ok[r"Decoy/Contaminant/Target"]=="T"
 C=ok[r"QValue_notch"]
@@ -44,31 +45,3 @@ plt.title("Decoy")
 #ok.set_yscale("log", nonposy='clip')
 
 plt.show()
-##
-##
-##fileName = r"C:/Users/stepa/Desktop/ppm.txt"
-##
-##nice = np.genfromtxt(fileName, delimiter='\t', unpack = True)
-##plt.figure(1)
-##
-##plt.subplot(131)
-##A = nice[0]
-##plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
-##plt.ylim([0,2000])
-##plt.vlines(0,0,2000,zorder = 10)
-##plt.title("prior")
-##
-####plt.subplot(132)
-####A = nice[1]
-####plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
-####plt.ylim([0,2000])
-####plt.title("after smooth")
-##
-##plt.subplot(122)
-##A = nice[2]
-##plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
-##plt.ylim([0,2000])
-##plt.vlines(0,0,2000,zorder = 10)
-##plt.title("after rf")
-##
-##plt.show()

### 2017-02-28 paper.tex
okay

@@
 
-Correctly identifying protein posttranslational-modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
-Methods to identify and localize PTMs from complex biological samples have been limited, but a few techniques are emerging.
+Correct identification of protein posttranslational-modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
 G-PTM-D\cite{Li_2016} is a recently developed tool for global identification of PTMs.
-Spectra file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of new PTM identification.
-We enhanced G-PTM-D by using non-standard mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
-The number of identified modifications increased by 20\% and the search time decreased by an order of magnitude.
-We also show the advantages of using mass-window searches in other contexts, when PTM discovery is not the primary goal.
+Spectral file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of PTM identification.
+We enhanced G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
+The proposed changes lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
+Additional advantages of using targeted mass-windows searches in context of novel modification discovery are shown.
 \end{abstract}

### 2017-02-28 paper.tex
changes

@@
 
-We have recently described a bioinformatics tool, G-PTM-D, for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets\cite{Li_2016}.
+A recently described bioinformatics tool for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets, G-PTM-D\cite{Li_2016} is the foundation for the work described in this paper.
 The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015,Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
@@
 
-Before describing the proposed changes to G-PTM-D, we introduce the database search mode definitions we use throughout.
+Before describing the proposed changes to G-PTM-D, a few definitions of the various search modes must be introduced.
 A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e.
@@
 
-We recommend using the mass-window search in other procedures as well, beyond the G-PTM-D workflow.
+The mass-window search is useful in other procedures as well, beyond the G-PTM-D workflow.
 Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on a wide-mass or an open-mass search.
-We demonstrate that a combination of specialized notch and interval searches produces superior identification and discovery results.
+A combination of specialized notch and interval searches produces superior identification and discovery results.
 
-Along with algorithmic improvements, we recommend using spectral calibration prior to running the G-PTM-D procedure.
+Along with algorithmic improvements, spectral calibration prior to running the G-PTM-D procedure is essential.
 A critical parameter for peptide identification and PTM localization is mass accuracy\cite{Scherl_2008}.

### 2017-03-01 paper.tex
add more ignored files, fix figure

--- 2017-02-28 paper.tex
@@
 We ran a completely open precursor mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
-The plots in Figure~\ref{fgr:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
+The plots in Figure~\ref{fig:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
 It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.

### 2017-03-01 paper.tex
make sentences on each line

--- 2017-03-01 paper.tex
@@
 
-\title[An \textsf{achemso} demo]
+\title[Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows]
   {Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows}
@@
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database.
-3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g.
-1\% FDR).
+3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR).
 
 Before describing the proposed changes to G-PTM-D, we introduce the database search mode definitions we use throughout.
-A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e.
-5ppm, or 2.1 Da).
-A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e.
-$\pm 500$ Da) between the observed precursor mass and the peptide mass.
+A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).
+A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
 An even less restrictive \textit{open-mass} search, matches peptides from a database to  fragmentation spectra without taking the precursor mass into account.
@@
 
-The final search with the augmented database is conducted with notches: the final notch search is specific to each dataset: we only recommend including the common missed monoisotopic mass values in the final search, the effect of using these notches is shown in Table~\ref{tab:table2}.
+The final search with the augmented database is conducted with notches: we only recommend including the common missed monoisotopic mass values in the final search,.
+The effect of using these notches is shown in Table~\ref{tab:table2}.
 
@@
 
-An automated script analyzes each identified peak, and provides clues to the nature of the peak.
-For every peak, a profile is automatically generated.
+An automated script analyzes each identified peak, and provides clues to the nature of the peak. For every peak, a profile is automatically generated.
 It includes the total number of unique peptides associated with the mass shift, the fraction of decoys, mass match with any known entry in the Unimod or UniProt database, mass match to an amino acid addition/removal combination, mass match to a combination of higher frequency peaks, fraction of localizable targets, localization residues and/or termini, and presence of any modifications in the matched peptides.
@@
 
-The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g.
-10 ppm) database search is performed on the dataset.
+The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g. 10 ppm) database search is performed on the dataset.
 This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy (ref)).
@@
 A listing of the contents of each file supplied as Supporting Information
-should be included. For instructions on what should be included in the
+should be included.
+For instructions on what should be included in the
 Supporting Information as well as how to prepare this material for

### 2017-03-01 paper.tex
fixup

--- 2017-03-01 paper.tex
@@
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database.
-3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g.
-1\% FDR).
+3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR).
 
 Before describing the proposed changes to G-PTM-D, a few definitions of the various search modes must be introduced.
-A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e.
-5ppm, or 2.1 Da).
-A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e.
-$\pm 500$ Da) between the observed precursor mass and the peptide mass.
+A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).
+A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
 An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account.

### 2017-03-01 paper.tex
test
diagram

@@
 \usepackage{multirow}
+\usepackage{subcaption}
 
@@
 
-We propose an extension to G-PTM-D (Enhanced G-PTM-D) that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
+We propose an enhancement to G-PTM-D, that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
 The wide-mass search is replaced by a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
@@
 \begin{figure}
+\end{figure}
+
+\begin{figure}
+\begin{subfigure}[b]{.5\linewidth}
   \includegraphics{figureDifferentSearchModes.eps}
@@
   \label{fgr:differentSearchModes}
+\end{subfigure}%
+\begin{subfigure}[b]{.5\linewidth}
+  \includegraphics[scale=0.5]{diagram.eps}
+  \caption{Workflows}
+  \label{fgr:diagram}
+\end{subfigure}
+\caption{A figure}\label{fig:1}
 \end{figure}
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%% The same is true for Supporting Information, which should use the
-%% suppinfo environment.
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 \begin{suppinfo}

### 2017-03-01 paper.tex
stuff

@@
 
-The input into the workflow is a standard bottom-up tandem-MS dataset obtained from a tryptically digested protein sample along with a corresponding protein sequence database, and the output is a comprehensive set of peptide and PTM identifications.
+We first contrast mass-window searches with narrow-mass and open-mass searches, and demonstrate the advantages of using mass-window searches.
+The we demonstrate the improvements in the performance of the enhanced G-PTM-D over the original proposed methods, that are due to using mass-window searches.
+Finally, we discuss the application of mass-window searches to the discovery of novel modifications.
+
+\subsection{Mass-Window Searches}
+
+\subsubsection{Notch Search vs Narrow-Mass Search}
+
+The purpose of a narrow-mass search is to use the provided precursor mass along with fragmentation spectra information to find an exact match in a protein database.
+
+\subsubsection{Mass Window Searches vs Open-Mass Search}
+
+Restricting the allowed precursor masses is beneficial in searches that are intended to discover mass shift peaks.
+We ran a completely open precursor mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
+The plots in Figure~\ref{fig:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
+It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.
+
+This effect can be explained by the fact that it's easier to match long sequences to short peptide fragmentation profiles than it is to match short sequences to fragmentation pattern of a long peptide.
+
+\begin{figure}
+\caption{Upper and Lower bound cutoffs in Mouse Data}
+\label{fig:figure2-upperlowerbounds}
+\includegraphics[scale=0.5]{figure_2-upperlowerbounds}
+\end{figure}
+
+MetaMorpheus includes a capability to identify peaks in a mass shift histogram obtained from a search, using a recently described peak finding algorithm\cite{Rodriguez_2014}.
+We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
+Some histogram peaks are illustrated in Figure~\ref{fig:figure3}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
+
+\begin{figure}
+\caption{Some Histogram Peaks For Mouse Data}
+\label{fig:figure3}
+\includegraphics[scale=0.6]{figure_3peaks}
+\end{figure}
+
+We identified the property of some specific mass shifts that correspond to anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
+
+Mass windows provide an mechanism to exclude these problematic mass shifts from a wide-mass search by employing a interval search.
 
@@
 \end{table}
-
-\subsection{Mass Windows in an Open Search}
-
-Restricting the allowed precursor masses is beneficial in searches that are intended to discover mass shift peaks.
-We ran a completely open precursor mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
-The plots in Figure~\ref{fig:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
-It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.
-
-This effect can be explained by the fact that it's easier to match long sequences to short peptide fragmentation profiles than it is to match short sequences to fragmentation pattern of a long peptide.
-
-\begin{figure}
-\caption{Upper and Lower bound cutoffs in Mouse Data}
-\label{fig:figure2-upperlowerbounds}
-\includegraphics[scale=0.5]{figure_2-upperlowerbounds}
-\end{figure}
-
-MetaMorpheus includes a capability to identify peaks in a mass shift histogram obtained from a search, using a recently described peak finding algorithm\cite{Rodriguez_2014}.
-We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some histogram peaks are illustrated in Figure~\ref{fig:figure3}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
-
-\begin{figure}
-\caption{Some Histogram Peaks For Mouse Data}
-\label{fig:figure3}
-\includegraphics[scale=0.6]{figure_3peaks}
-\end{figure}
-
-We identified the property of some specific mass shifts that correspond to anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
-
-Mass windows provide an mechanism to exclude these problematic mass shifts from a wide-mass search by employing a interval search.
 

### 2017-03-01 GraphGeneration/HistogramsAroundZero.py

@@ -1,0 +1,156 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+
+massDiffs=data_sorted['MassDiff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(231)
+
+halfWidth = 2.2
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+plt.xlabel("Mass Diff")
+plt.ylabel("Count")
+plt.title("PSMs around zero")
+ok.set_yscale("log", nonposy='clip')
+
+ok = plt.subplot(233)
+
+halfWidth = 0.2
+min1 = 1-halfWidth
+max1 = 1+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+
+
[…]

### 2017-03-01 paper.tex
more text

@@
 The purpose of a narrow-mass search is to use the provided precursor mass along with fragmentation spectra information to find an exact match in a protein database.
+
+One has to be extremely careful in using a narrow-mass search beyond a nominal tolerance of 5 or 10 ppm. 
+While it does provide more identifications, a search within say $\pm 3.1$ Daltons should be done with great care.
+We only recommend such a search after peaks that correspond to modifications/sequence differences are excluded.
 

### 2017-03-02 paper.tex
add some text

@@
                       & Mouse    & 170743   &   171432& 174783 
+\end{tabular}
+\end{table}
+
+The initial notch search is significantly faster than a corresponding wide mass search, and this suggests that multiple search rounds could be conducted, iteratively adding more and more modifications.
+This recursive nature of the algorithm is well suited to add multiple modifications of different types on the same peptide.
+Due to these enhancements, we test the performance of our methods
+
+\begin{table}[]
+\centering
+\caption{My caption}
+\label{my-label}
+\begin{tabular}{|l|l|l|l|l|l|}
+\hline
+                      &        & \multicolumn{2}{l|}{Single Round}                                                                              & \multicolumn{2}{l|}{Multiple Rounds}                                                                           \\ \hline
+                      &        & \begin{tabular}[c]{@{}l@{}}Mods:\\ none\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mods:\\ ptmlist\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mods:\\ none\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mods:\\ ptmlist\end{tabular} \\ \hline
+\multirow{2}{*}{PSMs} & Jurkat &                                                      &                                                         &                                                      &                                                         \\ \cline{2-6}
+                      & Mouse  &                                                      &                                                         &                                                      &                                                         \\ \hline
+\multirow{2}{*}{Mods} & Jurkat &                                                      &                                                         &                                                      &                                                         \\ \cline{2-6}
+                      & Mouse  &                                                      &                                                         &                                                      &                                                         \\ \hline
+\multirow{2}{*}{Time} & Jurkat &                                                      &                                                         &                                                      &                                                         \\ \cline{2-6}
+                      & Mouse  &                                                      &                                                         &                                                      &                                                         \\ \hline
 \end{tabular}

### 2017-03-08 appendixNovelMods.tex
yeah

--- 2017-03-01 paper.tex
@@
-\documentclass[journal=jprobs,manuscript=article]{achemso}
+\documentclass[]{article}
 
-\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
-\usepackage[T1]{fontenc}       % Use modern font encodings
-\usepackage{multirow}
-\usepackage{subcaption}
-
-\newcommand*\mycommand[1]{\texttt{\emph{#1}}}
-
-\author{Stefan K. Solntsev}
-\email{solntsev@wisc.edu}
-\author{Michael R. Shortreed}
-\author{Brian L. Frey}
-\author{Lloyd M. Smith}
-\affiliation[UwMadison]
-{University of Wisconsin-Madison}
-
-\title[Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows]
-  {Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows}
+%opening
+\title{Novel Modification Discovery Heuristic}
+\author{}
 
@@
 
+\maketitle
+
 \begin{abstract}
 
-Correct identification of protein posttranslational-modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
-G-PTM-D\cite{Li_2016} is a recently developed tool for global identification of PTMs.
-Spectral file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of PTM identification.
-We enhanced G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
-The proposed changes lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Additional advantages of using targeted mass-windows searches in context of novel modification discovery are shown.
 \end{abstract}
 
-\section{Introduction}
-
-Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available.
-Information about a modification can include the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
-While this data is almost comprehensive, information about the localization of these modification to certain residues in specific proteins is scarce.
-
-A recently described bioinformatics tool for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets, G-PTM-D\cite{Li_2016} is the foundation for the work described in this paper.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015,Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
-2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database.
-3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR).
-
-Before describing the proposed changes to G-PTM-D, a few definitions of the various search modes must be introduced.
-A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).
-A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
-An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account.
-We propose alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
-The notch search is an extension of the narrow-mass search: it allows exact matches (within a tolerance) to specific mass differences between the theoretical peptide and observed precursor masses.
-The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
-The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum).
-Figure~\ref{fgr:differentSearchModes} provides a convenient way to visualize the various search modes.
-
-We propose an enhancement to G-PTM-D, that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
-The wide-mass search is replaced by a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
-This change improves the specificity and significantly decreases search time.
-The final narrow-mass search is also replaced by a notch search, further increasing the number of confidently identified peptides.
-
-The mass-window search is useful in other procedures as well, beyond the G-PTM-D workflow.
-Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on a wide-mass or an open-mass search.
-A combination of specialized notch and interval searches produces superior identification and discovery results.
-
-Along with algorithmic improvements, spectral calibration prior to running the G-PTM-D procedure is essential.
-A critical parameter for peptide identification and PTM localization is mass accuracy\cite{Scherl_2008}.
-Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
-Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.
-Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories:  External calibration prior to the MS experiment (e.g.
-standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
-real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).
-We use a post-acquisition calibration procedure, that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
-In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
-The increased mass accuracy of the recalibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
-
-\begin{figure}
-\end{figure}
-
-\begin{figure}
-\begin{subfigure}[b]{.5\linewidth}
-  \includegraphics{figureDifferentSearchModes.eps}
-  \caption{Search Modes}
-  \label{fgr:differentSearchModes}
-\end{subfigure}%
-\begin{subfigure}[b]{.5\linewidth}
-  \includegraphics[scale=0.5]{diagram.eps}
-  \caption{Workflows}
-  \label{fgr:diagram}
-\end{subfigure}
-\caption{A figure}\label{fig:1}
-\end{figure}
-
-\section{Experimental Procedures}
-We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the Enhanced G-PTM-D workflow.
-Two multi-fraction mammalian datasets (\textit{Mouse} and \textit{Jurkat}) with deep proteome coverage were used to evaluate the performance of the methods: these datasets are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
-Uniprot XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
-
-A 10ppm precursor mass tolerance was used for the initial calibration step, and then reduced to 5ppm for subsequent notch searches.
-For the database augmentation step, we allowed 120 modifications including PTMs, adducts, and chemical modifications.
-Final notch search used notches centered at \{0, 1.003, 2.005, 3.008\} in order to capture missed identifications due to missed monoisotopic peaks.
-
-For the discovery of novel modifications, we used a combination of two searches: A notch search with notches at all combinations of up to 2 residue mass additions or removals, and an interval search with lower bound of -187 Daltons, and with exclusions at masses corresponding to the notch search.
-
-\section{Results and Discussion}
-
-We first contrast mass-window searches with narrow-mass and open-mass searches, and demonstrate the advantages of using mass-window searches.
-The we demonstrate the improvements in the performance of the enhanced G-PTM-D over the original proposed methods, that are due to using mass-window searches.
-Finally, we discuss the application of mass-window searches to the discovery of novel modifications.
-
-\subsection{Mass-Window Searches}
-
-\subsubsection{Notch Search vs Narrow-Mass Search}
-
-The purpose of a narrow-mass search is to use the provided precursor mass along with fragmentation spectra information to find an exact match in a protein database.
-
-\subsubsection{Mass Window Searches vs Open-Mass Search}
-
-Restricting the allowed precursor masses is beneficial in searches that are intended to discover mass shift peaks.
-We ran a completely open precursor mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
-The plots in Figure~\ref{fig:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
-It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.
-
-This effect can be explained by the fact that it's easier to match long sequences to short peptide fragmentation profiles than it is to match short sequences to fragmentation pattern of a long peptide.
-
-\begin{figure}
-\caption{Upper and Lower bound cutoffs in Mouse Data}
-\label{fig:figure2-upperlowerbounds}
-\includegraphics[scale=0.5]{figure_2-upperlowerbounds}
-\end{figure}
-
-MetaMorpheus includes a capability to identify peaks in a mass shift histogram obtained from a search, using a recently described peak finding algorithm\cite{Rodriguez_2014}.
-We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some histogram peaks are illustrated in Figure~\ref{fig:figure3}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
-
-\begin{figure}
-\caption{Some Histogram Peaks For Mouse Data}
-\label{fig:figure3}
-\includegraphics[scale=0.6]{figure_3peaks}
-\end{figure}
-
-We identified the property of some specific mass shifts that correspond to anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
-
-Mass windows provide an mechanism to exclude these problematic mass shifts from a wide-mass search by employing a interval search.
-
-\subsection{Enhanced G-PTM-D}
-
-By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PSMs increased by 6\%.
-The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-label}.
-
-\begin{table}[]
-\centering
-\caption{Effects of Replacing Initial Wide-Mass Search by a Notch Search}
-\label{my-label}
-\begin{tabular}{ll|l|l}
-                      &        & Wide-Mass Search & Notch Search\\
-\hline
-\multirow{2}{*}{Time} & Jurkat & 22.1 hrs         & 2.9 hrs    \\
-                      & Mouse  & 13.5 hrs         & 1.2 hrs   \\
-\hline
-\multirow{2}{*}{PSMs} & Jurkat & 203237           & 210566    \\
-                      & Mouse  & 153674           & 162473   
-\end{tabular}
-\end{table}
-
-The final search with the augmented database is conducted with notches: we only recommend including the common missed monoisotopic mass values in the final search,.
-The effect of using these notches is shown in Table~\ref{tab:table2}.
-
-\begin{table}[]
-\centering
-\caption{Effects of Replacing Final Narrow-Mass Search by a Notch Search}
-\label{tab:table2}
-\begin{tabular}{ll|l|l}
-                      &        & Narrow-Mass Search & Notch Search\\
-\hline
-\multirow{2}{*}{PSMs} & Jurkat  & 210566   &  220574  \\
-                      & Mouse    & 162473   &   170743
-\end{tabular}
-\end{table}
-
-Since the XML database is annotated with PTMs from the UniProt PTM list, it is natural to use it in the augmentation step as well.
-It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many PTMs are missing, including adducts and glycosylations.
-Notch searches allow using expanded modification lists that include chemical modifications such as adducts and large mass PTMs such as glycosylations.
-Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS.
-A curated set of notches and the corresponding modifications is given in Appendix A.
-The overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%.
-We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Daltons.
-
-\begin{table}[]
-\centering
-\caption{Effects of Using Different Modification Lists}
-\label{tab:table3}
-\begin{tabular}{ll|l|l|l}
-                      &        & Only Uniprot & Uniprot+Glyco & Uniprot+Glyco+Adducts\\
-\hline
-\multirow{2}{*}{PSMs} & Jurkat  & 220574   &  222985 & 223578\\
-                      & Mouse    & 170743   &   171432& 174783 
-\end{tabular}
-\end{table}
-
-\subsection{Discovery of Modifications not in Database}
-
-Previously described wide mass search strategies for discovery and localization of unknown modifications benefit from using alternative, carefully designed search modes based on interval and notch searches.
-We discovered that limiting mass shifts to a lower bound of -187 Da (corresponding to the largest mass difference that could be attributed to loss of a single residue, Tryptophan), and no upper bound, is an important step in eliminating spurious PSMs.
-Furthermore, treating highly suspect mass shifts corresponding to residue additions/substitutions/deletions in a separate notch search with individualized false discovery rate estimates is beneficial.
-An automated tool built into MetaMorpheus allows confidently identifying novel modifications based on these search results.
+\section{The Heuristic}
 
@@
 
-\subsection{Calibration}
-
-The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g. 10 ppm) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy (ref)).
-Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
-The calibration algorithm is described in detail in Supplementary Material 1.
-A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value.
-We withheld 30\% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states.
-The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
-The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively.
-Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
-
-A 10 ppm narrow tolerance search produces matches with mass differences in ppm given in~\ref{fgr:figure1}.
-The average and standard deviations along with 1, 2, and 3 standard deviation widths are given in Table~\ref{tbl:calib}.
-
-\begin{table}
-  \caption{Calibration Results: ppm Mass Differences}
-  \label{tbl:calib}
-  \begin{tabular}{llll}
-    \hline
-    & prior& after smooth & after rf  \\
-    \hline
-    stdev&	1.553	&1.039&	0.898 \\
-	avg&	-3.148&	0.204&0.215\\
-	68.3\% within	& [-4.701,-1.595 ]&	[-0.834,1.243	]&	[-0.683,1.114]\\
-	95.4\% within&[	-6.254,	-0.042]&[	-1.873,2.282]&[	-1.581,	2.012]\\
-	99.7\% within&	[-7.807,	1.511]&[	-2.912,3.321]&[	-2.479,	2.910]\\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \includegraphics[scale=0.6]{figure_1.eps}
-  \caption{An example figure}
-  \label{fgr:figure1}
-\end{figure}
-
-%
-%\subsection{References}
-%
-%The class makes various changes to the way that references are
-%handled.  The class loads \textsf{natbib}, and also the
-%appropriate bibliography style.  References can be made using
-%the normal method; the citation should be placed before any
-%punctuation, as the class will move it if using a superscript
-%citation style
-%\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
-%The use of \textsf{natbib} allows the use of the various citation
-%commands of that package: \citeauthor{Abernethy2003} have shown
-%something, in \citeyear{Cotton1999}, or as given by
-%Ref.~\citenum{Mena2000}.  Long lists of authors will be
-%automatically truncated in most article formats, but not in
-%supplementary information or reviews \cite{Pople2003}. If you
-%encounter problems with the citation macros, please check that
-%your copy of \textsf{natbib} is up to date. The demonstration
-%database file \texttt{achemso-demo.bib} shows how to complete
-%entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
-%using the \texttt{\textbackslash latin} command.
-%
-%Multiple citations to be combined into a list can be given as
-%a single citation.  This uses the \textsf{mciteplus} package
-%\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
-%Citations other than the first of the list should be indicated
-%with a star. If the \textsf{mciteplus} package is not installed,
-%the standard bibliography tools will still work but starred
-%references will be ignored. Individual references can be referred
-%to using \texttt{\textbackslash mciteSubRef}:
-%``ref.~\mciteSubRef{Eisenstein2005}''.
-%
-%The class also handles notes to be added to the bibliography.  These
-%should be given in place in the document \bibnote{This is a note.
-%The text will be moved the the references section.  The title of the
-%section will change to ``Notes and References''.}.  As with
-%citations, the text should be placed before punctuation.  A note is
-%also generated if a citation has an optional note.  This assumes that
-%the whole work has already been cited: odd numbering will result if
-%this is not the case \cite[p.~1]{Cotton1999}.
-
-%\newpage
-%
-%\subsection{Floats}
-%
-%The example file also loads the optional \textsf{mhchem} package, so
-%that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
-%gives \ce{H2SO4}.  See the use in the bibliography file (when using
-%titles in the references section).
-%
-\begin{acknowledgement}
-
-The authors thank \ldots
-\end{acknowledgement}
-
-\begin{suppinfo}
-
-A listing of the contents of each file supplied as Supporting Information
-should be included.
-For instructions on what should be included in the
-Supporting Information as well as how to prepare this material for
-publications, refer to the journal's Instructions for Authors.
-
-The following files are available free of charge.
-\begin{itemize}
-  \item Filename: brief description
-  \item Filename: brief description
-\end{itemize}
-
-\end{suppinfo}
-
-\newpage
-
-\bibliography{citations}
-
 \end{document}

### 2017-03-08 paper.tex
oka

@@
 Correct identification of protein posttranslational-modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
-G-PTM-D\cite{Li_2016} is a recently developed tool for global identification of PTMs.
+G-PTM-D\cite{Li_2016} is a recently developed tool for global identification and localization of PTMs.
 Spectral file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of PTM identification.
@@
 The proposed changes lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Additional advantages of using targeted mass-windows searches in context of novel modification discovery are shown.
+Targeted mass-windows searches are also shown to be useful as a replacement to a standard wide-mass search, resulting in a 10\% increase in the number of confidently identified peptides: this leads to improvements in novel modification discovery heuristics originally based on the wide-mass search.
 \end{abstract}
@@
 
-A recently described bioinformatics tool for the discovery and localization of new PTMs from "bottom-up" tandem mass spectrometric datasets, G-PTM-D\cite{Li_2016} is the foundation for the work described in this paper.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015,Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
+Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, e.g. MODa\cite{Na_2011}.
+G-PTM-D\cite{Li_2016}, a recently described bioinformatics tool for the discovery and localization of new PTMs, is the foundation for the work described in this paper.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database.
@@
 
-Before describing the proposed changes to G-PTM-D, a few definitions of the various search modes must be introduced.
-A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).
-A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
-An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account.
-We propose alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
-The notch search is an extension of the narrow-mass search: it allows exact matches (within a tolerance) to specific mass differences between the theoretical peptide and observed precursor masses.
-The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
-The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum).
-Figure~\ref{fgr:differentSearchModes} provides a convenient way to visualize the various search modes.
+As described, G-PTM-D has a few significant limitations.
+The initial wide-mass search procedure is slow, taking hours for modest size datasets.
+Due to this, high mass modifications, such as glycosolations are out of reach, since widening the search window from the suggested $\pm 200$ Daltons increases the search time even more.
+Furthermore, in G-PTM-D, modification discovery is limited to PTMs in the Uniprot database. 
+Finally, PTMs that are similar in mass, such as Phosphorylation and Sulfonation are problematic since they are virtually indistinguishable in unprocessed spectra files.
+
+Spectral calibration prior to running the G-PTM-D procedure is essential for distinguishing modifications with similar mass.
+As a side effect, calibration signinficantly increases the overall number of confidently identified peptides and modifications.
 
@@
 
-The mass-window search is useful in other procedures as well, beyond the G-PTM-D workflow.
+The idea of restricting the search space by only allowing certain mass differences between the precursor and the identified peptide mass is useful in other procedures as well, beyond the G-PTM-D workflow.
 Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on a wide-mass or an open-mass search.
@@
 
-Along with algorithmic improvements, spectral calibration prior to running the G-PTM-D procedure is essential.
-A critical parameter for peptide identification and PTM localization is mass accuracy\cite{Scherl_2008}.
-Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
-Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.
-Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories:  External calibration prior to the MS experiment (e.g.
-standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
-real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).
-We use a post-acquisition calibration procedure, that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
-In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
-The increased mass accuracy of the recalibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
-
 \begin{figure}
-\end{figure}
-
-\begin{figure}
-\begin{subfigure}[b]{.5\linewidth}
-  \includegraphics{figureDifferentSearchModes.eps}
-  \caption{Search Modes}
-  \label{fgr:differentSearchModes}
-\end{subfigure}%
-\begin{subfigure}[b]{.5\linewidth}
   \includegraphics[scale=0.5]{diagram.eps}
@@
   \label{fgr:diagram}
-\end{subfigure}
-\caption{A figure}\label{fig:1}
 \end{figure}
@@
 
-We first contrast mass-window searches with narrow-mass and open-mass searches, and demonstrate the advantages of using mass-window searches.
-The we demonstrate the improvements in the performance of the enhanced G-PTM-D over the original proposed methods, that are due to using mass-window searches.
-Finally, we discuss the application of mass-window searches to the discovery of novel modifications.
+Post-acquisition calibration is a useful tool that is shown to increase the number of confidently identified peptides, and enables qualitatively distinguishing modifications with similar mass.
+Then, mass-window searches are contrasted with narrow-mass and open-mass searches.
+The improvements in the performance of the enhanced G-PTM-D over the originally proposed G-PTM-D, that are due to using mass-window searches are shown.
+Finally, the application of mass-window searches to the discovery of novel modifications is discussed.
+
+\subsection{Calibration}
+
+A critical parameter for peptide identification and PTM localization is mass accuracy\cite{Scherl_2008}.
+Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
+Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.
+Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories:  External calibration prior to the MS experiment (e.g.
+standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
+real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).
+We use a post-acquisition calibration procedure, that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
+In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
+The increased mass accuracy of the recalibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
+
+The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g. 10 ppm) database search is performed on the dataset.
+This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy (ref)).
+Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
+The calibration algorithm is described in detail in Supplementary Material 1.
+A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value.
+We withheld 30\% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states.
+The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
+The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively.
+Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
+
+A 10 ppm narrow tolerance search produces matches with mass differences in ppm given in~\ref{fgr:figure1}.
+The average and standard deviations along with 1, 2, and 3 standard deviation widths are given in Table~\ref{tbl:calib}.
+
+\begin{table}
+  \caption{Calibration Results: ppm Mass Differences}
+  \label{tbl:calib}
+  \begin{tabular}{llll}
+    \hline
+    & prior& after smooth & after rf  \\
+    \hline
+    stdev&	1.553	&1.039&	0.898 \\
+	avg&	-3.148&	0.204&0.215\\
+	68.3\% within	& [-4.701,-1.595 ]&	[-0.834,1.243	]&	[-0.683,1.114]\\
+	95.4\% within&[	-6.254,	-0.042]&[	-1.873,2.282]&[	-1.581,	2.012]\\
+	99.7\% within&	[-7.807,	1.511]&[	-2.912,3.321]&[	-2.479,	2.910]\\
+    \hline
+  \end{tabular}
+\end{table}
+
+\begin{figure}
+  \includegraphics[scale=0.6]{figure_1.eps}
+  \caption{An example figure}
+  \label{fgr:figure1}
+\end{figure}
 
 \subsection{Mass-Window Searches}
+
+A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).
+A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
+An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account.
+We propose alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
+The notch search is an extension of the narrow-mass search: it allows exact matches (within a tolerance) to specific mass differences between the theoretical peptide and observed precursor masses.
+The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
+The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum).
+Figure~\ref{fgr:differentSearchModes} provides a convenient way to visualize the various search modes.
+
+\begin{figure}
+  \includegraphics{figureDifferentSearchModes.eps}
+  \caption{Search Modes}
+  \label{fgr:differentSearchModes}
+\end{figure}
 
@@
 
-We describe the heuristic approach for augmenting the modification database and the final notch list.
-After the conclusion of the search, the PSMs within 1\% FDR are selected, and the corresponding mass differences are grouped into peaks, which are categorized into one of the following categories:
-\begin{itemize}
-\item \textit{No mass error} Exact match (within a certain noise tolerance) between the peptide corresponding to the MS2 fragmentation and the reported precursor monoisotopic mass.
-\item \textit{Monoisotopic error} Peaks around values such as 1.003, 2.006, 3.009 (average differences in mass between the monoisotopic peaks and the first few isotope peks).
-These correspond to having misidentified the monoisotopic peak in the preprocessing deconvolution step.
-In this case, the identified peptide is still correct.
-\item \textit{Exact PTM or Adduct Masses}e.g.
-42.010 for Acetylation or 21.981 for Sodium adduct.
-In this analysis, there is no conceptual difference between PTMs and adducts: both are observed simply as mass differences between the identified peptide and the precursor mass.
-These peaks are further classified into
-\begin{itemize}
-\item Localizeable - e.g. Methylation. There is evidence for the modification in the MS2 spectrum: the corresponding peaks are shifted by an appropriate amount. 
-\item Labile - e.g. Sulfonation. There is no evidence of the modification in the fragmentation spectrum. 
-\item Either Localizeable or Labile, e.g. Phosphorylation. Sometimes there is evidence of it in the fragmentation spectrum, and sometimes there is not. 
-\end{itemize}
-\item Amino acid removals, additions or substitutions.
-\item Modification dependent mass shifts - these peaks occur only in presence of certain modifications in the identified peptide.
-	E.g.
--15.995 in presence of an identified oxidation, or -79.966 in presence on an identified phosphorylation.
-\item Combinations of any of the above, e.g. 1.987 for combination of a monoisotopic error and a deamidation. Some of these combinations occur frequently, and it is crucial to account for them. 
-\end{itemize}
-
-An automated script analyzes each identified peak, and provides clues to the nature of the peak. For every peak, a profile is automatically generated.
-It includes the total number of unique peptides associated with the mass shift, the fraction of decoys, mass match with any known entry in the Unimod or UniProt database, mass match to an amino acid addition/removal combination, mass match to a combination of higher frequency peaks, fraction of localizable targets, localization residues and/or termini, and presence of any modifications in the matched peptides.
-We follow the following procedure to classify mass shift peaks. 
-1.	A z score is computed for the fraction of decoys in the peak, comparing with 1\%FDR.
-If it corresponds to p>0.05, we do not include this peak in the analysis, because there is no sufficient confidence it produces legitimate results.
-2.
-
-An attempted localization of the mass difference is automatically performed on every peptide-spectrum match.
-We carefully examine the mass differences with a localization fraction greater than 0.2, and attempt to deduce the chemical formula, specificity sites and positions within a peptide.
-Once a determination has been made, we add the new modification type to the modifications list.
-An example output of this step is:
-
-The improvements of using a Comb search instead of an Open search are two-fold. We summarize the differences in Table 2.
-	Open Search	Comb Search
-Search Time	54.36 hrs	29.66 hrs
-
-The improvement in the discernibility of PTMs is apparent when running a comb search on calibrated vs uncalibrated spectra files.
-Specifically, the discernibility of PTMs with similar mass errors (ones that are only different because of the mass defect).
-This was not really possible previously.
-In the current run, Sulfation and Phosphorylation are readily distinguishable.
-We present the numbers for the example below in the step 5 numerical validation section.
-OPEN SEARCH CASE STUDY
-Introduction of the comb search instead of an open search is motivated by a careful review of the Unimod database.
-Known modifications with mass difference within 200 Daltons have values that are within [-0.1, 0.2] of every integer.
-This interval choice stems from the mass defect in the primary isotope of the common elements.
-PTM combinations also have this property.
-We tested multiple search strategies in a search for one that gives the smallest number of false positives. 
-\subsubsection{Known Modification Confirmation Case Study}
-
-Confirmation of the properties of Sodium Adducts, Phosphorylation and Sulfonation.
-\subsubsection{Glycans Case Study}
-Proteins that contain modifications such as Hexose, HexNAC, and similar ones are interesting.
-The uniprot database only lists some possible sites for such modifications, but does not specify the actual type, thus creating ambiguity regarding the type of the modification.
-The new modification discovery workflow is well suited to identify such proteins.
-Table~\ref{tbl:glyco} lists the automatically identified glyco modification masses and identities, sorted by frequency of occurence.
-
-\begin{table}[]
-\centering
-\caption{Glyco}
-\label{tbl:glyco}
-\begin{tabular}{lll}
-MassShift & Count & UnimodID                        \\
-\hline
-203.079   & 30    & HexNAc                          \\
-162.052   & 18    & Hex                             \\
-876.322   & 10    & Hex(2)HexNAc(2)dHex(1)          \\
-656.228   & 8     & Hex(1)HexNAc(1)NeuAc(1)         \\
-365.132   & 8     & Hex(1)HexNAc(1)                 \\
-1216.422  & 8     & Hex(5)HexNAc(2)                 \\
-730.265   & 6     & Hex(2)HexNAc(2)                 \\
-1378.475  & 4     & Hex(6)HexNAc(2)                 \\
-947.323   & 4     & Hex(1)HexNAc(1)NeuAc(2)         \\
-963.317   & 3     & Hex(1)HexNAc(1)NeuAc(1)NeuGc(1) \\
-178.047   & 3     & Galactosyl                      \\
-1038.375  & 3     & dHex(1)Hex(3)HexNAc(2)          \\
-146.058   & 2     & dHex                            \\
-340.1     & 2     & Glucosylgalactosyl              \\
-406.158   & 2     & HexNAc(2)                       \\
-697.256   & 2     & HexNAc(2)NeuAc(1)               \\
-672.223   & 2     & Hex(1)HexNAc(1)NeuGc(1)         \\
-860.326   & 2     & Hex(1)HexNAc(2)dHex(2)         
-\end{tabular}
-\end{table}
-
-\subsubsection{Novel Modification Case Study}
-
-\subsection{Calibration}
-
-The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g. 10 ppm) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy (ref)).
-Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
-The calibration algorithm is described in detail in Supplementary Material 1.
-A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value.
-We withheld 30\% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states.
-The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
-The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively.
-Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
-
-A 10 ppm narrow tolerance search produces matches with mass differences in ppm given in~\ref{fgr:figure1}.
-The average and standard deviations along with 1, 2, and 3 standard deviation widths are given in Table~\ref{tbl:calib}.
-
-\begin{table}
-  \caption{Calibration Results: ppm Mass Differences}
-  \label{tbl:calib}
-  \begin{tabular}{llll}
-    \hline
-    & prior& after smooth & after rf  \\
-    \hline
-    stdev&	1.553	&1.039&	0.898 \\
-	avg&	-3.148&	0.204&0.215\\
-	68.3\% within	& [-4.701,-1.595 ]&	[-0.834,1.243	]&	[-0.683,1.114]\\
-	95.4\% within&[	-6.254,	-0.042]&[	-1.873,2.282]&[	-1.581,	2.012]\\
-	99.7\% within&	[-7.807,	1.511]&[	-2.912,3.321]&[	-2.479,	2.910]\\
-    \hline
-  \end{tabular}
-\end{table}
-
-\begin{figure}
-  \includegraphics[scale=0.6]{figure_1.eps}
-  \caption{An example figure}
-  \label{fgr:figure1}
-\end{figure}
-
-%
-%\subsection{References}
-%
-%The class makes various changes to the way that references are
-%handled.  The class loads \textsf{natbib}, and also the
-%appropriate bibliography style.  References can be made using
-%the normal method; the citation should be placed before any
-%punctuation, as the class will move it if using a superscript
-%citation style
-%\cite{Mena2000,Abernethy2003,Friedman-Hill2003,EuropeanCommission2008}.
-%The use of \textsf{natbib} allows the use of the various citation
-%commands of that package: \citeauthor{Abernethy2003} have shown
-%something, in \citeyear{Cotton1999}, or as given by
-%Ref.~\citenum{Mena2000}.  Long lists of authors will be
-%automatically truncated in most article formats, but not in
-%supplementary information or reviews \cite{Pople2003}. If you
-%encounter problems with the citation macros, please check that
-%your copy of \textsf{natbib} is up to date. The demonstration
-%database file \texttt{achemso-demo.bib} shows how to complete
-%entries correctly. Notice that ``\latin{et al.}'' is auto-formatted
-%using the \texttt{\textbackslash latin} command.
-%
-%Multiple citations to be combined into a list can be given as
-%a single citation.  This uses the \textsf{mciteplus} package
-%\cite{Johnson1972,*Arduengo1992,*Eisenstein2005,*Arduengo1994}.
-%Citations other than the first of the list should be indicated
-%with a star. If the \textsf{mciteplus} package is not installed,
-%the standard bibliography tools will still work but starred
-%references will be ignored. Individual references can be referred
-%to using \texttt{\textbackslash mciteSubRef}:
-%``ref.~\mciteSubRef{Eisenstein2005}''.
-%
-%The class also handles notes to be added to the bibliography.  These
-%should be given in place in the document \bibnote{This is a note.
-%The text will be moved the the references section.  The title of the
-%section will change to ``Notes and References''.}.  As with
-%citations, the text should be placed before punctuation.  A note is
-%also generated if a citation has an optional note.  This assumes that
-%the whole work has already been cited: odd numbering will result if
-%this is not the case \cite[p.~1]{Cotton1999}.
-
-%\newpage
-%
-%\subsection{Floats}
-%
-%The example file also loads the optional \textsf{mhchem} package, so
-%that formulas are easy to input: \texttt{\textbackslash ce\{H2SO4\}}
-%gives \ce{H2SO4}.  See the use in the bibliography file (when using
-%titles in the references section).
 %
@@
 \begin{suppinfo}
-
-A listing of the contents of each file supplied as Supporting Information
-should be included.
-For instructions on what should be included in the
-Supporting Information as well as how to prepare this material for
-publications, refer to the journal's Instructions for Authors.
 

### 2017-03-08 GraphGeneration/HistogramsAround80and113.py

@@ -16,9 +16,9 @@ TDhighConf = B[C<0.01]
 
 ok = plt.subplot(122)
 
-halfWidth = 0.01
-min1 = 113.08406-halfWidth
-max1 = 113.08406+halfWidth
+halfWidth = 0.05
+min1 = 21.96-halfWidth
+max1 = 21.96+halfWidth
 
 massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
 massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
@@ -26,8 +26,10 @@ massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiff
 print(massDiffsHighConf1)
 
 
-a = plt.hist(massDiffsHighConf1, bins  = 25, range=[min1, max1], label = "Targets")
-b = plt.hist(massDiffsHighConf1decoy, bins  =25, range=[min1, max1], label = "Decoys")
+a = plt.hist(massDiffsHighConf1, bins  = 250, range=[min1, max1], label = "Targets")
+b = plt.hist(massDiffsHighConf1decoy, bins  =250, range=[min1, max1], label = "Decoys")
+
+ok.set_yscale("log", nonposy='clip')
 
 plt.legend()
 

### 2017-03-08 GraphGeneration/HistogramsAroundZero.py

@@ -2,7 +2,9 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
 
 data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
@@ -14,9 +16,9 @@ C=data_sorted['QValue_notch'].values
 massDiffsHighConf = massDiffs[C<0.01]
 TDhighConf = B[C<0.01]
 
-ok = plt.subplot(231)
+ok = plt.subplot(321)
 
-halfWidth = 2.2
+halfWidth = 2.5
 min1 = -halfWidth
 max1 = halfWidth
 
@@ -30,93 +32,16 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max
 
 plt.legend()
 
-plt.xlabel("Mass Diff")
+#plt.xlabel("Mass Diff")
 plt.ylabel("Count")
-plt.title("PSMs around zero")
+plt.title("PSMs within $\pm 2.5$")
 ok.set_yscale("log", nonposy='clip')
 
-ok = plt.subplot(233)
-
-halfWidth = 0.2
-min1 = 1-halfWidth
-max1 = 1+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
[…]

### 2017-03-08 GraphGeneration/fig1-SearchTypesDiagram.py

@@ -3,48 +3,67 @@ import matplotlib.pyplot as plt
 
 plt.figure(1)
 
+plt.subplot(411)
+
+dates = [-100,-1,0,1,100]
+values = [1,1,2,1,1]
+
+plt.plot(dates, values, '-')
+
 ylabels = ['Reject', 'Accept']
 yTicks = [1,2]
 
-plt.subplot(411)
-plt.plot([0],[2], 'go')
-plt.hlines(1,-100,-2, colors = 'r')
-plt.hlines(1,2,100, colors = 'r')
-plt.ylim([0,3])
+plt.ylim([0.5,2.5])
 plt.yticks(yTicks, ylabels)
 plt.title("Narrow-Mass Search")
 
 plt.subplot(412)
-plt.plot([0],[2], 'go')
-plt.plot([16],[2], 'go')
-plt.plot([-17],[2], 'go')
-plt.plot([42],[2], 'go')
-plt.hlines(1,-100,-19, colors = 'r')
-plt.hlines(1,-15,-2, colors = 'r')
-plt.hlines(1,2,14, colors = 'r')
-plt.hlines(1,18,40, colors = 'r')
-plt.hlines(1,44,100, colors = 'r')
-plt.ylim([0,3])
-plt.yticks(yTicks, ylabels)
+
+dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
+values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
+
+plt.plot(dates, values, '-')
+
+ylabels = ['Reject', 'Accept']
+yTicks = [1,2]
+
 plt.title("Notch Search")
+plt.ylim([0.5,2.5])
 
+plt.yticks(yTicks, ylabels)
 plt.subplot(413)
-plt.hlines(1,-100,-57, colors = 'r')
-plt.plot([-16],[1], 'ro')
-plt.plot([57],[1], 'ro')
-plt.hlines(2,-57,-18, colors = 'g')
-plt.hlines(2,-14,55, colors = 'g')
-plt.hlines(2,59,100, colors = 'g')
-plt.ylim([0,3])
-plt.yticks(yTicks, ylabels)
-plt.title("Interval Search")
 
-plt.subplot(414)
-plt.hlines(2,-100,100, colors = 'g')
-plt.ylim([0,3])
+
+
+
[…]

### 2017-03-17 paper.tex
oka

@@
 \subsection{Mass-Window Searches}
-
 A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).

### 2017-03-17 GraphGeneration/test2.py

@@ -1,27 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-fileName = r"C:\Users\stepa\Source\CalibrationPaper\GraphGeneration\ppm.txt"
-
-nice = np.genfromtxt(fileName, delimiter='\t', unpack = True)
-plt.figure(1)
-
-plt.subplot(131)
-A = nice[0]
-plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
-plt.ylim([0,2000])
-plt.title("prior")
-
-plt.subplot(132)
-A = nice[1]
-plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
-plt.ylim([0,2000])
-plt.title("after smooth")
-
-plt.subplot(133)
-A = nice[2]
-plt.hist(A[~np.isnan(A)], bins  = 100, range=[-10,10])
-plt.ylim([0,2000])
-plt.title("after rf")
-
-plt.show()

### 2017-03-17 GraphGeneration/test3.py

@@ -1,7 +1,8 @@
 import numpy as np
 import matplotlib.pyplot as plt
 import pandas
-maxi = 15000
+
+maxi = 6000 
 
 #ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
 #ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
@@ -13,35 +14,94 @@ maxi = 15000
 #ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Br
 
 #ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-01-16\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Cl
-ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-27-09-26-26\Task2Search\aggregate20ppmAroundZero.psmtsv", sep='\t')  #Cl
+#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-27-09-26-26\Task2Search\aggregate20ppmAroundZero.psmtsv", sep='\t')  #Cl
+
+
+
+#start = r"C:\Users\stepa\Desktop\QE-HF_mass_cali_tests\2017-03-09-16-07-50\Task1Calibrate\c\03-07-17_YL_3e6_30K10ppmAroundZero"
+#start = r"C:\Users\stepa\Desktop\QE-HF_mass_cali_tests\2017-03-09-16-07-50\Task1Calibrate\d\03-07-17_YL_3e6_60K10ppmAroundZero"
+#start = r"C:\Users\stepa\Desktop\QE-HF_mass_cali_tests\2017-03-09-16-07-50\Task1Calibrate\e\03-07-17_YL_5e6_30K10ppmAroundZero"
[…]

### 2017-03-27 paper.tex
some changes

@@
 
-Correct identification of protein posttranslational-modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
+Correct identification of protein post-translational modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
 G-PTM-D\cite{Li_2016} is a recently developed tool for global identification and localization of PTMs.
-Spectral file calibration prior to applying the tool, and algorithmic improvements in the peptide database search greatly improve the accuracy and efficiency of PTM identification.
-We enhanced G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
+Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search significantly increase the accuracy and efficiency of PTM identification.
+We enhance G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
 The proposed changes lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Targeted mass-windows searches are also shown to be useful as a replacement to a standard wide-mass search, resulting in a 10\% increase in the number of confidently identified peptides: this leads to improvements in novel modification discovery heuristics originally based on the wide-mass search.
+Targeted mass-window searches are also shown to be useful as replacements to standard wide-mass searches, resulting in a 10\% increase in the overall number of confidently identified peptides, and a 5\% increase in the number of mass shifts observed.
+Using a modification discovery heuristic originally based on the wide-mass search, targeted mass-window searches lead to an increase in the number of novel modifications discovered.
 \end{abstract}
@@
 Spectral calibration prior to running the G-PTM-D procedure is essential for distinguishing modifications with similar mass.
-As a side effect, calibration signinficantly increases the overall number of confidently identified peptides and modifications.
+As a side effect, calibration significantly increases the overall number of confidently identified peptides and modifications.
 
-We propose an enhancement to G-PTM-D, that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
+We present an enhancement to G-PTM-D, that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
 The wide-mass search is replaced by a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
@@
 The calibration algorithm is described in detail in Supplementary Material 1.
-A simple test of the calibration quality is to withhold some known peak matches from the inputs of the calibration software, and to determine if they have been shifted closer to the theoretically correct value.
-We withheld 30\% of identifications, and measured the standard deviation and the average errors of the peak differences for the monoisotopic MS1 peaks with known charge states.
+
 The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
@@
 
-\subsection{Enhanced G-PTM-D}
+\subsection{G-PTM-D}
 

### 2017-04-13 paper.tex
many

@@
 \usepackage{subcaption}
+\usepackage{booktabs}
 
@@
 
-Post-acquisition calibration is a useful tool that is shown to increase the number of confidently identified peptides, and enables qualitatively distinguishing modifications with similar mass.
-Then, mass-window searches are contrasted with narrow-mass and open-mass searches.
-The improvements in the performance of the enhanced G-PTM-D over the originally proposed G-PTM-D, that are due to using mass-window searches are shown.
+The first modification to the G-PTM-D workflow, post-acquisition calibration, is shown to increase the number of confidently identified peptides, and enables qualitatively distinguishing modifications with similar mass.
+The other major enhancement, the mass-window search, is contrasted with the narrow-mass and open-mass search.
+The superior performance of G-PTM-D achieved by the use of mass-window searches is demonstrated.
 Finally, the application of mass-window searches to the discovery of novel modifications is discussed.
@@
 standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
-real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration, REF).
+real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration).
 We use a post-acquisition calibration procedure, that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
-The increased mass accuracy of the recalibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
+The increased mass accuracy of the calibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
 
-The spectral calibration procedure consists of two steps: Step 1: A standard tolerance (e.g. 10 ppm) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy (ref)).
-Step 2:  The identifications from step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
+The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
+This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy \cite{Elias_2007}).
+Step 2:  The identifications from Step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
 The calibration algorithm is described in detail in Supplementary Material 1.
 
-The average mass error went down from -3.67ppm to -0.15ppm in the jurkat dataset and from -3.32ppm for the mouse dataset.
-The standard deviations decreased from 3.64ppm to 2.82ppm, and from 3.07ppm to 2.54ppm for the jurkat and mouse datasets respectively.
-Figure 2 shows the clear improvement in the error for the withheld identified peptides, as a result of applying spectral calibration.
+Calibration quality is accessed through the results of a standard 10ppm precursor tolerance database search.
+PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure, see Table~\ref{tbl:calib}.
+The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
+Furthermore, the systemic noise introduced by various other factors is addressed: this is evident by the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
+Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.
 
-A 10 ppm narrow tolerance search produces matches with mass differences in ppm given in~\ref{fgr:figure1}.
-The average and standard deviations along with 1, 2, and 3 standard deviation widths are given in Table~\ref{tbl:calib}.
-
-\begin{table}
-  \caption{Calibration Results: ppm Mass Differences}
+\begin{table}[]
+\centering
+\caption{PSMs Before and After Calibration}
 \label{tbl:calib}
-  \begin{tabular}{llll}
-    \hline
-    & prior& after smooth & after rf  \\
-    \hline
-    stdev&	1.553	&1.039&	0.898 \\
-	avg&	-3.148&	0.204&0.215\\
-	68.3\% within	& [-4.701,-1.595 ]&	[-0.834,1.243	]&	[-0.683,1.114]\\
-	95.4\% within&[	-6.254,	-0.042]&[	-1.873,2.282]&[	-1.581,	2.012]\\
-	99.7\% within&	[-7.807,	1.511]&[	-2.912,3.321]&[	-2.479,	2.910]\\
-    \hline
+\begin{tabular}{@{}llllll@{}}
+                                                                                &        & \multicolumn{2}{l}{Mouse} & \multicolumn{2}{l}{Jurkat} \\ \cmidrule(l){3-6} 
+                                                                                &        & Orig        & Calib       & Orig        & Calib        \\ \cmidrule(l){3-6} 
+\multicolumn{2}{l}{PSM Count}                                                            &    152208       &        155617 &         217612   &   219172        \\
+\multicolumn{2}{l}{Average Error (ppm)}                                                           &    -1.09         &       -0.16      &      -2.10      &        -0.16      \\
+\multicolumn{2}{l}{St. Dev of Error}                                                     &     3.35        &      0.90       &       2.15      &        0.91      \\
+\multirow{3}{*}{\begin{tabular}[c]{@{}l@{}}Confidence\\ Intervals\end{tabular}} & 68.3\% &   [-4.44,2.27]          &    [-1.06,0.74]          &    [-4.26,0.05]          &  [-1.08,0.75]            \\
+                                                                                & 95.4\% &   [-7.80,5.62]          &    [-1.97,1.64]         &   [-6.41,2.20]          &      [-1.99,1.61]         \\
+                                                                                & 99.7\% &   [-11.15,8.98]          &      [-2.87,2.54]        &    [-8.56,4.36]          &      [-2.90,2.58]         \\ \cmidrule(l){3-6}
 \end{tabular}
@@
 \begin{figure}
-  \includegraphics[scale=0.6]{figure_1.eps}
-  \caption{An example figure}
+  \includegraphics[scale=0.85]{fig2-calibrationQuality.eps}
+  \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
   \label{fgr:figure1}
@@
 
+\newpage
+
 \subsection{Mass-Window Searches}
+
 A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).

### 2017-04-13 GraphGeneration/HistogramsAround80and113.py

@@ -1,65 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-
-massDiffs=data_sorted['MassDiff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-ok = plt.subplot(122)
-
-halfWidth = 0.05
-min1 = 21.96-halfWidth
-max1 = 21.96+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-
-print(massDiffsHighConf1)
-
-
-a = plt.hist(massDiffsHighConf1, bins  = 250, range=[min1, max1], label = "Targets")
-b = plt.hist(massDiffsHighConf1decoy, bins  =250, range=[min1, max1], label = "Decoys")
-
-ok.set_yscale("log", nonposy='clip')
-
-plt.legend()
-
-plt.xlabel("Mass Diff")
-plt.ylabel("Count")
-plt.title("PSMs around 113.084")
-plt.annotate('Leucine Mass', xy=(113.08406, 50), xytext=(113.074, 40),
-            arrowprops=dict(facecolor='black', shrink=0.05))
- 
-ok = plt.subplot(121)
-
-halfWidth = 0.01
-min1 = 79.961573-halfWidth
-max1 = 79.961573+halfWidth
-
-massDiffsHighConf2 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-04-13 GraphGeneration/fig2-calibQuality.py

@@ -1,0 +1,93 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+maxi = 25000
+
+f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row')
+ax1.set_title('Uncalibrated')
+ax2.set_title('Calibrated')
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-12-15-20-37\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+A=ok[r"MassDiff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue_notch"]
+
+ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+ax1.set_ylim([1,maxi])
+ax1.vlines(0,0,maxi,zorder = 10)
+ax1.set_ylabel('PSM count')
+ax1.annotate(r"Mouse", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
+                size='large', ha='right', va='center', xytext=(0,0.5))
+
+mu = np.mean(A[(C<0.01) & B])
+sigma = np.std(A[(C<0.01) & B])
+textToDisplay = 'count = %d\n$\mu=%.2f$\n$\sigma=%.2f$\n68.3%%: [%.2f,%.2f]\n95.4%%: [%.2f,%.2f]\n99.7%%: [%.2f,%.2f]'%(len(A[(C<0.01) & B]), mu, sigma,mu-sigma, mu+sigma,mu-2*sigma, mu+2*sigma,mu-3*sigma, mu+3*sigma)
+print(textToDisplay)
+
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-12-15-21-10\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+
+A=ok[r"MassDiff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue_notch"]
+
+ax2.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
[…]

### 2017-04-13 GraphGeneration/fig3-HistogramsAround80and113.py

@@ -1,0 +1,54 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.mlab as mlab
+import scipy
+import pandas
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+
+massDiffs=data_sorted['MassDiff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+fig, axes = plt.subplots(nrows=1, ncols=2)
+
+halfWidth = 0.01
+center = 113.08406
+min1 = center-halfWidth
+max1 = center+halfWidth
+
+bins = 25
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+axes[0].hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
+axes[0].hist(massDiffsHighConf1decoy, bins  =bins, range=[min1, max1], label = "Decoys")
+
+from scipy.interpolate import UnivariateSpline
+s = massDiffsHighConf1decoy
+p, x = np.histogram(s, bins=bins) # bin it into bins
+x = x[:-1] + (x[1] - x[0])/2   # convert bin edges to centers
+f = UnivariateSpline(x, p, s=222)
+axes[0].plot(x, f(x))
+
+halfWidth = 0.01
+center = 79.961573 
+min1 = center-halfWidth
+max1 = center+halfWidth
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
[…]

### 2017-04-13 GraphGeneration/genericCalibQuality.py

@@ -2,32 +2,15 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-maxi = 6000 
-
-#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
-#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
-
-#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
-#ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t')
-
-#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Br
-#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task3Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Br
-
-#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-01-16\Task1Search\aggregate20ppmAroundZero.psmtsv", sep='\t') #Cl
-#ok = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-27-09-26-26\Task2Search\aggregate20ppmAroundZero.psmtsv", sep='\t')  #Cl
+maxi = 3000
 
 
+#start =  r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task1Calibrate\120426_Jurkat_highLC_Frac19_allPSMs_10ppmAroundZero"
[…]

### 2017-04-13 GraphGeneration/genericCoIsolationHistograms.py

@@ -3,13 +3,17 @@ import matplotlib.pyplot as plt
 import pandas
 
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
 
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
 
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task2Calibrate\2017-03-31-12-50-27\Task1Search\04-30-13_CAST_Frac5_4uL-Calibrated2.1aroundZero.psmtsv", sep='\t')
+
 data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
 
-massDiffs=data_sorted['MassDiff (Da)'].values
+#massDiffs=data_sorted['MassDiff (Da)'].values
+massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
 B=data_sorted['Decoy/Contaminant/Target'].values
 C=data_sorted['QValue_notch'].values
 

### 2017-04-13 GraphGeneration/genericCustomPeakPlot.py

@@ -1,0 +1,43 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.mlab as mlab
+import pandas
+
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Yeast\2017-03-13-14-25-38\Task2Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+
+massDiffs=data_sorted['MassDiff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+center = 57.021464
+fig, axes = plt.subplots(nrows=1, ncols=2)
+
+halfWidth = 0.05
+min1 = center-halfWidth
+max1 = center+halfWidth
+
+bins = 25
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+axes[0].hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
+axes[0].hist(massDiffsHighConf1decoy, bins  =bins, range=[min1, max1], label = "Decoys")
+
+
+halfWidth = 0.01
+min1 = center-halfWidth
+max1 = center+halfWidth
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+axes[1].hist(massDiffsHighConf1, bins  = 25, range=[min1, max1], label = "Targets")
+axes[1].hist(massDiffsHighConf1decoy, bins  =25, range=[min1, max1], label = "Decoys")
+
+
+
+
+plt.show()

### 2017-04-13 GraphGeneration/genericErrorVsVariousParams.py

@@ -1,0 +1,42 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLbeforesc1.ms1dptsv"
+#fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLbeforesc2.ms1dptsv"
+#fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLbeforesc4.ms1dptsv"
+#fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLafterfc2after.ms1dptsv"
+
+
+nice = np.loadtxt(fileName, skiprows=1, unpack = True)
+
+print("average error", sum(nice[-1])/len(nice[-1]))
+
+plt.figure(1)
+plt.subplot(231)
+plt.plot(nice[0],nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(nice[0]), max(nice[0]), zorder = 10)
+plt.xlabel("mz")
+plt.subplot(232)
+plt.plot(nice[1],nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(nice[1]), max(nice[1]), zorder = 10)
+plt.xlabel("rt")
+plt.subplot(233)
+plt.plot(np.log(nice[2]),nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(np.log(nice[2])), max(np.log(nice[2])), zorder = 10)
+plt.xlabel("log(intensity)")
+plt.subplot(234)
+plt.plot(np.log(nice[3]),nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(np.log(nice[3])), max(np.log(nice[3])), zorder = 10)
+plt.xlabel("log(TIC)")
+plt.subplot(235)
+plt.plot(np.log(nice[4]),nice[-1], 'bo',markersize = .3)
+plt.hlines(0, min(np.log(nice[4])), max(np.log(nice[4])), zorder = 10)
[…]

### 2017-04-13 GraphGeneration/genericHistogramsAroundZero.py

@@ -1,0 +1,178 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+
+
+data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
+
+massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(321)
+
+halfWidth = 2.5
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+#plt.xlabel("Mass Diff")
+plt.ylabel("Count")
+plt.title("PSMs within $\pm 2.5$")
+ok.set_yscale("log", nonposy='clip')
+
+ok = plt.subplot(322)
+
+
+
+halfWidth = 0.1
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-04-13 GraphGeneration/test.py

@@ -1,74 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-fileName = r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-13-26-00\02-14-17_Cl-1_rep1beforesc1.ms1dptsv"
-
-nice = np.loadtxt(fileName, skiprows=1, unpack = True)
-
-print(sum(nice[-1])/len(nice[-1]))
-
-plt.figure(1)
-plt.subplot(231)
-plt.plot(nice[0],nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(nice[0]), max(nice[0]), zorder = 10)
-plt.xlabel("mz")
-plt.subplot(232)
-plt.plot(nice[1],nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(nice[1]), max(nice[1]), zorder = 10)
-plt.xlabel("rt")
-plt.subplot(233)
-plt.plot(np.log(nice[2]),nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(np.log(nice[2])), max(np.log(nice[2])), zorder = 10)
-plt.xlabel("log(intensity)")
-plt.subplot(234)
-plt.plot(np.log(nice[3]),nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(np.log(nice[3])), max(np.log(nice[3])), zorder = 10)
-plt.xlabel("log(TIC)")
-plt.subplot(235)
-plt.plot(np.log(nice[4]),nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(np.log(nice[4])), max(np.log(nice[4])), zorder = 10)
-plt.xlabel("log(injection time)")
-
-if len(nice)==7:
-    plt.subplot(236)
-    plt.plot(nice[5],nice[-1], 'bo',markersize = .3)
-    plt.hlines(0, min(nice[5]), max(nice[5]), zorder = 10)
-    plt.xlabel("isolation mz")
-
-
-plt.figure(2)
-
-fileName = r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-13-26-00\02-14-17_Cl-1_rep1afterfc2after.ms1dptsv"
-
-nice = np.loadtxt(fileName, skiprows=1, unpack = True)
-
-print(sum(nice[-1])/len(nice[-1]))
-
-plt.subplot(231)
[…]

### 2017-04-14 paper.tex
fixed

@@
 
-A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e. 5ppm, or 2.1 Da).
+Prior to discussing the impact mass-windows searches have on the G-PTM-D workflow, we need to define the search modes discussed.
+A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e.
+5ppm, or 2.1 Da).
 A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
@@
 The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
-The smaller the fraction of the mass space, and the narrower the mass window per search element, the faster the search will run (as fewer searches are executed per mass spectrum).
 Figure~\ref{fgr:differentSearchModes} provides a convenient way to visualize the various search modes.
@@
 \begin{figure}
-  \includegraphics{figureDifferentSearchModes.eps}
+  \includegraphics[scale=0.7]{figureDifferentSearchModes.eps}
   \caption{Search Modes}
@@
 
-The purpose of a narrow-mass search is to use the provided precursor mass along with fragmentation spectra information to find an exact match in a protein database.
+The purpose of a narrow-mass search is to use the provided precursor mass along with fragmentation spectra information to find an exact match to the observed peptide in a protein database.
 
-One has to be extremely careful in using a narrow-mass search beyond a nominal tolerance of 5 or 10 ppm. 
-While it does provide more identifications, a search within say $\pm 3.1$ Daltons should be done with great care.
-We only recommend such a search after peaks that correspond to modifications/sequence differences are excluded.
+Using a calibration or instrument noise percursor tolerance, such as 5ppm or 10ppm is a valid approach, since it identifies only exact matches.
+A major drawback of this approach is that it misses monoisotope errors and co-isolated peptides.
+To avoid those complications, and to get a more comprehensive list of matches, a wider search tolerance is used, e.g. $\pm 3.5$ Da.
+This gives a substantially larger list of matches, and all the mass shifts could be attributed to co-isolation or missed monoisotopes.
+A post-processing step may be employed to confidently identify the co-isolated and missed monoisotope matches, but a wide range of high-confidence matches with un-identified mass shifts remain.
+These often correspond to modifications on the identified peptides, and require an extensive framework (such as G-PTM-D) to identify and label them.
+More importantly, those modified peptides \textit{do not} correspond to the identified one, thus arguably defeating the purpose of a narrow-mass search.
 
-\subsubsection{Mass Window Searches vs Open-Mass Search}
+We observe that when using an incomplete database that does not include comprehensive modification information, a notch search with notches at monoisotopic peaks is superior to either of the narrow-mass searches.
+
+\begin{table}[]
+\centering
+\caption{Notch Search vs Narrow-Mass Search comparison of PSMs for Mouse}
+\label{my-label}
+\begin{tabular}{@{}llll@{}}
+                    & 5ppm Search & 3.5 Da Search & Notch Search \\ \cmidrule(l){2-4}
+PSMs Within 1\% FDR & 157188      & 178003        & 173483       \\
+MisAssigned PSMs    & 0           & 7783          & 0            \\
+Good PSMs           & 157188      & 170220        & 173483       \\ 
+\end{tabular}
+\end{table}
+
+\subsubsection{Mass-Window Searches vs Open-Mass Search}
 

### 2017-04-14 GraphGeneration/genericHistogramsAroundZero.py

@@ -4,7 +4,9 @@ import pandas
 
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
 
 
 data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)

### 2017-04-19 paper.tex
many

@@
 G-PTM-D\cite{Li_2016} is a recently developed tool for global identification and localization of PTMs.
-Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search significantly increase the accuracy and efficiency of PTM identification.
+Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and efficiency of PTM identification.
 We enhance G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
@@
 \begin{figure}
-  \includegraphics[scale=0.5]{diagram.eps}
+  \includegraphics{diagram.eps}
   \caption{Workflows}
@@
 \begin{figure}
-  \includegraphics[scale=0.85]{fig2-calibrationQuality.eps}
+  \includegraphics{fig2-calibrationQuality.eps}
   \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
@@
 \begin{figure}
-  \includegraphics[scale=0.7]{figureDifferentSearchModes.eps}
+\includegraphics{fig1-searchTypes.eps}
 \caption{Search Modes}
-  \label{fgr:differentSearchModes}
+\label{fig:fig1-searchtypes}
 \end{figure}
@@
 Using a calibration or instrument noise percursor tolerance, such as 5ppm or 10ppm is a valid approach, since it identifies only exact matches.
-A major drawback of this approach is that it misses monoisotope errors and co-isolated peptides.
-To avoid those complications, and to get a more comprehensive list of matches, a wider search tolerance is used, e.g. $\pm 3.5$ Da.
-This gives a substantially larger list of matches, and all the mass shifts could be attributed to co-isolation or missed monoisotopes.
-A post-processing step may be employed to confidently identify the co-isolated and missed monoisotope matches, but a wide range of high-confidence matches with un-identified mass shifts remain.
-These often correspond to modifications on the identified peptides, and require an extensive framework (such as G-PTM-D) to identify and label them.
+A major drawback of this strategy is that it misses monoisotope errors.
+To avoid this complication, and to get a more comprehensive list of matches, a wider search tolerance is often used, e.g. $\pm 3.5$ Da.
+This gives a substantially larger list of matches, and the resulting PSMs with non-zero mass shifts could be attributed to missed monoisotopes or co-isolated peptides.
+While this approach provides a larger list of confident PSMs, a substantial fraction of those is wrong, since certain mass shifts in the chosen Dalton range arise due to sequence variations or PTMs.
+These require an extensive framework (such as G-PTM-D) to identify and label them.
 More importantly, those modified peptides \textit{do not} correspond to the identified one, thus arguably defeating the purpose of a narrow-mass search.
 
-We observe that when using an incomplete database that does not include comprehensive modification information, a notch search with notches at monoisotopic peaks is superior to either of the narrow-mass searches.
+We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the narrow-mass searches.
 
@@
                     & 5ppm Search & 3.5 Da Search & Notch Search \\ \cmidrule(l){2-4}
-PSMs Within 1\% FDR & 157188      & 178003        & 173483       \\
-MisAssigned PSMs    & 0           & 7783          & 0            \\
-Good PSMs           & 157188      & 170220        & 173483       \\ 
+PSMs Within 1\% FDR & 157188      & 178003        &        \\
+MisAssigned PSMs    & 0           &               & 0            \\
+Good PSMs           & 157188      &         &        \\ 
 \end{tabular}
 \end{table}
+
+\begin{figure}
+\caption{Mouse histograms around 0}
+\label{fig:fig3mouse-1012}
+\includegraphics{fig3mouse-1012}
+\end{figure}
+
+\begin{figure}
+\caption{Jurkat histograms around 0}
+\label{fig:fig3jurkat-1012}
+\includegraphics{fig3jurkat-1012}
+\end{figure}
 
@@
 \caption{Upper and Lower bound cutoffs in Mouse Data}
-\label{fig:figure2-upperlowerbounds}
-\includegraphics[scale=0.5]{figure_2-upperlowerbounds}
+\label{fig:fig4-limitsOnOpenSearch}
+\includegraphics{fig4-limitsOnOpenSearch}
 \end{figure}
@@
 \caption{Some Histogram Peaks For Mouse Data}
-\label{fig:figure3}
-\includegraphics[scale=0.6]{figure_3peaks}
+\label{fig:fig5-HistogramsAround80and113}
+\includegraphics{fig5-HistogramsAround80and113}
 \end{figure}

### 2017-04-19 GraphGeneration/fig1-SearchTypesDiagram.py

@@ -1,7 +1,11 @@
 import numpy as np
 import matplotlib.pyplot as plt
 
-plt.figure(1)
+
+
+height = 4
+width = 3.25 
+plt.figure(figsize=(width, height))
 
 plt.subplot(411)
 
@@ -65,5 +69,10 @@ plt.title("Wide-Mass Search")
 
 plt.yticks(yTicks, ylabels)
 plt.tight_layout()
+
+
+plt.savefig('fig1-searchTypes.eps', format='eps', dpi=1200)
+plt.savefig('fig1-searchTypes.png', format='png', dpi=1200)
+
 
 plt.show()

### 2017-04-19 GraphGeneration/fig2-calibQuality.py

@@ -4,7 +4,11 @@ import pandas
 
 maxi = 25000
 
-f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row')
+
+height = 6
+width = 7
+
+f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row', figsize=(width,height))
 ax1.set_title('Uncalibrated')
 ax2.set_title('Calibrated')
 

### 2017-04-19 GraphGeneration/fig3-HistogramsAround80and113.py

@@ -1,54 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.mlab as mlab
-import scipy
-import pandas
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-
-massDiffs=data_sorted['MassDiff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-fig, axes = plt.subplots(nrows=1, ncols=2)
-
-halfWidth = 0.01
-center = 113.08406
-min1 = center-halfWidth
-max1 = center+halfWidth
-
-bins = 25
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-axes[0].hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
-axes[0].hist(massDiffsHighConf1decoy, bins  =bins, range=[min1, max1], label = "Decoys")
-
-from scipy.interpolate import UnivariateSpline
-s = massDiffsHighConf1decoy
-p, x = np.histogram(s, bins=bins) # bin it into bins
-x = x[:-1] + (x[1] - x[0])/2   # convert bin edges to centers
-f = UnivariateSpline(x, p, s=222)
-axes[0].plot(x, f(x))
-
-halfWidth = 0.01
-center = 79.961573 
-min1 = center-halfWidth
-max1 = center+halfWidth
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
[…]

### 2017-04-19 GraphGeneration/fig3-jurkat;-1;0;1;2.py

@@ -1,0 +1,141 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+
+
+
+
+height = 6
+width = 7
+plt.figure(figsize=(width, height))
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+massDiffs=data_sorted['MassDiff (Da)'].values
+
+#data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
+#massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(221)
+
+halfWidth = 3.5
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+plt.ylabel("Count")
+plt.title("PSMs within $\pm 2.5$ Da")
+ok.set_yscale("log", nonposy='clip')
+
+ok = plt.subplot(222)
+#ok.set_yscale("log", nonposy='clip')
+halfWidth = 0.1
+min1 = -1-halfWidth
+max1 = -1+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-04-19 GraphGeneration/fig3-mouse;-1;0;1;2.py

@@ -1,0 +1,143 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task3Gptmd\aggregate_allPSMs_coisolation3.500.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-14-16-50-03\Task1Search\allPSMs_-187andUp.psmtsv", sep='\t')
+
+
+
+
+height = 6
+width = 7
+plt.figure(figsize=(width, height))
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+massDiffs=data_sorted['MassDiff (Da)'].values
+
+#data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
+#massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(221)
+
+halfWidth = 3.5
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+plt.ylabel("Count")
+plt.title("PSMs within $\pm 2.5$ Da")
[…]

### 2017-04-19 GraphGeneration/fig4-limitsOnOpenSearch.py

@@ -2,10 +2,14 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
+height = 3
+width = 7
+plt.figure(figsize=(width, height))
+
+
 offset = 200
 
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateOpenSearch.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateOpenSearch.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-14-16-50-03\Task1Search\allPSMs_OpenSearch.psmtsv", sep='\t')
 
 
 data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
@@ -80,6 +84,7 @@ print(backwardDiv[6])
 print(backwardDiv[7])
 
 
+plt.tick_params(axis='both', which='major', labelsize=8)
 
 ok = plt.subplot(122)
 
@@ -90,5 +95,14 @@ plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
 plt.ylabel("FDR")
 plt.xlabel("Mass Diff")
 plt.title("Limit From Above")
+
+plt.tick_params(axis='both', which='major', labelsize=8)
+
+plt.tight_layout() 
  
+
+plt.savefig('fig4-limitsOnOpenSearch.eps', format='eps', dpi=1200)
+plt.savefig('fig4-limitsOnOpenSearch.png', format='png', dpi=1200)
+
+
 plt.show()

### 2017-04-19 GraphGeneration/fig5-HistogramsAround80and113.py

@@ -1,0 +1,68 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.mlab as mlab
+#import scipy
+import pandas
+
+
+height = 3
+width = 7
+
+
+plt.figure(figsize=(width, height))
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-14-16-50-03\Task1Search\allPSMs_-187andUp.psmtsv", sep='\t')
+
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+
+massDiffs=data_sorted['MassDiff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue_notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(121)
+
+halfWidth = 0.01
+center = 113.084064
+min1 = center-halfWidth
+max1 = center+halfWidth
+
+bins = 21 
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+plt.hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =bins, range=[min1, max1], label = "Decoys")
+
+
+plt.legend()
+plt.annotate('Leucine\nResidue\nMass', xy=(113.08406, 35), xytext=(113.08406-0.00857, 40), arrowprops=dict(arrowstyle="->"))
+
+ok = plt.subplot(122)
+
+halfWidth = 0.01
+center = 79.961573 
+min1 = center-halfWidth
+max1 = center+halfWidth
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+plt.hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
[…]

### 2017-04-19 GraphGeneration/genericHistogramsAroundZero.py

@@ -2,16 +2,16 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
 
 #data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
 
 
-data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
+data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
 
-massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
+massDiffs=data_sorted['MassDiff (Da)'].values
 B=data_sorted['Decoy/Contaminant/Target'].values
 C=data_sorted['QValue_notch'].values
 
@@ -20,7 +20,7 @@ TDhighConf = B[C<0.01]
 
 ok = plt.subplot(321)
 
-halfWidth = 2.5
+halfWidth = 3.5
 min1 = -halfWidth
 max1 = halfWidth
 

### 2017-04-24 paper.tex
changes

@@
 \usepackage{booktabs}
+\usepackage{hyperref}
 
@@
 G-PTM-D\cite{Li_2016} is a recently developed tool for global identification and localization of PTMs.
-Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and efficiency of PTM identification.
+Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and speed of PTM identification.
 We enhance G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
-The proposed changes lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Targeted mass-window searches are also shown to be useful as replacements to standard wide-mass searches, resulting in a 10\% increase in the overall number of confidently identified peptides, and a 5\% increase in the number of mass shifts observed.
-Using a modification discovery heuristic originally based on the wide-mass search, targeted mass-window searches lead to an increase in the number of novel modifications discovered.
+The changes described in this work lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
+Targeted mass-window searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-mass and wide-mass searches.
 \end{abstract}
@@
 
-Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available.
-Information about a modification can include the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
-While this data is almost comprehensive, information about the localization of these modification to certain residues in specific proteins is scarce.
+Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available\cite{Creasy_2004}.
+Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
+Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identifications in complex biological samples has continued to pose a difficult challenge for proteomics\cite{Olsen_2013}.
 
-Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, e.g. MODa\cite{Na_2011}.
-G-PTM-D\cite{Li_2016}, a recently described bioinformatics tool for the discovery and localization of new PTMs, is the foundation for the work described in this paper.
+Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those concentrate on easily studied modifications (e.g.
+phosphorylations) in well studied systems.
+Global discovery tools, such as MODa\cite{Na_2011} are available.
+G-PTM-D\cite{Li_2016}, is one recently described bioinformatics tool for the discovery and localization of new PTMs in unenriched samples that enables high confidence identification and discovery of a wide variety of different PTMs at the same time.
 The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
@@
 Furthermore, in G-PTM-D, modification discovery is limited to PTMs in the Uniprot database. 
-Finally, PTMs that are similar in mass, such as Phosphorylation and Sulfonation are problematic since they are virtually indistinguishable in unprocessed spectra files.
+Finally, PTMs that are similar in mass, such as phosphorylation and sulfonation are problematic since they are virtually indistinguishable in unprocessed spectra files.
+These three limitations motivated the work described here.
 
-Spectral calibration prior to running the G-PTM-D procedure is essential for distinguishing modifications with similar mass.
-As a side effect, calibration significantly increases the overall number of confidently identified peptides and modifications.
-
+Spectral calibration is essential for distinguishing modifications with similar mass, so we recommend running the G-PTM-D procedure only on calibrated spectra files.
 We present an enhancement to G-PTM-D, that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
-The wide-mass search is replaced by a notch search in the discovery phase, using notches centered at the monoisotopic mass differences introduced by known modifications.
+MSFragger, a method recently described in\cite{Kong_2017}, significantly decreases the runtime of wide-mass searches, but lacks specificity.
+Thererfore, we replace the wide-mass search in the first stage by a specialized search that only allows specific mass differences which correspond to pre-selected modification masses.
 This change improves the specificity and significantly decreases search time.
-The final narrow-mass search is also replaced by a notch search, further increasing the number of confidently identified peptides.
+The final narrow-mass search is also replaced by a specialized missed-monoisotope identifying search, further increasing the number of confidently identified peptides.
 
 The idea of restricting the search space by only allowing certain mass differences between the precursor and the identified peptide mass is useful in other procedures as well, beyond the G-PTM-D workflow.
-Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on a wide-mass or an open-mass search.
+Its efficacy is demonstrated in context of a search for novel modifications, that is usually based on a wide-mass or an open-mass search.
 A combination of specialized notch and interval searches produces superior identification and discovery results.
@@
 \section{Experimental Procedures}
-We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the Enhanced G-PTM-D workflow.
-Two multi-fraction mammalian datasets (\textit{Mouse} and \textit{Jurkat}) with deep proteome coverage were used to evaluate the performance of the methods: these datasets are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
+We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the G-PTM-D workflow.
+Two multi-fraction mammalian datasets (\textit{Mouse} and \textit{Jurkat}) with deep proteome coverage were used to evaluate performance.
+These tryptically digested cell lysates correspond to 18 and 27 spectra files (fractions) respectively, and are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
 Uniprot XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
@@
 An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account.
-We propose alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
+We describe alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
 The notch search is an extension of the narrow-mass search: it allows exact matches (within a tolerance) to specific mass differences between the theoretical peptide and observed precursor masses.

### 2017-05-11 paper.tex
yea

@@
 For the database augmentation step, we allowed 120 modifications including PTMs, adducts, and chemical modifications.
-Final notch search used notches centered at \{0, 1.003, 2.005, 3.008\} in order to capture missed identifications due to missed monoisotopic peaks.
+Final notch search used notches centered at \{0, 1.0029, 2.0052, 3.0077\} in order to capture missed identifications due to missed monoisotopic peaks.
 
@@
 
-The first modification to the G-PTM-D workflow, post-acquisition calibration, is shown to increase the number of confidently identified peptides, and enables qualitatively distinguishing modifications with similar mass.
-The other major enhancement, the mass-window search, is contrasted with the narrow-mass and open-mass search.
-The superior performance of G-PTM-D achieved by the use of mass-window searches is demonstrated.
-Finally, the application of mass-window searches to the discovery of novel modifications is discussed.
+This work recommends introducing two major changes to the standard G-PTM-D workflow: a spectral calibration, and a replacement of the standard searches with custom mass-window type searches.
+In order to clearly show the advantage of each proposed change, they are initially discussed in isolation, and then the following section shows the combined effect on the overall PTM identification.
+
+%The first modification to the G-PTM-D workflow, post-acquisition calibration, is shown to increase the number of confidently identified peptides, and enables qualitatively distinguishing modifications with similar mass.
+%The other major enhancement, the mass-window search, is contrasted with the narrow-mass and open-mass search.
+%The superior performance of G-PTM-D achieved by the use of mass-window searches is demonstrated.
+%Finally, the application of mass-window searches to the discovery of novel modifications is discussed.
 
@@
 
-A critical parameter for peptide identification and PTM localization is mass accuracy\cite{Scherl_2008}.
+A critical parameter for peptide identification is mass accuracy\cite{Scherl_2008}.
 Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
@@
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
-The increased mass accuracy of the calibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
+%The increased mass accuracy of the calibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
 
@@
 This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy \cite{Elias_2007}).
-Step 2:  The identifications from Step 1 are used to extract peak matches from the spectra, and a calibration procedure is performed.
+Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
 The calibration algorithm is described in detail in Supplementary Material 1.
@@
 The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
-Furthermore, the systemic noise introduced by various other factors is addressed: this is evident by the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
+	Furthermore, the systemic noise introduced by various other factors is addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
 Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.
@@
 \centering
-\caption{PSMs Before and After Calibration}
+\caption{PSMs Within 1\% FDR Before and After Calibration}
 \label{tbl:calib}
@@
 
+Besides simply increasing the number of confidently identified peptides, calibration is instrumental in increasing the discernibility between different mass shifts.
+A search that allows mass differences in the interval [79.942541, 79.980605] was conducted: the resultant mass shifts are aggregated in the histogram plots below.
+It is evident that the phosphorylation and sulfonation peaks are clearly separated and identifyable when the search is conducted on calibrated data.
+
+\begin{figure}
+  \includegraphics{fig2-calibrationQualityPhospho.eps}
+  \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
+  \label{fgr:figure1}
+\end{figure}
 \newpage

### 2017-05-11 GraphGeneration/fig2a-calibQuality.py

@@ -1,0 +1,98 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+maxi = 25000
+
+
+height = 6
+width = 7
+
+f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row', figsize=(width,height))
+ax1.set_title('Uncalibrated')
+ax2.set_title('Calibrated')
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+A=ok[r"Mass Diff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue Notch"]
+
+ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+ax1.set_ylim([1,maxi])
+ax1.vlines(0,0,maxi,zorder = 10)
+ax1.set_ylabel('PSM count')
+ax1.annotate(r"Mouse", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
+                size='large', ha='right', va='center', xytext=(0,0.5))
+
+mu = np.mean(A[(C<0.01) & B])
+sigma = np.std(A[(C<0.01) & B])
+textToDisplay = 'count = %d\n$\mu=%.2f$\n$\sigma=%.2f$\n68.3%%: [%.2f,%.2f]\n95.4%%: [%.2f,%.2f]\n99.7%%: [%.2f,%.2f]'%(len(A[(C<0.01) & B]), mu, sigma,mu-sigma, mu+sigma,mu-2*sigma, mu+2*sigma,mu-3*sigma, mu+3*sigma)
+print(textToDisplay)
+
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task3Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+
+A=ok[r"Mass Diff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue Notch"]
+
[…]

### 2017-05-11 GraphGeneration/fig3-jurkat;-1;0;1;2.py

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task1Calibrate\2017-05-04-17-26-29\Task1Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
 
 
 
@@ -11,14 +11,14 @@ height = 6
 width = 7
 plt.figure(figsize=(width, height))
 
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-massDiffs=data_sorted['MassDiff (Da)'].values
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
 
 #data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
 #massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
 
 B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
+C=data_sorted['QValue Notch'].values
 
 massDiffsHighConf = massDiffs[C<0.01]
 TDhighConf = B[C<0.01]
@@ -62,10 +62,11 @@ plt.ylabel("Count")
 
 plt.title("PSMs around -1 Da")
 
-plt.annotate('Glu->Lys', xy=(-0.947630, 18), xytext=(-0.947630+0.015, 33), arrowprops=dict(arrowstyle="->"))
-plt.annotate('Asn->Xle', xy=(-0.958863, 25), xytext=(-0.958863+0.01, 63), arrowprops=dict(arrowstyle="->"))
-plt.annotate('Asp->Asn or\nGlu->Gln\nor Amidation', xy=(-0.984016, 40), xytext=(-0.984016+0.015, 80), arrowprops=dict(arrowstyle="->"))
-plt.annotate('H(-1)', xy=(-1.007825, 100), xytext=(-1.007825-0.01, 120), arrowprops=dict(arrowstyle="->"))
[…]

### 2017-05-11 GraphGeneration/fig3-mouse;-1;0;1;2.py

@@ -2,9 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task3Gptmd\aggregate_allPSMs_coisolation3.500.psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-14-16-50-03\Task1Search\allPSMs_-187andUp.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task4Search\allPSMs_-187andUp.psmtsv", sep='\t')
 
 
 
@@ -13,14 +11,14 @@ height = 6
 width = 7
 plt.figure(figsize=(width, height))
 
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-massDiffs=data_sorted['MassDiff (Da)'].values
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
 
 #data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
 #massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
 
 B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
+C=data_sorted['QValue Notch'].values
 
 massDiffsHighConf = massDiffs[C<0.01]
 TDhighConf = B[C<0.01]
@@ -65,10 +63,11 @@ plt.ylabel("Count")
 plt.title("PSMs around -1 Da")
 
 #plt.annotate('Glu->Lys', xy=(-0.947630, 18), xytext=(-0.947630+0.015, 33), arrowprops=dict(arrowstyle="->"))
[…]

### 2017-05-11 GraphGeneration/genericErrorVsVariousParams.py

@@ -1,42 +1,55 @@
 import numpy as np
 import matplotlib.pyplot as plt
+import pandas
 
-fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLbeforesc1.ms1dptsv"
-#fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLbeforesc2.ms1dptsv"
-#fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLbeforesc4.ms1dptsv"
-#fileName = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uLafterfc2after.ms1dptsv"
+nice= pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-08-14-03-16\Task1Calibrate\04-29-13_B6_Frac9_9p5uLafterfc4after.ms1dptsv", sep='\t')
+#nice= pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-08-14-03-16\Task1Calibrate\04-29-13_B6_Frac9_9p5uLafterfc4after.ms2dptsv", sep='\t')
 
-
-nice = np.loadtxt(fileName, skiprows=1, unpack = True)
-
-print("average error", sum(nice[-1])/len(nice[-1]))
 
 plt.figure(1)
-plt.subplot(231)
-plt.plot(nice[0],nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(nice[0]), max(nice[0]), zorder = 10)
+plt.subplot(331)
+plt.plot(nice['mz'],nice['Label'], 'bo',markersize = .3)
+plt.hlines(0, min(nice['mz']), max(nice['mz']), zorder = 10)
 plt.xlabel("mz")
-plt.subplot(232)
-plt.plot(nice[1],nice[-1], 'bo',markersize = .3)
-plt.hlines(0, min(nice[1]), max(nice[1]), zorder = 10)
-plt.xlabel("rt")
-plt.subplot(233)
[…]

### 2017-05-11 GraphGeneration/genericHistogramsAroundZero.py

@@ -2,18 +2,16 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task2Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task4Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-05-04-11-56-07\Task1Search\04-30-13_CAST_Frac5_4uL-Calibrated_allPSMs_3.5aroundZero.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-05-04-11-56-07\Task2Search\04-30-13_CAST_Frac5_4uL-Calibrated_allPSMs_3.5aroundZero.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-05-04-11-56-07\Task3Search\04-30-13_CAST_Frac5_4uL-Calibrated_allPSMs_3.5aroundZero.psmtsv", sep='\t')
 
 
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
 
-massDiffs=data_sorted['MassDiff (Da)'].values
+massDiffs=data_sorted['Mass Diff (Da)'].values
 B=data_sorted['Decoy/Contaminant/Target'].values
[…]

### 2017-05-12 GraphGeneration/figAposter.py
yeah

@@ -1,10 +1,13 @@
 import numpy as np
+import matplotlib
 import matplotlib.pyplot as plt
 
 
+matplotlib.rcParams.update({'font.size': 24})
 
-height = 5
-width = 3.25 
+
+height = 10 
+width = 7 
 plt.figure(figsize=(width, height))
 
 plt.subplot(411)
@@ -12,7 +15,7 @@ plt.subplot(411)
 dates = [-100,-1,0,1,100]
 values = [1,1,2,1,1]
 
-plt.plot(dates, values, '-')
+plt.plot(dates, values, '-', linewidth=5)
 
 ylabels = ['Reject', 'Accept']
 yTicks = [1,2]
@@ -26,7 +29,7 @@ plt.subplot(412)
 dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
 values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
 
-plt.plot(dates, values, '-')
+plt.plot(dates, values, '-', linewidth=5)
 
 ylabels = ['Reject', 'Accept']
 yTicks = [1,2]
@@ -42,7 +45,7 @@ plt.subplot(413)
 
 dates =  [-100,-87,-86,-15,-14,-13,56,57,58,100]
 values = [1   ,  1, 2, 2,  1,  2, 2,1 ,2 ,2]
-plt.plot(dates, values, '-')
+plt.plot(dates, values, '-', linewidth=5)
 
 ylabels = ['Reject', 'Accept']
 yTicks = [1,2]
@@ -51,19 +54,22 @@ plt.ylim([0.5,2.5])
 
 plt.title("Interval Search")
 plt.yticks(yTicks, ylabels)
-plt.subplot(414)
+hehe =plt.subplot(414)
 
 
 
 dates =  [-100,100]
 values = [2 ,2]
 
-plt.plot(dates, values, '-')
+plt.plot(dates, values, '-', linewidth=5)
+
 
 ylabels = ['Reject', 'Accept']
 yTicks = [1,2]
 
 plt.ylim([0.5,2.5])
+
+hehe.set_xlabel('Mass Diff (Da)')
 
 plt.title("Wide-Mass Search")
 
@@ -71,8 +77,8 @@ plt.yticks(yTicks, ylabels)
 plt.tight_layout()
 
 
-plt.savefig('fig1-searchTypes.eps', format='eps', dpi=1200)
[…]

### 2017-05-12 GraphGeneration/figBposter.py

@@ -1,0 +1,48 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pandas
+
+
+matplotlib.rcParams.update({'font.size': 24})
+
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task4Search\allPSMs_-187andUp.psmtsv", sep='\t')
+
+height = 5
+width = 7
+plt.figure(figsize=(width, height))
+
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+halfWidth = 0.1 
+min1 = 1-halfWidth
+max1 = 1+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
+
+plt.ylabel("PSM Count")
+plt.axis([0.9, 1.1, 0, 1200])
+
+plt.title("Mass Diffs Around 1 Da")
+
+plt.annotate('Deamidation', xy=(0.984016, 900), xytext=(0.984016+0.02, 700), arrowprops=dict(arrowstyle="->"))
+plt.annotate('1 Mm', xy=(1.0029, 150), xytext=(1.027, 180), arrowprops=dict(arrowstyle="->"))
+
+
+plt.tight_layout()
+
+plt.savefig('figBposter.eps', format='eps', dpi=1200)
+plt.savefig('figBposter.png', format='png', dpi=1200)
+
+
+plt.show()

### 2017-05-12 GraphGeneration/figCalibrationPoster.py

@@ -1,0 +1,56 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pandas
+
+maxi = 30000
+
+matplotlib.rcParams.update({'font.size': 24})
+
+height = 5
+width = 14 
+
+f, ((ax1, ax2)) = plt.subplots(1, 2, sharex='col', sharey='row', figsize=(width,height))
+ax1.set_title('Uncalibrated')
+ax2.set_title('Calibrated')
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+A=ok[r"Mass Diff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue Notch"]
+
+ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+ax1.set_ylim([1,maxi])
+ax1.vlines(0,0,maxi,zorder = 10)
+ax1.set_ylabel('PSM count')
+ax1.set_xlabel('Error (PPM)')
+
+mu = np.mean(A[(C<0.01) & B])
+sigma = np.std(A[(C<0.01) & B])
+textToDisplay = 'count = %d\n$\mu=%.2f$\n$\sigma=%.2f$\n68.3%%: [%.2f,%.2f]\n95.4%%: [%.2f,%.2f]\n99.7%%: [%.2f,%.2f]'%(len(A[(C<0.01) & B]), mu, sigma,mu-sigma, mu+sigma,mu-2*sigma, mu+2*sigma,mu-3*sigma, mu+3*sigma)
+print(textToDisplay)
+
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task3Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+
+A=ok[r"Mass Diff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue Notch"]
+
+ax2.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+ax2.set_ylim([1,maxi])
+ax2.vlines(0,0,maxi,zorder = 10)
[…]

### 2017-05-12 GraphGeneration/figCposter.py

@@ -1,0 +1,100 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pandas
+
+
+matplotlib.rcParams.update({'font.size': 24})
+
+
+height = 5
+width = 7
+plt.figure(figsize=(width, height))
+
+
+offset = 200
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task4Search\allPSMs_OpenSearch.psmtsv", sep='\t')
+
+
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+
+A=data_sorted['Mass Diff (Da)'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+print(A[0])
+print(A[-1])
+
+AhighConf = A[C<0.01]
+BhighConf = B[C<0.01]
+
+print(AhighConf[0])
+print(AhighConf[-1])
+
+print(BhighConf[0])
+print(BhighConf[1])
+print(BhighConf[2])
+print(BhighConf[3])
+print(BhighConf[4])
+
+totalCount = len(AhighConf)
+   
+forward = [0]
+forwardDiv = []
+forwardInd= []
+for i in range(totalCount):
+    forward.append(forward[-1]+1 if BhighConf[i]=='D' else forward[-1])
+    forwardDiv.append(forward[-1]/(i+1))
+    forwardInd.append(AhighConf[i])
+
+print(forwardDiv[0])
+print(forwardDiv[1])
+print(forwardDiv[2])
+print(forwardDiv[3])
+print(forwardDiv[4])
+print(forwardDiv[5])
+print(forwardDiv[6])
+print(forwardDiv[7])
+print(forwardDiv[8])
+
+plt.scatter(forwardInd[offset:totalCount-offset:100], forwardDiv[offset:totalCount-offset:100])
+plt.xscale('symlog')
+plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
+plt.ylabel("FDR")
+plt.xlabel("Mass Diff Lower Limit")
+plt.title("Restricting the Open Search")
+
+
+backward = [0]
+backwardDiv= []
+backwardInd= []
+for i in range(totalCount-1,-1,-1):
+    backward.append(backward[-1]+1 if BhighConf[i]=='D' else backward[-1])
[…]

### 2017-06-16 GraphGeneration/figPhosphorylationPoster.py
Yahhhh

@@ -1,0 +1,47 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pandas
+
+
+matplotlib.rcParams.update({'font.size': 24})
+
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-30-10-25-53\Task1Search\allPSMs_name.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-05-30-10-26-30\Task1Search\allPSMs_name.psmtsv", sep='\t')
+
+
+height = 5
+width = 7
+plt.figure(figsize=(width, height))
+
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+halfWidth = 0.03 
+min1 = 79.961573-halfWidth
+max1 = 79.961573+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="T")]
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
+
+plt.ylabel("PSM Count")
+plt.xlabel("Mass Diff (Da)")
+plt.axis([min1, max1, 0, 150])
+
+plt.title("Original")
+
+plt.tight_layout()
+
+plt.savefig('figPhosphorylationPre.eps', format='eps', dpi=1200)
+plt.savefig('figPhosphorylationPre.png', format='png', dpi=1200)
+
+
+plt.show()

### 2017-06-16 GraphGeneration/figPhosphorylationPosterCalib.py

@@ -1,0 +1,52 @@
+import numpy as np
+import matplotlib
+import matplotlib.pyplot as plt
+import pandas
+
+
+matplotlib.rcParams.update({'font.size': 24})
+
+
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-30-10-25-53\Task1Search\allPSMs_name.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-05-30-10-26-30\Task1Search\allPSMs_name.psmtsv", sep='\t')
+
+
+height = 5
+width = 7
+plt.figure(figsize=(width, height))
+
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+halfWidth = 0.03 
+min1 = 79.961573-halfWidth
+max1 = 79.961573+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="T")]
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
+
+plt.ylabel("PSM Count")
+plt.xlabel("Mass Diff (Da)")
+
+
+plt.annotate('Sulfo', xy=(79.956815, 50), xytext=(79.94, 50), arrowprops=dict(arrowstyle="->"))
+plt.annotate('Phospho', xy=( 79.966331, 100), xytext=(79.97, 100), arrowprops=dict(arrowstyle="->"))
+
+plt.axis([min1, max1, 0, 150])
+
+plt.title("Calibrated")
+
+plt.tight_layout()
+
+plt.savefig('figPhosphorylationCalib.eps', format='eps', dpi=1200)
+plt.savefig('figPhosphorylationCalib.png', format='png', dpi=1200)
+
+
+plt.show()

### 2017-06-20 paper.tex
yeah
Yahhhh

@@
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
-Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identifications in complex biological samples has continued to pose a difficult challenge for proteomics\cite{Olsen_2013}.
+\textbf{Protein databases such as UniProt may include localization information such as the observation of a modification on a specific amino acid in a protein, but this information is incomplete.}
+Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identifications in complex biological samples has continued to pose a challenge for proteomics\cite{Olsen_2013}.
 
-Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those concentrate on easily studied modifications (e.g.
+Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those concentrate on common modifications (e.g.
 phosphorylations) in well studied systems.
-Global discovery tools, such as MODa\cite{Na_2011} are available.
-G-PTM-D\cite{Li_2016}, is one recently described bioinformatics tool for the discovery and localization of new PTMs in unenriched samples that enables high confidence identification and discovery of a wide variety of different PTMs at the same time.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011} that provides spectral matches to unmodified peptides along with the mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM).
-2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database.
+Global discovery tools, such as MODa\cite{Na_2011} are also available, \textbf{but those have some drawbacks}.
+G-PTM-D\cite{Li_2016}, is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs at the same time in unenriched samples.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011} that provides mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM)
+2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database
 3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR).
@@
 Due to this, high mass modifications, such as glycosolations are out of reach, since widening the search window from the suggested $\pm 200$ Daltons increases the search time even more.
+The final narrow-mass search is also problematic, since it does not account for mass errors due to missing the monoisotopic peak.
 Furthermore, in G-PTM-D, modification discovery is limited to PTMs in the Uniprot database.
+This is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those significantly degrades the overall final identification rates.
 Finally, PTMs that are similar in mass, such as phosphorylation and sulfonation are problematic since they are virtually indistinguishable in unprocessed spectra files.
-These three limitations motivated the work described here.
+These limitations motivated the work described here.
 
+We enhance the G-PTM-D procedure in the following ways.
 Spectral calibration is essential for distinguishing modifications with similar mass, so we recommend running the G-PTM-D procedure only on calibrated spectra files.
-We present an enhancement to G-PTM-D, that addresses the long search times of the wide-mass search, and the limitations of the standard final narrow-mass search.
-MSFragger, a method recently described in\cite{Kong_2017}, significantly decreases the runtime of wide-mass searches, but lacks specificity.
-Thererfore, we replace the wide-mass search in the first stage by a specialized search that only allows specific mass differences which correspond to pre-selected modification masses.
+To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search that only allows specific mass differences which correspond to pre-selected modification masses.
 This change improves the specificity and significantly decreases search time.
 The final narrow-mass search is also replaced by a specialized missed-monoisotope identifying search, further increasing the number of confidently identified peptides.
+Finally, we expand the limited Uniprot PTM list with a customizeable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
+This change not only expands the list of identified and localized modifications, but also indirectly improves the confident PSM count.
 
@@
 
-\begin{figure}
+\begin{figure}[H]
   \includegraphics{diagram.eps}
@@
 We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the G-PTM-D workflow.
-Two multi-fraction mammalian datasets (\textit{Mouse} and \textit{Jurkat}) with deep proteome coverage were used to evaluate performance.
-These tryptically digested cell lysates correspond to 18 and 27 spectra files (fractions) respectively, and are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
+Two multi-fraction mammalian datasets (\textit{Jurkat} and \textit{Mouse}) with deep proteome coverage were used to evaluate performance.
+These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells, and 18 fractions from Mouse islets, and are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
+
 Uniprot XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
 
-A 10ppm precursor mass tolerance was used for the initial calibration step, and then reduced to 5ppm for subsequent notch searches.
+A 10ppm precursor mass tolerance is used for the initial calibration step, and then reduced to 5ppm for subsequent notch searches.
 For the database augmentation step, we allowed 120 modifications including PTMs, adducts, and chemical modifications.
@@
 real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration).
-We use a post-acquisition calibration procedure, that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
+We use a post-acquisition calibration procedure that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
@@
 \begin{figure}
-  \includegraphics{fig2-calibrationQualityPhospho.eps}
+  \includegraphics{fig2-calibrationQuality.eps}
   \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
@@
 Mass windows provide an mechanism to exclude these problematic mass shifts from a wide-mass search by employing a interval search.
+
+MSFragger, a method recently described in\cite{Kong_2017}, significantly decreases the runtime of wide-mass searches, but lacks specificity.
 

### 2017-06-20 paper.tex
ahh

@@
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
-\textbf{Protein databases such as UniProt may include localization information such as the observation of a modification on a specific amino acid in a protein, but this information is incomplete.}
+Protein databases such as UniProt include information on the localization of specific modifications to specific residues in the protein, but this information is incomplete.
 Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identifications in complex biological samples has continued to pose a challenge for proteomics\cite{Olsen_2013}.
 
-Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those concentrate on common modifications (e.g.
-phosphorylations) in well studied systems.
-Global discovery tools, such as MODa\cite{Na_2011} are also available, \textbf{but those have some drawbacks}.
-G-PTM-D\cite{Li_2016}, is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs at the same time in unenriched samples.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011} that provides mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM)
+Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those mostly concentrate on common modifications (e.g.
+phosphorylations) in well-studied systems.
+Global discovery tools, such as MODa\cite{Na_2011} are also available.
+G-PTM-D\cite{Li_2016}, is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs at the same time.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011}, that provides mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM)
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database
@@
 Due to this, high mass modifications, such as glycosolations are out of reach, since widening the search window from the suggested $\pm 200$ Daltons increases the search time even more.
-The final narrow-mass search is also problematic, since it does not account for mass errors due to missing the monoisotopic peak.
-Furthermore, in G-PTM-D, modification discovery is limited to PTMs in the Uniprot database.
-This is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those significantly degrades the overall final identification rates.
+The final narrow-mass search is also problematic, since it does not account for peptide-spectral matches(PSMs) with mass differences arising due to mistaking the monoisotopic peak.
+In G-PTM-D, modification discovery is limited to PTMs in the Uniprot database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
 Finally, PTMs that are similar in mass, such as phosphorylation and sulfonation are problematic since they are virtually indistinguishable in unprocessed spectra files.
@@
 
-We enhance the G-PTM-D procedure in the following ways.
 Spectral calibration is essential for distinguishing modifications with similar mass, so we recommend running the G-PTM-D procedure only on calibrated spectra files.
 To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search that only allows specific mass differences which correspond to pre-selected modification masses.
-This change improves the specificity and significantly decreases search time.
-The final narrow-mass search is also replaced by a specialized missed-monoisotope identifying search, further increasing the number of confidently identified peptides.
-Finally, we expand the limited Uniprot PTM list with a customizeable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
+This change improves the specificity of the search and significantly decreases search time.
+The final narrow-mass search is also replaced by a specialized missed-monoisotope identifying search, further increasing the number of confidently identified peptides and PTMs.
+Finally, we replace the limited Uniprot PTM list with a customizeable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
 This change not only expands the list of identified and localized modifications, but also indirectly improves the confident PSM count.
+The ehanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
 
-The idea of restricting the search space by only allowing certain mass differences between the precursor and the identified peptide mass is useful in other procedures as well, beyond the G-PTM-D workflow.
-Its efficacy is demonstrated in context of a search for novel modifications, that is usually based on a wide-mass or an open-mass search.
+The idea of restricting the search space by only allowing certain mass differences between the precursor and the identified peptide mass is useful in procedures beyond the G-PTM-D workflow.
+Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on unrestricted searches.
 A combination of specialized notch and interval searches produces superior identification and discovery results.
@@
 \begin{figure}[H]
-  \includegraphics{diagram.eps}
+  \includegraphics[scale=0.5]{diagram.eps}
   \caption{Workflows}
@@
 \section{Experimental Procedures}
-We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the G-PTM-D workflow.
+We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
 Two multi-fraction mammalian datasets (\textit{Jurkat} and \textit{Mouse}) with deep proteome coverage were used to evaluate performance.
@@
 
-A 10ppm precursor mass tolerance is used for the initial calibration step, and then reduced to 5ppm for subsequent notch searches.
-For the database augmentation step, we allowed 120 modifications including PTMs, adducts, and chemical modifications.
-Final notch search used notches centered at \{0, 1.0029, 2.0052, 3.0077\} in order to capture missed identifications due to missed monoisotopic peaks.
+A 10ppm precursor mass tolerance is used for the initial calibration step, and then reduced to 5ppm for subsequent searches.
+We allowed 120 modifications including PTMs, adducts, and chemical modifications in the database augmentation step.
+%Final notch search used notches centered at \{0, 1.0029, 2.0052, 3.0077\} in order to capture missed identifications due to missed monoisotopic peaks.
 
-For the discovery of novel modifications, we used a combination of two searches: A notch search with notches at all combinations of up to 2 residue mass additions or removals, and an interval search with lower bound of -187 Daltons, and with exclusions at masses corresponding to the notch search.
+%For the discovery of novel modifications, we used a combination of two searches: A notch search with notches at all combinations of up to 2 residue mass additions or removals, and an interval search with lower bound of -187 Daltons, and with exclusions at masses corresponding to the notch search.
 
@@
 This work recommends introducing two major changes to the standard G-PTM-D workflow: a spectral calibration, and a replacement of the standard searches with custom mass-window type searches.
-In order to clearly show the advantage of each proposed change, they are initially discussed in isolation, and then the following section shows the combined effect on the overall PTM identification.
-
-%The first modification to the G-PTM-D workflow, post-acquisition calibration, is shown to increase the number of confidently identified peptides, and enables qualitatively distinguishing modifications with similar mass.
-%The other major enhancement, the mass-window search, is contrasted with the narrow-mass and open-mass search.
-%The superior performance of G-PTM-D achieved by the use of mass-window searches is demonstrated.
-%Finally, the application of mass-window searches to the discovery of novel modifications is discussed.
+In order to clearly show the advantage of each proposed change, they are initially discussed in isolation, and then the following section shows the combined effect on the overall G-PTM-D performance.
 
@@
 The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
-Furthermore, the systemic noise introduced by various other factors is addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
+The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
 Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.
@@
 
-Besides simply increasing the number of confidently identified peptides, calibration is instrumental in increasing the discernibility between different mass shifts.
-A search that allows mass differences in the interval [79.942541, 79.980605] was conducted: the resultant mass shifts are aggregated in the histogram plots below.
-It is evident that the phosphorylation and sulfonation peaks are clearly separated and identifyable when the search is conducted on calibrated data.
-
-\begin{figure}
-  \includegraphics{fig2-calibrationQuality.eps}
-  \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
-  \label{fgr:figure1}
-\end{figure}
 \newpage
@@
 The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
-Figure~\ref{fgr:differentSearchModes} provides a convenient way to visualize the various search modes.
+Figure~\ref{fig:fig1-searchtypes} provides a convenient way to visualize the various search modes.
 
@@
 
-\subsection{G-PTM-D}
+\subsection{Enhanced G-PTM-D Results}
+
+A major improvement in the G-PTM-D performance arises from using calibrated spectra as inputs to the procedure.
+Calibration is instrumental in increasing the discernibility between different mass shifts.
+Mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:phosphoHistograms}.
+It is evident that the phosphorylation and sulfonation peaks are clearly separated and identifyable when the search is conducted on calibrated data.
+
+\begin{figure}	
+	\centering
+	\begin{subfigure}[t]{3.1in}
+		\centering
+		\includegraphics[scale=0.5]{figPhosphorylationPre.eps}
+		\label{fig:1a}		
+	\end{subfigure}
+	\quad
+	\begin{subfigure}[t]{3.1in}
+		\centering
+		\includegraphics[scale=0.5]{figPhosphorylationCalib.eps}
+		\label{fig:1b}
+	\end{subfigure}
+	\caption{Histograms of mass differences observed around 80 Daltons when using calibrated and uncalibrated spectra}\label{fig:phosphoHistograms}
+\end{figure}
 
 By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PSMs increased by 6\%.
-The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-label}.
+The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
 
@@
 \caption{Effects of Replacing Initial Wide-Mass Search by a Notch Search}
-\label{my-label}
+\label{my-labelff}
 \begin{tabular}{ll|l|l}

### 2017-06-20 paper.tex
yea

@@
 We allowed 120 modifications including PTMs, adducts, and chemical modifications in the database augmentation step.
-%Final notch search used notches centered at \{0, 1.0029, 2.0052, 3.0077\} in order to capture missed identifications due to missed monoisotopic peaks.
 
@@
 
+\subsubsection{Discovery of Modifications not in Database}
+
+Previously described wide mass search strategies for discovery and localization of unknown modifications benefit from using alternative, carefully designed search modes based on interval and notch searches.
+We discovered that limiting mass shifts to a lower bound of -187 Da (corresponding to the largest mass difference that could be attributed to loss of a single residue, Tryptophan), and no upper bound, is an important step in eliminating spurious PSMs.
+Furthermore, treating highly suspect mass shifts corresponding to residue additions/substitutions/deletions in a separate notch search with individualized false discovery rate estimates is beneficial.
+An automated tool built into MetaMorpheus allows confidently identifying novel modifications based on these search results.
+
 \subsection{Enhanced G-PTM-D Results}
@@
 
-The final search with the augmented database is conducted with notches: we only recommend including the common missed monoisotopic mass values in the final search,.
+The final notch search uses notches centered at \{0, 1.0029, 2.0052, 3.0077\}, in order to capture missed identifications due to missed monoisotopic peaks.
+The results are contrasted with the standard narrow-mass final search.
 The effect of using these notches is shown in Table~\ref{tab:table2}.
@@
 
-The initial notch search is significantly faster than a corresponding wide mass search, and this suggests that multiple search rounds could be conducted, iteratively adding more and more modifications.
-This recursive nature of the algorithm is well suited to add multiple modifications of different types on the same peptide.
-Due to these enhancements, we test the performance of our methods
-
-\begin{table}[]
-\centering
-\caption{My caption}
-\label{my-label}
-\begin{tabular}{|l|l|l|l|l|l|}
-\hline
-                      &        & \multicolumn{2}{l|}{Single Round}                                                                              & \multicolumn{2}{l|}{Multiple Rounds}                                                                           \\ \hline
-                      &        & \begin{tabular}[c]{@{}l@{}}Mods:\\ none\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mods:\\ ptmlist\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mods:\\ none\end{tabular} & \begin{tabular}[c]{@{}l@{}}Mods:\\ ptmlist\end{tabular} \\ \hline
-\multirow{2}{*}{PSMs} & Jurkat &                                                      &                                                         &                                                      &                                                         \\ \cline{2-6}
-                      & Mouse  &                                                      &                                                         &                                                      &                                                         \\ \hline
-\multirow{2}{*}{Mods} & Jurkat &                                                      &                                                         &                                                      &                                                         \\ \cline{2-6}
-                      & Mouse  &                                                      &                                                         &                                                      &                                                         \\ \hline
-\multirow{2}{*}{Time} & Jurkat &                                                      &                                                         &                                                      &                                                         \\ \cline{2-6}
-                      & Mouse  &                                                      &                                                         &                                                      &                                                         \\ \hline
-\end{tabular}
-\end{table}
-
-\subsection{Discovery of Modifications not in Database}
-
-Previously described wide mass search strategies for discovery and localization of unknown modifications benefit from using alternative, carefully designed search modes based on interval and notch searches.
-We discovered that limiting mass shifts to a lower bound of -187 Da (corresponding to the largest mass difference that could be attributed to loss of a single residue, Tryptophan), and no upper bound, is an important step in eliminating spurious PSMs.
-Furthermore, treating highly suspect mass shifts corresponding to residue additions/substitutions/deletions in a separate notch search with individualized false discovery rate estimates is beneficial.
-An automated tool built into MetaMorpheus allows confidently identifying novel modifications based on these search results.
-
 %

### 2017-07-10 paper.tex
many

@@
 
-Restricting the allowed precursor masses is beneficial in searches that are intended to discover mass shift peaks.
-We ran a completely open precursor mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
-The plots in Figure~\ref{fig:figure2-upperlowerbounds} show the FDR in the list filtered by using a lower or upper bound cutoff on the allowed mass shifts.
+Analogous to a notch search being an extension of the narrow-mass search, an interval search is a natural extension of the open-mass search.
+Restricting the allowed differences between the precursor and the identified peptide mass is often beneficial in searches that are intended to discover mass shift peaks.
+While this method is not used in the G-PTM-D pipeline, we include it here for completeness, and demonstrate the improved performance when it is used in a heuristic for the discovery of novel modifications.
+
+One way of restricting open mass searches is by introducing an upper or lower limit on the accepted mass differences.
+To illustrate the effect of such restrictions, we ran an open-mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
+The plots in Figure~\ref{fig:fig4-limitsOnOpenSearch} show the FDR in the list filtered by using an upper or lower bound cutoff on the allowed mass shifts.
 It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.
-
 This effect can be explained by the fact that it's easier to match long sequences to short peptide fragmentation profiles than it is to match short sequences to fragmentation pattern of a long peptide.
@@
 We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some histogram peaks are illustrated in Figure~\ref{fig:figure3}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
+Some histogram peaks are illustrated in Figure~\ref{fig:fig5-HistogramsAround80and113}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
 

### 2017-07-10 GraphGeneration/fig4-limitsOnOpenSearch.py

@@ -9,14 +9,14 @@ plt.figure(figsize=(width, height))
 
 offset = 200
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-14-16-50-03\Task1Search\allPSMs_OpenSearch.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-07-06-17-29-30\Task1Search\aggregatePSMs_OpenSearch.psmtsv", sep='\t')
 
 
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
 
-A=data_sorted['MassDiff (Da)'].values
+A=data_sorted['Mass Diff (Da)'].values
 B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
+C=data_sorted['QValue Notch'].values
 
 print(A[0])
 print(A[-1])
@@ -60,7 +60,7 @@ plt.xscale('symlog')
 plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
 plt.ylabel("FDR")
 plt.xlabel("Mass Diff")
-plt.title("Limit From Below")
+plt.title("Limit From Above")
 
 
 backward = [0]
@@ -94,7 +94,7 @@ plt.xscale('symlog')
 plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
 plt.ylabel("FDR")
 plt.xlabel("Mass Diff")
-plt.title("Limit From Above")
+plt.title("Limit From Below")
 
 plt.tick_params(axis='both', which='major', labelsize=8)
 

### 2017-07-10 paper.tex
euy

@@
 Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and speed of PTM identification.
-We enhance G-PTM-D by using targeted mass-window searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
+We enhance G-PTM-D by using targeted multi-notch searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
 The changes described in this work lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Targeted mass-window searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-mass and wide-mass searches.
+Targeted multi-notch searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-mass and wide-mass searches.
 \end{abstract}

### 2017-07-10 GraphGeneration/fig4-limitsOnOpenSearch.py

@@ -9,7 +9,8 @@ plt.figure(figsize=(width, height))
 
 offset = 200
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-07-06-17-29-30\Task1Search\aggregatePSMs_OpenSearch.psmtsv", sep='\t')
+#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-07-06-17-29-30\Task1Search\aggregatePSMs_OpenSearch.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-05-05-18-29-43\Task2Calibrate\2017-07-07-13-19-26\Task1Search\aggregatePSMs_OpenSearch.psmtsv", sep='\t')
 
 
 data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)

### 2017-07-10 paper.tex
yey

@@
 
-Many peptide residue modification types are known, and databases containing detailed information about such modifications are readily available\cite{Creasy_2004}.
+Many types of post-translational modification (PTM) are known, and databases containing detailed information about such modifications are readily available\cite{Creasy_2004}.
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
 Protein databases such as UniProt include information on the localization of specific modifications to specific residues in the protein, but this information is incomplete.
-Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identifications in complex biological samples has continued to pose a challenge for proteomics\cite{Olsen_2013}.
+Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identification in complex biological samples has continued to pose a challenge for proteomics\cite{Olsen_2013}.
 

### 2017-07-10 paper.tex
nice

@@
 
-Many types of post-translational modification (PTM) are known, and databases containing detailed information about such modifications are readily available\cite{Creasy_2004}.
+Many types of post-translational modifications (PTMs) are known, and databases containing detailed information about such modifications are readily available\cite{Creasy_2004}.
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
@@
 
-Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those mostly concentrate on common modifications (e.g.
+Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those address primarily common modifications (e.g.
 phosphorylations) in well-studied systems.
 Global discovery tools, such as MODa\cite{Na_2011} are also available.
-G-PTM-D\cite{Li_2016}, is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs at the same time.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011}, that provides mass differences between the identified peptides and the measured parent peptide masses (hypothesized to differ in mass due to the presence of a PTM)
-2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the proteins in the search database
+G-PTM-D\cite{Li_2016} is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011}, that provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
+2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
 3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR).

### 2017-07-10 paper.tex
yehafjk

@@
 
-As described, G-PTM-D has a few significant limitations.
+As described above, G-PTM-D has a few significant limitations.
 The initial wide-mass search procedure is slow, taking hours for modest size datasets.
-Due to this, high mass modifications, such as glycosolations are out of reach, since widening the search window from the suggested $\pm 200$ Daltons increases the search time even more.
-The final narrow-mass search is also problematic, since it does not account for peptide-spectral matches(PSMs) with mass differences arising due to mistaking the monoisotopic peak.
+Due to this, high mass modifications, such as glycosylations are out of reach, since widening the search window from the suggested $\pm 200$ Da increases the search time even more.
+The final narrow-mass search is also problematic, since it does not account for peptide-spectral matches(PSMs) with mass differences arising due to misidentifying the monoisotopic peak.
 In G-PTM-D, modification discovery is limited to PTMs in the Uniprot database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
-Finally, PTMs that are similar in mass, such as phosphorylation and sulfonation are problematic since they are virtually indistinguishable in unprocessed spectra files.
+Finally, PTMs that are similar in mass, such as phosphorylation (79.966331 Da) and sulfonation (79.956815 Da), are problematic since they are virtually indistinguishable in unprocessed spectral files.
 These limitations motivated the work described here.

### 2017-07-10 paper.tex
fhdkjfh

@@
 Spectral calibration is essential for distinguishing modifications with similar mass, so we recommend running the G-PTM-D procedure only on calibrated spectra files.
-To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search that only allows specific mass differences which correspond to pre-selected modification masses.
+To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected modification masses (e.g.
+PTMs and amino acids).
 This change improves the specificity of the search and significantly decreases search time.
-The final narrow-mass search is also replaced by a specialized missed-monoisotope identifying search, further increasing the number of confidently identified peptides and PTMs.
+The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in precursor mass deconvolution (+1.0029, + 2.0052 and + 3.0077 Da), further increasing the number of confidently identified peptides and PTMs.
 Finally, we replace the limited Uniprot PTM list with a customizeable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
 This change not only expands the list of identified and localized modifications, but also indirectly improves the confident PSM count.
-The ehanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
+The enhanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
 

### 2017-07-10 paper.tex
dfjakdf

@@
 
-\title[Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows]
-  {Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Defined Mass Windows}
+\title[Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches]
+  {Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches}
 
@@
 The enhanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
-
-The idea of restricting the search space by only allowing certain mass differences between the precursor and the identified peptide mass is useful in procedures beyond the G-PTM-D workflow.
-Its efficacy is demonstrated in context of a search for novel modifications, which is usually based on unrestricted searches.
-A combination of specialized notch and interval searches produces superior identification and discovery results.
 

### 2017-07-10 paper.tex
djfkajf

@@
 \section{Experimental Procedures}
+
 We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
-Two multi-fraction mammalian datasets (\textit{Jurkat} and \textit{Mouse}) with deep proteome coverage were used to evaluate performance.
+Three multi-fraction mammalian datasets (Jurkat cell, C57BL/6J (B6) and CAST/EiJ (CAST) mouse islet lysates) with deep proteome coverage were used to evaluate performance.
 These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells, and 18 fractions from Mouse islets, and are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
@@
 We allowed 120 modifications including PTMs, adducts, and chemical modifications in the database augmentation step.
-
-%For the discovery of novel modifications, we used a combination of two searches: A notch search with notches at all combinations of up to 2 residue mass additions or removals, and an interval search with lower bound of -187 Daltons, and with exclusions at masses corresponding to the notch search.
 

### 2017-07-10 paper.tex
fjkldsjf

@@
 
-This work recommends introducing two major changes to the standard G-PTM-D workflow: a spectral calibration, and a replacement of the standard searches with custom mass-window type searches.
-In order to clearly show the advantage of each proposed change, they are initially discussed in isolation, and then the following section shows the combined effect on the overall G-PTM-D performance.
+This work presents four major changes to the previously published G-PTM-D workflow:
+a spectral calibration;
+a replacement of the wide parent mass tolerance search with a custom multi-notch search for PTM discovery;
+replacement of the final traditional narrow mass search with a limited multi-notch search that permits identification of incorrectly deconvoluted precursor masses;
+and lastly, addition of capacity for identification of multiple cofragmented peptides from a single MS/MS scan.
+To clarify the advantage of each change, they are first discussed separately, followed by an analysis of their combined effect upon overall G-PTM-D performance.
 

### 2017-07-10 paper.tex
fjdkfj

@@
 standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
-real-time calibration using a lock mass standard\cite{Olsen_2005}); and subsequent to the MS experiment (post-acquisition spectral calibration).
-We use a post-acquisition calibration procedure that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
+real-time calibration using a lock mass standard\cite{Olsen_2005}); and spectral calibration subsequent to the MS experiment (post-acquisition spectral calibration).
+We developed a post-acquisition calibration procedure that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
-%The increased mass accuracy of the calibrated spectra leads directly to improved identifications of both modified and unmodified peptides, as well as to increased confidence for PTM localization.
-
 The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy \cite{Elias_2007}).
+This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy\cite{Elias_2007}).
 Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
@@
 
-Calibration quality is accessed through the results of a standard 10ppm precursor tolerance database search.
-PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure, see Table~\ref{tbl:calib}.
+Calibration quality is evaluated through the results of a standard 10ppm precursor tolerance database search.
+PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure (Table~\ref{tbl:calib}).
 The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
-The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which places calibrated data squarely in the sub-ppm mode.
+The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which show that the calibrated mass spectra have sub-ppm mass errors.
 Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.

### 2017-07-10 paper.tex
huge

@@
 
-\subsection{Mass-Window Searches}
+\subsection{Multi-Notch Searches}
 
-Prior to discussing the impact mass-windows searches have on the G-PTM-D workflow, we need to define the search modes discussed.
-A standard \textit{narrow-mass} search considers matching theoretical peptides to spectra if the peptide mass and the identified precursor mass are equal, within some tolerance (i.e.
+Prior to discussing the impact that multi-notch searches have on the G-PTM-D performance, it is useful to describe the variety of search tolerances used.
+Traditional searches in bottom-up proteomics score potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal, within some defined tolerance (e.g.
 5ppm, or 2.1 Da).
-A \textit{wide-mass} search matches peptide fragment masses to fragmentation spectra while allowing large differences (i.e. $\pm 500$ Da) between the observed precursor mass and the peptide mass.
-An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account.
-We describe alternatives to these search modes, namely a \textit{notch search} and an \textit{interval search} that fall under a broad category of \textit{mass-window} searches.
-The notch search is an extension of the narrow-mass search: it allows exact matches (within a tolerance) to specific mass differences between the theoretical peptide and observed precursor masses.
-The interval search is an extension of the wide-mass search: specific problematic mass differences are excluded from allowed matches.
-Figure~\ref{fig:fig1-searchtypes} provides a convenient way to visualize the various search modes.
+We'll refer to this type of search here as a \textit{single notch} search.
+The other common search mode, used in the original G-PTM-D publication, is a \textit{wide-parent tolerance search} ($\pm 500$ Da).
+In that search, the precursor mass and the mass of the best matching theoretical peptide could be many Da apart.
+An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account at all.
+We describe an alternative to these search modes: the multi-notch search.
+The \textit{multi-notch} search is an extension of the single-notch search to multiple specific mass differences between precursor mass and theoretical mass~\ref{fig:fig1-searchtypes} where the mass difference is easily interpretable.
+It can also be seen as a narrowing of the wide-parent tolerance search to only include pre-specified mass differences.
+The advantages and disadvantages of this alternative strategy are described below.
 

### 2017-07-10 paper.tex
ajsfdkdsjf

@@
 5ppm, or 2.1 Da).
-We'll refer to this type of search here as a \textit{single notch} search.
+We'll refer to this type of search here as a \textit{single-notch} search.
 The other common search mode, used in the original G-PTM-D publication, is a \textit{wide-parent tolerance search} ($\pm 500$ Da).
@@
 
-\subsubsection{Notch Search vs Narrow-Mass Search}
+\subsubsection{Traditional Single-Notch Search vs. Limited Multi-Notch Search}
 
-The purpose of a narrow-mass search is to use the provided precursor mass along with fragmentation spectra information to find an exact match to the observed peptide in a protein database.
+The traditional single-notch search limits the range of acceptable mass differences between a candidate peptide and the observed experimental precursor in order to find an exact match to the observed peptide in a protein database.
+A mass tolerance of 5-20ppm is a common precursor mass tolerance that allows only mass differences between theoretical and experimental of 5-20 ppm.
+This is a highly restrictive search.
+It is not uncommon for the precursor mass tolerance of single-notch searches to be reduced to as much as 2 or 3 Da.
+One rationale for this is that errors in the deconvolution of MS1 spectra frequently result in an experimental precursor mass that is in error by 1-3 Da, corresponding to missed monoisotopic errors.
+This relaxed tolerance increases the number of correctly identified peptides.
+A second rationale for this approach is that multiple peptides are frequently co-isolated and co-fragmented with all peptide fragments being reported in the same tandem mass spectrum.
+Frequently a peptide with a mass differing from the precursor mass reported by the instrument is the identified peptide.
+In case the co-fragmented peptides have the same charge, the mass difference between the identified peptide and the isolated precursor can be as much as 2 or 3 Da, so a lose 2 or 3 Da tolerance is able to identify those.
+The major limitation to this relaxed parent mass tolerance is a significant increase in the number of false positive identifications (due to a vast expansion in the number of acceptable theoretical peptides and simultaneous failure to accommodate low mass modifications such as amidation and deamidation), arguably defeating the purpose of a narrow-mass search.
 
-Using a calibration or instrument noise percursor tolerance, such as 5ppm or 10ppm is a valid approach, since it identifies only exact matches.
-A major drawback of this strategy is that it misses monoisotope errors.
-To avoid this complication, and to get a more comprehensive list of matches, a wider search tolerance is often used, e.g. $\pm 3.5$ Da.
-This gives a substantially larger list of matches, and the resulting PSMs with non-zero mass shifts could be attributed to missed monoisotopes or co-isolated peptides.
-While this approach provides a larger list of confident PSMs, a substantial fraction of those is wrong, since certain mass shifts in the chosen Dalton range arise due to sequence variations or PTMs.
-These require an extensive framework (such as G-PTM-D) to identify and label them.
-More importantly, those modified peptides \textit{do not} correspond to the identified one, thus arguably defeating the purpose of a narrow-mass search.
+An alternative to these two traditional single-notch searches is a limited multi-notch search that allows mass differences between precursor and theoretical peptide of 0, 1.0029, 2.0051 or 3.0077 Da but limits the width about any one of those three tolerances to a few ppm (0-20 ppm).
+This approach provides a means of getting an increased number of correct identifications but mitigates problems with false positive identifications.
+This alternative would appear to result in loss of identification of co-isolated peptides.
+However, MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolute all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
+This expanded list of precursors is used in every spectrum match frequently resulting in multiple peptide identifications from single tandem mass spectrum.
 
-We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the narrow-mass searches.
+INSERT PARAGRAPH DESCRIBING RESULTS OF SINGLE NOTCH VERSUS LIMITED MULTI NOTCH \& COISOLATION
+
+We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the single-notch searches.
 

### 2017-07-10 paper.tex
djfksdjf

@@
 
-\subsubsection{Mass-Window Searches vs Open-Mass Search}
+\subsubsection{Multi-Notch Search vs. Wide-Mass Search}
 
@@
 
-\subsubsection{Discovery of Modifications not in Database}
-
-Previously described wide mass search strategies for discovery and localization of unknown modifications benefit from using alternative, carefully designed search modes based on interval and notch searches.
-We discovered that limiting mass shifts to a lower bound of -187 Da (corresponding to the largest mass difference that could be attributed to loss of a single residue, Tryptophan), and no upper bound, is an important step in eliminating spurious PSMs.
-Furthermore, treating highly suspect mass shifts corresponding to residue additions/substitutions/deletions in a separate notch search with individualized false discovery rate estimates is beneficial.
-An automated tool built into MetaMorpheus allows confidently identifying novel modifications based on these search results.
-
 \subsection{Enhanced G-PTM-D Results}
 
-A major improvement in the G-PTM-D performance arises from using calibrated spectra as inputs to the procedure.
-Calibration is instrumental in increasing the discernibility between different mass shifts.
-Mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:phosphoHistograms}.
-It is evident that the phosphorylation and sulfonation peaks are clearly separated and identifyable when the search is conducted on calibrated data.
+A major improvement in G-PTM-D performance arises from using calibrated spectra as inputs to the procedure. 
+Calibration and the associated increase in mass accuracy is essential for discrimination of small differences between mass shifts due to different modifications.
+To illustrate this, mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:phosphoHistograms}.
+It is evident from these histograms that the phosphorylation and sulfonation peaks are clearly separated and identifiable when the search is conducted on well-calibrated spectra.
 

### 2017-07-10 paper.tex
djkfdjf

@@
 \centering
-\caption{Effects of Replacing Initial Wide-Mass Search by a Notch Search}
+\caption{Effects of Replacing Initial Wide-Mass Search by a Multi-Notch Search}
 \label{my-labelff}
@@
 Notch searches allow using expanded modification lists that include chemical modifications such as adducts and large mass PTMs such as glycosylations.
-Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMS.
+Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMs.
 A curated set of notches and the corresponding modifications is given in Appendix A.
 The overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%.
-We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Daltons.
+We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da.
 

### 2017-07-10 paper.tex
final

@@
 
-Analogous to a notch search being an extension of the narrow-mass search, an interval search is a natural extension of the open-mass search.
-Restricting the allowed differences between the precursor and the identified peptide mass is often beneficial in searches that are intended to discover mass shift peaks.
-While this method is not used in the G-PTM-D pipeline, we include it here for completeness, and demonstrate the improved performance when it is used in a heuristic for the discovery of novel modifications.
+Chick~\citep{Chick_2015} showed that by significantly relaxing the parent mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified.
+In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in precursor and parent mass.
+A major limitation of the Chick work was the extremely long search times because of the vast number of theoretical candidate peptides that were evaluated for each and every MS2 spectrum.
+Recently, Kong\cite{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for wide parent mass tolerance searches.
+Search times for +/- 500 Da parent tolerances went from six years to six seconds.
+The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
+The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, but interpretation of the mass difference was left to the user.
+PTMs corresponding to the reported mass difference can frequently be inferred; such as when the mass difference reported matches a known PTM mass (e.g. 79.966331 Da for phosphorylation).
+However, mass differences can arise from not only one modification, but frequently from two or more modifications, some of which increase the mass and the peptide and some of which decrease it.
+In addition, small missed cleavages are sometimes reported mistakenly as mass differences.
+Also, some mass differences, such as those matching to amino acid masses, are easily interpreted but suffer from high local FDR and should only be used with caution.
+Finally, seemingly uninterpretable mass differences can be reported when the best matching peptide has an actual precursor mass different from the precursor mass selected by the mass spectrometer.
+The range of masses selected for fragmentation in a tandem mass spectrometer is commonly +/- 2 Da.
+Therefore, multiple peptides can be introduced to the cell and co-fragmented.
+This often results in peptide identification of a less abundant peptide with mass slightly larger or smaller than the selected precursor.
+The unconstrained utilization of mass difference and acceptance of the theoretical best matching sequence as true can result in a high number of false positive assignments.
 
-One way of restricting open mass searches is by introducing an upper or lower limit on the accepted mass differences.
-To illustrate the effect of such restrictions, we ran an open-mass search, and restricted ourselves to looking at 1\% FDR matches, both targets and decoys.
-The plots in Figure~\ref{fig:fig4-limitsOnOpenSearch} show the FDR in the list filtered by using an upper or lower bound cutoff on the allowed mass shifts.
-It shows the fundamental difference between upper and lower bound cutoffs: while it is beneficial to restrict open searches to ignore large negative values, restricting matches with a large positive mass difference has a negative effect on the overall FDR.
-This effect can be explained by the fact that it's easier to match long sequences to short peptide fragmentation profiles than it is to match short sequences to fragmentation pattern of a long peptide.
+A third alternative, reported here, is the use of a multi-notch search (>100 discrete notches) offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs and the least false positive PSMs (see the TABLE i and TABLE j).
+	Table I (Count of theoretical peptide candidates for each search mode (single notch 5ppm, single notch 3.5, limited multi notch, 500 Da, and full multi-notch).
+	Table J (Total PSMs for each search type – 500 or full multi-notch). 
+One output of an MSFragger search is a list of PSMs with each entry containing the mass difference between the isolated precursor and the mass of the best matching theoretical peptide.
+MSFragger also doesn’t automatically calibrate spectra prior to searching.
+Therefore, mass differences from closely spaced PTMs (e.g.
+phosphorylation and sulfonation) can easily be misinterpreted.
+MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.
 
-\begin{figure}
-\caption{Upper and Lower bound cutoffs in Mouse Data}
-\label{fig:fig4-limitsOnOpenSearch}
-\includegraphics{fig4-limitsOnOpenSearch}
-\end{figure}
-
-MetaMorpheus includes a capability to identify peaks in a mass shift histogram obtained from a search, using a recently described peak finding algorithm\cite{Rodriguez_2014}.
+G-PTM-D originally added PTMs with masses that were included in the UniProt ptmlist… We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search.
+The new notches correspond to metal adducts (e.g.
+Na and K), glycoslyations, sample handling and processing artifacts (e.g.
+carbamylation, ammonia loss), lipidation, and etc.
+Msfragger has limited mass range as reported.
 We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some histogram peaks are illustrated in Figure~\ref{fig:fig5-HistogramsAround80and113}, including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR.
-
-\begin{figure}
-\caption{Some Histogram Peaks For Mouse Data}
-\label{fig:fig5-HistogramsAround80and113}
-\includegraphics{fig5-HistogramsAround80and113}
-\end{figure}
-
-We identified the property of some specific mass shifts that correspond to anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
-
-Mass windows provide an mechanism to exclude these problematic mass shifts from a wide-mass search by employing a interval search.
-
-MSFragger, a method recently described in\cite{Kong_2017}, significantly decreases the runtime of wide-mass searches, but lacks specificity.
+Some of these problematic histogram peaks are illustrated in Figure [fig:figure3], including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR .
+Some Histogram Peaks For Mouse Data
+Some Histogram Peaks For Mouse Data
+We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals .
 

### 2017-07-10 paper.tex
minor
dmkfdsjf
yeah

@@
 Correct identification of protein post-translational modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
-G-PTM-D\cite{Li_2016} is a recently developed tool for global identification and localization of PTMs.
+G-PTM-D\citep{Li_2016} is a recently developed tool for global identification and localization of PTMs.
 Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and speed of PTM identification.
@@
 
-Many types of post-translational modifications (PTMs) are known, and databases containing detailed information about such modifications are readily available\cite{Creasy_2004}.
+Many types of post-translational modifications (PTMs) are known, and databases containing detailed information about such modifications are readily available\citep{Creasy_2004}.
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
 Protein databases such as UniProt include information on the localization of specific modifications to specific residues in the protein, but this information is incomplete.
-Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identification in complex biological samples has continued to pose a challenge for proteomics\cite{Olsen_2013}.
+Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identification in complex biological samples has continued to pose a challenge for proteomics\citep{Olsen_2013}.
 
@@
 phosphorylations) in well-studied systems.
-Global discovery tools, such as MODa\cite{Na_2011} are also available.
-G-PTM-D\cite{Li_2016} is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\cite{Chick_2015, Na_2011}, that provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
+Global discovery tools, such as MODa\citep{Na_2011} are also available.
+G-PTM-D\citep{Li_2016} is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\citep{Chick_2015, Na_2011}, that provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
@@
 \begin{figure}[H]
-  \includegraphics[scale=0.5]{diagram.eps}
-  \caption{Workflows}
+  \includegraphics{paperDiagram.png}
+  \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color.}
   \label{fgr:diagram}
@@
 
-We developed a modified version of the Morpheus software for bottom-up spectral database searching\cite{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
+We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
 Three multi-fraction mammalian datasets (Jurkat cell, C57BL/6J (B6) and CAST/EiJ (CAST) mouse islet lysates) with deep proteome coverage were used to evaluate performance.
-These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells, and 18 fractions from Mouse islets, and are described in detail in~\cite{Shortreed_2015, Cesnik_2016}.
+These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells, and 18 fractions from Mouse islets, and are described in detail in~\citep{Shortreed_2015, Cesnik_2016}.
 
@@
 
-A critical parameter for peptide identification is mass accuracy\cite{Scherl_2008}.
+A critical parameter for peptide identification is mass accuracy\citep{Scherl_2008}.
 Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
@@
 standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
-real-time calibration using a lock mass standard\cite{Olsen_2005}); and spectral calibration subsequent to the MS experiment (post-acquisition spectral calibration).
-We developed a post-acquisition calibration procedure that builds upon the software lock mass concept\cite{Cox_2011} recently reported by the Mann group.
+real-time calibration using a lock mass standard\citep{Olsen_2005}); and spectral calibration subsequent to the MS experiment (post-acquisition spectral calibration).
+We developed a post-acquisition calibration procedure that builds upon the software lock mass concept\citep{Cox_2011} recently reported by the Mann group.
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
 The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy\cite{Elias_2007}).
+This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy\citep{Elias_2007}).
 Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
@@
 
-\begin{figure}
-  \includegraphics{fig2-calibrationQuality.eps}
-  \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
-  \label{fgr:figure1}
-\end{figure}
+%\begin{figure}
+%  \includegraphics{calib1.png}
+ % \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
+ % \label{fgr:figure1}
+%\end{figure}
 
@@
 
-\begin{figure}
-\includegraphics{fig1-searchTypes.eps}
-\caption{Search Modes}
-\label{fig:fig1-searchtypes}
-\end{figure}
+%\begin{figure}
+%\includegraphics{fig1-searchTypes.png}
+%\caption{Search Modes}
+%\label{fig:fig1-searchtypes}
+%\end{figure}
 
@@
 
-\begin{figure}
-\caption{Mouse histograms around 0}
-\label{fig:fig3mouse-1012}
-\includegraphics{fig3mouse-1012}
-\end{figure}
+%\begin{figure}
+%\caption{Mouse histograms around 0}
+%\label{fig:fig3mouse-1012}
+%\includegraphics{fig3mouse-1012.png}
+%\end{figure}
 
-\begin{figure}
-\caption{Jurkat histograms around 0}
-\label{fig:fig3jurkat-1012}
-\includegraphics{fig3jurkat-1012}
-\end{figure}
+%\begin{figure}
+%\caption{Jurkat histograms around 0}
+%\label{fig:fig3jurkat-1012}
+%\includegraphics{fig3jurkat-1012.eps}
+%\end{figure}
 
@@
 A major limitation of the Chick work was the extremely long search times because of the vast number of theoretical candidate peptides that were evaluated for each and every MS2 spectrum.
-Recently, Kong\cite{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for wide parent mass tolerance searches.
+Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for wide parent mass tolerance searches.
 Search times for +/- 500 Da parent tolerances went from six years to six seconds.
@@
 
-\begin{figure}	
-	\centering
-	\begin{subfigure}[t]{3.1in}
-		\centering
-		\includegraphics[scale=0.5]{figPhosphorylationPre.eps}
-		\label{fig:1a}		
-	\end{subfigure}
-	\quad
-	\begin{subfigure}[t]{3.1in}
-		\centering
-		\includegraphics[scale=0.5]{figPhosphorylationCalib.eps}
-		\label{fig:1b}
-	\end{subfigure}
-	\caption{Histograms of mass differences observed around 80 Daltons when using calibrated and uncalibrated spectra}\label{fig:phosphoHistograms}
-\end{figure}
+%\begin{figure}	
+%	\centering
+%	\begin{subfigure}[t]{3.1in}
+%		\centering
+%		\includegraphics[scale=0.5]{figPhosphorylationPre.png}
+	%	\label{fig:1a}		
+%	\end{subfigure}
+%	\quad
+%	\begin{subfigure}[t]{3.1in}
+%		\centering
+%		\includegraphics[scale=0.5]{figPhosphorylationCalib.png}
+%		\label{fig:1b}
+%	\end{subfigure}
+%	\caption{Histograms of mass differences observed around 80 Daltons when using calibrated and uncalibrated %spectra}\label{fig:phosphoHistograms}
+%\end{figure}
 
@@
 
-Since the XML database is annotated with PTMs from the UniProt PTM list, it is natural to use it in the augmentation step as well.
+Since the protein database is annotated with PTMs from the UniProt PTM list, it is natural to use it in the augmentation step as well.
 It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many PTMs are missing, including adducts and glycosylations.

### 2017-07-10 GraphGeneration/fig1-SearchTypesDiagram.py

@@ -1,78 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-
-
-
-height = 5
-width = 3.25 
-plt.figure(figsize=(width, height))
-
-plt.subplot(411)
-
-dates = [-100,-1,0,1,100]
-values = [1,1,2,1,1]
-
-plt.plot(dates, values, '-')
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-plt.yticks(yTicks, ylabels)
-plt.title("Narrow-Mass Search")
-
-plt.subplot(412)
-
-dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
-values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
-
-plt.plot(dates, values, '-')
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.title("Notch Search")
-plt.ylim([0.5,2.5])
-
-plt.yticks(yTicks, ylabels)
-plt.subplot(413)
-
-
-
-
-dates =  [-100,-87,-86,-15,-14,-13,56,57,58,100]
-values = [1   ,  1, 2, 2,  1,  2, 2,1 ,2 ,2]
-plt.plot(dates, values, '-')
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-
-plt.title("Interval Search")
-plt.yticks(yTicks, ylabels)
-plt.subplot(414)
-
-
-
-dates =  [-100,100]
-values = [2 ,2]
-
-plt.plot(dates, values, '-')
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-
-plt.title("Wide-Mass Search")
-
-plt.yticks(yTicks, ylabels)
-plt.tight_layout()
-
-
-plt.savefig('fig1-searchTypes.eps', format='eps', dpi=1200)
-plt.savefig('fig1-searchTypes.png', format='png', dpi=1200)
-
-
-plt.show()

### 2017-07-10 GraphGeneration/fig2-calibQuality.py

@@ -1,97 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-maxi = 25000
-
-
-height = 6
-width = 7
-
-f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row', figsize=(width,height))
-ax1.set_title('Uncalibrated')
-ax2.set_title('Calibrated')
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-12-15-20-37\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
-A=ok[r"MassDiff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue_notch"]
-
-ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-ax1.set_ylim([1,maxi])
-ax1.vlines(0,0,maxi,zorder = 10)
-ax1.set_ylabel('PSM count')
-ax1.annotate(r"Mouse", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
-                size='large', ha='right', va='center', xytext=(0,0.5))
-
-mu = np.mean(A[(C<0.01) & B])
-sigma = np.std(A[(C<0.01) & B])
-textToDisplay = 'count = %d\n$\mu=%.2f$\n$\sigma=%.2f$\n68.3%%: [%.2f,%.2f]\n95.4%%: [%.2f,%.2f]\n99.7%%: [%.2f,%.2f]'%(len(A[(C<0.01) & B]), mu, sigma,mu-sigma, mu+sigma,mu-2*sigma, mu+2*sigma,mu-3*sigma, mu+3*sigma)
-print(textToDisplay)
-
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-12-15-21-10\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
-
-A=ok[r"MassDiff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue_notch"]
-
[…]

### 2017-07-10 GraphGeneration/fig2b-calibQuality.py

@@ -1,98 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-maxi = 25000
-
-
-height = 6
-width = 7
-
-f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row', figsize=(width,height))
-ax1.set_title('Uncalibrated')
-ax2.set_title('Calibrated')
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
-A=ok[r"Mass Diff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue Notch"]
-
-ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-ax1.set_ylim([1,maxi])
-ax1.vlines(0,0,maxi,zorder = 10)
-ax1.set_ylabel('PSM count')
-ax1.annotate(r"Mouse", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
-                size='large', ha='right', va='center', xytext=(0,0.5))
-
-mu = np.mean(A[(C<0.01) & B])
-sigma = np.std(A[(C<0.01) & B])
-textToDisplay = 'count = %d\n$\mu=%.2f$\n$\sigma=%.2f$\n68.3%%: [%.2f,%.2f]\n95.4%%: [%.2f,%.2f]\n99.7%%: [%.2f,%.2f]'%(len(A[(C<0.01) & B]), mu, sigma,mu-sigma, mu+sigma,mu-2*sigma, mu+2*sigma,mu-3*sigma, mu+3*sigma)
-print(textToDisplay)
-
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task3Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
-
-A=ok[r"Mass Diff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue Notch"]
-
[…]

### 2017-07-10 GraphGeneration/fig3-jurkat;-1;0;1;2.py

@@ -1,142 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task1Calibrate\2017-05-04-17-26-29\Task1Search\allPSMs_3.5aroundZero.psmtsv", sep='\t')
-
-
-
-
-height = 6
-width = 7
-plt.figure(figsize=(width, height))
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-massDiffs=data_sorted['Mass Diff (Da)'].values
-
-#data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
-#massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
-
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-ok = plt.subplot(221)
-
-halfWidth = 3.5
-min1 = -halfWidth
-max1 = halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-
-
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
-plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
-
-plt.legend()
-
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 2.5$ Da")
-ok.set_yscale("log", nonposy='clip')
-
-ok = plt.subplot(222)
-#ok.set_yscale("log", nonposy='clip')
-halfWidth = 0.1
-min1 = -1-halfWidth
-max1 = -1+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-07-10 GraphGeneration/fig3-mouse;-1;0;1;2.py

@@ -1,142 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task4Search\allPSMs_-187andUp.psmtsv", sep='\t')
-
-
-
-
-height = 6
-width = 7
-plt.figure(figsize=(width, height))
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-massDiffs=data_sorted['Mass Diff (Da)'].values
-
-#data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
-#massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
-
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-ok = plt.subplot(221)
-
-halfWidth = 3.5
-min1 = -halfWidth
-max1 = halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-
-
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
-plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
-
-plt.legend()
-
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 2.5$ Da")
-ok.set_yscale("log", nonposy='clip')
-
-ok = plt.subplot(222)
-#ok.set_yscale("log", nonposy='clip')
-halfWidth = 0.1
-min1 = -1-halfWidth
-max1 = -1+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-07-10 GraphGeneration/fig4-limitsOnOpenSearch.py

@@ -1,109 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-height = 3
-width = 7
-plt.figure(figsize=(width, height))
-
-
-offset = 200
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-07-06-17-29-30\Task1Search\aggregatePSMs_OpenSearch.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-05-05-18-29-43\Task2Calibrate\2017-07-07-13-19-26\Task1Search\aggregatePSMs_OpenSearch.psmtsv", sep='\t')
-
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-
-A=data_sorted['Mass Diff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-print(A[0])
-print(A[-1])
-
-AhighConf = A[C<0.01]
-BhighConf = B[C<0.01]
-
-print(AhighConf[0])
-print(AhighConf[-1])
-
-print(BhighConf[0])
-print(BhighConf[1])
-print(BhighConf[2])
-print(BhighConf[3])
-print(BhighConf[4])
-
-totalCount = len(AhighConf)
-   
-forward = [0]
-forwardDiv = []
-forwardInd= []
-for i in range(totalCount):
-    forward.append(forward[-1]+1 if BhighConf[i]=='D' else forward[-1])
-    forwardDiv.append(forward[-1]/(i+1))
-    forwardInd.append(AhighConf[i])
-
-print(forwardDiv[0])
-print(forwardDiv[1])
-print(forwardDiv[2])
-print(forwardDiv[3])
-print(forwardDiv[4])
-print(forwardDiv[5])
-print(forwardDiv[6])
-print(forwardDiv[7])
-print(forwardDiv[8])
-
-ok = plt.subplot(121)
-
-plt.scatter(forwardInd[offset:totalCount-offset:100], forwardDiv[offset:totalCount-offset:100])
-plt.xscale('symlog')
-plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
-plt.ylabel("FDR")
-plt.xlabel("Mass Diff")
[…]

### 2017-07-10 GraphGeneration/fig5-HistogramsAround80and113.py

@@ -1,68 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.mlab as mlab
-#import scipy
-import pandas
-
-
-height = 3
-width = 7
-
-
-plt.figure(figsize=(width, height))
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-04-14-16-50-03\Task1Search\allPSMs_-187andUp.psmtsv", sep='\t')
-
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-
-massDiffs=data_sorted['MassDiff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-ok = plt.subplot(121)
-
-halfWidth = 0.01
-center = 113.084064
-min1 = center-halfWidth
-max1 = center+halfWidth
-
-bins = 21 
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-plt.hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
-plt.hist(massDiffsHighConf1decoy, bins  =bins, range=[min1, max1], label = "Decoys")
-
-
-plt.legend()
-plt.annotate('Leucine\nResidue\nMass', xy=(113.08406, 35), xytext=(113.08406-0.00857, 40), arrowprops=dict(arrowstyle="->"))
-
-ok = plt.subplot(122)
-
-halfWidth = 0.01
-center = 79.961573 
-min1 = center-halfWidth
-max1 = center+halfWidth
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-plt.hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
[…]

### 2017-07-10 GraphGeneration/figAposter.py

@@ -1,84 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-
-
-matplotlib.rcParams.update({'font.size': 24})
-
-
-height = 10 
-width = 7 
-plt.figure(figsize=(width, height))
-
-plt.subplot(411)
-
-dates = [-100,-1,0,1,100]
-values = [1,1,2,1,1]
-
-plt.plot(dates, values, '-', linewidth=5)
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-plt.yticks(yTicks, ylabels)
-plt.title("Narrow-Mass Search")
-
-plt.subplot(412)
-
-dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
-values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
-
-plt.plot(dates, values, '-', linewidth=5)
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.title("Notch Search")
-plt.ylim([0.5,2.5])
-
-plt.yticks(yTicks, ylabels)
-plt.subplot(413)
-
-
-
-
-dates =  [-100,-87,-86,-15,-14,-13,56,57,58,100]
-values = [1   ,  1, 2, 2,  1,  2, 2,1 ,2 ,2]
-plt.plot(dates, values, '-', linewidth=5)
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-
-plt.title("Interval Search")
-plt.yticks(yTicks, ylabels)
-hehe =plt.subplot(414)
-
-
-
-dates =  [-100,100]
-values = [2 ,2]
-
-plt.plot(dates, values, '-', linewidth=5)
-
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-
-hehe.set_xlabel('Mass Diff (Da)')
-
-plt.title("Wide-Mass Search")
-
-plt.yticks(yTicks, ylabels)
-plt.tight_layout()
-
-
-plt.savefig('figAposter.eps', format='eps', dpi=1200)
-plt.savefig('figAposter.png', format='png', dpi=1200)
-
-
-plt.show()

### 2017-07-10 GraphGeneration/figBposter.py

@@ -1,48 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-import pandas
-
-
-matplotlib.rcParams.update({'font.size': 24})
-
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task4Search\allPSMs_-187andUp.psmtsv", sep='\t')
-
-height = 5
-width = 7
-plt.figure(figsize=(width, height))
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-massDiffs=data_sorted['Mass Diff (Da)'].values
-
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-halfWidth = 0.1 
-min1 = 1-halfWidth
-max1 = 1+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
-
-plt.ylabel("PSM Count")
-plt.axis([0.9, 1.1, 0, 1200])
-
-plt.title("Mass Diffs Around 1 Da")
-
-plt.annotate('Deamidation', xy=(0.984016, 900), xytext=(0.984016+0.02, 700), arrowprops=dict(arrowstyle="->"))
-plt.annotate('1 Mm', xy=(1.0029, 150), xytext=(1.027, 180), arrowprops=dict(arrowstyle="->"))
-
-
-plt.tight_layout()
-
-plt.savefig('figBposter.eps', format='eps', dpi=1200)
-plt.savefig('figBposter.png', format='png', dpi=1200)
-
-
-plt.show()

### 2017-07-10 GraphGeneration/figCalibrationPoster.py

@@ -1,56 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-import pandas
-
-maxi = 30000
-
-matplotlib.rcParams.update({'font.size': 24})
-
-height = 5
-width = 14 
-
-f, ((ax1, ax2)) = plt.subplots(1, 2, sharex='col', sharey='row', figsize=(width,height))
-ax1.set_title('Uncalibrated')
-ax2.set_title('Calibrated')
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
-A=ok[r"Mass Diff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue Notch"]
-
-ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-ax1.set_ylim([1,maxi])
-ax1.vlines(0,0,maxi,zorder = 10)
-ax1.set_ylabel('PSM count')
-ax1.set_xlabel('Error (PPM)')
-
-mu = np.mean(A[(C<0.01) & B])
-sigma = np.std(A[(C<0.01) & B])
-textToDisplay = 'count = %d\n$\mu=%.2f$\n$\sigma=%.2f$\n68.3%%: [%.2f,%.2f]\n95.4%%: [%.2f,%.2f]\n99.7%%: [%.2f,%.2f]'%(len(A[(C<0.01) & B]), mu, sigma,mu-sigma, mu+sigma,mu-2*sigma, mu+2*sigma,mu-3*sigma, mu+3*sigma)
-print(textToDisplay)
-
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task3Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
-
-A=ok[r"Mass Diff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue Notch"]
-
-ax2.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-ax2.set_ylim([1,maxi])
-ax2.vlines(0,0,maxi,zorder = 10)
[…]

### 2017-07-10 GraphGeneration/figCposter.py

@@ -1,100 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-import pandas
-
-
-matplotlib.rcParams.update({'font.size': 24})
-
-
-height = 5
-width = 7
-plt.figure(figsize=(width, height))
-
-
-offset = 200
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task4Search\allPSMs_OpenSearch.psmtsv", sep='\t')
-
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-
-A=data_sorted['Mass Diff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-print(A[0])
-print(A[-1])
-
-AhighConf = A[C<0.01]
-BhighConf = B[C<0.01]
-
-print(AhighConf[0])
-print(AhighConf[-1])
-
-print(BhighConf[0])
-print(BhighConf[1])
-print(BhighConf[2])
-print(BhighConf[3])
-print(BhighConf[4])
-
-totalCount = len(AhighConf)
-   
-forward = [0]
-forwardDiv = []
-forwardInd= []
-for i in range(totalCount):
-    forward.append(forward[-1]+1 if BhighConf[i]=='D' else forward[-1])
-    forwardDiv.append(forward[-1]/(i+1))
-    forwardInd.append(AhighConf[i])
-
-print(forwardDiv[0])
-print(forwardDiv[1])
-print(forwardDiv[2])
-print(forwardDiv[3])
-print(forwardDiv[4])
-print(forwardDiv[5])
-print(forwardDiv[6])
-print(forwardDiv[7])
-print(forwardDiv[8])
-
-plt.scatter(forwardInd[offset:totalCount-offset:100], forwardDiv[offset:totalCount-offset:100])
-plt.xscale('symlog')
-plt.xticks([-1e4, -1e3, -1e2, -1e1, 0,1e1, 1e2, 1e3, 1e4])
-plt.ylabel("FDR")
-plt.xlabel("Mass Diff Lower Limit")
-plt.title("Restricting the Open Search")
-
-
-backward = [0]
-backwardDiv= []
-backwardInd= []
-for i in range(totalCount-1,-1,-1):
-    backward.append(backward[-1]+1 if BhighConf[i]=='D' else backward[-1])
[…]

### 2017-07-10 GraphGeneration/figPhosphorylationPoster.py

@@ -1,47 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-import pandas
-
-
-matplotlib.rcParams.update({'font.size': 24})
-
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-30-10-25-53\Task1Search\allPSMs_name.psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-05-30-10-26-30\Task1Search\allPSMs_name.psmtsv", sep='\t')
-
-
-height = 5
-width = 7
-plt.figure(figsize=(width, height))
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-massDiffs=data_sorted['Mass Diff (Da)'].values
-
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-halfWidth = 0.03 
-min1 = 79.961573-halfWidth
-max1 = 79.961573+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="T")]
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
-
-plt.ylabel("PSM Count")
-plt.xlabel("Mass Diff (Da)")
-plt.axis([min1, max1, 0, 150])
-
-plt.title("Original")
-
-plt.tight_layout()
-
-plt.savefig('figPhosphorylationPre.eps', format='eps', dpi=1200)
-plt.savefig('figPhosphorylationPre.png', format='png', dpi=1200)
-
-
-plt.show()

### 2017-07-10 GraphGeneration/figPhosphorylationPosterCalib.py

@@ -1,52 +1,0 @@
-import numpy as np
-import matplotlib
-import matplotlib.pyplot as plt
-import pandas
-
-
-matplotlib.rcParams.update({'font.size': 24})
-
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-30-10-25-53\Task1Search\allPSMs_name.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-05-30-10-26-30\Task1Search\allPSMs_name.psmtsv", sep='\t')
-
-
-height = 5
-width = 7
-plt.figure(figsize=(width, height))
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-massDiffs=data_sorted['Mass Diff (Da)'].values
-
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-halfWidth = 0.03 
-min1 = 79.961573-halfWidth
-max1 = 79.961573+halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="T")]
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
-
-plt.ylabel("PSM Count")
-plt.xlabel("Mass Diff (Da)")
-
-
-plt.annotate('Sulfo', xy=(79.956815, 50), xytext=(79.94, 50), arrowprops=dict(arrowstyle="->"))
-plt.annotate('Phospho', xy=( 79.966331, 100), xytext=(79.97, 100), arrowprops=dict(arrowstyle="->"))
-
-plt.axis([min1, max1, 0, 150])
-
-plt.title("Calibrated")
-
-plt.tight_layout()
-
-plt.savefig('figPhosphorylationCalib.eps', format='eps', dpi=1200)
-plt.savefig('figPhosphorylationCalib.png', format='png', dpi=1200)
-
-
-plt.show()

### 2017-07-10 GraphGeneration/genericCalibQuality.py

@@ -1,90 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-maxi = 3000
-
-
-#start =  r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task1Calibrate\120426_Jurkat_highLC_Frac19_allPSMs_10ppmAroundZero"
-start =  r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-04-10-15-00-59\Task1Calibrate\120426_Jurkat_highLC_Frac14_allPSMs_10ppmAroundZero"
-
-#start =  r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac3_6uL_allPSMs_10ppmAroundZero"
-#start =  r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-30-13_CAST_Frac9_9p5uL_allPSMs_10ppmAroundZero"
-#start =  r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\04-29-13_B6_Frac9_9p5uL_allPSMs_10ppmAroundZero"
-
-
-plt.figure(num=1,figsize=(10.9,6))
-
-plt.suptitle(start.split('\\')[-1], fontsize=14)
-
-plt.figure(1)
-
-
-ok = pandas.read_csv(start+".psmtsv", sep='\t')  
-A=ok[r"MassDiff (ppm)"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue_notch"]
-
-
-ok = plt.subplot(131)
-plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-plt.ylim([1,maxi])
-plt.vlines(0,0,maxi*.65,zorder = 10)
-
-mu = np.mean(A[(C<0.01) & B])
-sigma = np.std(A[(C<0.01) & B])
[…]

### 2017-07-10 GraphGeneration/genericCoIsolationHistograms.py

@@ -1,175 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Desktop\02-15-17_Cys-tag_light\2017-02-25-14-00-36\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-
-
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-02-25-13-59-40\Task2Calibrate\2017-03-31-12-50-27\Task1Search\04-30-13_CAST_Frac5_4uL-Calibrated2.1aroundZero.psmtsv", sep='\t')
-
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-
-#massDiffs=data_sorted['MassDiff (Da)'].values
-massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-ok = plt.subplot(321)
-
-halfWidth = 2.5
-min1 = -halfWidth
-max1 = halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-
-
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
-plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
-
-plt.legend()
-
-#plt.xlabel("Mass Diff")
[…]

### 2017-07-10 GraphGeneration/genericCustomPeakPlot.py

@@ -1,43 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import matplotlib.mlab as mlab
-import pandas
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Jurkat\2017-02-25-13-57-45\Task4Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Yeast\2017-03-13-14-25-38\Task2Search\aggregateintervals[-187.000-Infinity].psmtsv", sep='\t')
-
-data_sorted = data.sort_values(['MassDiff (Da)'], ascending=True)
-
-massDiffs=data_sorted['MassDiff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue_notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-center = 57.021464
-fig, axes = plt.subplots(nrows=1, ncols=2)
-
-halfWidth = 0.05
-min1 = center-halfWidth
-max1 = center+halfWidth
-
-bins = 25
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-axes[0].hist(massDiffsHighConf1, bins  = bins, range=[min1, max1], label = "Targets")
-axes[0].hist(massDiffsHighConf1decoy, bins  =bins, range=[min1, max1], label = "Decoys")
-
-
-halfWidth = 0.01
-min1 = center-halfWidth
-max1 = center+halfWidth
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-axes[1].hist(massDiffsHighConf1, bins  = 25, range=[min1, max1], label = "Targets")
-axes[1].hist(massDiffsHighConf1decoy, bins  =25, range=[min1, max1], label = "Decoys")
-
-
-
-
-plt.show()

### 2017-07-10 GraphGeneration/genericErrorVsVariousParams.py

@@ -1,55 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-nice= pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-08-14-03-16\Task1Calibrate\04-29-13_B6_Frac9_9p5uLafterfc4after.ms1dptsv", sep='\t')
-#nice= pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-08-14-03-16\Task1Calibrate\04-29-13_B6_Frac9_9p5uLafterfc4after.ms2dptsv", sep='\t')
-
-
-plt.figure(1)
-plt.subplot(331)
-plt.plot(nice['mz'],nice['Label'], 'bo',markersize = .3)
-plt.hlines(0, min(nice['mz']), max(nice['mz']), zorder = 10)
-plt.xlabel("mz")
-
-plt.subplot(332)
-plt.plot(nice['rt'],nice['Label'], 'bo',markersize = .3)
-plt.hlines(0, min(nice['rt']), max(nice['rt']), zorder = 10)
-plt.xlabel("rt")
-
-plt.subplot(333)
-plt.plot(np.log(nice['intensity']),nice['Label'], 'bo',markersize = .3)
-plt.hlines(0, min(np.log(nice['intensity'])), max(np.log(nice['intensity'])), zorder = 10)
-plt.xlabel("intensity")
-
-plt.subplot(334)
-plt.plot(np.log(nice['TIC']),nice['Label'], 'bo',markersize = .3)
-plt.hlines(0, min(np.log(nice['TIC'])), max(np.log(nice['TIC'])), zorder = 10)
-plt.xlabel("TIC")
-
-plt.subplot(335)
-plt.plot(np.log(nice['InjectionTime']),nice['Label'], 'bo',markersize = .3)
-plt.hlines(0, min(np.log(nice['InjectionTime'])), max(np.log(nice['InjectionTime'])), zorder = 10)
-plt.xlabel("InjectionTime")
-
-plt.subplot(336)
-plt.plot(nice['Score'],nice['Label'], 'bo',markersize = .3)
-plt.hlines(0, min(nice['Score']), max(nice['Score']), zorder = 10)
-plt.xlabel("score")
-
-plt.subplot(337)
-plt.plot(nice['Peptide Monoisotopic Mass'],nice['Label'], 'bo',markersize = .3)
[…]

### 2017-07-10 GraphGeneration/genericHistogramsAroundZero.py

@@ -1,178 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-05-04-11-56-07\Task1Search\04-30-13_CAST_Frac5_4uL-Calibrated_allPSMs_3.5aroundZero.psmtsv", sep='\t')
-#data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-05-04-11-56-07\Task2Search\04-30-13_CAST_Frac5_4uL-Calibrated_allPSMs_3.5aroundZero.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-04-07-18-09-31\Task1Calibrate\2017-05-04-11-56-07\Task3Search\04-30-13_CAST_Frac5_4uL-Calibrated_allPSMs_3.5aroundZero.psmtsv", sep='\t')
-
-
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-
-massDiffs=data_sorted['Mass Diff (Da)'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-massDiffsHighConf = massDiffs[C<0.01]
-TDhighConf = B[C<0.01]
-
-ok = plt.subplot(321)
-
-halfWidth = 3.5
-min1 = -halfWidth
-max1 = halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-
-
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
-plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
-
-plt.legend()
-
-#plt.xlabel("Mass Diff")
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 2.5$")
-ok.set_yscale("log", nonposy='clip')
-
[…]

### 2017-07-10 paper.tex
jfkdjf

@@
 The final narrow-mass search is also problematic, since it does not account for peptide-spectral matches(PSMs) with mass differences arising due to misidentifying the monoisotopic peak.
-In G-PTM-D, modification discovery is limited to PTMs in the Uniprot database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
+In G-PTM-D, modification discovery is limited to PTMs in the UniProt database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
 Finally, PTMs that are similar in mass, such as phosphorylation (79.966331 Da) and sulfonation (79.956815 Da), are problematic since they are virtually indistinguishable in unprocessed spectral files.
@@
 The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in precursor mass deconvolution (+1.0029, + 2.0052 and + 3.0077 Da), further increasing the number of confidently identified peptides and PTMs.
-Finally, we replace the limited Uniprot PTM list with a customizeable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
+Finally, we replace the limited UniProt PTM list with a customizable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
 This change not only expands the list of identified and localized modifications, but also indirectly improves the confident PSM count.
@@
 
-Uniprot XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
+UniProt XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
 
@@
 replacement of the final traditional narrow mass search with a limited multi-notch search that permits identification of incorrectly deconvoluted precursor masses;
-and lastly, addition of capacity for identification of multiple cofragmented peptides from a single MS/MS scan.
+and lastly, addition of capacity for identification of multiple co-fragmented peptides from a single MS/MS scan.
 To clarify the advantage of each change, they are first discussed separately, followed by an analysis of their combined effect upon overall G-PTM-D performance.
@@
 \label{tbl:calib}
-\begin{tabular}{@{}llllll@{}}
-                                                                                &        & \multicolumn{2}{l}{Mouse} & \multicolumn{2}{l}{Jurkat} \\ \cmidrule(l){3-6} 
-                                                                                &        & Orig        & Calib       & Orig        & Calib        \\ \cmidrule(l){3-6} 
-\multicolumn{2}{l}{PSM Count}                                                            &    152208       &        155617 &         217612   &   219172        \\
-\multicolumn{2}{l}{Average Error (ppm)}                                                           &    -1.09         &       -0.16      &      -2.10      &        -0.16      \\
-\multicolumn{2}{l}{St. Dev of Error}                                                     &     3.35        &      0.90       &       2.15      &        0.91      \\
-\multirow{3}{*}{\begin{tabular}[c]{@{}l@{}}Confidence\\ Intervals\end{tabular}} & 68.3\% &   [-4.44,2.27]          &    [-1.06,0.74]          &    [-4.26,0.05]          &  [-1.08,0.75]            \\
+\begin{tabular}{llllll}
+
+                                                                                &        & Orig        & Calib       & Orig        & Calib        \\
+ &PSM Count                                                           &    152208       &        155617 &         217612   &   219172        \\
+ &Average Error (ppm)                                                           &    -1.09         &       -0.16      &      -2.10      &        -0.16      \\
+ &St. Dev of Error                                                     &     3.35        &      0.90       &       2.15      &        0.91      \\
+Confidence Intervals& 68.3\% &   [-4.44,2.27]          &    [-1.06,0.74]          &    [-4.26,0.05]          &  [-1.08,0.75]            \\
                                                                                 & 95.4\% &   [-7.80,5.62]          &    [-1.97,1.64]         &   [-6.41,2.20]          &      [-1.99,1.61]         \\
-                                                                                & 99.7\% &   [-11.15,8.98]          &      [-2.87,2.54]        &    [-8.56,4.36]          &      [-2.90,2.58]         \\ \cmidrule(l){3-6}
+                                                                                & 99.7\% &   [-11.15,8.98]          &      [-2.87,2.54]        &    [-8.56,4.36]          &      [-2.90,2.58]         \\
 \end{tabular}
@@
 The \textit{multi-notch} search is an extension of the single-notch search to multiple specific mass differences between precursor mass and theoretical mass~\ref{fig:fig1-searchtypes} where the mass difference is easily interpretable.
-It can also be seen as a narrowing of the wide-parent tolerance search to only include pre-specified mass differences.
+It can also be seen as a narrowing of the wide-parent tolerance search to only include prespecified mass differences.
 The advantages and disadvantages of this alternative strategy are described below.
@@
 
-INSERT PARAGRAPH DESCRIBING RESULTS OF SINGLE NOTCH VERSUS LIMITED MULTI NOTCH \& COISOLATION
+INSERT PARAGRAPH DESCRIBING RESULTS OF SINGLE NOTCH VERSUS LIMITED MULTI NOTCH \& CO-ISOLATION
 
@@
 \label{my-label}
-\begin{tabular}{@{}llll@{}}
-                    & 5ppm Search & 3.5 Da Search & Notch Search \\ \cmidrule(l){2-4}
+\begin{tabular}{llll}
+                    & 5ppm Search & 3.5 Da Search & Notch Search \\
 PSMs Within 1\% FDR & 157188      & 178003        &        \\
-MisAssigned PSMs    & 0           &               & 0            \\
+misassigned PSMs    & 0           &               & 0            \\
 Good PSMs           & 157188      &         &        \\ 
@@
 One output of an MSFragger search is a list of PSMs with each entry containing the mass difference between the isolated precursor and the mass of the best matching theoretical peptide.
-MSFragger also doesn’t automatically calibrate spectra prior to searching.
-Therefore, mass differences from closely spaced PTMs (e.g.
-phosphorylation and sulfonation) can easily be misinterpreted.
+MSFragger also doesn't automatically calibrate spectra prior to searching.
+Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
 MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.
 
-G-PTM-D originally added PTMs with masses that were included in the UniProt ptmlist… We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search.
-The new notches correspond to metal adducts (e.g.
-Na and K), glycoslyations, sample handling and processing artifacts (e.g.
-carbamylation, ammonia loss), lipidation, and etc.
-Msfragger has limited mass range as reported.
+G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list.
+We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search.
+The new notches correspond to metal adducts (e.g. Na and K), glycosylations., sample handling and processing artifacts (e.g. carbamylation, ammonia loss), lipidation, and etc.
+MSFragger has limited mass range as reported.
 We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some of these problematic histogram peaks are illustrated in Figure [fig:figure3], including peaks corresponding to masses of Sulfo and Phospho, and a problematic peak at mass of Leucine, with a high FDR .
+Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of Sulfonation and Phosphorylation, and a problematic peak at mass of Leucine, with a high FDR.
 Some Histogram Peaks For Mouse Data
@@
 
-%\begin{figure}	
-%	\centering
-%	\begin{subfigure}[t]{3.1in}
-%		\centering
-%		\includegraphics[scale=0.5]{figPhosphorylationPre.png}
-	%	\label{fig:1a}		
-%	\end{subfigure}
-%	\quad
-%	\begin{subfigure}[t]{3.1in}
-%		\centering
-%		\includegraphics[scale=0.5]{figPhosphorylationCalib.png}
-%		\label{fig:1b}
-%	\end{subfigure}
-%	\caption{Histograms of mass differences observed around 80 Daltons when using calibrated and uncalibrated %spectra}\label{fig:phosphoHistograms}
-%\end{figure}
-
-By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in Uniprot, the number of confidently identified PSMs increased by 6\%.
+By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in UniProt, the number of confidently identified PSMs increased by 6\%.
 The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
@@
 \begin{tabular}{ll|l|l|l}
-                      &        & Only Uniprot & Uniprot+Glyco & Uniprot+Glyco+Adducts\\
+                      &        & Only UniProt & UniProt+Glycosylations & UniProt+Glycosylations+Adducts\\
 \hline

### 2017-07-12 paper.tex
eyreur

@@
 A 10ppm precursor mass tolerance is used for the initial calibration step, and then reduced to 5ppm for subsequent searches.
-We allowed 120 modifications including PTMs, adducts, and chemical modifications in the database augmentation step.
+We allowed 120 modifications including PTMs, adducts, and chemical modifications in the database augmentation step, see Supplementary Material 1 for a complete listing.
 
@@
 Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
-The calibration algorithm is described in detail in Supplementary Material 1.
+The calibration algorithm is described in detail in Supplementary Material 2.
 
-Calibration quality is evaluated through the results of a standard 10ppm precursor tolerance database search.
+We demonstrated the effectiveness of the calibration approach using a single CAST Mouse raw file as an example by displaying the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, before and after the calibration procedure is applied.
+In Figure [fgr:fig1-calibErrors] these errors are plotted versus two important variables: the m/z value of the peak and the retention time.
+It can be seen in the graphs of Figure ### that calibration successfully addresses errors in non-linearity as a function of both retention time and m/z.
+We note that in addition to these two parameters, the calibration function native to MetaMorpheus also considers non-linearities in peak intensity, injection time, total ion count, etc.
+In addition, MetaMorpheus calibrates both MS1 and MS2 (using MS2 fragment matches), but only results from MS1 are shown here.
+The complete list of parameters analyzed in the calibration is provided as Table $$$ in the supplement.
+
+\begin{figure}[H]
+ \includegraphics{fig1-calibErrors.png}
+ \caption{Errors in m/z plotted vs the m/z value and the retention time}
+ \label{fgr:fig1-calibErrors}
+\end{figure}
+
+Calibration quality is further evaluated through the results of a standard 10ppm precursor tolerance database search.
 PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure (Table~\ref{tbl:calib}).
@@
 The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which show that the calibrated mass spectra have sub-ppm mass errors.
-Histograms of the mass errors before and after calibration are plotted in Figure~\ref{fgr:figure1}.
+Histograms of the mass errors in identified peptides before and after calibration are plotted in Figure~\ref{fgr:fig2-10ppmSearchCalib}.
 
@@
 \label{tbl:calib}
-\begin{tabular}{llllll}
+\begin{tabular}{llll}
 
-                                                                                &        & Orig        & Calib       & Orig        & Calib        \\
- &PSM Count                                                           &    152208       &        155617 &         217612   &   219172        \\
- &Average Error (ppm)                                                           &    -1.09         &       -0.16      &      -2.10      &        -0.16      \\
- &St. Dev of Error                                                     &     3.35        &      0.90       &       2.15      &        0.91      \\
-Confidence Intervals& 68.3\% &   [-4.44,2.27]          &    [-1.06,0.74]          &    [-4.26,0.05]          &  [-1.08,0.75]            \\
-                                                                                & 95.4\% &   [-7.80,5.62]          &    [-1.97,1.64]         &   [-6.41,2.20]          &      [-1.99,1.61]         \\
-                                                                                & 99.7\% &   [-11.15,8.98]          &      [-2.87,2.54]        &    [-8.56,4.36]          &      [-2.90,2.58]         \\
+                                                                                &      & Orig        & Calib        \\
+ &PSM Count                                                           &      217612   &   219172        \\
+ &Average Error (ppm)                                                              &      -2.10      &        -0.16      \\
+ &St. Dev of Error                                                    &       2.15      &        0.91      \\
+Confidence Intervals& 68.3\%     &    [-4.26,0.05]          &  [-1.08,0.75]            \\
+                                                                                & 95.4\%      &   [-6.41,2.20]          &      [-1.99,1.61]         \\
+                                                                                & 99.7\%       &    [-8.56,4.36]          &      [-2.90,2.58]         \\
 \end{tabular}
@@
 
-%\begin{figure}
-%  \includegraphics{calib1.png}
- % \caption{Histograms demonstrating the effect of calibration on PSM mass errors}
- % \label{fgr:figure1}
-%\end{figure}
+\begin{figure}[H]
+ \includegraphics{fig2-10ppmSearchCalib.png}
+ \caption{Errors in m/z plotted vs the m/z value and the retention time}
+ \label{fgr:fig2-10ppmSearchCalib}
+\end{figure}
 
@@
 We describe an alternative to these search modes: the multi-notch search.
-The \textit{multi-notch} search is an extension of the single-notch search to multiple specific mass differences between precursor mass and theoretical mass~\ref{fig:fig1-searchtypes} where the mass difference is easily interpretable.
+The \textit{multi-notch} search is an extension of the single-notch search to multiple specific mass differences between precursor mass and theoretical mass~\ref{fig:fig3-searchTypes} where the mass difference is easily interpretable.
 It can also be seen as a narrowing of the wide-parent tolerance search to only include prespecified mass differences.
@@
 
-%\begin{figure}
-%\includegraphics{fig1-searchTypes.png}
-%\caption{Search Modes}
-%\label{fig:fig1-searchtypes}
-%\end{figure}
+\begin{figure}[H]
+\includegraphics{fig3-searchTypes.png}
+\caption{Search Modes}
+\label{fig:fig3-searchTypes}
+\end{figure}
 

### 2017-07-12 paper.tex
jdfkdjf

@@
 We demonstrated the effectiveness of the calibration approach using a single CAST Mouse raw file as an example by displaying the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, before and after the calibration procedure is applied.
-In Figure [fgr:fig1-calibErrors] these errors are plotted versus two important variables: the m/z value of the peak and the retention time.
-It can be seen in the graphs of Figure ### that calibration successfully addresses errors in non-linearity as a function of both retention time and m/z.
+In Figure\ref{fgr:fig1-calibErrors} these errors are plotted versus two important variables: the m/z value of the peak and the retention time.
+It can be seen in the graphs of Figure\ref{fgr:fig1-calibErrors} that calibration successfully addresses errors in non-linearity as a function of both retention time and m/z.
 We note that in addition to these two parameters, the calibration function native to MetaMorpheus also considers non-linearities in peak intensity, injection time, total ion count, etc.
 In addition, MetaMorpheus calibrates both MS1 and MS2 (using MS2 fragment matches), but only results from MS1 are shown here.
-The complete list of parameters analyzed in the calibration is provided as Table $$$ in the supplement.
+The complete list of parameters analyzed in the calibration is provided as Table lala in the supplement.
 
@@
 
-Calibration quality is further evaluated through the results of a standard 10ppm precursor tolerance database search.
+Calibration quality improvements are further shown through the results of a standard 10ppm precursor tolerance database search (Table\ref{tbl:calib}).
 PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure (Table~\ref{tbl:calib}).
@@
 Histograms of the mass errors in identified peptides before and after calibration are plotted in Figure~\ref{fgr:fig2-10ppmSearchCalib}.
+The main results of calibration in summary are increased PSMs at 1\% FDR, decreased error in precursor mass, and decreased standard deviation in average mass error.
 
@@
 
-Prior to discussing the impact that multi-notch searches have on the G-PTM-D performance, it is useful to describe the variety of search tolerances used.
-Traditional searches in bottom-up proteomics score potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal, within some defined tolerance (e.g.
+It is useful to describe the variety of search tolerances used, prior to discussing the impact that multi-notch searches have on the G-PTM-D performance.
+Traditional searches in bottom-up proteomics score potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal within some defined tolerance (e.g.
 5ppm, or 2.1 Da).
 We'll refer to this type of search here as a \textit{single-notch} search.
-The other common search mode, used in the original G-PTM-D publication, is a \textit{wide-parent tolerance search} ($\pm 500$ Da).
-In that search, the precursor mass and the mass of the best matching theoretical peptide could be many Da apart.
-An even less restrictive \textit{open-mass} search, matches peptides from a database to fragmentation spectra without taking the precursor mass into account at all.
+The other common search mode is a \textit{wide-parent tolerance search} (e.g. +/- 500Da)[Chick, Kong, Qiyao].
+In searches of that type, the precursor mass and the mass of the best matching theoretical peptide could be many Da apart.
 We describe an alternative to these search modes: the multi-notch search.

### 2017-07-12 FigureGeneration/Extras/SingleFileCalibQuality.py

@@ -1,0 +1,70 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+maxi = 3000
+
+
+start = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-07-10-17-50-26\Task1Calibrate\04-30-13_CAST_Frac5_4uL"
+
+plt.figure(num=1,figsize=(10.9,6))
+
+plt.suptitle(start.split('\\')[-1], fontsize=14)
+
+plt.figure(1)
+
+
+ok = pandas.read_csv(start+"PSMsBeforeLinearCalib.psmtsv", sep='\t')  
+#A=ok[r"Mass Diff (ppm)"]
+A=ok[r"Improvement Possible"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue"]
+
+
+ok = plt.subplot(131)
+plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+plt.ylim([1,maxi])
+plt.vlines(0,0,maxi*.65,zorder = 10)
+
+plt.title("Uncalibrated")
+
+
+ok = pandas.read_csv(start+"PSMsBeforeNonLinearCalib.psmtsv", sep='\t')  
+
+#A=ok[r"Mass Diff (ppm)"]
+A=ok[r"Improvement Possible"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue"]
+
+
+
+
+
+ok = plt.subplot(132)
+plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+plt.ylim([1,maxi])
+plt.vlines(0,0,maxi*.65,zorder = 10)
+
+
+plt.title("Linear Calibration")
+
+ok = pandas.read_csv(start+"PSMsAfterCalib.psmtsv", sep='\t')  
+
+#A=ok[r"Mass Diff (ppm)"]
+A=ok[r"Improvement Possible"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue"]
+
+
+
+ok = plt.subplot(133)
+plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+plt.ylim([1,maxi])
+plt.vlines(0,0,maxi*.65,zorder = 10)
+
+
+plt.title("Nice Calibration")
+
+plt.savefig('myimage.png', format='png', dpi=1200)
+
+plt.show()

### 2017-07-12 FigureGeneration/fig1-calibErrors.py

@@ -1,0 +1,43 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+nice= pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-07-10-17-50-26\Task1Calibrate\04-30-13_CAST_Frac5_4uLinLinearbeforesc1.ms1dptsv", sep='\t')
+nice2= pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-07-10-17-50-26\Task1Calibrate\04-30-13_CAST_Frac5_4uLinNonLinearafterfc4.ms1dptsv", sep='\t')
+
+
+height = 6
+width = 7
+
+f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row', figsize=(width,height))
+
+ax1.plot(nice['mz'],nice['Label'], 'bo',markersize = .3)
+ax1.hlines(0, min(nice['mz']), max(nice['mz']), zorder = 10)
+
+ax1.set_ylabel('m/z error')
+ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
+                size='large', ha='right', va='center', xytext=(0,0.5))
+
+ax2.plot(nice['rt'],nice['Label'], 'bo',markersize = .3)
+ax2.hlines(0, min(nice['rt']), max(nice['rt']), zorder = 10)
+
+ax3.plot(nice2['mz'],nice2['Label'], 'bo',markersize = .3)
+ax3.hlines(0, min(nice2['mz']), max(nice2['mz']), zorder = 10)
+
+ax3.set_ylabel('m/z error')
+ax3.set_xlabel('m/z')
+
+ax3.annotate(r"Calibrated", xy=(0, 0.5), xycoords=ax3.yaxis.label, textcoords='offset points',
+                size='large', ha='right', va='center', xytext=(0,0.5))
+
+ax4.plot(nice2['rt'],nice2['Label'], 'bo',markersize = .3)
+ax4.hlines(0, min(nice2['rt']), max(nice2['rt']), zorder = 10)
+ax4.set_xlabel('rt')
+
+plt.tight_layout()
+
+plt.subplots_adjust(left=0.3)
+
+plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig1-calibErrors.png', format='png', dpi=600)
+
+plt.show()

### 2017-07-12 FigureGeneration/fig2-10ppmSearchCalib.py

@@ -1,0 +1,49 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+maxi = 25000
+
+
+height = 6
+width = 7 
+
+f, ((ax1), (ax3)) = plt.subplots(2, 1, sharex='col', sharey='row', figsize=(width,height))
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task1Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+A=ok[r"Mass Diff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue Notch"]
+
+ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+ax1.set_ylim([1,maxi])
+ax1.vlines(0,0,maxi,zorder = 10)
+ax1.set_ylabel('PSM count')
+ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
+                size='large', ha='right', va='center', xytext=(0,0.5))
+
+maxi = 35000
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task3Search\allPSMs_10ppmAroundZero.psmtsv", sep='\t')  
+A=ok[r"Mass Diff (ppm)"]
+B=ok[r"Decoy/Contaminant/Target"]=="T"
+C=ok[r"QValue Notch"]
+
+ax3.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
+ax3.set_ylim([1,maxi])
+ax3.vlines(0,0,maxi,zorder = 10)
+ax3.set_ylabel('PSM count')
+ax3.set_xlabel('Mass Error (ppm)')
+
+ax3.annotate(r"Calibrated", xy=(0, 0.5), xycoords=ax3.yaxis.label, textcoords='offset points',
+                size='large', ha='right', va='center', xytext=(0,0.5))
+
+
+plt.tight_layout()
+
+plt.subplots_adjust(left=0.3)
+
+
+plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig2-10ppmSearchCalib.png', format='png', dpi=600)
+
+plt.show()

### 2017-07-12 FigureGeneration/fig3-searchTypes.py

@@ -1,0 +1,62 @@
+import numpy as np
+import matplotlib.pyplot as plt
+
+
+
+height = 5
+width = 3.25 
+plt.figure(figsize=(width, height))
+
+plt.subplot(311)
+
+dates = [-100,-1,0,1,100]
+values = [1,1,2,1,1]
+
+plt.plot(dates, values, '-')
+
+ylabels = ['Reject', 'Accept']
+yTicks = [1,2]
+
+plt.ylim([0.5,2.5])
+plt.yticks(yTicks, ylabels)
+plt.title("Narrow-Mass Search")
+
+plt.subplot(312)
+
+dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
+values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
+
+plt.plot(dates, values, '-')
+
+ylabels = ['Reject', 'Accept']
+yTicks = [1,2]
+
+plt.title("Notch Search")
+plt.ylim([0.5,2.5])
+
+plt.yticks(yTicks, ylabels)
+
+plt.subplot(313)
+
+
+
+dates =  [-100,100]
+values = [2 ,2]
+
+plt.plot(dates, values, '-')
+
+ylabels = ['Reject', 'Accept']
+yTicks = [1,2]
+
+plt.ylim([0.5,2.5])
+
+plt.title("Wide-Mass Search")
+
+plt.yticks(yTicks, ylabels)
+plt.tight_layout()
+
+
+plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig3-searchTypes.png', format='png', dpi=600)
+
+
+plt.show()

### 2017-07-12 paper.tex
afsdhkj

@@
 This is a highly restrictive search.
-It is not uncommon for the precursor mass tolerance of single-notch searches to be reduced to as much as 2 or 3 Da.
+It is not uncommon for the precursor mass tolerance of single-notch searches to be expanded to as much as 2 or 3 Da.
 One rationale for this is that errors in the deconvolution of MS1 spectra frequently result in an experimental precursor mass that is in error by 1-3 Da, corresponding to missed monoisotopic errors.
@@
 Frequently a peptide with a mass differing from the precursor mass reported by the instrument is the identified peptide.
-In case the co-fragmented peptides have the same charge, the mass difference between the identified peptide and the isolated precursor can be as much as 2 or 3 Da, so a lose 2 or 3 Da tolerance is able to identify those.
+In case the co-fragmented peptides have the same charge, the mass difference between the identified peptide and the isolated precursor can be as much as 2 or 3 Da, so a loose 2 or 3 Da tolerance is able to identify those.
 The major limitation to this relaxed parent mass tolerance is a significant increase in the number of false positive identifications (due to a vast expansion in the number of acceptable theoretical peptides and simultaneous failure to accommodate low mass modifications such as amidation and deamidation), arguably defeating the purpose of a narrow-mass search.
 
-An alternative to these two traditional single-notch searches is a limited multi-notch search that allows mass differences between precursor and theoretical peptide of 0, 1.0029, 2.0051 or 3.0077 Da but limits the width about any one of those three tolerances to a few ppm (0-20 ppm).
+An alternative to these two traditional single-notch searches is a limited multi-notch search that allows mass differences between precursor and theoretical peptide of 0, 1.0029, 2.0051 or 3.0077 Da but limits the width about any one of those three tolerances to a few ppm (<20 ppm).
 This approach provides a means of getting an increased number of correct identifications but mitigates problems with false positive identifications.
 This alternative would appear to result in loss of identification of co-isolated peptides.
-However, MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolute all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
+However, MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
 This expanded list of precursors is used in every spectrum match frequently resulting in multiple peptide identifications from single tandem mass spectrum.
@@
 INSERT PARAGRAPH DESCRIBING RESULTS OF SINGLE NOTCH VERSUS LIMITED MULTI NOTCH \& CO-ISOLATION
-
-We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the single-notch searches.
 
@@
 
-Chick~\citep{Chick_2015} showed that by significantly relaxing the parent mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified.
+Chick\citep{Chick_2015} showed that by significantly relaxing the parent mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified.
 In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in precursor and parent mass.
@@
 The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
-The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, but interpretation of the mass difference was left to the user.
-PTMs corresponding to the reported mass difference can frequently be inferred; such as when the mass difference reported matches a known PTM mass (e.g. 79.966331 Da for phosphorylation).
-However, mass differences can arise from not only one modification, but frequently from two or more modifications, some of which increase the mass and the peptide and some of which decrease it.
-In addition, small missed cleavages are sometimes reported mistakenly as mass differences.
-Also, some mass differences, such as those matching to amino acid masses, are easily interpreted but suffer from high local FDR and should only be used with caution.
-Finally, seemingly uninterpretable mass differences can be reported when the best matching peptide has an actual precursor mass different from the precursor mass selected by the mass spectrometer.
-The range of masses selected for fragmentation in a tandem mass spectrometer is commonly +/- 2 Da.
-Therefore, multiple peptides can be introduced to the cell and co-fragmented.
-This often results in peptide identification of a less abundant peptide with mass slightly larger or smaller than the selected precursor.
-The unconstrained utilization of mass difference and acceptance of the theoretical best matching sequence as true can result in a high number of false positive assignments.
+The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, with interpretation of the mass difference left as an exercise the user.
 
-A third alternative, reported here, is the use of a multi-notch search (>100 discrete notches) offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs and the least false positive PSMs (see the TABLE i and TABLE j).
+There are two major problems with using the results from unrestricted precursor mass tolerance searches to infer peptide modifications regardless of whether or not they are of the Chick or Kong variety.
+
+First, the freedom in precursor mass difference results in a multitude of false positive identifications.
+For example, small missed cleavages are often reported as mass differences.
+The peptide PEPKK is often identified as PEPK with a mass error of 128 Da (the converse is also true, PEPK may be identified with a mass error of -128 Da).
+This error is relatively easy to spot but when the missed cleavages contain multiple amino acids, the interpretation and discovery of the error is much more difficult.
+Another example of this relates to peptide modifications.
+The peptide PEPSphosphoK could be reported with a mass error of -79.96Da when the better identification would have been PEPSK.
+Constraining the search using the multi-notch approach described below, directly prohibits assignment of the best match for a peak erroneously, and instead produces an alternative peak assignment with a much lower false positive rate (see below).
+
+Second, unrestricted precursor mass search struggle to correctly identify co-isolated/co-fragmented peptides as we described above for single notch searches with tolerance of 2-3Da.
+
+The alternative, described here, is the use of a multi-notch search (>100 discrete notches), which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the least false positive PSM identifications (see the TABLE i and TABLE j).
 Table I (Count of theoretical peptide candidates for each search mode (single notch 5ppm, single notch 3.5, limited multi notch, 500 Da, and full multi-notch).
 Table J (Total PSMs for each search type – 500 or full multi-notch).
-One output of an MSFragger search is a list of PSMs with each entry containing the mass difference between the isolated precursor and the mass of the best matching theoretical peptide.
-MSFragger also doesn't automatically calibrate spectra prior to searching.
-Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
-MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.
+The results of the multi-notch search used in conjunction with G-PTM-D, provide and even further refinement of the results with actual assignment of PTMs and other modifications (see below).
 
-G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list.
-We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search.
-The new notches correspond to metal adducts (e.g. Na and K), glycosylations., sample handling and processing artifacts (e.g. carbamylation, ammonia loss), lipidation, and etc.
-MSFragger has limited mass range as reported.
-We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of Sulfonation and Phosphorylation, and a problematic peak at mass of Leucine, with a high FDR.
-Some Histogram Peaks For Mouse Data
-Some Histogram Peaks For Mouse Data
-We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals .
+\subsection{Co-isolation}
 

### 2017-07-12 paper.tex
coisolation

@@
 
+[Cite Bo Zhang if paper exists].
+We showed above in some amazing figure an example where several peptides were co-eluting from the column into that mass-spec within the tolerance used for isolation for fragmentation.
+In that particular case, the precursor with m/z dsfadf was chosen for isolation and fragmentation.
+In a conventional search, only that precursor mass would be used to select a range of theoretical for comparison.
+However, all peptides in that range are co-fragmented and a better solution would be to use all co-isolated precursor masses.
+In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor (+/- 3 Da).
+Each deconvoluted mass is used to select a range of theoreticals to use for identification.
+The result of this is that often, more than one peptide can be identified from a single MS2 scan.
+Below, in Table 3434, we should the number of PSMs for a raw file where we use only a single precursor mass and we compare that to the case where we allow for multiple different precursors.
+You can see a significant increase in the number of identified PSMs.
+Importantly, you can see in Supplemental Figure erf3e, that the precursor mass errors go down significantly and the standard deviation is much reduced.
+This is extremely benficial in both open mass and multi-notch searches because it reduces the number the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
+
+Get the story about how precurors with similar m/z with different z can in their deconvoluted state be many (10s, 100s of Da apart).
+This is a big problem in general searches but also for open mass and multi-notch searches which are used to discover possible PTMs.
+
 \subsection{Enhanced G-PTM-D Results}

### 2017-07-12 paper.tex
ajdsfk

@@
 
+In the original G-PTM-D manuscript, an open mass search (+/- 200-500Da) was used to create a histogram of mass difference peaks, from which a number of PTMs could be inferred.
+A perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
+This database was used for a final search and a number of PTMs were identified.
+There were several limitations to that work.
+One, the open mass search was really slow.
+Two, uncalibrated data made it hard to discern closely spaced modifications.
+Three, certain mass differences with high FDRs were incorrectly allowed.
+Four, missed monoisotopic peaks and co-isolated peaks were not handled.
+Five, separate programs were required (both Morpheus and a perl script).
+MetaMorpheus resolves all of these issues and more.
+Below we will show how each individual enhancement to G-PTM-D improved performance over the original work and then we will show at the end, the combined effect of all enhancements.
+
 A major improvement in G-PTM-D performance arises from using calibrated spectra as inputs to the procedure. 
@@
 It is evident from these histograms that the phosphorylation and sulfonation peaks are clearly separated and identifiable when the search is conducted on well-calibrated spectra.
+
+[THERE IS SUPPOSED TO BE A COOL FIGURE HERE]
+
+G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list. 
+We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search.
+The new notches correspond to metal adducts (e.g.
+Na and K), glycosylations., sample handling and processing artifacts (e.g.
+carbamylation, ammonia loss), lipidation, and etc.
+MSFragger has limited mass range as reported.
+We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
+Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of Sulfonation and Phosphorylation, and a problematic peak at mass of Leucine, with a high FDR.
+Some Histogram Peaks For Mouse Data Some Histogram Peaks For Mouse Data We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals .
+MSFragger also doesn’t automatically calibrate spectra prior to searching.
+Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
+MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.
 
@@
 The effect of using these notches is shown in Table~\ref{tab:table2}.
+
+We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the single-notch searches.
 

### 2017-07-13 paper.tex
calib

@@
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
-Protein databases such as UniProt include information on the localization of specific modifications to specific residues in the protein, but this information is incomplete.
+Protein databases such as UniProt\citep{Uniprot_2017} include information on the localization of specific modifications to specific residues in the protein, but this information is incomplete.
 Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identification in complex biological samples has continued to pose a challenge for proteomics\citep{Olsen_2013}.
@@
 
-We demonstrated the effectiveness of the calibration approach using a single CAST Mouse raw file as an example by displaying the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, before and after the calibration procedure is applied.
-In Figure\ref{fgr:fig1-calibErrors} these errors are plotted versus two important variables: the m/z value of the peak and the retention time.
-It can be seen in the graphs of Figure\ref{fgr:fig1-calibErrors} that calibration successfully addresses errors in non-linearity as a function of both retention time and m/z.
-We note that in addition to these two parameters, the calibration function native to MetaMorpheus also considers non-linearities in peak intensity, injection time, total ion count, etc.
+We demonstrate the effectiveness of the calibration approach using a single CAST Mouse raw file as an example by displaying the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, before and after the calibration procedure is applied.
+In Figure~\ref{fgr:fig1-calibErrors} these errors are plotted versus two important variables: the m/z value of the peak and the retention time of the corresponding scan.
+Calibration successfully removes both linear and non-linear dependencies between the m/z errors and these variables.
+We note that in addition to these two parameters, the calibration function native to MetaMorpheus also considers  peak intensity, injection time, total ion count, and other variables as potentially influencing the m/z errors.
 In addition, MetaMorpheus calibrates both MS1 and MS2 (using MS2 fragment matches), but only results from MS1 are shown here.
@@
 
-Calibration quality improvements are further shown through the results of a standard 10ppm precursor tolerance database search (Table\ref{tbl:calib}).
+Calibration quality improvements are further shown through the results of a standard 10ppm precursor tolerance database search (Table~\ref{tbl:calib}).
 PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure (Table~\ref{tbl:calib}).

### 2017-07-13 paper.tex
calib

@@
 
-Calibration quality improvements are further shown through the results of a standard 10ppm precursor tolerance database search (Table~\ref{tbl:calib}).
-PSMs within 1\% FDR prior to calibration are compared with PSMs after the calibration procedure (Table~\ref{tbl:calib}).
-The calibration procedure successfully addresses instrument miscalibration by centering the average error around zero.
+Calibration quality improvements are further shown through the results of standard 10ppm precursor tolerance database search, conducted on uncalibrated vs calibrated spectra files.
+Table~\ref{tbl:calib} details the results of these searches, showing the number of confidently identified PSMs, the average ppm differences between the identified peptide and the precursor mass, and the standard deviation of those differences.
+The calibration procedure successfully addresses instrument miscalibration by centering the average ppm error around zero.
 The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which show that the calibrated mass spectra have sub-ppm mass errors.
@@
 \label{tbl:calib}
-\begin{tabular}{llll}
+\begin{tabular}{lll}
 
-                                                                                &      & Orig        & Calib        \\
- &PSM Count                                                           &      217612   &   219172        \\
- &Average Error (ppm)                                                              &      -2.10      &        -0.16      \\
- &St. Dev of Error                                                    &       2.15      &        0.91      \\
-Confidence Intervals& 68.3\%     &    [-4.26,0.05]          &  [-1.08,0.75]            \\
-                                                                                & 95.4\%      &   [-6.41,2.20]          &      [-1.99,1.61]         \\
-                                                                                & 99.7\%       &    [-8.56,4.36]          &      [-2.90,2.58]         \\
+                                                                                      & Orig        & Calib        \\
+ PSM Count                                                           &      217612   &   219172        \\
+ Average Error (ppm)                                                              &      -2.10      &        -0.16      \\
+ St. Dev of Error                                                    &       2.15      &        0.91      \\
 \end{tabular}
@@
  \includegraphics{fig2-10ppmSearchCalib.png}
- \caption{Errors in m/z plotted vs the m/z value and the retention time}
+ \caption{Histograms of the mass errors in identified peptides before and after calibration}
  \label{fgr:fig2-10ppmSearchCalib}

### 2017-07-13 paper.tex
search modes

@@
 
-It is useful to describe the variety of search tolerances used, prior to discussing the impact that multi-notch searches have on the G-PTM-D performance.
-Traditional searches in bottom-up proteomics score potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal within some defined tolerance (e.g.
+It is useful to carefully describe the two commonly used database search approaches, prior to defining multi-notch searches and the impact that they have on G-PTM-D performance.
+A traditional search in bottom-up proteomics considers potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal within some defined tolerance (e.g.
 5ppm, or 2.1 Da).
-We'll refer to this type of search here as a \textit{single-notch} search.
-The other common search mode is a \textit{wide-parent tolerance search} (e.g. +/- 500Da)[Chick, Kong, Qiyao].
+We refer to this type of search here as a \textit{single-notch} search.
+The other common search is a \textit{wide-parent tolerance search} (e.g. +/- 500Da)\citep{Chick_2015,Kong_2017,Li_2016}.
 In searches of that type, the precursor mass and the mass of the best matching theoretical peptide could be many Da apart.
-We describe an alternative to these search modes: the multi-notch search.
-The \textit{multi-notch} search is an extension of the single-notch search to multiple specific mass differences between precursor mass and theoretical mass~\ref{fig:fig3-searchTypes} where the mass difference is easily interpretable.
+We describe an alternative to these search modes: the \textit{multi-notch} search.
+This search is an extension of the single-notch search that allows a variety of multiple specific mass differences between precursor mass and theoretical mass (See Figure~\ref{fig:fig3-searchTypes}) where the mass difference is easily interpretable.
 It can also be seen as a narrowing of the wide-parent tolerance search to only include prespecified mass differences.

### 2017-07-14 paper.tex
search modes
djfkajdf

@@
 This change improves the specificity of the search and significantly decreases search time.
-The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in precursor mass deconvolution (+1.0029, + 2.0052 and + 3.0077 Da), further increasing the number of confidently identified peptides and PTMs.
+The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in precursor mass deconvolution (+1.0029, + 2.0054 and + 3.0078 Da), further increasing the number of confidently identified peptides and PTMs.
 Finally, we replace the limited UniProt PTM list with a customizable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
@@
 The traditional single-notch search limits the range of acceptable mass differences between a candidate peptide and the observed experimental precursor in order to find an exact match to the observed peptide in a protein database.
-A mass tolerance of 5-20ppm is a common precursor mass tolerance that allows only mass differences between theoretical and experimental of 5-20 ppm.
-This is a highly restrictive search.
-It is not uncommon for the precursor mass tolerance of single-notch searches to be expanded to as much as 2 or 3 Da.
-One rationale for this is that errors in the deconvolution of MS1 spectra frequently result in an experimental precursor mass that is in error by 1-3 Da, corresponding to missed monoisotopic errors.
-This relaxed tolerance increases the number of correctly identified peptides.
+A mass tolerance of 1-20ppm is a commonly used, resulting in a highly restrictive search.
+It is not uncommon for the precursor mass tolerance of single-notch searches to be relaxed to as much as 2 or 3 Da.
+One rationale for this is that errors in the deconvolution of MS1 spectra frequently result in an experimental precursor mass reported being different by 1.0029 or 2.0054 Da from the true isolated peptide mass.
+The relaxed tolerance approach identifies those peptides correctly.
 A second rationale for this approach is that multiple peptides are frequently co-isolated and co-fragmented with all peptide fragments being reported in the same tandem mass spectrum.
-Frequently a peptide with a mass differing from the precursor mass reported by the instrument is the identified peptide.
-In case the co-fragmented peptides have the same charge, the mass difference between the identified peptide and the isolated precursor can be as much as 2 or 3 Da, so a loose 2 or 3 Da tolerance is able to identify those.
-The major limitation to this relaxed parent mass tolerance is a significant increase in the number of false positive identifications (due to a vast expansion in the number of acceptable theoretical peptides and simultaneous failure to accommodate low mass modifications such as amidation and deamidation), arguably defeating the purpose of a narrow-mass search.
+Frequently the only well-fragmented peptide is different from the one corresponding to the isolated peak reported by the instrument, and the relaxed tolerance approach may identify this peptide, provided the reported charge state matches.
+The major limitation of this approach is that a 2 or 3 Da tolerance includes mass shifts that, when present, may indicate that the matched peptide is not the correct identification, arguably defeating the purpose of a narrow-mass search.
+This may happen when the mass difference arises due to identified peptide being different from the actual isolated one by a low mass modification or an amino acid substitution.
 
-An alternative to these two traditional single-notch searches is a limited multi-notch search that allows mass differences between precursor and theoretical peptide of 0, 1.0029, 2.0051 or 3.0077 Da but limits the width about any one of those three tolerances to a few ppm (<20 ppm).
-This approach provides a means of getting an increased number of correct identifications but mitigates problems with false positive identifications.
-This alternative would appear to result in loss of identification of co-isolated peptides.
-However, MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
-This expanded list of precursors is used in every spectrum match frequently resulting in multiple peptide identifications from single tandem mass spectrum.
+The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4mouse-1012}, generated by compiling the results of a 2.5 Da tolerance single-notch search.
 
-INSERT PARAGRAPH DESCRIBING RESULTS OF SINGLE NOTCH VERSUS LIMITED MULTI NOTCH \& CO-ISOLATION
+\begin{figure}
+\caption{Jurkat histograms around 0}
+\label{fig:fig4mouse-1012}
+\includegraphics{fig4mouse-1012.png}
+\end{figure}
+
+An alternative to these two traditional single-notch searches is a limited multi-notch search that allows mass differences between precursor and theoretical peptide of 0, 1.0029 or 2.0054 Da but limits the width about any one of those three tolerances to a few ppm (1-20 ppm).
+This approach provides a means of getting an increased number of correct identifications but mitigates problems with wrong identifications.
+This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to mitigate this problem by identifying multiple co-isolated precursor masses, see below.
+
+To qualitatively illustrate the described advantages, we conducted the three searches on all 28 calibrated Jurkat fractions.
+The numbers of confidently identified PSMs, misassigned PSMs are reported in Table~\ref{tbl:singleVsMultiNotch}.
 
@@
 \caption{Notch Search vs Narrow-Mass Search comparison of PSMs for Mouse}
-\label{my-label}
+\label{tbl:singleVsMultiNotch}
 \begin{tabular}{llll}
                     & 5ppm Search & 3.5 Da Search & Notch Search \\
-PSMs Within 1\% FDR & 157188      & 178003        &        \\
-misassigned PSMs    & 0           &               & 0            \\
-Good PSMs           & 157188      &         &        \\ 
+PSMs Within 1\% FDR & 157188      & 178003        & 177213       \\
+misassigned PSMs    & 0           & 1521          & 0            \\
 \end{tabular}
 \end{table}
-
-%\begin{figure}
-%\caption{Mouse histograms around 0}
-%\label{fig:fig3mouse-1012}
-%\includegraphics{fig3mouse-1012.png}
-%\end{figure}
-
-%\begin{figure}
-%\caption{Jurkat histograms around 0}
-%\label{fig:fig3jurkat-1012}
-%\includegraphics{fig3jurkat-1012.eps}
-%\end{figure}
 
@@
 This is a big problem in general searches but also for open mass and multi-notch searches which are used to discover possible PTMs.
+
+MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
+This expanded list of precursors is used in every spectrum match,frequently resulting in multiple peptide identifications from single tandem mass spectrum.
 
@@
 
-The final notch search uses notches centered at \{0, 1.0029, 2.0052, 3.0077\}, in order to capture missed identifications due to missed monoisotopic peaks.
+The final notch search uses notches centered at \{0, 1.0029, 2.0054, 3.0078\}, in order to capture missed identifications due to missed monoisotopic peaks.
 The results are contrasted with the standard narrow-mass final search.

### 2017-07-14 FigureGeneration/fig4-mouse;-1;0;1;2.py

@@ -1,0 +1,132 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-05-05-18-28-01\Task2Calibrate\2017-07-06-17-29-30\Task1Search\aggregatePSMs_-187andUp.psmtsv", sep='\t')
+
+
+
+
+height = 6
+width = 7
+plt.figure(figsize=(width, height))
+
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+#data_sorted = data.sort_values(['MassDiffToBestMass (Da)'], ascending=True)
+#massDiffs=data_sorted['MassDiffToBestMass (Da)'].values
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+massDiffsHighConf = massDiffs[C<0.01]
+TDhighConf = B[C<0.01]
+
+ok = plt.subplot(221)
+
+halfWidth = 2.505
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.01), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+plt.ylabel("Count")
+plt.title("PSMs within $\pm 2.5$ Da")
+ok.set_yscale("log", nonposy='clip')
+
+ok = plt.subplot(223)
+
+halfWidth = 2.505
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
[…]

### 2017-07-14 paper.tex
djfkdjf

@@
 This approach provides a means of getting an increased number of correct identifications but mitigates problems with wrong identifications.
-This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to mitigate this problem by identifying multiple co-isolated precursor masses, see below.
+This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to mitigate this problem by identifying multiple co-isolated precursor masses, see section below.
 
 To qualitatively illustrate the described advantages, we conducted the three searches on all 28 calibrated Jurkat fractions.
-The numbers of confidently identified PSMs, misassigned PSMs are reported in Table~\ref{tbl:singleVsMultiNotch}.
+The numbers of confidently identified PSMs and misassigned PSMs are reported in Table~\ref{tbl:singleVsMultiNotch}.
 

### 2017-07-14 paper.tex
fjkdjf

@@
 
-There are two major problems with using the results from unrestricted precursor mass tolerance searches to infer peptide modifications regardless of whether or not they are of the Chick or Kong variety.
-
-First, the freedom in precursor mass difference results in a multitude of false positive identifications.
+There is a major problem with using the results from unrestricted precursor mass tolerance searches to infer peptide modifications regardless of whether or not they are of the Chick or Kong variety.
+The freedom in precursor mass difference results in a multitude of false positive \textit{mass difference identifications}.
 For example, small missed cleavages are often reported as mass differences.
@@
 
-Second, unrestricted precursor mass search struggle to correctly identify co-isolated/co-fragmented peptides as we described above for single notch searches with tolerance of 2-3Da.
+The alternative we recommend, is the use of a multi-notch search, which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the least false positive PSM identifications.
 
-The alternative, described here, is the use of a multi-notch search (>100 discrete notches), which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the least false positive PSM identifications (see the TABLE i and TABLE j).
-Table I (Count of theoretical peptide candidates for each search mode (single notch 5ppm, single notch 3.5, limited multi notch, 500 Da, and full multi-notch).
-Table J (Total PSMs for each search type – 500 or full multi-notch).
-The results of the multi-notch search used in conjunction with G-PTM-D, provide and even further refinement of the results with actual assignment of PTMs and other modifications (see below).
+To compare the multi-notch search with the wide-mass search, we conducted searches of all of Mouse fractions, with a multi-notch search with notches at 0 and 79.96, and a +/-500 Da wide mass search.
+The advantage of the multi-notch search over the wide-mass search is shown in Table~\ref{tbl:multiVsWide} by comparing mass difference histogram peak membership numbers.
+
+\begin{table}[]
+\centering
+\caption{Multi-Notch Search vs. Wide-Mass Search}
+\label{tbl:multiVsWide}
+\begin{tabular}{lll}
+                    & Multi-Notch & Wide-Mass\\
+0 peak & 139374      & 137483        \\
+79.96 peak & 3402           & 3298          \\
+\end{tabular}
+\end{table}
 
@@
 
-The final notch search uses notches centered at \{0, 1.0029, 2.0054, 3.0078\}, in order to capture missed identifications due to missed monoisotopic peaks.
+The final notch search uses notches centered at \{0, 1.0029, 2.0054\}, in order to capture missed identifications due to missed monoisotopic peaks.
 The results are contrasted with the standard narrow-mass final search.

### 2017-10-16 paper.tex
FirstTwoSections

@@
 
-Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist, and those address primarily common modifications (e.g.
-phosphorylations) in well-studied systems.
-Global discovery tools, such as MODa\citep{Na_2011} are also available.
-G-PTM-D\citep{Li_2016} is one recently described bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\citep{Chick_2015, Na_2011}, that provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
+Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist.
+These primarily address common modifications (e.g. phosphorylations) in well-studied systems, but global discovery tools, such as MODa\citep{Na_2011} and G-PTM-D\citep{Li_2016}, are also available.
+G-PTM-D is a bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs.
+The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
@@
 
-As described above, G-PTM-D has a few significant limitations.
+While G-PTM-D delivers confident identification of numerous types of PTMs, it has a few significant limitations.
 The initial wide-mass search procedure is slow, taking hours for modest size datasets.
-Due to this, high mass modifications, such as glycosylations are out of reach, since widening the search window from the suggested $\pm 200$ Da increases the search time even more.
-The final narrow-mass search is also problematic, since it does not account for peptide-spectral matches(PSMs) with mass differences arising due to misidentifying the monoisotopic peak.
-In G-PTM-D, modification discovery is limited to PTMs in the UniProt database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
+Thus, high mass modifications such as glycosylations are out of reach, since widening the search window from the suggested $\pm 200$ Da increases the already-slow search time.
+The final narrow-mass search is also problematic, as it does not account for peptide-spectral matches (PSMs) with mass differences arising from misassignment of the monoisotopic peak.
+Discovery of modifications is limited to PTMs in the UniProt database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
 Finally, PTMs that are similar in mass, such as phosphorylation (79.966331 Da) and sulfonation (79.956815 Da), are problematic since they are virtually indistinguishable in unprocessed spectral files.
@@
 
-Spectral calibration is essential for distinguishing modifications with similar mass, so we recommend running the G-PTM-D procedure only on calibrated spectra files.
-To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected modification masses (e.g.
+Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files.
+To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
 PTMs and amino acids).
 This change improves the specificity of the search and significantly decreases search time.
-The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in precursor mass deconvolution (+1.0029, + 2.0054 and + 3.0078 Da), further increasing the number of confidently identified peptides and PTMs.
-Finally, we replace the limited UniProt PTM list with a customizable expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
-This change not only expands the list of identified and localized modifications, but also indirectly improves the confident PSM count.
+The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and + 2.0054), further increasing the number of confidently identified peptides and PTMs.
+Finally, we replace the limited UniProt PTM list with a custom expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
+This change not only expands the list of identified and localized modifications, but also increases the number of confidently identified PSMs.
 The enhanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
@@
 
-We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus\footnote{Free and Open Source software available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}}, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
-Three multi-fraction mammalian datasets (Jurkat cell, C57BL/6J (B6) and CAST/EiJ (CAST) mouse islet lysates) with deep proteome coverage were used to evaluate performance.
-These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells, and 18 fractions from Mouse islets, and are described in detail in~\citep{Shortreed_2015, Cesnik_2016}.
+We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
+MetaMorpheus is freely available as a standalone Windows executable program from GitHub \url{https://github.com/smith-chem-wisc/MetaMorpheus}.
+All of the source code is also available at that link.
+MetaMorpheus heavily utilizes a NuGet package called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
+mzLib can be found at link \url{https://github.com/smith-chem-wisc/mzLib}.
 
-UniProt XML protein databases acquired on Feb 27, 2017 containing only reviewed human/mouse proteins were employed.
+Three multi-fraction mammalian datasets with deep proteome coverage were used to evaluate performance.
+These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells and 9 fractions from each of two different pancreatic islets from two species of mice.
+These datasets are available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215} and \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00470} repositories, and are described in detail in~\citep{Shortreed_2015, Cesnik_2016}.
 
-A 10ppm precursor mass tolerance is used for the initial calibration step, and then reduced to 5ppm for subsequent searches.
-We allowed 120 modifications including PTMs, adducts, and chemical modifications in the database augmentation step, see Supplementary Material 1 for a complete listing.
+UniProt XML protein databases used were downloaded on Sept 14, 2017, and contained only reviewed human/mouse proteins.
+
+209 modifications including PTMs, adducts, and chemical modifications were allowed in the database augmentation step, see Supplementary Table 1 for a complete listing.
 
@@
 The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4mouse-1012}, generated by compiling the results of a 2.5 Da tolerance single-notch search.
-
-\begin{figure}
-\caption{Jurkat histograms around 0}
-\label{fig:fig4mouse-1012}
-\includegraphics{fig4mouse-1012.png}
-\end{figure}
 

### 2017-10-16 paper.tex
calib

@@
 Higher mass accuracy provides increased specificity and thus confidence in peptide identifications, decreasing the false discovery rate.
-Instrument noise, systemic drift and miscalibration limit the mass accuracy in acquired spectra.
-Multiple calibration strategies to improve mass accuracy have been devised, and fall into three general categories: External calibration prior to the MS experiment (e.g.
+Instrument noise, systematic drift and miscalibration all limit the mass accuracy in acquired spectra.
+Multiple calibration strategies to improve mass accuracy have been devised and fall into three general categories: external calibration prior to the MS experiment (e.g.
 standard instrument calibration protocols); internal calibration during the MS experiment (e.g.
@@
 We developed a post-acquisition calibration procedure that builds upon the software lock mass concept\citep{Cox_2011} recently reported by the Mann group.
-In our strategy, the m/z differences between expected and observed peaks in the peptide tandem ms spectra are compiled, and then used to recalibrate the spectra.
+In our strategy, the m/z differences between expected and observed peaks in the peptide tandem MS spectra are compiled, and then used to recalibrate the spectra.
 The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g 1\%; based upon the target:decoy strategy\citep{Elias_2007}).
+This yields a set of peptide identifications subject to a desired false discovery rate (e.g. 1\% based upon the target:decoy strategy\citep{Elias_2007}).
 Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
@@
 
-We demonstrate the effectiveness of the calibration approach using a single CAST Mouse raw file as an example by displaying the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, before and after the calibration procedure is applied.
-In Figure~\ref{fgr:fig1-calibErrors} these errors are plotted versus two important variables: the m/z value of the peak and the retention time of the corresponding scan.
+We demonstrate the effectiveness of the calibration approach using a single CAST Mouse raw file as an example.
+Figure~\ref{fgr:fig1-calibErrors} displays the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, both before and after the calibration procedure is applied.
+These errors are plotted versus two important variables: the m/z value of the peak and the retention time of the corresponding scan.
 Calibration successfully removes both linear and non-linear dependencies between the m/z errors and these variables.
 We note that in addition to these two parameters, the calibration function native to MetaMorpheus also considers peak intensity, injection time, total ion count, and other variables as potentially influencing the m/z errors.
-In addition, MetaMorpheus calibrates both MS1 and MS2 (using MS2 fragment matches), but only results from MS1 are shown here.
-The complete list of parameters analyzed in the calibration is provided as Table lala in the supplement.
+MetaMorpheus calibrates both MS1 and MS2 scans (using MS2 fragment matches), but only the results from MS1 are shown here.
+MetaMorpheus calibrated files are verified to be compatible with a variety of tools and workflows, including TPP, Mascot, and MsFragger.
 
@@
 
-Calibration quality improvements are further shown through the results of standard 10ppm precursor tolerance database search, conducted on uncalibrated vs calibrated spectra files.
-Table~\ref{tbl:calib} details the results of these searches, showing the number of confidently identified PSMs, the average ppm differences between the identified peptide and the precursor mass, and the standard deviation of those differences.
-The calibration procedure successfully addresses instrument miscalibration by centering the average ppm error around zero.
-The systemic noise introduced by various other factors is also addressed: this is evident in the decrease in the standard deviation of the errors in the precursor mass estimates, which show that the calibrated mass spectra have sub-ppm mass errors.
+Calibration quality improvements are also shown by the results of a standard database search, conducted on uncalibrated versus calibrated spectra using a 10 ppm precursor mass tolerance.
+Table~\ref{tbl:calib} details the results of these searches.
+The calibration procedure successfully addresses instrument miscalibration by centering the mass errors around zero ppm.
+The systematic noise introduced by various other factors has also been corrected, as evidenced by the decrease in the standard deviation of the errors in the precursor masses, which shows that the calibrated mass spectra predominantly have sub-ppm mass errors.
+A beneficial side effect of calibration is a corresponding increase in the number of confidently identified PSMs by around 1\%, when using the same wide 10 ppm precursor mass tolerance.
 Histograms of the mass errors in identified peptides before and after calibration are plotted in Figure~\ref{fgr:fig2-10ppmSearchCalib}.
-The main results of calibration in summary are increased PSMs at 1\% FDR, decreased error in precursor mass, and decreased standard deviation in average mass error.
+In summary, the main results of calibration are: more PSMs at 1\% FDR, decreased error in precursor mass, and decreased standard deviation in average mass error.
 

### 2017-10-16 paper.tex
tsnslmns

@@
 We refer to this type of search here as a \textit{single-notch} search.
-The other common search is a \textit{wide-parent tolerance search} (e.g. +/- 500Da)\citep{Chick_2015,Kong_2017,Li_2016}.
+The other common search is a \textit{wide-parent tolerance search} (e.g. $\pm 500$ Da)\citep{Chick_2015,Kong_2017,Li_2016}.
 In searches of that type, the precursor mass and the mass of the best matching theoretical peptide could be many Da apart.
@@
 The traditional single-notch search limits the range of acceptable mass differences between a candidate peptide and the observed experimental precursor in order to find an exact match to the observed peptide in a protein database.
-A mass tolerance of 1-20ppm is a commonly used, resulting in a highly restrictive search.
+A mass tolerance of <20ppm is commonly used on high-resolution MS1 data, resulting in a highly restrictive search.
 It is not uncommon for the precursor mass tolerance of single-notch searches to be relaxed to as much as 2 or 3 Da.
-One rationale for this is that errors in the deconvolution of MS1 spectra frequently result in an experimental precursor mass reported being different by 1.0029 or 2.0054 Da from the true isolated peptide mass.
+One rationale for this is that errors in the deisotoping of MS1 spectra frequently result in a reported experimental precursor mass being different by 1.0029 or 2.0054 Da from the true isolated peptide mass.
 The relaxed tolerance approach identifies those peptides correctly.
-A second rationale for this approach is that multiple peptides are frequently co-isolated and co-fragmented with all peptide fragments being reported in the same tandem mass spectrum.
-Frequently the only well-fragmented peptide is different from the one corresponding to the isolated peak reported by the instrument, and the relaxed tolerance approach may identify this peptide, provided the reported charge state matches.
+A second rationale for this approach is that multiple peptides are frequently co-isolated and co-fragmented with all peptide fragments being observed in the same tandem mass spectrum.
+Frequently, the best-fragmented peptide is different from the one corresponding to the precursor mass reported by the instrument, and the relaxed tolerance approach may identify this peptide, provided the reported charge state matches.
 The major limitation of this approach is that a 2 or 3 Da tolerance includes mass shifts that, when present, may indicate that the matched peptide is not the correct identification, arguably defeating the purpose of a narrow-mass search.
-This may happen when the mass difference arises due to identified peptide being different from the actual isolated one by a low mass modification or an amino acid substitution.
+This may happen when the mass difference arises due to the identified peptide being different from the actual isolated precursor by a low mass modification or amino acid substitutions.
 
 The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4mouse-1012}, generated by compiling the results of a 2.5 Da tolerance single-notch search.
+These plots clearly show that numerous peptides have missed monoisotopic (Mm) peaks, deamidation, or an amino acid substitution.
 
-An alternative to these two traditional single-notch searches is a limited multi-notch search that allows mass differences between precursor and theoretical peptide of 0, 1.0029 or 2.0054 Da but limits the width about any one of those three tolerances to a few ppm (1-20 ppm).
-This approach provides a means of getting an increased number of correct identifications but mitigates problems with wrong identifications.
-This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to mitigate this problem by identifying multiple co-isolated precursor masses, see section below.
+We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wider (few Da) precursor mass tolerance during a traditional single-notch search.
+This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (1-20 ppm).
+This approach provides a means of getting an increased number of correct identifications while mitigating problems with incorrect identifications.
+This alternative would appear to result in loss of identification of co-isolated peptides, but MetaMorpheus is programmed to handle this problem by identifying multiple co-isolated precursor masses (see section below).
 
-To qualitatively illustrate the described advantages, we conducted the three searches on all 28 calibrated Jurkat fractions.
-The numbers of confidently identified PSMs and misassigned PSMs are reported in Table~\ref{tbl:singleVsMultiNotch}.
+To illustrate the described advantages, we compared the three types of searches using all 28 calibrated Jurkat fractions (Table~\ref{tbl:singleVsMultiNotch}).
+The limited notch search yields the largest number of confidently identified PSMs with only a small fraction being misassinged.
+We consider PSMs to be misassigned if the G-PTM-D procedure determines a higher-scoring modified peptide for a PSM, see the last section for details.
 

### 2017-10-16 paper.tex
mmswms

@@
 
-Chick\citep{Chick_2015} showed that by significantly relaxing the parent mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified.
+Chick et al.
+showed that by significantly relaxing the parent mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified\citep{Chick_2015}.
 In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in precursor and parent mass.
@@
 Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for wide parent mass tolerance searches.
-Search times for +/- 500 Da parent tolerances went from six years to six seconds.
+Search times for $\pm 500$ Da parent tolerances went from six years to six seconds.
 The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
-The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, with interpretation of the mass difference left as an exercise the user.
+The mass differences can be clustered into histogram peaks\citep{Rodriguez_2014}, with interpretation of the mass difference left as an exercise for the user.
 
-There is a major problem with using the results from unrestricted precursor mass tolerance searches to infer peptide modifications regardless of whether or not they are of the Chick or Kong variety.
+There is a major problem with using the results from unrestricted precursor mass tolerance searches to infer peptide modifications regardless of whether they are of the Chick or Kong variety.
 The freedom in precursor mass difference results in a multitude of false positive \textit{mass difference identifications}.
@@
 Another example of this relates to peptide modifications.
-The peptide PEPSphosphoK could be reported with a mass error of -79.96Da when the better identification would have been PEPSK.
-Constraining the search using the multi-notch approach described below, directly prohibits assignment of the best match for a peak erroneously, and instead produces an alternative peak assignment with a much lower false positive rate (see below).
+The phosphopeptide PEPpSK could be reported with a mass error of -79.96 Da when the better identification would have been PEPSK.
+Constraining the search using the multi-notch approach described below, directly prohibits erroneous assignment of the best match for a peak, and instead produces an alternative peak assignment with a much lower false positive rate (see below).
 
-The alternative we recommend, is the use of a multi-notch search, which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the least false positive PSM identifications.
+The alternative we recommend, is the use of a multi-notch search, which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the fewest false positive PSM identifications.
 
-To compare the multi-notch search with the wide-mass search, we conducted searches of all of Mouse fractions, with a multi-notch search with notches at 0 and 79.96, and a +/-500 Da wide mass search.
-The advantage of the multi-notch search over the wide-mass search is shown in Table~\ref{tbl:multiVsWide} by comparing mass difference histogram peak membership numbers.
+We compared searches of all Mouse islet fractions using a $\pm 500$ Da wide mass search versus a multi-notch search with notches at 0 and 79.96.
+The advantage of the multi-notch search over the wide-mass search is shown in Table~\ref{tbl:multiVsWide} by an increase in confidently identified PSMs (1\% FDR) for both regions (0 Da no-PTM peptides and +79.96 Da phosphorylated peptides).
 

### 2017-10-16 paper.tex
coisolation

@@
 However, all peptides in that range are co-fragmented and a better solution would be to use all co-isolated precursor masses.
-In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor (+/- 3 Da).
+In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor ($\pm 3$ Th).
 Each deconvoluted mass is used to select a range of theoreticals to use for identification.
@@
 Importantly, you can see in Supplemental Figure erf3e, that the precursor mass errors go down significantly and the standard deviation is much reduced.
-This is extremely benficial in both open mass and multi-notch searches because it reduces the number the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
+This is extremely beneficial in both open mass and multi-notch searches because it reduces the number the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
 
@@
 MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
-This expanded list of precursors is used in every spectrum match,frequently resulting in multiple peptide identifications from single tandem mass spectrum.
+This expanded list of precursors is used in every spectrum match, frequently resulting in multiple peptide identifications from single tandem mass spectrum.
 

### 2017-10-17 paper.tex
jdfkj

@@
 
-In the original G-PTM-D manuscript, an open mass search (+/- 200-500Da) was used to create a histogram of mass difference peaks, from which a number of PTMs could be inferred.
-A perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
+In the original G-PTM-D manuscript, an open mass search ($\pm 200-500 $ Da) was used to create a histogram of mass difference peaks, from which a number of PTMs could be inferred.
+A Perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
 This database was used for a final search and a number of PTMs were identified.
@@
 Four, missed monoisotopic peaks and co-isolated peaks were not handled.
-Five, separate programs were required (both Morpheus and a perl script).
+Five, separate programs were required (both Morpheus and a Perl script).
 MetaMorpheus resolves all of these issues and more.

### 2017-10-17 paper.tex
fjdkf

@@
 G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list.
-We have since increased the variety of PTMs added but including a number of new notches in the G-PTM-D first pass search.
-The new notches correspond to metal adducts (e.g.
-Na and K), glycosylations., sample handling and processing artifacts (e.g.
-carbamylation, ammonia loss), lipidation, and etc.
-MSFragger has limited mass range as reported.
-We identified peaks with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of Sulfonation and Phosphorylation, and a problematic peak at mass of Leucine, with a high FDR.
-Some Histogram Peaks For Mouse Data Some Histogram Peaks For Mouse Data We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals .
+We have since greatly increased the variety of PTMs added by including a number of new notches in the G-PTM-D first pass search.
+The new notches correspond to metal adducts (e.g. Na and K), glycosylations., sample handling and processing artifacts (e.g. carbamylation, ammonia loss), lipidation, and etc.
+The improvement in accuracy and lowering of mass error were sufficient that many new modifications at large mass and many PTM combinations could be confidently added and discovered using the newly enhanced G-PTM-D approach.
+We show, in TABLE/FIGURE below a list of PTMs used in the original G-PTM-D search along with the list of PTMs used in the enhanced G-PTM-D search.
+The number of different modifications increased from 22 to 333.
+The total number of PSMs increased by . We also show in the table the associated local FDR.
+
+We identified peaks in the original G-PTM-D approach with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
+Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of sulfonation and phosphorylation, and a problematic peak at mass of Leucine, with a high FDR (see Table above).
+INSERT A GREAT PLOT HERE We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
 MSFragger also doesn’t automatically calibrate spectra prior to searching.
@@
 
-Since the protein database is annotated with PTMs from the UniProt PTM list, it is natural to use it in the augmentation step as well.
-It has a few significant drawbacks: Some mass values do not correspond to actual mass shifts, lability of modifications is not taken into account, and many PTMs are missing, including adducts and glycosylations.
-Notch searches allow using expanded modification lists that include chemical modifications such as adducts and large mass PTMs such as glycosylations.
 Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMs.
@@
 
-%
+In the final table of this manuscript we show the combined effect of all G-PTM-D enhancements (calibration, mult-notch search for PTMs, limited multi-notch final search and coisolation.
+The number of PSMs, PTMs, unique peptides, PTM-containing proteins all went up a bunch.
+The total search time dropped by an order of magnitude.
+These enhancements demonstrate that this approach can and should be used whenever bottom-up proteomics is done. It is fast enough to be part of every regular work-flow.
+And, the significant contribution of signal from PTM peptides is clearly beneficial to biologists seeking to understand their data better.
+
 \begin{acknowledgement}

### 2017-10-18 paper.tex
noice

@@
 The changes described in this work lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Targeted multi-notch searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-mass and wide-mass searches.
+Targeted multi-notch searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-window and open database searches.
 \end{abstract}
@@
 
-Many types of post-translational modifications (PTMs) are known, and databases containing detailed information about such modifications are readily available\citep{Creasy_2004}.
+Many types of protein post-translational modifications (PTMs) are known, and databases containing detailed information about such modifications are readily available\citep{Creasy_2004}.
 Information about a modification often includes the chemical or isotopic composition, mass, specificity to certain residues, and possible restriction of placement to peptide or protein termini.
-Protein databases such as UniProt\citep{Uniprot_2017} include information on the localization of specific modifications to specific residues in the protein, but this information is incomplete.
-Despite the extensive cataloging of the different possible protein modifications, their comprehensive discovery and identification in complex biological samples has continued to pose a challenge for proteomics\citep{Olsen_2013}.
+Protein databases such as UniProt\citep{Uniprot_2017} include information on the localization of specific modifications to specific residues in the protein, but this information is far from complete.
+Despite the extensive cataloging of the different protein modifications, their comprehensive discovery and identification in complex biological samples has continued to pose a challenge for proteomics\citep{Olsen_2013}.
 
 Numerous procedures for identification and localization of PTMs from "bottom-up" tandem mass spectrometric datasets exist.
-These primarily address common modifications (e.g. phosphorylations) in well-studied systems, but global discovery tools, such as MODa\citep{Na_2011} and G-PTM-D\citep{Li_2016}, are also available.
-G-PTM-D is a bioinformatics tool for high confidence identification and localization of a wide variety of different PTMs.
-The G-PTM-D workflow consists of three stages: 1) A wide-mass database search\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
+These primarily address common modifications (e.g.
+phosphorylations) in well-studied systems, but global discovery tools, such as MODa\citep{Na_2011} and G-PTM-D\citep{Li_2016}, for identification and localization of a wide variety of different PTMs are also available.
+The G-PTM-D workflow consists of three stages: 1) A wide precursor mass tolerance database search ('open' search)\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
-3) A final standard narrow-mass search of the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g. 1\% FDR).
+3) A final standard tight precursor tolerance search ('narrow-window' search) using the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g.
+1\% FDR).
 
 While G-PTM-D delivers confident identification of numerous types of PTMs, it has a few significant limitations.
-The initial wide-mass search procedure is slow, taking hours for modest size datasets.
+The initial open search is slow, taking hours for modest size datasets.
 Thus, high mass modifications such as glycosylations are out of reach, since widening the search window from the suggested $\pm 200$ Da increases the already-slow search time.
-The final narrow-mass search is also problematic, as it does not account for peptide-spectral matches (PSMs) with mass differences arising from misassignment of the monoisotopic peak.
+The final narrow-window search is also problematic, as it does not account for peptide-spectral matches (PSMs) with mass differences arising from misassignment of the monoisotopic peak during precursor deisotoping.
 Discovery of modifications is limited to PTMs in the UniProt database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
@@
 Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files.
-To address the long search times of the wide-mass search, we replace the wide-mass search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
+To address the long search times of the open search, we replace the search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
 PTMs and amino acids).
 This change improves the specificity of the search and significantly decreases search time.
-The final narrow-mass search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and + 2.0054), further increasing the number of confidently identified peptides and PTMs.
+The final narrow-window search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and + 2.0054), further increasing the number of confidently identified peptides and PTMs.
 Finally, we replace the limited UniProt PTM list with a custom expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
@@
 We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
-MetaMorpheus is freely available as a standalone Windows executable program from GitHub \url{https://github.com/smith-chem-wisc/MetaMorpheus}.
-All of the source code is also available at that link.
-MetaMorpheus heavily utilizes a NuGet package called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
-mzLib can be found at link \url{https://github.com/smith-chem-wisc/mzLib}.
+MetaMorpheus is freely accessible at \url{https://github.com/smith-chem-wisc/MetaMorpheus} as a standalone Windows executable program.
+.NET Core versions, as well as the complete source code are available.
+MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
+mzLib can be found at \url{https://github.com/smith-chem-wisc/mzLib}.
 
@@
 These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells and 9 fractions from each of two different pancreatic islets from two species of mice.
-These datasets are available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215} and \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00470} repositories, and are described in detail in~\citep{Shortreed_2015, Cesnik_2016}.
+These datasets are available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215} and \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00470}, and are described in detail in~\citep{Shortreed_2015, Cesnik_2016}.
 
-UniProt XML protein databases used were downloaded on Sept 14, 2017, and contained only reviewed human/mouse proteins.
+UniProt XML protein databases used were downloaded on Sept 14, 2017, and contain only reviewed human/mouse/contaminant proteins.
 
@@
 a spectral calibration;
-a replacement of the wide parent mass tolerance search with a custom multi-notch search for PTM discovery;
+a replacement of the open search with a custom multi-notch search for PTM discovery;
 replacement of the final traditional narrow mass search with a limited multi-notch search that permits identification of incorrectly deconvoluted precursor masses;
@@
 It is useful to carefully describe the two commonly used database search approaches, prior to defining multi-notch searches and the impact that they have on G-PTM-D performance.
-A traditional search in bottom-up proteomics considers potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal within some defined tolerance (e.g.
+A traditional narrow-window search in bottom-up proteomics considers potential matches between theoretical peptides and experimental spectra if the candidate peptide mass and the observed precursor mass are equal within some defined tolerance (e.g.
 5ppm, or 2.1 Da).
-We refer to this type of search here as a \textit{single-notch} search.
-The other common search is a \textit{wide-parent tolerance search} (e.g. $\pm 500$ Da)\citep{Chick_2015,Kong_2017,Li_2016}.
+The other common search is an open search (e.g. $\pm 500$ Da)\citep{Chick_2015,Kong_2017,Li_2016}.
 In searches of that type, the precursor mass and the mass of the best matching theoretical peptide could be many Da apart.
 We describe an alternative to these search modes: the \textit{multi-notch} search.
-This search is an extension of the single-notch search that allows a variety of multiple specific mass differences between precursor mass and theoretical mass (See Figure~\ref{fig:fig3-searchTypes}) where the mass difference is easily interpretable.
-It can also be seen as a narrowing of the wide-parent tolerance search to only include prespecified mass differences.
+This search is an extension of the narrow-window search that allows a variety of multiple specific mass differences between precursor mass and theoretical mass (See Figure~\ref{fig:fig3-searchTypes}) where the mass difference is easily interpretable.
+It can also be seen as a narrowing of the open search to only allow prespecified mass differences.
 The advantages and disadvantages of this alternative strategy are described below.
@@
 
-\subsubsection{Traditional Single-Notch Search vs. Limited Multi-Notch Search}
+\subsubsection{Traditional Narrow-Window Search vs. Limited Multi-Notch Search}
 
-The traditional single-notch search limits the range of acceptable mass differences between a candidate peptide and the observed experimental precursor in order to find an exact match to the observed peptide in a protein database.
+The traditional narrow-window search limits the range of acceptable mass differences between a candidate peptide and the observed experimental precursor in order to find an exact match to the observed peptide in a protein database.
 A mass tolerance of <20ppm is commonly used on high-resolution MS1 data, resulting in a highly restrictive search.
-It is not uncommon for the precursor mass tolerance of single-notch searches to be relaxed to as much as 2 or 3 Da.
+It is not uncommon for the precursor mass tolerance of narrow-window searches to be relaxed to as much as 2 or 3 Da.
 One rationale for this is that errors in the deisotoping of MS1 spectra frequently result in a reported experimental precursor mass being different by 1.0029 or 2.0054 Da from the true isolated peptide mass.
@@
 Frequently, the best-fragmented peptide is different from the one corresponding to the precursor mass reported by the instrument, and the relaxed tolerance approach may identify this peptide, provided the reported charge state matches.
-The major limitation of this approach is that a 2 or 3 Da tolerance includes mass shifts that, when present, may indicate that the matched peptide is not the correct identification, arguably defeating the purpose of a narrow-mass search.
+The major limitation of this approach is that a 2 or 3 Da tolerance includes mass shifts that, when present, may indicate that the matched peptide is not the correct identification, arguably defeating the purpose of a narrow-window search.
 This may happen when the mass difference arises due to the identified peptide being different from the actual isolated precursor by a low mass modification or amino acid substitutions.
 
-The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4mouse-1012}, generated by compiling the results of a 2.5 Da tolerance single-notch search.
+The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4mouse-1012}, generated by compiling the results of a 2.5 Da tolerance narrow-window search.
 These plots clearly show that numerous peptides have missed monoisotopic (Mm) peaks, deamidation, or an amino acid substitution.
 
-We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wider (few Da) precursor mass tolerance during a traditional single-notch search.
+We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wider (few Da) precursor mass tolerance during a traditional narrow-window search.
 This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (1-20 ppm).
@@
 \centering
-\caption{Notch Search vs Narrow-Mass Search comparison of PSMs for Mouse}
+\caption{Notch Search vs Narrow-Window Search comparison of PSMs for Mouse}
 \label{tbl:singleVsMultiNotch}
@@
 
-\subsubsection{Multi-Notch Search vs. Wide-Mass Search}
+\subsubsection{Multi-Notch Search vs. Open Search}
 
 Chick et al.
-showed that by significantly relaxing the parent mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified\citep{Chick_2015}.
-In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in precursor and parent mass.
+showed that by significantly relaxing the precursor mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified\citep{Chick_2015}.
+In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in identified peptide mass and precursor mass.
 A major limitation of the Chick work was the extremely long search times because of the vast number of theoretical candidate peptides that were evaluated for each and every MS2 spectrum.
-Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for wide parent mass tolerance searches.
-Search times for $\pm 500$ Da parent tolerances went from six years to six seconds.
+Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for open searches.
+Search times for $\pm 500$ Da precursor tolerance went from six years to six seconds.
 The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
@@
 We compared searches of all Mouse islet fractions using a $\pm 500$ Da wide mass search versus a multi-notch search with notches at 0 and 79.96.
-The advantage of the multi-notch search over the wide-mass search is shown in Table~\ref{tbl:multiVsWide} by an increase in confidently identified PSMs (1\% FDR) for both regions (0 Da no-PTM peptides and +79.96 Da phosphorylated peptides).
+The advantage of the multi-notch search over the open search is shown in Table~\ref{tbl:multiVsWide} by an increase in confidently identified PSMs (1\% FDR) for both regions (0 Da no-PTM peptides and +79.96 Da phosphorylated peptides).
 
@@
 \centering
-\caption{Multi-Notch Search vs. Wide-Mass Search}
+\caption{Multi-Notch Search vs. Open Search}
 \label{tbl:multiVsWide}
 \begin{tabular}{lll}
-                    & Multi-Notch & Wide-Mass\\
+                    & Multi-Notch & Open\\
 0 peak & 139374      & 137483        \\
@@
 
-By employing the novel notch search strategy instead of the wide-mass search in the G-PTM-D process, using only the modifications listed in UniProt, the number of confidently identified PSMs increased by 6\%.
+By employing the novel multi-notch search strategy instead of the open search in the G-PTM-D process, using only the modifications listed in UniProt, the number of confidently identified PSMs increased by 6\%.
 The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
@@
 The final notch search uses notches centered at \{0, 1.0029, 2.0054\}, in order to capture missed identifications due to missed monoisotopic peaks.
-The results are contrasted with the standard narrow-mass final search.
+The results are contrasted with the standard narrow-window final search.
 The effect of using these notches is shown in Table~\ref{tab:table2}.
 
-We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the single-notch searches.
+We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the narrow-window searches.
 
@@
 \centering
-\caption{Effects of Replacing Final Narrow-Mass Search by a Notch Search}
+\caption{Effects of Replacing Final Narrow-Window Search by a Notch Search}
 \label{tab:table2}
 \begin{tabular}{ll|l|l}
-                      &        & Narrow-Mass Search & Notch Search\\
+                      &        & Narrow-Window Search & Notch Search\\
 \hline

### 2017-10-19 paper.tex
plots

@@
 \begin{figure}[H]
-  \includegraphics{paperDiagram.png}
+  \includegraphics[scale=0.1]{paperDiagram.png}
   \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color.}
@@
 a replacement of the open search with a custom multi-notch search for PTM discovery;
-replacement of the final traditional narrow mass search with a limited multi-notch search that permits identification of incorrectly deconvoluted precursor masses;
+replacement of the final traditional narrow mass search with a limited multi-notch search that permits exact matches even when incorrectly deisotoped precursor masses are present;
 and lastly, addition of capacity for identification of multiple co-fragmented peptides from a single MS/MS scan.

### 2017-10-19 FigureGeneration/fig1-calibErrors.py

@@ -2,25 +2,30 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-nice= pandas.read_csv(r"C:\Users\stepa\Data\PaperData\Jurkat\2017-07-31-17-44-19\Task1Calibrate\120426_Jurkat_highLC_Frac21inLinearbeforesc1.ms1dptsv", sep='\t')
-nice2= pandas.read_csv(r"C:\Users\stepa\Data\PaperData\Jurkat\2017-07-31-17-44-19\Task1Calibrate\120426_Jurkat_highLC_Frac21inNonLinearafterfc9.ms1dptsv", sep='\t')
+nice = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\2017-10-12-17-10-32\Task1-CalibratingAllJurkat\120426_Jurkat_highLC_Frac16init.ms1dptsv", sep='\t')
+nice2 = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\2017-10-12-17-10-32\Task1-CalibratingAllJurkat\120426_Jurkat_highLC_Frac16round7inter-scan.ms1dptsv", sep='\t')
 
 height = 6
 width = 7
 
+minY=-0.02
+maxY=0.02
+
 f, ((ax1, ax2), (ax3, ax4)) = plt.subplots(2, 2, sharex='col', sharey='row', figsize=(width,height))
 
-ax1.plot(nice['mz'],nice['Label'], 'bo',markersize = .3)
+ax1.plot(nice['mz'],nice['LabelTh'], 'bo',markersize = .3)
 ax1.hlines(0, min(nice['mz']), max(nice['mz']), zorder = 10)
+ax1.set_ylim([minY,maxY])
 
 ax1.set_ylabel('m/z error')
 ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
                 size='large', ha='right', va='center', xytext=(0,0.5))
 
-ax2.plot(nice['rt'],nice['Label'], 'bo',markersize = .3)
+ax2.plot(nice['rt'],nice['LabelTh'], 'bo',markersize = .3)
 ax2.hlines(0, min(nice['rt']), max(nice['rt']), zorder = 10)
+ax2.set_ylim([minY,maxY])
 
[…]

### 2017-10-19 FigureGeneration/fig2-10ppmSearchCalib.py

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-maxi = 40000
+maxi = 35000 
 
 
 height = 6
@@ -10,7 +10,7 @@ width = 7
 
 f, ((ax1), (ax3)) = plt.subplots(2, 1, sharex='col', sharey='row', figsize=(width,height))
 
-ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\Jurkat\2017-08-01-09-15-26\Task1Search\aggregatePSMs_10ppmAroundZero.psmtsv", sep='\t')  
+ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\1mm10ppmSearch\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')  
 
 ok['Mass Diff (ppm)'] = pandas.to_numeric(ok['Mass Diff (ppm)'], errors='coerce')
 
@@ -25,9 +25,7 @@ ax1.set_ylabel('PSM count')
 ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
                 size='large', ha='right', va='center', xytext=(0,0.5))
 
-maxi = 40000
-
-ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\Jurkat\2017-07-31-17-44-19\Task1Calibrate\2017-08-01-09-28-39\Task1Search\aggregatePSMs_10ppmAroundZero.psmtsv", sep='\t')  
+ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\1mm10ppmSearch\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')  
 
 ok['Mass Diff (ppm)'] = pandas.to_numeric(ok['Mass Diff (ppm)'], errors='coerce')
 

### 2017-10-19 FigureGeneration/fig3-searchTypes.py

@@ -19,7 +19,7 @@ yTicks = [1,2]
 
 plt.ylim([0.5,2.5])
 plt.yticks(yTicks, ylabels)
-plt.title("Narrow-Mass Search")
+plt.title("Narrow-Window Search")
 
 plt.subplot(312)
 
@@ -31,7 +31,7 @@ plt.plot(dates, values, '-')
 ylabels = ['Reject', 'Accept']
 yTicks = [1,2]
 
-plt.title("Notch Search")
+plt.title("Multi-Notch Search")
 plt.ylim([0.5,2.5])
 
 plt.yticks(yTicks, ylabels)
@@ -50,7 +50,7 @@ yTicks = [1,2]
 
 plt.ylim([0.5,2.5])
 
-plt.title("Wide-Mass Search")
+plt.title("Open Search")
 
 plt.yticks(yTicks, ylabels)
 plt.tight_layout()

### 2017-10-20 paper.tex
nice

@@
 
-The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4mouse-1012}, generated by compiling the results of a 2.5 Da tolerance narrow-window search.
+The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4jurkat-1012}, generated by compiling the results of a 2.5 Da tolerance narrow-window search.
 These plots clearly show that numerous peptides have missed monoisotopic (Mm) peaks, deamidation, or an amino acid substitution.
+
+\begin{figure}[H]
+\includegraphics{fig4jurkat-1012.png}
+\caption{Histograms PSM mass differences found in various dalton ranges}
+\label{fig:fig4jurkat-1012}
+\end{figure}
 
@@
 \begin{tabular}{llll}
-                    & 5ppm Search & 3.5 Da Search & Notch Search \\
-PSMs Within 1\% FDR & 157188      & 178003        & 177213       \\
-misassigned PSMs    & 0           & 1521          & 0            \\
+                    & 5ppm Search & 2.5 Da Search & Notch Search \\
+PSMs Within 1\% FDR & 275359      & 262543        & 279895       \\
+misassigned PSMs    & ?           & ?          & ?            \\
 \end{tabular}
 \end{table}
+
+A specific example of a misassigned PSM is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 10.
+With G-PTM-D added modifications, it is clear that the correct identification has a mass difference of 0 and a score 20.
 
@@
 The phosphopeptide PEPpSK could be reported with a mass error of -79.96 Da when the better identification would have been PEPSK.
-Constraining the search using the multi-notch approach described below, directly prohibits erroneous assignment of the best match for a peak, and instead produces an alternative peak assignment with a much lower false positive rate (see below).
 
@@
 
-[Cite Bo Zhang if paper exists].
+Consider the scan JurkatFrac12 scan 12336. Isolation m/z  = 726.65.
+Isolation width is 2.5 Th.
+It is shown in Figure~\ref{fig:fig5-coIsolationSpectrum}.
+
+\begin{figure}[H]
+\includegraphics{fig5-coIsolationSpectrum.png}
+\caption{m/z peaks in scan 12333}
+\label{fig:fig5-coIsolationSpectrum}
+\end{figure}
+
 We showed above in some amazing figure an example where several peptides were co-eluting from the column into that mass-spec within the tolerance used for isolation for fragmentation.

### 2017-10-20 FigureGeneration/Extras/SingleFileCalibQuality.py

@@ -1,70 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-maxi = 3000
-
-
-start = r"C:\Users\stepa\Data\CalibrationPaperData\Mouse\2017-07-10-17-50-26\Task1Calibrate\04-30-13_CAST_Frac5_4uL"
-
-plt.figure(num=1,figsize=(10.9,6))
-
-plt.suptitle(start.split('\\')[-1], fontsize=14)
-
-plt.figure(1)
-
-
-ok = pandas.read_csv(start+"PSMsBeforeLinearCalib.psmtsv", sep='\t')  
-#A=ok[r"Mass Diff (ppm)"]
-A=ok[r"Improvement Possible"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue"]
-
-
-ok = plt.subplot(131)
-plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-plt.ylim([1,maxi])
-plt.vlines(0,0,maxi*.65,zorder = 10)
-
-plt.title("Uncalibrated")
-
-
-ok = pandas.read_csv(start+"PSMsBeforeNonLinearCalib.psmtsv", sep='\t')  
-
-#A=ok[r"Mass Diff (ppm)"]
-A=ok[r"Improvement Possible"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue"]
-
-
-
-
-
-ok = plt.subplot(132)
-plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-plt.ylim([1,maxi])
-plt.vlines(0,0,maxi*.65,zorder = 10)
-
-
-plt.title("Linear Calibration")
-
-ok = pandas.read_csv(start+"PSMsAfterCalib.psmtsv", sep='\t')  
-
-#A=ok[r"Mass Diff (ppm)"]
-A=ok[r"Improvement Possible"]
-B=ok[r"Decoy/Contaminant/Target"]=="T"
-C=ok[r"QValue"]
-
-
-
-ok = plt.subplot(133)
-plt.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
-plt.ylim([1,maxi])
-plt.vlines(0,0,maxi*.65,zorder = 10)
-
-
-plt.title("Nice Calibration")
-
-plt.savefig('myimage.png', format='png', dpi=1200)
-
-plt.show()

### 2017-10-20 FigureGeneration/fig4-jurkat;-1;0;1;2.py

@@ -2,7 +2,7 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\Jurkat\2017-07-31-17-44-19\Task2Search\aggregatePSMs_3.5aroundZero.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\twoPointFiveAroundZero\Task1-SearchTask\aggregatePSMs_2.5daltonsAroundZero.psmtsv", sep='\t')
 
 
 
@@ -26,7 +26,7 @@ TDhighConf = B[C<0.01]
 
 ok = plt.subplot(221)
 
-halfWidth = 3.505
+halfWidth = 2.505
 min1 = -halfWidth
 max1 = halfWidth
 
@@ -41,10 +41,13 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.01), range=[min1, max
 plt.legend()
 
 plt.ylabel("Count")
-plt.title("PSMs within $\pm 3.5$ Da")
+plt.title("PSMs within $\pm 2.5$ Da")
 ok.set_yscale("log", nonposy='clip')
+ok.set_xlim([-2.5,2.5])
 
 ok = plt.subplot(223)
+ok.set_ylim([0,200])
+ok.set_xlim([-1.1,-0.9])
 
 
 #ok.set_yscale("log", nonposy='clip')
@@ -66,11 +69,12 @@ plt.legend()
 plt.ylabel("Count")
 plt.title("PSMs around -1 Da")
 
-plt.annotate('Oxidation\n+Ammonia Loss', xy=(-1.031634,50), xytext=(-1.031634-0.075, 80), arrowprops=dict(arrowstyle="->"))
-plt.annotate('Dehydro', xy=(-1.007825, 60), xytext=(-1,80), arrowprops=dict(arrowstyle="->"))
-plt.annotate('Amidation', xy=(-0.984016, 30), xytext=(-0.984016+0.03, 50), arrowprops=dict(arrowstyle="->"))
+plt.annotate('Oxidation\n+Ammonia\nLoss', xy=(-1.031634,90), xytext=(-1.031634-0.065, 140), arrowprops=dict(arrowstyle="->"))
[…]

### 2017-10-20 FigureGeneration/fig5-coIsolationSpectrum.py

@@ -1,0 +1,36 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+def frange(start, stop, step):
+    i = start
+    while i < stop:
+        yield i
+        i += step
+
+# This is from Jurkat 12 scan 12333
+
+mz = [0, 0, 0, 0, 0, 0, 162124.125, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 98809.40625, 0, 0, 0, 2548705.75, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 142358.0469, 0, 0, 195410.3438, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 322681.6563, 0, 0, 1411281.375, 0, 0, 1936653.75, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 333969, 0, 0, 202836.7031, 0, 0, 0, 0, 0, 96753.42188, 0, 0, 84754.38281, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 328649.5313, 0, 94499.35156, 0, 889372.375, 0, 1566347.875, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 5150338, 0, 0, 0, 0, 0, 223941.7188, 0, 569038, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 258916.7969, 0, 0, 0, 0, 0, 0, 5426417.5, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 84718.61719, 0, 986879.5625, 0, 0, 0, 173286.3438, 0, 0, 0, 0, 0, 0, 0, 0, 3149313.25, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0,
[…]

### 2017-10-20 FigureGeneration/table2threeNarrow.py

@@ -1,0 +1,48 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\twoPointFiveAroundZero\Task1-SearchTask\aggregatePSMs_2.5daltonsAroundZero.psmtsv", sep='\t')
+
+data['Mass Diff (Da)'] = pandas.to_numeric(data['Mass Diff (Da)'], errors='coerce')
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+len1 = len(massDiffs[C<0.01])
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2mm5ppm\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')
+
+
+data['Mass Diff (Da)'] = pandas.to_numeric(data['Mass Diff (Da)'], errors='coerce')
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+
+len2 = len(massDiffs[C<0.01])
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\exact5ppm\Task1-SearchTask\aggregatePSMs_5ppmAroundZero.psmtsv", sep='\t')
+
+
+
+
+data['Mass Diff (Da)'] = pandas.to_numeric(data['Mass Diff (Da)'], errors='coerce')
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+massDiffs=data_sorted['Mass Diff (Da)'].values
+
+
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+len3 = len(massDiffs[C<0.01])
+
+print(len3)
+print(len1) 
+print(len2)

### 2017-10-25 paper.tex
okay

@@
 A specific example of a misassigned PSM is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 10.
+It is Frac9 scan 21230.
 With G-PTM-D added modifications, it is clear that the correct identification has a mass difference of 0 and a score 20.
+
+\newpage
 
@@
 
-We compared searches of all Mouse islet fractions using a $\pm 500$ Da wide mass search versus a multi-notch search with notches at 0 and 79.96.
-The advantage of the multi-notch search over the open search is shown in Table~\ref{tbl:multiVsWide} by an increase in confidently identified PSMs (1\% FDR) for both regions (0 Da no-PTM peptides and +79.96 Da phosphorylated peptides).
+We compared searches of all Jurkat fractions using a $\pm 500$ Da wide mass search versus a multi-notch search with notches at 0 and 79.961 with 0.02 Dalton tolerance.
+The advantage of the multi-notch search over the open search is shown in Table~\ref{tbl:multiVsWide} by at least a $10\%$ increase in the number of target PSMs with score of at least 9.079, for both regions (0 Da no-PTM peptides and +79.96 Da phosphorylated and sulfonated peptides).
+It is important to point out that the difference in the numbers of PSMs is not due to FDR considerations, but rather to the erroneous peak assignment in open searches.
 
@@
 \begin{tabular}{lll}
-                    & Multi-Notch & Open\\
-0 peak & 139374      & 137483        \\
-79.96 peak & 3402           & 3298          \\
+                    & Open & Multi-Notch\\
+0 peak    & 176073  (0.15\% FDR)     & 193328 (0.23\% FDR)     \\
+79.96 peak     & 732   (0.68\% FDR)   & 957  (0.84\% FDR)           \\
 \end{tabular}
 \end{table}
+
+The same results are presented in a graph format in Figure~\ref{fig:fig6-ZeroAndPhospho}.
+
+\begin{figure}[H]
+\includegraphics{fig6-ZeroAndPhospho.png}
+\caption{Zero and Phoshpo and Sulfo}
+\label{fig:fig6-ZeroAndPhospho}
+\end{figure}
+
+A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat frac 5.
+The open search identifies ATPARAPESPPSADPALVAGPAEEAEC[Common Fixed:Carbamidomethyl of C]PPPR as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
+The identified b ion matches are 172.08479,269.13756,880.44028.
+The better match is identified by the multi-notch search, as APESPPSADPALVAGPAEEAEC[Common Fixed:Carbamidomethyl of C]PPPR with a score of 16.246 and mass difference of 79.9655, and no missed cleavages.
+The identified b ion matches are 168.08988,297.13247.
+So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of ATPAR and addition of phosphorylation), we have an easily identifyable mass shift corresponding to a phosphoyrlation site.
 
@@
 
-We showed above in some amazing figure an example where several peptides were co-eluting from the column into that mass-spec within the tolerance used for isolation for fragmentation.
-In that particular case, the precursor with m/z dsfadf was chosen for isolation and fragmentation.
+We showed above in Figure~\ref{fig:fig5-coIsolationSpectrum} an example where several peptides were co-eluting from the column into that mass-spec within the tolerance used for isolation for fragmentation.
+In that particular case, the precursor with m/z 726.65 was chosen for isolation and fragmentation.
 In a conventional search, only that precursor mass would be used to select a range of theoretical for comparison.
 However, all peptides in that range are co-fragmented and a better solution would be to use all co-isolated precursor masses.
-In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor ($\pm 3$ Th).
+In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor ($\pm 1.25$ Th).
 Each deconvoluted mass is used to select a range of theoreticals to use for identification.

### 2017-10-25 FigureGeneration/fig6-ZeroAndPhoshpo.py

@@ -1,0 +1,158 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+height = 6
+width = 7
+plt.figure(figsize=(width, height))
+
+#################
+
+# data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2017-10-20-12-48-57-zeroAndMidPhosphoSulfo\Task1-SearchTask\aggregatePSMs_custom.psmtsv", sep='\t')
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2017-10-24-12-02-12-noDeisotopeZeroSulfoPhospho\Task1-SearchTask\aggregatePSMs_zeroAndMidSulfoPhospho.psmtsv", sep='\t')
+
+data['Mass Diff (Da)'] = pandas.to_numeric(data['Mass Diff (Da)'], errors='coerce')
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+
+massDiffs=data_sorted['Mass Diff (Da)'].values
+scores=data_sorted['Score'].values
+notches=data_sorted['Notch'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+scoreCutoff = min(scores[(C<0.01) &(notches=="1")])
+
+print(scoreCutoff)
+
+massDiffsHighConf = massDiffs[scores>scoreCutoff]
+TDhighConf = B[scores>scoreCutoff]
+
+#################
+
+ok = plt.subplot(222)
+
+halfWidth = 0.02 
+min1 = -halfWidth
+max1 = halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+print(len(massDiffsHighConf1))
+print(len(massDiffsHighConf1decoy))
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, max1], label = "Decoys")
+
[…]

### 2017-10-27 paper.tex
exampleOfWrongId

@@
 
-A specific example of a misassigned PSM is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 10.
-It is Frac9 scan 21230.
-With G-PTM-D added modifications, it is clear that the correct identification has a mass difference of 0 and a score 20.
+A specific example of a misassigned PSM is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence QAFDDAIAELDTLNEDSYK, Jurkat Frac9 scan 20962.
+With G-PTM-D added modifications, the correct peptide is identified with sequence [PeptideTermMod:Ammonia loss on Q]QAFDDAIAELDTLNEDSY[Mod:Oxidation on Y]K.
+It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
 

### 2017-11-08 FigureGeneration/table2threeNarrow.py
noice

@@ -2,6 +2,76 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
+
+
+print('Loading orig...')
+
+#5ppm search
+#data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\exact5ppm\Task1-SearchTask\aggregatePSMs_5ppmAroundZero.psmtsv", sep='\t')
+
+#2.5 da search
+#data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\twoPointFiveAroundZero\Task1-SearchTask\aggregatePSMs_2.5daltonsAroundZero.psmtsv", sep='\t')
+
+#notch search
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2mm5ppm\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')
+
+print('Loading final...')
+
+dataFinal = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2017-10-20-16-43-18-final\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')
+
+print('Starting loop...')
+
+
+with open('workfileC.tsv', 'w') as f:
+  for index, row in data.iterrows():
+    final = dataFinal.loc[(dataFinal['File Name'] == row['File Name']) & (dataFinal['Scan Number'] == row['Scan Number']) & (dataFinal['Precursor MZ'] == row['Precursor MZ'])]
+    initScore = row['Score']
+    if(len(final['Score']) == 0):
+      continue
+    finalScore = final['Score'].iat[0]
+    if(int(initScore)<int(finalScore) and 'Phospho' not in row['Full Sequence'] and final['Notch'].iat[0] =="0"):
+      if(row['QValue Notch'] >= 0.01):
+        break
[…]

### 2017-11-09 paper.tex
noice
okay

@@
 Recently, Kong\citep{Kong_2017} reported a new index-based search strategy (MSFragger) that radically improved search speeds for open searches.
-Search times for $\pm 500$ Da precursor tolerance went from six years to six seconds.
+$\pm 500$ Da precursor tolerance searches saw a 100-fold improvement in speed over most existing proteome database search tools.
 The output of both search strategies is a list of PSMs, where the best matching unmodified theoretical peptide was found for each MS2 fragmentation spectrum and the mass difference between the theoretical and experimental precursor masses was determined.
@@
 INSERT A GREAT PLOT HERE We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
-MSFragger also doesn’t automatically calibrate spectra prior to searching.
+MSFragger also doesn't automatically calibrate spectra prior to searching.
 Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.

### 2017-11-09 paper.tex
noice

@@
 PSMs Within 1\% FDR & 275359      & 262543        & 279895       \\
-misassigned PSMs    & ?           & ?          & ?            \\
+misassigned PSMs    & 2041           & 9041          & 1371            \\
 \end{tabular}
@@
 
-A specific example of a misassigned PSM is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence QAFDDAIAELDTLNEDSYK, Jurkat Frac9 scan 20962.
+One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence QAFDDAIAELDTLNEDSYK, Jurkat Frac9 scan 20962.
 With G-PTM-D added modifications, the correct peptide is identified with sequence [PeptideTermMod:Ammonia loss on Q]QAFDDAIAELDTLNEDSY[Mod:Oxidation on Y]K.
 It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
+
+While this kind of misassgnment is not possible in the 5ppm narrow-window search, often missed monoistopes are assigned to a wrong sequence.
+
+Finally, even the multi-notch search has some misassignments.
+A specific example is
+
+25956	0	25.126	20.112	AEDGHEEAMETEASTSGEV[Mod:Oxidation on V]AEVAEEAMETESSEK	AEDGHEEAM[Common Variable:Oxidation of M]ETEASTSGEVAEVAEEAMETESSEK	3614.45197	3614.45197
+
+130059	0.000192	18.168	13.096	[PeptideTermMod:Ammonia loss on Q]QLLC[Common Fixed:Carbamidomethyl of C]DLVGISPSSEHDVLK	LQALQSQVEFLEQSMVDK	2092.05106	2092.05106	0
+
+131801	0.000192	18.193	13.065	RN[Mod:Deamidation on N]DDISELEDLSELEDLK	LEEELSVEGNTDINTDLKD	2132.9961	2132.9961	0
+
+156807	0.000278	20.257	12.033	DLGLSESGE[Metal:Sodium on E]DVNAAILDESGK	PVTVEPMDQLDDEEGLPEK	2139.97838	2139.98818	-0.0098
 
@@
 The identified b ion matches are 168.08988,297.13247.
-So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of ATPAR and addition of phosphorylation), we have an easily identifyable mass shift corresponding to a phosphoyrlation site.
+So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of ATPAR and addition of phosphorylation), we have an easily identifiable mass shift corresponding to a phosphorylation site.
 
 \subsection{Co-isolation}
+
+Peptide co-isolation and co-fragmentation complicates mass spec data analysis, but carefully taking co-isolation into account provides substantially better results~\cite{Zhang2014}.
 
@@
 
-In the original G-PTM-D manuscript, an open mass search ($\pm 200-500 $ Da) was used to create a histogram of mass difference peaks, from which a number of PTMs could be inferred.
+In the original G-PTM-D manuscript, an open mass search ($\pm 200-500$ Da) was used to create a histogram of mass difference peaks, from which PTMs could be inferred.
 A Perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
@@
 There were several limitations to that work.
-One, the open mass search was really slow.
+One, the open mass search was unacceptably slow.
 Two, uncalibrated data made it hard to discern closely spaced modifications.
@@
 MetaMorpheus resolves all of these issues and more.
-Below we will show how each individual enhancement to G-PTM-D improved performance over the original work and then we will show at the end, the combined effect of all enhancements.
+In the preceding sections, we detailed the performance improvements introduced by each enhancements: in this section we show their combined effect.
 
@@
 
-We identified peaks in the original G-PTM-D approach with anomalously high false discovery rates, which could be excluded from wide mass searches in order to improve the overall identification quality.
-Some of these problematic histogram peaks are illustrated in Figure, including peaks corresponding to masses of sulfonation and phosphorylation, and a problematic peak at mass of Leucine, with a high FDR (see Table above).
-INSERT A GREAT PLOT HERE We identified some specific mass shifts that are characterized by anomalously high false discovery rates: these mass shifts correspond precisely to various combinations of residue mass additions and removals.
-MSFragger also doesn't automatically calibrate spectra prior to searching.
-Therefore, mass differences from closely spaced PTMs (e.g. phosphorylation and sulfonation) can easily be misinterpreted.
-MetaMorpheus has built in functionality to identify multiple peptides from each tandem mass spectrum resulting in improved identification of peptides and peptide modifications.
-
-By employing the novel multi-notch search strategy instead of the open search in the G-PTM-D process, using only the modifications listed in UniProt, the number of confidently identified PSMs increased by 6\%.
-The overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
+By employing the novel multi-notch search strategy instead of the open search in the initial G-PTM-D search, the overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
 
@@
 \centering
-\caption{Effects of Replacing Initial Wide-Mass Search by a Multi-Notch Search}
+\caption{Effects of Replacing Initial Open Search by a Multi-Notch Search}
 \label{my-labelff}
 \begin{tabular}{ll|l|l}
-                      &        & Wide-Mass Search & Notch Search\\
+                      &        & Open Search & Multi-Notch Search\\
 \hline
@@
                       & Mouse  & 13.5 hrs         & 1.2 hrs   \\
-\hline
-\multirow{2}{*}{PSMs} & Jurkat & 203237           & 210566    \\
-                      & Mouse  & 153674           & 162473   
 \end{tabular}
@@
 
-The final notch search uses notches centered at \{0, 1.0029, 2.0054\}, in order to capture missed identifications due to missed monoisotopic peaks.
+In order to capture missed identifications due to missed monoisotopic peaks, the final mulit-notch search uses notches centered at \{0, 1.0029, 2.0054\}, .
 The results are contrasted with the standard narrow-window final search.
 The effect of using these notches is shown in Table~\ref{tab:table2}.
-
-We observe that when using an incomplete database that does not include comprehensive modification and sequence variation information, a notch search with notches at monoisotopic peaks is superior to either of the narrow-window searches.
 
@@
 \end{table}
+
+\subsubsection{Variations}
+
+The numbers reported above reflect the best way of running G-PTM-D.
+There are two natural questions that arise: 1) What if we use a FASTA file instead of a unimod XML? 2) What if we don't include the modifications not interesting to biologists?
+In this section we attempt to convince the reader that the way we do G-PTM-D is the best.
 

### 2017-11-13 FigureGeneration/fig7-phosphoSulfoUncalibCalib.py
figs

@@ -1,0 +1,103 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import pandas
+
+height = 3
+width = 7
+plt.figure(figsize=(width, height))
+
+#################
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\CAST\gptmd\Task1-GPTMDTask\PSMs.psmtsv", sep='\t')
+
+data['Mass Diff (Da)'] = pandas.to_numeric(data['Mass Diff (Da)'], errors='coerce')
+data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
+
+massDiffs=data_sorted['Mass Diff (Da)'].values
+scores=data_sorted['Score'].values
+notches=data_sorted['Notch'].values
+B=data_sorted['Decoy/Contaminant/Target'].values
+C=data_sorted['QValue Notch'].values
+
+scoreCutoff = min(scores[(C<0.01) &(notches=="1")])
+
+print(scoreCutoff)
+
+massDiffsHighConf = massDiffs[scores>scoreCutoff]
+TDhighConf = B[scores>scoreCutoff]
+
+#################
+
+ok = plt.subplot(121)
+
+halfWidth = 0.02 
+centerMass = 79.961573
+min1 = centerMass-halfWidth
+max1 = centerMass+halfWidth
+
+massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
+massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
+
+print(len(massDiffsHighConf1))
+print(len(massDiffsHighConf1decoy))
+
+plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
+plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, max1], label = "Decoys")
+
+plt.legend()
+
+plt.ylabel("Count")
+plt.title("Uncalibrated")
+
+ok.set_xlim([min1, max1])
+ok.set_ylim([0,60])
+
+#################
+
+data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\CAST\Calibrated\gptmd\Task1-GPTMDTask\PSMs.psmtsv", sep='\t')
+
[…]

### 2017-11-13 paper.tex
stuff

@@
 \begin{figure}[H]
-  \includegraphics[scale=0.1]{paperDiagram.png}
+  \includegraphics[scale=0.5]{paperDiagram.png}
   \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color.}
@@
 \centering
-\caption{Notch Search vs Narrow-Window Search comparison of PSMs for Mouse}
+\caption{Multi-Notch Search vs Narrow-Window Search comparison of PSMs for Jurkat}
 \label{tbl:singleVsMultiNotch}
 \begin{tabular}{llll}
-                    & 5ppm Search & 2.5 Da Search & Notch Search \\
+                    & 5ppm Search & 2.5 Da Search & Multi-Notch Search \\
 PSMs Within 1\% FDR & 275359      & 262543        & 279895       \\
@@
 
-One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence QAFDDAIAELDTLNEDSYK, Jurkat Frac9 scan 20962.
-With G-PTM-D added modifications, the correct peptide is identified with sequence [PeptideTermMod:Ammonia loss on Q]QAFDDAIAELDTLNEDSY[Mod:Oxidation on Y]K.
-It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
+One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence QAFDDAIAELDTLNEDSYK, in Jurkat Frac9 scan 20962.
+With G-PTM-D added modifications, the correct peptide is identified with sequence [Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K. It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
 
 While this kind of misassgnment is not possible in the 5ppm narrow-window search, often missed monoistopes are assigned to a wrong sequence.
+This explains the higher number of misassignments in the 5ppm search over the number in the multi-notch search.
 
 Finally, even the multi-notch search has some misassignments.
-A specific example is
+One example is the wrong identification of AEDGHEEAM[Oxidation]ETEASTSGEVAEVAEEAMETESSEK with a score of 20.112, when the correct identification, after the G-PTM-D procedure is AEDGHEEAMETEASTSGEV[Oxidation]AEVAEEAMETESSEK, where the masses are identical.
 
-25956	0	25.126	20.112	AEDGHEEAMETEASTSGEV[Mod:Oxidation on V]AEVAEEAMETESSEK	AEDGHEEAM[Common Variable:Oxidation of M]ETEASTSGEVAEVAEEAMETESSEK	3614.45197	3614.45197
-
-130059	0.000192	18.168	13.096	[PeptideTermMod:Ammonia loss on Q]QLLC[Common Fixed:Carbamidomethyl of C]DLVGISPSSEHDVLK	LQALQSQVEFLEQSMVDK	2092.05106	2092.05106	0
-
-131801	0.000192	18.193	13.065	RN[Mod:Deamidation on N]DDISELEDLSELEDLK	LEEELSVEGNTDINTDLKD	2132.9961	2132.9961	0
-
-156807	0.000278	20.257	12.033	DLGLSESGE[Metal:Sodium on E]DVNAAILDESGK	PVTVEPMDQLDDEEGLPEK	2139.97838	2139.98818	-0.0098
-
-\newpage
+%25956	0	25.126	20.112	AEDGHEEAMETEASTSGEV[Mod:Oxidation on V]AEVAEEAMETESSEK	AEDGHEEAM[Common Variable:Oxidation of M]ETEASTSGEVAEVAEEAMETESSEK	3614.45197	3614.45197
+%130059	0.000192	18.168	13.096	[PeptideTermMod:Ammonia loss on Q]QLLC[Common Fixed:Carbamidomethyl of C]DLVGISPSSEHDVLK	LQALQSQVEFLEQSMVDK	2092.05106	2092.05106	0
+%131801	0.000192	18.193	13.065	RN[Mod:Deamidation on N]DDISELEDLSELEDLK	LEEELSVEGNTDINTDLKD	2132.9961	2132.9961	0
+%156807	0.000278	20.257	12.033	DLGLSESGE[Metal:Sodium on E]DVNAAILDESGK	PVTVEPMDQLDDEEGLPEK	2139.97838	2139.98818	-0.0098
 
@@
 A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat frac 5.
-The open search identifies ATPARAPESPPSADPALVAGPAEEAEC[Common Fixed:Carbamidomethyl of C]PPPR as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
-The identified b ion matches are 172.08479,269.13756,880.44028.
-The better match is identified by the multi-notch search, as APESPPSADPALVAGPAEEAEC[Common Fixed:Carbamidomethyl of C]PPPR with a score of 16.246 and mass difference of 79.9655, and no missed cleavages.
-The identified b ion matches are 168.08988,297.13247.
+The open search identifies ATPARAPESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
+The better match is identified by the multi-notch search, as APESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR with a score of 16.246 and mass difference of 79.9655, and no missed cleavages.
 So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of ATPAR and addition of phosphorylation), we have an easily identifiable mass shift corresponding to a phosphorylation site.
@@
 
-Consider the scan JurkatFrac12 scan 12336. Isolation m/z  = 726.65.
+Consider the Jurkat Frac12 scan 12336.
+Isolation m/z  = 726.65.
 Isolation width is 2.5 Th.
 It is shown in Figure~\ref{fig:fig5-coIsolationSpectrum}.
+There are at least four different peptides being co-isolated and co-fragmented, and there is evidence in the MS/MS spectra for all of them.
 
@@
 
-We showed above in Figure~\ref{fig:fig5-coIsolationSpectrum} an example where several peptides were co-eluting from the column into that mass-spec within the tolerance used for isolation for fragmentation.
-In that particular case, the precursor with m/z 726.65 was chosen for isolation and fragmentation.
-In a conventional search, only that precursor mass would be used to select a range of theoretical for comparison.
+In a conventional search, only the isolation precursor mass would be used to select a range of theoretical for comparison.
 However, all peptides in that range are co-fragmented and a better solution would be to use all co-isolated precursor masses.
@@
 The result of this is that often, more than one peptide can be identified from a single MS2 scan.
-Below, in Table 3434, we should the number of PSMs for a raw file where we use only a single precursor mass and we compare that to the case where we allow for multiple different precursors.
-You can see a significant increase in the number of identified PSMs.
-Importantly, you can see in Supplemental Figure erf3e, that the precursor mass errors go down significantly and the standard deviation is much reduced.
 This is extremely beneficial in both open mass and multi-notch searches because it reduces the number the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
+%Below, in Table 3434, we should the number of PSMs for a raw file where we use only a single precursor mass and we compare that to the case where we allow for multiple different precursors.
+%You can see a significant increase in the number of identified PSMs.
+%Importantly, you can see in Supplemental Figure erf3e, that the precursor mass errors go down significantly and the standard deviation is much reduced.
 
-Get the story about how precurors with similar m/z with different z can in their deconvoluted state be many (10s, 100s of Da apart).
-This is a big problem in general searches but also for open mass and multi-notch searches which are used to discover possible PTMs.
+%Get the story about how precurors with similar m/z with different z can in their deconvoluted state be many (10s, 100s of Da apart).
+%This is a big problem in general searches but also for open mass and multi-notch searches which are used to discover possible PTMs.
 
-MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
-This expanded list of precursors is used in every spectrum match, frequently resulting in multiple peptide identifications from single tandem mass spectrum.
+%MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
+%This expanded list of precursors is used in every spectrum match, frequently resulting in multiple peptide identifications from single tandem mass spectrum.
 
@@
 
-In the original G-PTM-D manuscript, an open mass search ($\pm 200-500$ Da) was used to create a histogram of mass difference peaks, from which PTMs could be inferred.
+In the original G-PTM-D manuscript, an open mass search ($\pm 200$ Da) was used to create a histogram of mass difference peaks, from which PTMs could be inferred.
 A Perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
@@
 MetaMorpheus resolves all of these issues and more.
-In the preceding sections, we detailed the performance improvements introduced by each enhancements: in this section we show their combined effect.
+In the preceding sections, we detailed the performance improvements introduced by each enhancement: in this section we show their combined effect.
+Every comparison is a contrast of results from the original G-PTM-D workflow with the fully enhanced one.
 
@@
 Calibration and the associated increase in mass accuracy is essential for discrimination of small differences between mass shifts due to different modifications.
-To illustrate this, mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:phosphoHistograms}.
+To illustrate this, mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:fig7-ZeroAndPhospho}.
 It is evident from these histograms that the phosphorylation and sulfonation peaks are clearly separated and identifiable when the search is conducted on well-calibrated spectra.
 
-[THERE IS SUPPOSED TO BE A COOL FIGURE HERE]
+\begin{figure}[H]
+\includegraphics{fig7-ZeroAndPhospho.png}
+\caption{Peaks around phospho and sulfo in uncalibrated and calibrated data}
+\label{fig:fig7-ZeroAndPhospho}
+\end{figure}
 
-G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list.
-We have since greatly increased the variety of PTMs added by including a number of new notches in the G-PTM-D first pass search.
-The new notches correspond to metal adducts (e.g. Na and K), glycosylations., sample handling and processing artifacts (e.g. carbamylation, ammonia loss), lipidation, and etc.
-The improvement in accuracy and lowering of mass error were sufficient that many new modifications at large mass and many PTM combinations could be confidently added and discovered using the newly enhanced G-PTM-D approach.
-We show, in TABLE/FIGURE below a list of PTMs used in the original G-PTM-D search along with the list of PTMs used in the enhanced G-PTM-D search.
-The number of different modifications increased from 22 to 333.
-The total number of PSMs increased by . We also show in the table the associated local FDR.
-
-By employing the novel multi-notch search strategy instead of the open search in the initial G-PTM-D search, the overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
+The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass shifts in the open and the multi-notch searches.
+We do not show this difference here, since it is not informative due to the fact that the modifications added to the database are plausible ptms and not actually discovered ones.
+Nevertheless, one advantage of the enhanced approach is clear: the overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
 
@@
 
-In order to capture missed identifications due to missed monoisotopic peaks, the final mulit-notch search uses notches centered at \{0, 1.0029, 2.0054\}, .
-The results are contrasted with the standard narrow-window final search.
-The effect of using these notches is shown in Table~\ref{tab:table2}.
+After the final search using the augmented database is conducted (narrow-window in the original, and limited multi-notch in the enhanced G-PTM-D), the numbers of PSMs, peptides, and proteins are shown in Table~\ref{tab:table2}.
+The numbers all went up.
 
@@
 \centering
-\caption{Effects of Replacing Final Narrow-Window Search by a Notch Search}
+\caption{Enhanced G-PTM-D overall improvement numbers}
 \label{tab:table2}
 \begin{tabular}{ll|l|l}
-                      &        & Narrow-Window Search & Notch Search\\
+                      &        & Original G-PTM-D & Enhanced G-PTM-D\\
 \hline
 \multirow{2}{*}{PSMs} & Jurkat  & 210566   &  220574  \\
-                      & Mouse    & 162473   &   170743
+                      & Mouse    & 162473   &   170743 \\
+\hline
+\multirow{2}{*}{Unique Peptides} & Jurkat  & 85542   &  101980  \\
+                      & Mouse    & ?   &   ? \\
+\hline
+\multirow{2}{*}{Proteins} & Jurkat  & 9254   &  9587  \\
+                      & Mouse    & ?   &   ? \\
 \end{tabular}
 \end{table}
+
+We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da.
+
+These enhancements demonstrate that this approach can and should be used whenever bottom-up proteomics is done.
+It is fast enough to be part of every regular work-flow.
+And, the significant contribution of signal from PTM peptides is clearly beneficial to biologists seeking to understand their data better.
 
@@
 The numbers reported above reflect the best way of running G-PTM-D.
-There are two natural questions that arise: 1) What if we use a FASTA file instead of a unimod XML? 2) What if we don't include the modifications not interesting to biologists?
+There are two natural questions that arise: 1) What if we use a FASTA file instead of a unimod XML?
+2) What if we don't include the modifications not interesting to biologists, for example sample handling artifacts?
 In this section we attempt to convince the reader that the way we do G-PTM-D is the best.
 
+G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list.
+We have since greatly increased the variety of PTMs added by including a number of new notches in the G-PTM-D first pass search: a curated set of notches and the corresponding modifications is given in Supplementary Table 1.
 Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMs.
-A curated set of notches and the corresponding modifications is given in Appendix A.
-The overall identification rate increased from 164,697 to 179,345 peptide-spectral matches (PSMs), which in turn increased the number of modified peptides by an additional 20\%.
-We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da.
+The overall identification rate decreases from 220,574 to 217,597 peptide-spectral matches (PSMs), which in turn decreased the number of identified modified peptides by 4\%.
 
@@
 \hline
-\multirow{2}{*}{PSMs} & Jurkat  & 220574   &  222985 & 223578\\
+\multirow{2}{*}{PSMs} & Jurkat  & 217597   &  218547 & 220574\\
                       & Mouse    & 170743   &   171432& 174783 
@@
 \end{table}
-
-In the final table of this manuscript we show the combined effect of all G-PTM-D enhancements (calibration, mult-notch search for PTMs, limited multi-notch final search and coisolation.
-The number of PSMs, PTMs, unique peptides, PTM-containing proteins all went up a bunch.
-The total search time dropped by an order of magnitude.
-These enhancements demonstrate that this approach can and should be used whenever bottom-up proteomics is done. It is fast enough to be part of every regular work-flow.
-And, the significant contribution of signal from PTM peptides is clearly beneficial to biologists seeking to understand their data better.
 
@@
 
-\newpage
-
 \bibliography{citations}

### 2017-11-21 paper.tex
yah

@@
 
-\title[Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches]
-  {Rapid and Accurate Global PTM Discovery (G-PTM-D) Using Post-Acquisition Spectral Calibration and Multi-Notch Searches}
+\title[Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus]
+  {Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus}
+
+% Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus
 
 \begin{document}
+
+\sloppy
 
@@
 Correct identification of protein post-translational modifications (PTMs) is crucial to understanding many aspects of protein function in biological processes.
-G-PTM-D\citep{Li_2016} is a recently developed tool for global identification and localization of PTMs.
+G-PTM-D is a recently developed technique for global identification and localization of PTMs.
 Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and speed of PTM identification.
-We enhance G-PTM-D by using targeted multi-notch searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
-The changes described in this work lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in the search time.
-Targeted multi-notch searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-window and open database searches.
+We enhance G-PTM-D by using multi-notch searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
+The changes described in this work lead to a 20\% increase in the number of identified modifications and an order of magnitude decrease in search time.
+The complete workflow is implemented in MetaMorpheus, a software tool which integrates the database search procedure, identification of co-isolated peptides, spectral calibration and the enhanced G-PTM-D workflow.
+Multi-notch searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-window and open database searches.
 \end{abstract}
@@
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
-3) A final standard tight precursor tolerance search ('narrow-window' search) using the augmented database to identify both modified and unmodified peptides subject to the standard FDR threshold (e.g.
-1\% FDR).
+3) A final standard tight precursor tolerance search ('narrow-window' search) using the augmented database to identify both modified and unmodified peptides subject to a standard false discovery rate (FDR) threshold based upon the target:decoy strategy\citep{Elias_2007}.
 
@@
 
-Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files.
+The enhanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
+Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files: for this purpose we implemented a calibration procedure in MetaMorpheus.
 To address the long search times of the open search, we replace the search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
@@
 This change improves the specificity of the search and significantly decreases search time.
-The final narrow-window search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and + 2.0054), further increasing the number of confidently identified peptides and PTMs.
-Finally, we replace the limited UniProt PTM list with a custom expanded list of modifications that includes chemical adducts, glycosylations, lipids, and other important modifications.
-This change not only expands the list of identified and localized modifications, but also increases the number of confidently identified PSMs.
-The enhanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
+The database augmentation step is similar to the original G-PTM-D procedure, but the modification list is significantly expanded to accommodate modifications such as chemical adducts, glycosylations, lipids, and other important modifications.
+The final narrow-window search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and + 2.0054), and the capability for identifying co-isolated peptides, further increasing the number of confidently identified peptides and PTMs.
 
@@
   \includegraphics[scale=0.5]{paperDiagram.png}
-  \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color.}
+  \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color. The dark orange color corresponds to the inputs.}
   \label{fgr:diagram}
@@
 We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
-MetaMorpheus is freely accessible at \url{https://github.com/smith-chem-wisc/MetaMorpheus} as a standalone Windows executable program.
-.NET Core versions, as well as the complete source code are available.
-MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
+MetaMorpheus is freely available at \url{https://github.com/smith-chem-wisc/MetaMorpheus} as standalone Windows and Linux programs.
+MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.x
 mzLib can be found at \url{https://github.com/smith-chem-wisc/mzLib}.
@@
 These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells and 9 fractions from each of two different pancreatic islets from two species of mice.
-These datasets are available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215} and \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00470}, and are described in detail in~\citep{Shortreed_2015, Cesnik_2016}.
+These datasets are available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215} and \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00470}, and are described in detail elsewhere~\citep{Sheynkman2013, Shortreed_2015, Cesnik_2016}.
 
@@
 
-209 modifications including PTMs, adducts, and chemical modifications were allowed in the database augmentation step, see Supplementary Table 1 for a complete listing.
+The database augmentation step allowed 209 modification types, including PTMs, adducts, and chemical modifications. See Supplementary Table 1 for a complete listing.
 
@@
 In our strategy, the m/z differences between expected and observed peaks in the peptide tandem MS spectra are compiled, and then used to recalibrate the spectra.
-The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 0.01 Dalton fragment) database search is performed on the dataset.
-This yields a set of peptide identifications subject to a desired false discovery rate (e.g. 1\% based upon the target:decoy strategy\citep{Elias_2007}).
-Step 2: The identifications from Step 1 are used to extract peak matches from the spectra, and a peak shifting procedure is performed.
+The spectral calibration procedure consists of two steps: Step 1: A narrow tolerance (e.g. 10 ppm precursor, 20 ppm fragment) database search is performed on the dataset.
+This yields a set of peptide identifications subject to a desired FDR.
+Step 2: The identifications from Step 1 provide a theoretical list of m/z peaks that should match closely to observed m/z peaks in the spectra file, and the discrepancies between the theoretical and observed peaks are used to inform a peak shifting procedure.
 The calibration algorithm is described in detail in Supplementary Material 2.
 
-We demonstrate the effectiveness of the calibration approach using a single CAST Mouse raw file as an example.
+We demonstrate the effectiveness of the calibration approach using a single Jurkat fraction (16) as an example.
 Figure~\ref{fgr:fig1-calibErrors} displays the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, both before and after the calibration procedure is applied.
 These errors are plotted versus two important variables: the m/z value of the peak and the retention time of the corresponding scan.
+In addition to these two parameters, the calibration function native to MetaMorpheus also considers peak intensity, injection time, total ion count, and other variables as potentially influencing the m/z errors.
 Calibration successfully removes both linear and non-linear dependencies between the m/z errors and these variables.
-We note that in addition to these two parameters, the calibration function native to MetaMorpheus also considers peak intensity, injection time, total ion count, and other variables as potentially influencing the m/z errors.
+Prior to calibration, 95\% of the errors fell between -8.4 and 3.1 ppm, and after calibration 95\% of the errors fell between -1.7 and 2.0 ppm.
 MetaMorpheus calibrates both MS1 and MS2 scans (using MS2 fragment matches), but only the results from MS1 are shown here.
 MetaMorpheus calibrated files are verified to be compatible with a variety of tools and workflows, including TPP, Mascot, and MsFragger.
+Some people use extra software for identifying good search parameters: the calibration procedure in MetaMorpheus outputs those file-specific tolerances as a side product of the calibration method.
 
@@
  \includegraphics{fig1-calibErrors.png}
- \caption{Errors in m/z plotted vs the m/z value and the retention time}
+ \caption{Errors in m/z value of select peaks before (top panels) and after (bottom panels) calibration. These errors are plotted vs two variables: m/z value (left) and the retention time (right).}
  \label{fgr:fig1-calibErrors}
@@
 Table~\ref{tbl:calib} details the results of these searches.
-The calibration procedure successfully addresses instrument miscalibration by centering the mass errors around zero ppm.
+The calibration procedure successfully addresses instrument miscalibration by centering the mass errors around zero ppm (-2.08 ppm vs -0.10 ppm).
 The systematic noise introduced by various other factors has also been corrected, as evidenced by the decrease in the standard deviation of the errors in the precursor masses, which shows that the calibrated mass spectra predominantly have sub-ppm mass errors.
-A beneficial side effect of calibration is a corresponding increase in the number of confidently identified PSMs by around 1\%, when using the same wide 10 ppm precursor mass tolerance.
+A beneficial side effect of calibration is a corresponding increase of about 1\% in the number of confidently identified PSMs , when using the same 10 ppm precursor mass tolerance.
 Histograms of the mass errors in identified peptides before and after calibration are plotted in Figure~\ref{fgr:fig2-10ppmSearchCalib}.
@@
                                                                                       & Orig        & Calib        \\
- PSM Count                                                           &      217612   &   219172        \\
- Average Error (ppm)                                                              &      -2.10      &        -0.16      \\
- St. Dev of Error                                                    &       2.15      &        0.91      \\
+ PSM Count                                                           &      270548   &   273825        \\
+ Average Error (ppm)                                                              &      -2.08      &        -0.10      \\
+ St. Dev of Error                                                    &       2.24      &        1.06      \\
 \end{tabular}
@@
 It can also be seen as a narrowing of the open search to only allow prespecified mass differences.
+One example of multi-notch search is the \textit{limited multi-notch} search, with notches at masses corresponding to Dalton differences between the monoisotopic peak and other peaks in averagine isotopic envelopes (e.g.
+1.0029, 2.0054).
 The advantages and disadvantages of this alternative strategy are described below.
@@
 This may happen when the mass difference arises due to the identified peptide being different from the actual isolated precursor by a low mass modification or amino acid substitutions.
-
 The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4jurkat-1012}, generated by compiling the results of a 2.5 Da tolerance narrow-window search.
@@
 We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wider (few Da) precursor mass tolerance during a traditional narrow-window search.
-This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (1-20 ppm).
+This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (e.g.
+1-20 ppm).
 This approach provides a means of getting an increased number of correct identifications while mitigating problems with incorrect identifications.
@@
 We consider PSMs to be misassigned if the G-PTM-D procedure determines a higher-scoring modified peptide for a PSM, see the last section for details.
+One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat Frac9 scan 20962.
+With G-PTM-D added modifications, the correct peptide is identified with sequence \texttt{[Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K}.
+It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
+While this kind of misassignment is not possible in the 5ppm narrow-window search, often missed monoistopes are assigned to a wrong sequence.
+This explains the higher number of misassignments in the 5ppm search over the number in the multi-notch search.
+Finally, even the multi-notch search has some misassignments.
+One example, in Jurkat Frac6 scan 15602, is the wrong identification of \texttt{SIQEIQELDK[acetyl]DDESLR} with a score of 10.153, when the correct identification, after the G-PTM-D procedure is \texttt{S[acetyl]IQEIQELDKDDESLR} with score 16.249, where the masses are identical.
+While all of the described searches have some misassigned false positive identifications, the multi-notch search minimizes them while also yielding the most PSMs.
 
@@
 \begin{tabular}{llll}
-                    & 5ppm Search & 2.5 Da Search & Multi-Notch Search \\
+                    & 5ppm Search & 2.5 Da Search & Limited Multi-Notch Search \\
 PSMs Within 1\% FDR & 275359      & 262543        & 279895       \\
-misassigned PSMs    & 2041           & 9041          & 1371            \\
+Misassigned PSMs    & 2041           & 9041          & 1371            \\
 \end{tabular}
 \end{table}
-
-One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence QAFDDAIAELDTLNEDSYK, in Jurkat Frac9 scan 20962.
-With G-PTM-D added modifications, the correct peptide is identified with sequence [Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K. It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
-
-While this kind of misassgnment is not possible in the 5ppm narrow-window search, often missed monoistopes are assigned to a wrong sequence.
-This explains the higher number of misassignments in the 5ppm search over the number in the multi-notch search.
-
-Finally, even the multi-notch search has some misassignments.
-One example is the wrong identification of AEDGHEEAM[Oxidation]ETEASTSGEVAEVAEEAMETESSEK with a score of 20.112, when the correct identification, after the G-PTM-D procedure is AEDGHEEAMETEASTSGEV[Oxidation]AEVAEEAMETESSEK, where the masses are identical.
-
-%25956	0	25.126	20.112	AEDGHEEAMETEASTSGEV[Mod:Oxidation on V]AEVAEEAMETESSEK	AEDGHEEAM[Common Variable:Oxidation of M]ETEASTSGEVAEVAEEAMETESSEK	3614.45197	3614.45197
-%130059	0.000192	18.168	13.096	[PeptideTermMod:Ammonia loss on Q]QLLC[Common Fixed:Carbamidomethyl of C]DLVGISPSSEHDVLK	LQALQSQVEFLEQSMVDK	2092.05106	2092.05106	0
-%131801	0.000192	18.193	13.065	RN[Mod:Deamidation on N]DDISELEDLSELEDLK	LEEELSVEGNTDINTDLKD	2132.9961	2132.9961	0
-%156807	0.000278	20.257	12.033	DLGLSESGE[Metal:Sodium on E]DVNAAILDESGK	PVTVEPMDQLDDEEGLPEK	2139.97838	2139.98818	-0.0098
 
@@
 showed that by significantly relaxing the precursor mass tolerance in bottom-up searches, an incredible variety of unforeseen peptides were being fragmented and could be identified\citep{Chick_2015}.
-In particular, PTM-containing peptides could be identified from those MS2 fragments that did not possess the modification and the type of PTM could be inferred from differences in identified peptide mass and precursor mass.
+In particular, PTM-containing peptides could be identified from MS2 fragments that did not possess the modification, and the type of PTM could be inferred from differences in identified peptide mass and precursor mass.
 A major limitation of the Chick work was the extremely long search times because of the vast number of theoretical candidate peptides that were evaluated for each and every MS2 spectrum.
@@
 For example, small missed cleavages are often reported as mass differences.
-The peptide PEPKK is often identified as PEPK with a mass error of 128 Da (the converse is also true, PEPK may be identified with a mass error of -128 Da).
-This error is relatively easy to spot but when the missed cleavages contain multiple amino acids, the interpretation and discovery of the error is much more difficult.
+The peptide \texttt{PEPKK} is often identified as \texttt{PEPK} with a mass error of 128 Da (the converse is also true, \texttt{PEPK} may be identified with a mass error of -128 Da).
 Another example of this relates to peptide modifications.
-The phosphopeptide PEPpSK could be reported with a mass error of -79.96 Da when the better identification would have been PEPSK.
+The phosphopeptide \texttt{PEP[Phospho]SK} could be reported with a mass error of -79.96 Da when the better identification would have been \texttt{PEPSK}.
+These simple errors are relatively easy to spot, but not when there are missed cleavages with multiple amino acids and/or mulitple PTMs.
 
 The alternative we recommend, is the use of a multi-notch search, which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the fewest false positive PSM identifications.
-
-We compared searches of all Jurkat fractions using a $\pm 500$ Da wide mass search versus a multi-notch search with notches at 0 and 79.961 with 0.02 Dalton tolerance.
+We compared searches of all Jurkat fractions using a $\pm 500$ Da wide mass search to a multi-notch search with notches at 0 and 79.961 with 0.02 Dalton tolerance.
 The advantage of the multi-notch search over the open search is shown in Table~\ref{tbl:multiVsWide} by at least a $10\%$ increase in the number of target PSMs with score of at least 9.079, for both regions (0 Da no-PTM peptides and +79.96 Da phosphorylated and sulfonated peptides).
-It is important to point out that the difference in the numbers of PSMs is not due to FDR considerations, but rather to the erroneous peak assignment in open searches.
+It is important to point out that the difference in the numbers of PSMs is not due to FDR considerations, but rather to erroneous peak assignments in open searches, that misassign PSMs due to the extreme freedom in precursor masses allowed in open searches.
+A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat frac 5.
+The open search identifies \texttt{ATPARAPESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR} as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
+The better match is identified by the multi-notch search, as \texttt{APESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR} with a score of 16.246 and mass difference of 79.9655, and no missed cleavages.
+So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of \texttt{ATPAR} and addition of phosphorylation), we have an easily identifiable mass shift corresponding to a phosphorylation site.
 
@@
 \centering
-\caption{Multi-Notch Search vs. Open Search}
+\caption{Number of PSMs in Multi-Notch Search vs. Open Search for select mass differences. These PSMs are shown as mass difference histograms in Figure~\ref{fig:fig6-ZeroAndPhospho}}
 \label{tbl:multiVsWide}
@@
 
-The same results are presented in a graph format in Figure~\ref{fig:fig6-ZeroAndPhospho}.
-
 \begin{figure}[H]
 \includegraphics{fig6-ZeroAndPhospho.png}
-\caption{Zero and Phoshpo and Sulfo}
+\caption{Histogram of PSM counts for select mass differences around zero, sulfo, and phospho masses. The Multi-Notch searches produce visibly more PSMs.}
 \label{fig:fig6-ZeroAndPhospho}
 \end{figure}
-
-A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat frac 5.
-The open search identifies ATPARAPESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
-The better match is identified by the multi-notch search, as APESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR with a score of 16.246 and mass difference of 79.9655, and no missed cleavages.
-So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of ATPAR and addition of phosphorylation), we have an easily identifiable mass shift corresponding to a phosphorylation site.
 
@@
 
-Consider the Jurkat Frac12 scan 12336.
-Isolation m/z  = 726.65.
-Isolation width is 2.5 Th.
+Consider the Jurkat Frac 12 scan 12336 with precursor scan 12333.
+Isolation m/z is 726.65 Th and isolation width is 2.5 Th.
 It is shown in Figure~\ref{fig:fig5-coIsolationSpectrum}.
-There are at least four different peptides being co-isolated and co-fragmented, and there is evidence in the MS/MS spectra for all of them.
+There are at least three different peptides being co-isolated and co-fragmented, and there is evidence in the MS/MS spectrum for all of them.
 
@@
 \includegraphics{fig5-coIsolationSpectrum.png}
-\caption{m/z peaks in scan 12333}
+\caption{m/z peaks in MS1 scan 12333}
 \label{fig:fig5-coIsolationSpectrum}
@@
 
-In a conventional search, only the isolation precursor mass would be used to select a range of theoretical for comparison.
-However, all peptides in that range are co-fragmented and a better solution would be to use all co-isolated precursor masses.
+In a conventional search, only the isolation m/z value of 726.65 would be used to select a range of theoretical peptides for comparison.
+Assuming correct de-isotoping and charge state calculation, the mass 1447.74 Da would be searched.
+However, all peptides in the range[725.4, 727.9] are co-fragmented and a better solution would be to use all co-isolated precursor masses.
 In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor ($\pm 1.25$ Th).
@@
 The result of this is that often, more than one peptide can be identified from a single MS2 scan.
-This is extremely beneficial in both open mass and multi-notch searches because it reduces the number the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
-%Below, in Table 3434, we should the number of PSMs for a raw file where we use only a single precursor mass and we compare that to the case where we allow for multiple different precursors.
-%You can see a significant increase in the number of identified PSMs.
-%Importantly, you can see in Supplemental Figure erf3e, that the precursor mass errors go down significantly and the standard deviation is much reduced.
-
-%Get the story about how precurors with similar m/z with different z can in their deconvoluted state be many (10s, 100s of Da apart).
-%This is a big problem in general searches but also for open mass and multi-notch searches which are used to discover possible PTMs.
-
-%MetaMorpheus was programmed to ignore the precursor mass reported by the mass spectrometer in the mass spectrum and instead, deconvolutes all peaks in the MS1 spectrum associated with the tandem mass spectrum to recover all possible precursor masses.
-%This expanded list of precursors is used in every spectrum match, frequently resulting in multiple peptide identifications from single tandem mass spectrum.
+This is extremely beneficial in both open mass and multi-notch searches because it reduces the number of the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
 
@@
 A Perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
-This database was used for a final search and a number of PTMs were identified.
-There were several limitations to that work.
+This database was used for a final search and numerous PTMs were identified.
+While that work represented a substantial conceptual advance in identifying PTMs, there were several limitations to its utility.
 One, the open mass search was unacceptably slow.
@@
 Five, separate programs were required (both Morpheus and a Perl script).
+Six, the small list of 13 PTMs does not include many important modifications.
 MetaMorpheus resolves all of these issues and more.
@@
 The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass shifts in the open and the multi-notch searches.
-We do not show this difference here, since it is not informative due to the fact that the modifications added to the database are plausible ptms and not actually discovered ones.
-Nevertheless, one advantage of the enhanced approach is clear: the overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
+At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
 
@@
 
-After the final search using the augmented database is conducted (narrow-window in the original, and limited multi-notch in the enhanced G-PTM-D), the numbers of PSMs, peptides, and proteins are shown in Table~\ref{tab:table2}.
-The numbers all went up.
+After the final search using the augmented database is conducted (narrow-window in the original, and limited multi-notch in the enhanced G-PTM-D), the numbers of PSMs, peptides, and proteins all increase as shown in Table~\ref{tab:table2}.
 
@@
 \multirow{2}{*}{Unique Peptides} & Jurkat  & 85542   &  101980  \\
-                      & Mouse    & ?   &   ? \\
+                      & Mouse    & 56235   &   66365 \\
 \hline
 \multirow{2}{*}{Proteins} & Jurkat  & 9254   &  9587  \\
-                      & Mouse    & ?   &   ? \\
+                      & Mouse    & 7874   &   8151 \\
+\hline
+\multirow{2}{*}{PTM containing peptides} & Jurkat  & 12041   &  22173  \\
+                      & Mouse    & 14547   &   18487 \\
 \end{tabular}
@@
 
-These enhancements demonstrate that this approach can and should be used whenever bottom-up proteomics is done.
+These enhancements demonstrate that the enhanced G-PTM-D approach can and should be used whenever bottom-up proteomics is done.
 It is fast enough to be part of every regular work-flow.
-And, the significant contribution of signal from PTM peptides is clearly beneficial to biologists seeking to understand their data better.
+And, the significant contribution of correctly identified PTM peptides is clearly beneficial to biologists seeking to understand their data better.
 
-\subsubsection{Variations}
+%\begin{acknowledgement}
 
-The numbers reported above reflect the best way of running G-PTM-D.
-There are two natural questions that arise: 1) What if we use a FASTA file instead of a unimod XML?
-2) What if we don't include the modifications not interesting to biologists, for example sample handling artifacts?
-In this section we attempt to convince the reader that the way we do G-PTM-D is the best.
+%The authors thank \ldots
+%\end{acknowledgement}
 
-G-PTM-D originally added PTMs with masses that were included in the UniProt PTM list.
-We have since greatly increased the variety of PTMs added by including a number of new notches in the G-PTM-D first pass search: a curated set of notches and the corresponding modifications is given in Supplementary Table 1.
-Even though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMs.
-The overall identification rate decreases from 220,574 to 217,597 peptide-spectral matches (PSMs), which in turn decreased the number of identified modified peptides by 4\%.
+%\begin{suppinfo}
+
+%The following files are available free of charge.
+%\begin{itemize}
+%  \item Filename: brief description
+%  \item Filename: brief description
+%\end{itemize}
+
+%\end{suppinfo}
+
+\bibliography{citations}
+
+\end{document}
+ven though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMs.
+The overall identification rate decreases from 220,574 to 217,597 peptide-spectral matches (PSMs), which in turn decreased the number of identified modified peptides by 4\%, see Table~\ref{tab:table3};
 
@@
 \end{document}
+ent}
+ocument}

### 2017-11-21 FigureGeneration/fig1-calibErrors.py

@@ -2,8 +2,30 @@ import numpy as np
 import matplotlib.pyplot as plt
 import pandas
 
-nice = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\2017-10-12-17-10-32\Task1-CalibratingAllJurkat\120426_Jurkat_highLC_Frac16init.ms1dptsv", sep='\t')
-nice2 = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\2017-10-12-17-10-32\Task1-CalibratingAllJurkat\120426_Jurkat_highLC_Frac16round7inter-scan.ms1dptsv", sep='\t')
+nice = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\CalibrationTask\Task1-CalibratingAllJurkat\120426_Jurkat_highLC_Frac16init.ms1dptsv", sep='\t')
+
+print("mean")
+print(np.mean(nice['labelPPM']))
+print("st dev")
+print(np.std(nice['labelPPM']))
+print("95% min")
+print(np.mean(nice['labelPPM'])-2*np.std(nice['labelPPM']))
+print("95% max")
+print(np.mean(nice['labelPPM'])+2*np.std(nice['labelPPM']))
+
+
+nice2 = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\CalibrationTask\Task1-CalibratingAllJurkat\120426_Jurkat_highLC_Frac16round7inter-scan.ms1dptsv", sep='\t')
+
+print("mean")
+print(np.mean(nice2['labelPPM']))
+print("st dev")
+print(np.std(nice2['labelPPM']))
+print("95% min")
+print(np.mean(nice2['labelPPM'])-2*np.std(nice2['labelPPM']))
+print("95% max")
+print(np.mean(nice2['labelPPM'])+2*np.std(nice2['labelPPM']))
+
+
 
 height = 6
 width = 7
@@ -17,7 +39,7 @@ ax1.plot(nice['mz'],nice['LabelTh'], 'bo',markersize = .3)
 ax1.hlines(0, min(nice['mz']), max(nice['mz']), zorder = 10)
 ax1.set_ylim([minY,maxY])
 
-ax1.set_ylabel('m/z error')
+ax1.set_ylabel('m/z error (Th)')
[…]

### 2017-11-21 FigureGeneration/fig2-10ppmSearchCalib.py

@@ -17,6 +17,14 @@ ok['Mass Diff (ppm)'] = pandas.to_numeric(ok['Mass Diff (ppm)'], errors='coerce'
 A=ok[r"Mass Diff (ppm)"]
 B=ok[r"Decoy/Contaminant/Target"]=="T"
 C=ok[r"QValue Notch"]
+D=ok[r"Notch"]
+
+print("count")
+print(len(A[(C<0.01) & B]))
+print("mean")
+print(np.mean(A[(C<0.01) & B & (D==r'0')]))
+print("st dev")
+print(np.std(A[(C<0.01) & B & (D==r'0')]))
 
 ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
 ax1.set_ylim([1,maxi])
@@ -25,13 +33,24 @@ ax1.set_ylabel('PSM count')
 ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
                 size='large', ha='right', va='center', xytext=(0,0.5))
 
-ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\1mm10ppmSearch\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')  
+
+
+
+ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\1mm10ppm\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')  
 
 ok['Mass Diff (ppm)'] = pandas.to_numeric(ok['Mass Diff (ppm)'], errors='coerce')
 
 A=ok[r"Mass Diff (ppm)"]
 B=ok[r"Decoy/Contaminant/Target"]=="T"
 C=ok[r"QValue Notch"]
+D=ok[r"Notch"]
+
+print("count")
+print(len(A[(C<0.01) & B]))
+print("mean")
+print(np.mean(A[(C<0.01) & B & (D==r'0')]))
+print("st dev")
+print(np.std(A[(C<0.01) & B & (D==r'0')]))
 
 ax3.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
 ax3.set_ylim([1,maxi])

### 2017-11-21 FigureGeneration/fig3-searchTypes.py

@@ -3,11 +3,11 @@ import matplotlib.pyplot as plt
 
 
 
-height = 5
-width = 3.25 
+height = 2
+width = 7 
 plt.figure(figsize=(width, height))
 
-plt.subplot(311)
+plt.subplot(131)
 
 dates = [-100,-1,0,1,100]
 values = [1,1,2,1,1]
@@ -19,9 +19,10 @@ yTicks = [1,2]
 
 plt.ylim([0.5,2.5])
 plt.yticks(yTicks, ylabels)
+plt.xlabel("Mass Diff (Da)")
 plt.title("Narrow-Window Search")
 
-plt.subplot(312)
+plt.subplot(133)
 
 dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
 values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
@@ -35,8 +36,9 @@ plt.title("Multi-Notch Search")
 plt.ylim([0.5,2.5])
 
 plt.yticks(yTicks, ylabels)
+plt.xlabel("Mass Diff (Da)")
 
-plt.subplot(313)
+plt.subplot(132)
 
 
 
@@ -53,6 +55,11 @@ plt.ylim([0.5,2.5])
 plt.title("Open Search")
 
 plt.yticks(yTicks, ylabels)
+plt.xlabel("Mass Diff (Da)")
+
+
+
+
 plt.tight_layout()
 
 

### 2017-11-21 FigureGeneration/fig5-coIsolationSpectrum.py

@@ -19,18 +19,30 @@ print(len(mz))
 print(len(objects))
 print(len(y_pos))
 print(y_pos[0])
-plt.figure(figsize=(5,3.25))
+
+
+
+height = 3
+width = 7 
+plt.figure(figsize=(width, height))
+
 plt.bar(y_pos, mz)
 plt.xticks(np.arange(35,400,50), [725,725.5, 726,726.5,727,727.5,728,728.5])
 plt.xlim([0,400])
-plt.xlabel('m/z')
+plt.ylim([0,6000000])
+plt.xlabel('m/z (Th)')
 plt.ylabel('Intensity')
-plt.title('Scan 12333 in m/z range [724.64,728.65]')
- 
+plt.title('Scan 12333 in m/z range [724.65,728.64]')
+
+plt.plot([(724.87-724.65)*100,(724.87-724.65+1/2)*100,(724.87-724.65+2/2)*100, (724.87-724.65+3/2)*100], [2800000,1700000,1050000,350000], 'r*', label="Mass 1447.74 charge 2")
+plt.plot([(725.4-724.65)*100,(725.4-724.65+1/2)*100,(725.4-724.65+2/2)*100, (725.4-724.65+3/2)*100], [2200000,1800000,900000,500000], 'rs', label="Mass 1448.8 charge 2")
+plt.plot([(726.32-724.65)*100,(726.32-724.65+1/3)*100,(726.32-724.65+2/3)*100, (726.32-724.65+3/3)*100, (726.32-724.65+4/3)*100], [5500000,5800000,3700000,1800000,500000], 'ro', label="Mass 2175.96 charge 3")
+
+plt.legend()
 
 plt.tight_layout()
 
-plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig5-coIsolationSpectrum.png', format='png', dpi=600)
+plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig5-coIsolationSpectrum.png', format='png', dpi=1200)
 
 plt.show()
 

### 2017-11-21 FigureGeneration/fig6-ZeroAndPhoshpo.py

@@ -46,8 +46,8 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, ma
 
 plt.legend()
 
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 0.02$ Da")
+plt.ylabel("PSM Count")
+plt.title("Muti-Notch Search Mass Diff Histograms")
 
 ok.set_xlim([min1, max1])
 ok.set_ylim([0,75000])
@@ -72,8 +72,7 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, ma
 
 plt.legend()
 
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 0.02$ Da")
+plt.ylabel("PSM Count")
 
 ok.set_xlim([min1, max1])
 ok.set_ylim([0,290])
@@ -116,8 +115,8 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, ma
 
 plt.legend()
 
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 0.02$ Da")
+plt.ylabel("PSM Count")
+plt.title("Open Search Mass Diff Histograms")
 
 ok.set_xlim([min1, max1])
 ok.set_ylim([0,75000])
@@ -142,8 +141,7 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, ma
 
 plt.legend()
 
-plt.ylabel("Count")
-plt.title("PSMs within $\pm 0.02$ Da")
+plt.ylabel("PSM Count")
 
 ok.set_xlim([min1, max1])
 ok.set_ylim([0,290])

### 2017-11-21 FigureGeneration/fig7-phosphoSulfoUncalibCalib.py

@@ -44,6 +44,9 @@ print(len(massDiffsHighConf1decoy))
 plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
 plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, max1], label = "Decoys")
 
+plt.annotate('?', xy=(79.962, 22), xytext=(79.955,30), arrowprops=dict(arrowstyle="->"))
+plt.annotate('?', xy=(79.971, 20), xytext=(79.975,30), arrowprops=dict(arrowstyle="->"))
+
 plt.legend()
 
 plt.ylabel("Count")
@@ -85,6 +88,9 @@ print(len(massDiffsHighConf1decoy))
 
 plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
 plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, max1], label = "Decoys")
+
+plt.annotate('sulfo', xy=(79.956, 23), xytext=(79.951,30), arrowprops=dict(arrowstyle="->"))
+plt.annotate('phospho', xy=(79.967, 28), xytext=(79.971,30), arrowprops=dict(arrowstyle="->"))
 
 plt.legend()
 

### 2017-11-21 FigureGeneration/table2threeNarrow.py

@@ -22,7 +22,7 @@ dataFinal = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibr
 print('Starting loop...')
 
 
-with open('workfileC.tsv', 'w') as f:
+with open('workfile.tsv', 'w') as f:
   for index, row in data.iterrows():
     final = dataFinal.loc[(dataFinal['File Name'] == row['File Name']) & (dataFinal['Scan Number'] == row['Scan Number']) & (dataFinal['Precursor MZ'] == row['Precursor MZ'])]
     initScore = row['Score']

### 2017-11-27 paper.tex
author

@@
 \author{Stefan K. Solntsev}
-\email{solntsev@wisc.edu}
 \author{Michael R. Shortreed}
@@
 \author{Lloyd M. Smith}
+\email{smith@chem.wisc.edu}
 \affiliation[UwMadison]
@@
   {Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus}
-
-% Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus
 
@@
 \end{document}
-ven though adducts arising from experimental artifacts are not ultimately interesting to biologists, not including them as an option decreases the total number of identified proteins and PTMs.
-The overall identification rate decreases from 220,574 to 217,597 peptide-spectral matches (PSMs), which in turn decreased the number of identified modified peptides by 4\%, see Table~\ref{tab:table3};
-
-\begin{table}[]
-\centering
-\caption{Effects of Using Different Modification Lists}
-\label{tab:table3}
-\begin{tabular}{ll|l|l|l}
-                      &        & Only UniProt & UniProt+Glycosylations & UniProt+Glycosylations+Adducts\\
-\hline
-\multirow{2}{*}{PSMs} & Jurkat  & 217597   &  218547 & 220574\\
-                      & Mouse    & 170743   &   171432& 174783 
-\end{tabular}
-\end{table}
-
-\begin{acknowledgement}
-
-The authors thank \ldots
-\end{acknowledgement}
-
-\begin{suppinfo}
-
-The following files are available free of charge.
-\begin{itemize}
-  \item Filename: brief description
-  \item Filename: brief description
-\end{itemize}
-
-\end{suppinfo}
-
-\bibliography{citations}
-
-\end{document}
-ent}
-ocument}

### 2017-11-27 paper.tex
lots

@@
 phosphorylations) in well-studied systems, but global discovery tools, such as MODa\citep{Na_2011} and G-PTM-D\citep{Li_2016}, for identification and localization of a wide variety of different PTMs are also available.
-The G-PTM-D workflow consists of three stages: 1) A wide precursor mass tolerance database search ('open' search)\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
+G-PTM-D was developed in our laboratory and consists of three stages: 1) A wide precursor mass tolerance database search ('open' search)\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
@@
 MetaMorpheus calibrates both MS1 and MS2 scans (using MS2 fragment matches), but only the results from MS1 are shown here.
-MetaMorpheus calibrated files are verified to be compatible with a variety of tools and workflows, including TPP, Mascot, and MsFragger.
-Some people use extra software for identifying good search parameters: the calibration procedure in MetaMorpheus outputs those file-specific tolerances as a side product of the calibration method.
+MetaMorpheus calibrated files are verified to be compatible with a variety of tools and workflows, including TPP\citep{Deutsch2010}, and MSFragger\citep{Kong_2017}.
+In addition to calibrated spectra files, the calibration procedure outputs the standard deviations and averages of the errors post-calibration, which could then be used as search parameters; this is an alternative to using dedicated tools such as Param-Medic\citep{May2017} for determining the best search parameters for a given spectra file.
 
@@
 It can also be seen as a narrowing of the open search to only allow prespecified mass differences.
-One example of multi-notch search is the \textit{limited multi-notch} search, with notches at masses corresponding to Dalton differences between the monoisotopic peak and other peaks in averagine isotopic envelopes (e.g.
+One example of multi-notch search is the \textit{limited multi-notch} search, with notches at masses corresponding to Dalton differences between the monoisotopic peak and other peaks in averagine\citep{Senko1995} isotopic envelopes (e.g.
 1.0029, 2.0054).
@@
 \includegraphics{fig3-searchTypes.png}
-\caption{Search Modes}
+\caption{The three searches illustrated. Blue regions signify the allowed mass differences, while the pink ones show the mass differences that are not considered.}
 \label{fig:fig3-searchTypes}
@@
 \includegraphics{fig4jurkat-1012.png}
-\caption{Histograms PSM mass differences found in various dalton ranges}
+\caption{Histograms of PSM mass differences found in various dalton ranges.
+MM, Missed Monoisotopic peak; Asp, Aspartic Acid; Asn, Asparagine; Glu, Glutamic Acid; Gln, Glutamine; Xle, Leucine/Isoleucine; Val, Valine; Thr, Threonine; Pro, Proline.}
 \label{fig:fig4jurkat-1012}
@@
 
-We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wider (few Da) precursor mass tolerance during a traditional narrow-window search.
+We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wide (few Da) precursor mass tolerance during a traditional narrow-window search.
 This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (e.g.
@@
 To illustrate the described advantages, we compared the three types of searches using all 28 calibrated Jurkat fractions (Table~\ref{tbl:singleVsMultiNotch}).
-The limited notch search yields the largest number of confidently identified PSMs with only a small fraction being misassinged.
-We consider PSMs to be misassigned if the G-PTM-D procedure determines a higher-scoring modified peptide for a PSM, see the last section for details.
+The limited notch search yields the largest number of confidently identified PSMs with only a small fraction being misassigned.
+We consider PSMs to be misassigned if the G-PTM-D procedure determines a higher-scoring modified peptide for a PSM (see the last section for details).
 One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat Frac9 scan 20962.
 With G-PTM-D added modifications, the correct peptide is identified with sequence \texttt{[Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K}.
-It has a mass difference of -0.00171(-0.79458 ppm) and a score 19.222.
-While this kind of misassignment is not possible in the 5ppm narrow-window search, often missed monoistopes are assigned to a wrong sequence.
+It has a mass difference of -0.00171(-0.79458 ppm) and a score of 19.222.
+While this kind of misassignment is not possible in the 5ppm narrow-window search, such missed monoisotopes are often assigned to a wrong sequence.
 This explains the higher number of misassignments in the 5ppm search over the number in the multi-notch search.
 Finally, even the multi-notch search has some misassignments.
-One example, in Jurkat Frac6 scan 15602, is the wrong identification of \texttt{SIQEIQELDK[acetyl]DDESLR} with a score of 10.153, when the correct identification, after the G-PTM-D procedure is \texttt{S[acetyl]IQEIQELDKDDESLR} with score 16.249, where the masses are identical.
+One example, in Jurkat Frac6 scan 15602, is the wrong identification of \texttt{SIQEIQELDK[acetyl]DDESLR} with a score of 10.153, when the correct identification, after the G-PTM-D procedure, is \texttt{S[acetyl]IQEIQELDKDDESLR} with score 16.249, where the masses are identical.
 While all of the described searches have some misassigned false positive identifications, the multi-notch search minimizes them while also yielding the most PSMs.
@@
 The phosphopeptide \texttt{PEP[Phospho]SK} could be reported with a mass error of -79.96 Da when the better identification would have been \texttt{PEPSK}.
-These simple errors are relatively easy to spot, but not when there are missed cleavages with multiple amino acids and/or mulitple PTMs.
+These simple errors are relatively easy to spot, but not when there are missed cleavages with multiple amino acids and/or multiple PTMs.
 
@@
 \centering
-\caption{Number of PSMs in Multi-Notch Search vs. Open Search for select mass differences. These PSMs are shown as mass difference histograms in Figure~\ref{fig:fig6-ZeroAndPhospho}}
+\caption{Number of PSMs in Multi-Notch Search vs. Open Search for select mass differences.}
 \label{tbl:multiVsWide}
 \begin{tabular}{lll}
-                    & Open & Multi-Notch\\
+               & Open                     & Multi-Notch         \\
 0 peak         & 176073  (0.15\% FDR)     & 193328 (0.23\% FDR) \\
@@
 \end{table}
-
-\begin{figure}[H]
-\includegraphics{fig6-ZeroAndPhospho.png}
-\caption{Histogram of PSM counts for select mass differences around zero, sulfo, and phospho masses. The Multi-Notch searches produce visibly more PSMs.}
-\label{fig:fig6-ZeroAndPhospho}
-\end{figure}
 
@@
 Assuming correct de-isotoping and charge state calculation, the mass 1447.74 Da would be searched.
-However, all peptides in the range[725.4, 727.9] are co-fragmented and a better solution would be to use all co-isolated precursor masses.
+However, all peptides in the range 725.4 to 727.9 Th are co-fragmented and a better solution would be to use all co-isolated precursor masses.
 In MetaMorpheus, for each MS2 scan, the corresponding MS1 scan is deconvoluted and the complete list of precursors is gathered for the range surrounding the isolated precursor ($\pm 1.25$ Th).
@@
 
-In the original G-PTM-D manuscript, an open mass search ($\pm 200$ Da) was used to create a histogram of mass difference peaks, from which PTMs could be inferred.
+In the original G-PTM-D manuscript\citep{Li_2016}, an open mass search ($\pm 200$ Da) was used to create a histogram of mass difference peaks, from which PTMs could be inferred.
 A Perl script was used to mine those mass differences to modify a proteomics database with potential PTMs.
@@
 Five, separate programs were required (both Morpheus and a Perl script).
-Six, the small list of 13 PTMs does not include many important modifications.
+Six, the small list of 13 PTMs did not include many important modifications.
 MetaMorpheus resolves all of these issues and more.
@@
 Calibration and the associated increase in mass accuracy is essential for discrimination of small differences between mass shifts due to different modifications.
-To illustrate this, mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:fig7-ZeroAndPhospho}.
+To illustrate this, mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:fig7-PhosphoAndSulfo}.
 It is evident from these histograms that the phosphorylation and sulfonation peaks are clearly separated and identifiable when the search is conducted on well-calibrated spectra.
@@
 \begin{figure}[H]
-\includegraphics{fig7-ZeroAndPhospho.png}
+\includegraphics{fig7-SulfoAndPhospho.png}
 \caption{Peaks around phospho and sulfo in uncalibrated and calibrated data}
-\label{fig:fig7-ZeroAndPhospho}
+\label{fig:fig7-PhosphoAndSulfo}
 \end{figure}

### 2017-11-27 FigureGeneration/fig3-searchTypes.py

@@ -1,69 +1,137 @@
 import numpy as np
 import matplotlib.pyplot as plt
-
+import matplotlib.patches as patches
 
 
 height = 2
 width = 7 
-plt.figure(figsize=(width, height))
+fig = plt.figure(figsize=(width, height))
 
-plt.subplot(131)
-
-dates = [-100,-1,0,1,100]
-values = [1,1,2,1,1]
-
-plt.plot(dates, values, '-')
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.ylim([0.5,2.5])
-plt.yticks(yTicks, ylabels)
-plt.xlabel("Mass Diff (Da)")
-plt.title("Narrow-Window Search")
-
-plt.subplot(133)
-
-dates =  [-100,-18,-17,-16,-1,0,1,41,42,43,79,80,81,100]
-values = [1   ,  1,  2,  1 ,1,2,1, 1, 2, 1,1 ,2 ,1 , 1]
-
-plt.plot(dates, values, '-')
-
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
-
-plt.title("Multi-Notch Search")
-plt.ylim([0.5,2.5])
-
-plt.yticks(yTicks, ylabels)
-plt.xlabel("Mass Diff (Da)")
-
-plt.subplot(132)
+#######
 
 
+ax = fig.add_subplot(131)
+for p in [
+    patches.Rectangle(
+        (-105, 0), 101, 1,
+        alpha=0.4,
+        facecolor="red"
+    ),
+    patches.Rectangle(
+        (-4, 0), 8, 1,
+    ),
+    patches.Rectangle(
+        (4, 0), 101, 1,
+        alpha=0.4,
+        facecolor="red"
+    )
+]:
+    ax.add_patch(p)
 
-dates =  [-100,100]
-values = [2 ,2]
 
-plt.plot(dates, values, '-')
+ax.set_xlim([-105,105])
+ax.set_ylim([-2, 9])
 
-ylabels = ['Reject', 'Accept']
-yTicks = [1,2]
+ax.set_xlabel('Mass Diff (Da)')
 
-plt.ylim([0.5,2.5])
-
-plt.title("Open Search")
-
-plt.yticks(yTicks, ylabels)
-plt.xlabel("Mass Diff (Da)")
+ax.set_xlabel('Mass Diff (Da)')
 
 
[…]

### 2017-11-27 FigureGeneration/fig4-jurkat;-1;0;1;2.py

@@ -129,7 +129,7 @@ plt.ylabel("Count")
 
 plt.title("PSMs around 2 Da")
 
-plt.annotate('Xle->D', xy=(1.942879, 25), xytext=(1.942879-0.04, 35), arrowprops=dict(arrowstyle="->"))
+plt.annotate('Xle->Asp', xy=(1.942879, 25), xytext=(1.942879-0.04, 35), arrowprops=dict(arrowstyle="->"))
 #plt.annotate('Thr->Cys', xy=(1.961506, 18), xytext=(1.961506-0.07, 35), arrowprops=dict(arrowstyle="->"))
 plt.annotate('Double\nDeamidation', xy=(1.968032, 40), xytext=(1.968032-0.067, 55), arrowprops=dict(arrowstyle="->"))
 plt.annotate('Val->Thr', xy=(1.979265, 29), xytext=(1.979265-0.01, 65), arrowprops=dict(arrowstyle="->"))

### 2017-11-27 FigureGeneration/fig5-coIsolationSpectrum.py

@@ -15,16 +15,13 @@ mz = [0, 0, 0, 0, 0, 0, 162124.125, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 0, 98809.40625
 objects = list(frange(724.65,728.64,0.01))
 y_pos = np.arange(len(objects))
 
-print(len(mz))
-print(len(objects))
-print(len(y_pos))
-print(y_pos[0])
-
-
-
-height = 3
+height = 6
 width = 7 
 plt.figure(figsize=(width, height))
+
+##############  MS1
+
+ok = plt.subplot(211)
 
 plt.bar(y_pos, mz)
 plt.xticks(np.arange(35,400,50), [725,725.5, 726,726.5,727,727.5,728,728.5])
@@ -34,11 +31,55 @@ plt.xlabel('m/z (Th)')
 plt.ylabel('Intensity')
 plt.title('Scan 12333 in m/z range [724.65,728.64]')
 
+plt.plot([(726.32-724.65)*100,(726.32-724.65+1/3)*100,(726.32-724.65+2/3)*100, (726.32-724.65+3/3)*100, (726.32-724.65+4/3)*100], [5500000,5800000,3700000,1800000,500000], 'ro', label="Mass 2175.96 charge 3")
 plt.plot([(724.87-724.65)*100,(724.87-724.65+1/2)*100,(724.87-724.65+2/2)*100, (724.87-724.65+3/2)*100], [2800000,1700000,1050000,350000], 'r*', label="Mass 1447.74 charge 2")
 plt.plot([(725.4-724.65)*100,(725.4-724.65+1/2)*100,(725.4-724.65+2/2)*100, (725.4-724.65+3/2)*100], [2200000,1800000,900000,500000], 'rs', label="Mass 1448.8 charge 2")
-plt.plot([(726.32-724.65)*100,(726.32-724.65+1/3)*100,(726.32-724.65+2/3)*100, (726.32-724.65+3/3)*100, (726.32-724.65+4/3)*100], [5500000,5800000,3700000,1800000,500000], 'ro', label="Mass 2175.96 charge 3")
 
 plt.legend()
+
+##############  MS2
+ok = plt.subplot(212)
+
[…]

### 2017-11-27 FigureGeneration/fig6-ZeroAndPhoshpo.py

@@ -1,156 +1,0 @@
-import numpy as np
-import matplotlib.pyplot as plt
-import pandas
-
-height = 6
-width = 7
-plt.figure(figsize=(width, height))
-
-#################
-
-# data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2017-10-20-12-48-57-zeroAndMidPhosphoSulfo\Task1-SearchTask\aggregatePSMs_custom.psmtsv", sep='\t')
-data = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\2017-10-24-12-02-12-noDeisotopeZeroSulfoPhospho\Task1-SearchTask\aggregatePSMs_zeroAndMidSulfoPhospho.psmtsv", sep='\t')
-
-data['Mass Diff (Da)'] = pandas.to_numeric(data['Mass Diff (Da)'], errors='coerce')
-data_sorted = data.sort_values(['Mass Diff (Da)'], ascending=True)
-
-massDiffs=data_sorted['Mass Diff (Da)'].values
-scores=data_sorted['Score'].values
-notches=data_sorted['Notch'].values
-B=data_sorted['Decoy/Contaminant/Target'].values
-C=data_sorted['QValue Notch'].values
-
-scoreCutoff = min(scores[(C<0.01) &(notches=="1")])
-
-print(scoreCutoff)
-
-massDiffsHighConf = massDiffs[scores>scoreCutoff]
-TDhighConf = B[scores>scoreCutoff]
-
-#################
-
-ok = plt.subplot(222)
-
-halfWidth = 0.02 
-min1 = -halfWidth
-max1 = halfWidth
-
-massDiffsHighConf1 = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1)]
-massDiffsHighConf1decoy = massDiffsHighConf[(massDiffsHighConf>min1) & (massDiffsHighConf<max1) & (TDhighConf=="D")]
-
-print(len(massDiffsHighConf1))
-print(len(massDiffsHighConf1decoy))
-
-plt.hist(massDiffsHighConf1, bins  = int((max1-min1)/0.001), range=[min1, max1], label = "Targets")
-plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, max1], label = "Decoys")
-
-plt.legend()
-
[…]

### 2017-11-27 FigureGeneration/fig7-phosphoSulfoUncalibCalib.py

@@ -103,7 +103,7 @@ ok.set_ylim([0,60])
 plt.tight_layout()
 
 
-plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig7-ZeroAndPhospho.png', format='png', dpi=600)
+plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig7-SulfoAndPhospho.png', format='png', dpi=600)
 
 
 plt.show()

### 2017-11-28 paper.tex
alright

@@
 \includegraphics{fig5-coIsolationSpectrum.png}
-\caption{m/z peaks in MS1 scan 12333}
+\caption{m/z peaks in MS1 scan 12333.
+The top plot shows the isotopic envelopes for three distinct peptides in a narrow isolation window.
+The bottom plot shows some identified fragment peaks in the MS/MS scan.}
 \label{fig:fig5-coIsolationSpectrum}
@@
 The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass shifts in the open and the multi-notch searches.
-At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 35 hours to 4 hours for all datasets, see Table~\ref{my-labelff}.
+At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 35 hours to 4 hours for all datasets (see Table~\ref{my-labelff}).
 
@@
                       &        & Open Search & Multi-Notch Search\\
-\hline
 \multirow{2}{*}{Time} & Jurkat & 22.1 hrs         & 2.9 hrs    \\
@@
 After the final search using the augmented database is conducted (narrow-window in the original, and limited multi-notch in the enhanced G-PTM-D), the numbers of PSMs, peptides, and proteins all increase as shown in Table~\ref{tab:table2}.
+We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da.
+These enhancements demonstrate that the enhanced G-PTM-D approach provides great value to bottom-up proteomic data analysis.
+It is fast enough included as part of every regular work-flow, and will yield significantly increased numbers of correctly identified PTM peptides.
 
@@
 \label{tab:table2}
-\begin{tabular}{ll|l|l}
+\begin{tabular}{llll}
                       &        & Original G-PTM-D & Enhanced G-PTM-D\\
-\hline
 \multirow{2}{*}{PSMs} & Jurkat  & 210566   &  220574  \\
                       & Mouse    & 162473   &   170743 \\
-\hline
 \multirow{2}{*}{Unique Peptides} & Jurkat  & 85542   &  101980  \\
                       & Mouse    & 56235   &   66365 \\
-\hline
 \multirow{2}{*}{Proteins} & Jurkat  & 9254   &  9587  \\
                       & Mouse    & 7874   &   8151 \\
-\hline
 \multirow{2}{*}{PTM containing peptides} & Jurkat  & 12041   &  22173  \\
@@
 \end{table}
-
-We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da.
-
-These enhancements demonstrate that the enhanced G-PTM-D approach can and should be used whenever bottom-up proteomics is done.
-It is fast enough to be part of every regular work-flow.
-And, the significant contribution of correctly identified PTM peptides is clearly beneficial to biologists seeking to understand their data better.
 

### 2017-11-28 FigureGeneration/fig3-searchTypes.py

@@ -13,7 +13,7 @@ fig = plt.figure(figsize=(width, height))
 ax = fig.add_subplot(131)
 for p in [
     patches.Rectangle(
-        (-105, 0), 101, 1,
+        (-110, 0), 106, 1,
         alpha=0.4,
         facecolor="red"
     ),
@@ -21,7 +21,7 @@ for p in [
         (-4, 0), 8, 1,
     ),
     patches.Rectangle(
-        (4, 0), 101, 1,
+        (4, 0), 106, 1,
         alpha=0.4,
         facecolor="red"
     )
@@ -29,15 +29,10 @@ for p in [
     ax.add_patch(p)
 
 
-ax.set_xlim([-105,105])
+ax.set_xlim([-110,110])
 ax.set_ylim([-2, 9])
 
-ax.set_xlabel('Mass Diff (Da)')
-
-ax.set_xlabel('Mass Diff (Da)')
-
-
-plt.annotate('Only mass differences\nclose to zero allowed', xy=(0,1), xytext=(-80,4), arrowprops=dict(arrowstyle="->"))
+plt.annotate('Only mass differences\nclose to zero allowed', xy=(0,1), xytext=(-106,4), arrowprops=dict(arrowstyle="->"))
 plt.title('Narrow-Window Search')
 
 
@@ -48,7 +43,7 @@ ax.get_yaxis().set_visible(False)
 ax = fig.add_subplot(132)
 for p in [
     patches.Rectangle(
-        (-105, 0), 5, 1,
+        (-110, 0), 10, 1,
         alpha=0.4,
         facecolor="red"
     ),
@@ -56,7 +51,7 @@ for p in [
         (-100, 0), 200, 1,
     ),
     patches.Rectangle(
-        (100, 0), 5, 1,
+        (100, 0), 10, 1,
         alpha=0.4,
         facecolor="red"
     )
@@ -64,14 +59,11 @@ for p in [
     ax.add_patch(p)
 
 
-ax.set_xlim([-105,105])
+ax.set_xlim([-110,110])
 ax.set_ylim([-2, 9])
 
-ax.set_xlabel('Mass Diff (Da)')
 
-ax.set_xlabel('Mass Diff (Da)')
-
-plt.annotate('Wide range of mass\ndifferences allowed', xy=(0,1), xytext=(-80,4), arrowprops=dict(arrowstyle="->"))
[…]

### 2017-11-28 FigureGeneration/fig5-coIsolationSpectrum.py

@@ -31,9 +31,9 @@ plt.xlabel('m/z (Th)')
 plt.ylabel('Intensity')
 plt.title('Scan 12333 in m/z range [724.65,728.64]')
 
-plt.plot([(726.32-724.65)*100,(726.32-724.65+1/3)*100,(726.32-724.65+2/3)*100, (726.32-724.65+3/3)*100, (726.32-724.65+4/3)*100], [5500000,5800000,3700000,1800000,500000], 'ro', label="Mass 2175.96 charge 3")
+plt.plot([(726.32-724.65)*100,(726.32-724.65+1/3)*100,(726.32-724.65+2/3)*100, (726.32-724.65+3/3)*100, (726.32-724.65+4/3)*100], [5500000,5800000,3700000,1800000,500000], 'bo', label="Mass 2175.96 charge 3")
 plt.plot([(724.87-724.65)*100,(724.87-724.65+1/2)*100,(724.87-724.65+2/2)*100, (724.87-724.65+3/2)*100], [2800000,1700000,1050000,350000], 'r*', label="Mass 1447.74 charge 2")
-plt.plot([(725.4-724.65)*100,(725.4-724.65+1/2)*100,(725.4-724.65+2/2)*100, (725.4-724.65+3/2)*100], [2200000,1800000,900000,500000], 'rs', label="Mass 1448.8 charge 2")
+plt.plot([(725.4-724.65)*100,(725.4-724.65+1/2)*100,(725.4-724.65+2/2)*100, (725.4-724.65+3/2)*100], [2200000,1800000,900000,500000], 'cs', label="Mass 1448.8 charge 2")
 
 plt.legend()
 
@@ -60,7 +60,7 @@ masses1 = [300.12224,401.16992,616.26052,703.29255,875.34096,1004.38355,1135.424
 mzs1 = [x+ 1 for x in masses1]
 converted1= [(x-100)/4 for x in mzs1]
 
-plt.plot(converted1, [45000,30000,15000,25000,30000,45000,25000,30000,35000,20000,30000,35000,105000,65000,55000,40000,15000], 'ro', label="YHTSQSGDEMTSLSEYVSR")
+plt.plot(converted1, [45000,30000,15000,25000,30000,45000,25000,30000,35000,20000,30000,35000,105000,65000,55000,40000,15000], 'bo', label="YHTSQSGDEMTSLSEYVSR")
 
[…]

### 2017-11-28 paper.tex
ajskf

@@
 G-PTM-D is a recently developed technique for global identification and localization of PTMs.
-Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy and speed of PTM identification.
+Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy, speed and scope of PTM identification.
 We enhance G-PTM-D by using multi-notch searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
@@
 phosphorylations) in well-studied systems, but global discovery tools, such as MODa\citep{Na_2011} and G-PTM-D\citep{Li_2016}, for identification and localization of a wide variety of different PTMs are also available.
-G-PTM-D was developed in our laboratory and consists of three stages: 1) A wide precursor mass tolerance database search ('open' search)\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses (hypothesized to differ in mass due to the presence of a PTM);
+G-PTM-D was developed in our laboratory and consists of three stages: 1) A wide precursor mass tolerance database search ('open' search)\citep{Chick_2015, Na_2011}, which provides mass differences between the identified peptides and the observed precursor peptide masses, hypothesized to differ in mass due to the presence of a PTM;
 2) For those peptides for which the mass difference corresponds to the mass of a known PTM, a database augmentation step adds plausible localized PTMs to the corresponding protein entries in the search database;
@@
 Discovery of modifications is limited to PTMs in the UniProt database: this is a significant limitation since that database does not include modifications arising from sample handling, and ignoring those modifications significantly degrades the final PTM identification rates.
-Finally, PTMs that are similar in mass, such as phosphorylation (79.966331 Da) and sulfonation (79.956815 Da), are problematic since they are virtually indistinguishable in unprocessed spectral files.
+Finally, PTMs that are similar in mass, such as phosphorylation (79.9663 Da) and sulfonation (79.9568 Da), are problematic since they are virtually indistinguishable in unprocessed spectral files.
 These limitations motivated the work described here.
@@
 The database augmentation step is similar to the original G-PTM-D procedure, but the modification list is significantly expanded to accommodate modifications such as chemical adducts, glycosylations, lipids, and other important modifications.
-The final narrow-window search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and + 2.0054), and the capability for identifying co-isolated peptides, further increasing the number of confidently identified peptides and PTMs.
+The final narrow-window search is also replaced by a limited multi-notch search that accommodates errors in the precursor mass deisotoping (+1.0029 and +2.0054), and the capability for identifying co-isolated peptides, further increasing the number of confidently identified peptides and PTMs.
 
@@
 We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
-MetaMorpheus is freely available at \url{https://github.com/smith-chem-wisc/MetaMorpheus} as standalone Windows and Linux programs.
-MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.x
+MetaMorpheus is freely available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}.
+MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
 mzLib can be found at \url{https://github.com/smith-chem-wisc/mzLib}.
@@
 a replacement of the open search with a custom multi-notch search for PTM discovery;
-replacement of the final traditional narrow mass search with a limited multi-notch search that permits exact matches even when incorrectly deisotoped precursor masses are present;
+replacement of the final traditional narrow-window search with a limited multi-notch search that permits exact matches even when incorrectly deisotoped precursor masses are present;
 and lastly, addition of capacity for identification of multiple co-fragmented peptides from a single MS/MS scan.
@@
 
-We demonstrate the effectiveness of the calibration approach using a single Jurkat fraction (16) as an example.
+We demonstrate the effectiveness of the calibration approach using Jurkat fraction 16 as an example.
 Figure~\ref{fgr:fig1-calibErrors} displays the errors in m/z values of MS1 peaks identified by Step 1 of the calibration, both before and after the calibration procedure is applied.
@@
 The described issues are apparent in mass difference histograms in Figure~\ref{fig:fig4jurkat-1012}, generated by compiling the results of a 2.5 Da tolerance narrow-window search.
-These plots clearly show that numerous peptides have missed monoisotopic (Mm) peaks, deamidation, or an amino acid substitution.
+These plots clearly show that numerous peptides have missed monoisotopic peaks, deamidation, or an amino acid substitution.
 
@@
 We consider PSMs to be misassigned if the G-PTM-D procedure determines a higher-scoring modified peptide for a PSM (see the last section for details).
-One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat Frac9 scan 20962.
+One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat fraction 9 scan 20962.
 With G-PTM-D added modifications, the correct peptide is identified with sequence \texttt{[Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K}.
@@
 Finally, even the multi-notch search has some misassignments.
-One example, in Jurkat Frac6 scan 15602, is the wrong identification of \texttt{SIQEIQELDK[acetyl]DDESLR} with a score of 10.153, when the correct identification, after the G-PTM-D procedure, is \texttt{S[acetyl]IQEIQELDKDDESLR} with score 16.249, where the masses are identical.
+One example, in Jurkat fraction 6 scan 15602, is the wrong identification of \texttt{SIQEIQELDK[acetyl]DDESLR} with a score of 10.153, when the correct identification, after the G-PTM-D procedure, is \texttt{S[acetyl]IQEIQELDKDDESLR} with score 16.249, where the masses are identical.
 While all of the described searches have some misassigned false positive identifications, the multi-notch search minimizes them while also yielding the most PSMs.
@@
 It is important to point out that the difference in the numbers of PSMs is not due to FDR considerations, but rather to erroneous peak assignments in open searches, that misassign PSMs due to the extreme freedom in precursor masses allowed in open searches.
-A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat frac 5.
+A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat fraction 5.
 The open search identifies \texttt{ATPARAPESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR} as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
@@
 
-Consider the Jurkat Frac 12 scan 12336 with precursor scan 12333.
+Consider the Jurkat fraction 12 scan 12336 with precursor scan 12333.
 Isolation m/z is 726.65 Th and isolation width is 2.5 Th.
@@
 \label{my-labelff}
-\begin{tabular}{ll|l|l}
+\begin{tabular}{llll}
                       &        & Open Search & Multi-Notch Search\\
@@
 After the final search using the augmented database is conducted (narrow-window in the original, and limited multi-notch in the enhanced G-PTM-D), the numbers of PSMs, peptides, and proteins all increase as shown in Table~\ref{tab:table2}.
-We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da.
+We identified hundreds of glycosylated peptides in these unenriched samples, with many of these modifications exceeding 1000 Da, see Table~\ref{tab:table3}.
 These enhancements demonstrate that the enhanced G-PTM-D approach provides great value to bottom-up proteomic data analysis.
@@
                       & Mouse    & 14547   &   18487 \\
+\end{tabular}
+\end{table}
+
+\begin{table}[]
+\centering
+\caption{Highest mass PTMs identified}
+\label{tab:table3}
+\begin{tabular}{ll}
+Mass (Daltons)&	Mod\\
+1378.474&	Hex(6)HexNAc(2)\\
+1216.4236&	Hex(5)HexNAc(2)\\
+1054.3719&	Hex(4)HexNAc(2)\\
+1038.375&	dHex(1)Hex(3)HexNAc(2)\\
+963.3177&	Hex(1)HexNAc(1)NeuAc(1)NeuGc(1)
 \end{tabular}

### 2017-11-28 paper.tex
jaskfdd

@@
 Frequently, the best-fragmented peptide is different from the one corresponding to the precursor mass reported by the instrument, and the relaxed tolerance approach may identify this peptide, provided the reported charge state matches.
-The major limitation of this approach is that a 2 or 3 Da tolerance includes mass shifts that, when present, may indicate that the matched peptide is not the correct identification, arguably defeating the purpose of a narrow-window search.
+The major limitation of this approach is that a 2 or 3 Da tolerance includes mass differences that, when present, may indicate that the matched peptide is not the correct identification, arguably defeating the purpose of a narrow-window search.
 This may happen when the mass difference arises due to the identified peptide being different from the actual isolated precursor by a low mass modification or amino acid substitutions.
@@
 We consider PSMs to be misassigned if the G-PTM-D procedure determines a higher-scoring modified peptide for a PSM (see the last section for details).
-One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.03335 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat fraction 9 scan 20962.
+One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.0334 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat fraction 9 scan 20962.
 With G-PTM-D added modifications, the correct peptide is identified with sequence \texttt{[Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K}.
-It has a mass difference of -0.00171(-0.79458 ppm) and a score of 19.222.
+It has a mass difference of -0.0017(-0.7946 ppm) and a score of 19.222.
 While this kind of misassignment is not possible in the 5ppm narrow-window search, such missed monoisotopes are often assigned to a wrong sequence.
@@
 A specific example of a PSM that has been erroneously assigned in a open search is scan 12835 in Jurkat fraction 5.
-The open search identifies \texttt{ATPARAPESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR} as the best match with a score of 17.231 and mass difference of -416.31028, and a single missed cleavage.
+The open search identifies \texttt{ATPARAPESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR} as the best match with a score of 17.231 and mass difference of -416.3103, and a single missed cleavage.
 The better match is identified by the multi-notch search, as \texttt{APESPPSADPALVAGPAEEAEC[Carbamidomethyl]PPPR} with a score of 16.246 and mass difference of 79.9655, and no missed cleavages.
-So, instead of having a PSM with an unidentifiable mass shift of -416.31028 (mass of removal of \texttt{ATPAR} and addition of phosphorylation), we have an easily identifiable mass shift corresponding to a phosphorylation site.
+So, instead of having a PSM with an unidentifiable mass difference of -416.3103 (mass of removal of \texttt{ATPAR} and addition of phosphorylation), we have an easily identifiable mass difference corresponding to a phosphorylation site.
 
@@
 The result of this is that often, more than one peptide can be identified from a single MS2 scan.
-This is extremely beneficial in both open mass and multi-notch searches because it reduces the number of the more difficult to interpret mass shifts and substantially improves G-PTM-D (see below).
+This is extremely beneficial in both open mass and multi-notch searches because it reduces the number of the more difficult to interpret mass differences and substantially improves G-PTM-D (see below).
 
@@
 A major improvement in G-PTM-D performance arises from using calibrated spectra as inputs to the procedure.
-Calibration and the associated increase in mass accuracy is essential for discrimination of small differences between mass shifts due to different modifications.
-To illustrate this, mass shifts around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:fig7-PhosphoAndSulfo}.
+Calibration and the associated increase in mass accuracy is essential for discrimination of small differences between mass differences due to different modifications.
+To illustrate this, mass differences around 80 Daltons from the initial notch search are aggregated in the histograms in Figure~\ref{fig:fig7-PhosphoAndSulfo}.
 It is evident from these histograms that the phosphorylation and sulfonation peaks are clearly separated and identifiable when the search is conducted on well-calibrated spectra.
@@
 
-The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass shifts in the open and the multi-notch searches.
+The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass differences in the open and the multi-notch searches.
 At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 35 hours to 4 hours for all datasets (see Table~\ref{my-labelff}).

### 2017-11-28 paper.tex
jfkdjf
dfjkdsjf

@@
 The enhanced G-PTM-D workflow is presented in Figure~\ref{fgr:diagram}.
-Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files: for this purpose we implemented a calibration procedure in MetaMorpheus.
+Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files: for this purpose we implemented a post-acquisition spectral calibration procedure.
 To address the long search times of the open search, we replace the search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
@@
 
-Three multi-fraction mammalian datasets with deep proteome coverage were used to evaluate performance.
-These tryptically digested cell lysates correspond to 28 fractions from human Jurkat cells and 9 fractions from each of two different pancreatic islets from two species of mice.
-These datasets are available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215} and \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00470}, and are described in detail elsewhere~\citep{Sheynkman2013, Shortreed_2015, Cesnik_2016}.
+A multi-fraction mammalian dataset with deep proteome coverage was used to evaluate performance.
+This tryptically digested cell lysate corresponds to 28 fractions from human Jurkat cells.
+The dataset is available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215}, and are described in detail elsewhere~\citep{Sheynkman2013, Shortreed_2015, Cesnik_2016}.
 
-UniProt XML protein databases used were downloaded on Sept 14, 2017, and contain only reviewed human/mouse/contaminant proteins.
+UniProt XML protein databases used were downloaded on Sept 14, 2017, and contain only reviewed human/contaminant proteins.
 
 The database augmentation step allowed 209 modification types, including PTMs, adducts, and chemical modifications. See Supplementary Table 1 for a complete listing.
+
+The data analysis was performed using MetaMorpheus Version: 0.0.201.0.
+The following search settings were used: protease = trypsin; maximum missed cleavages = 2; minimum peptide length = 5; maximum peptide length = unspecified; initiator methionine behavior = Variable; fixed modifications = Carbamidomethyl of C; variable modifications = Oxidation of M; max modification isoforms = 4096.
+Data files were processed on a Dell Inc. Precision Tower 7910 computer using Windows version 6.2.9200.0 with a 64-Bit processor and 24 cores operating at 2.993GHz and 128GB installed RAM.
 
@@
 \includegraphics{fig4jurkat-1012.png}
-\caption{Histograms of PSM mass differences found in various dalton ranges.
+\caption{Histograms of PSM mass differences found in various Dalton ranges.
 MM, Missed Monoisotopic peak; Asp, Aspartic Acid; Asn, Asparagine; Glu, Glutamic Acid; Gln, Glutamine; Xle, Leucine/Isoleucine; Val, Valine; Thr, Threonine; Pro, Proline.}
@@
 One example of a misassigned PSM, in the 2.5 Da Search, is the example where the correct identification has oxidation and ammonia loss, but was identified as an unmodified peptide with score 9.039 and mass difference of -1.0334 Da, with sequence \texttt{QAFDDAIAELDTLNEDSYK}, in Jurkat fraction 9 scan 20962.
-With G-PTM-D added modifications, the correct peptide is identified with sequence \texttt{[Ammonia los]QAFDDAIAELDTLNEDSY[Oxidation]K}.
+With G-PTM-D added modifications, the correct peptide is identified with sequence \texttt{[Ammonia loss]QAFDDAIAELDTLNEDSY[Oxidation]K}.
 It has a mass difference of -0.0017(-0.7946 ppm) and a score of 19.222.
-While this kind of misassignment is not possible in the 5ppm narrow-window search, such missed monoisotopes are often assigned to a wrong sequence.
+While this kind of misassignment is not possible in the 5ppm narrow-window search, a different misassignment is common: scans with precursor mass errors due to misidentified monoisotopic peaks are often assigned to a wrong sequence.
 This explains the higher number of misassignments in the 5ppm search over the number in the multi-notch search.
@@
 In a conventional search, only the isolation m/z value of 726.65 would be used to select a range of theoretical peptides for comparison.
-Assuming correct de-isotoping and charge state calculation, the mass 1447.74 Da would be searched.
+Assuming correct deisotoping and charge state calculation, the mass 1447.74 Da would be searched.
 However, all peptides in the range 725.4 to 727.9 Th are co-fragmented and a better solution would be to use all co-isolated precursor masses.
@@
 \includegraphics{fig7-SulfoAndPhospho.png}
-\caption{Peaks around phospho and sulfo in uncalibrated and calibrated data}
+\caption{Peaks around phosphorylation and sulfonation in uncalibrated and calibrated data}
 \label{fig:fig7-PhosphoAndSulfo}
@@
 The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass differences in the open and the multi-notch searches.
-At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 35 hours to 4 hours for all datasets (see Table~\ref{my-labelff}).
-
-\begin{table}[]
-\centering
-\caption{Effects of Replacing Initial Open Search by a Multi-Notch Search}
-\label{my-labelff}
-\begin{tabular}{llll}
-                      &        & Open Search & Multi-Notch Search\\
-\multirow{2}{*}{Time} & Jurkat & 22.1 hrs         & 2.9 hrs    \\
-                      & Mouse  & 13.5 hrs         & 1.2 hrs   \\
-\end{tabular}
-\end{table}
+At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 22.1 hours to 2.9 hours.
 
@@
 \label{tab:table2}
-\begin{tabular}{llll}
-                      &        & Original G-PTM-D & Enhanced G-PTM-D\\
-\multirow{2}{*}{PSMs} & Jurkat  & 210566   &  220574  \\
-                      & Mouse    & 162473   &   170743 \\
-\multirow{2}{*}{Unique Peptides} & Jurkat  & 85542   &  101980  \\
-                      & Mouse    & 56235   &   66365 \\
-\multirow{2}{*}{Proteins} & Jurkat  & 9254   &  9587  \\
-                      & Mouse    & 7874   &   8151 \\
-\multirow{2}{*}{PTM containing peptides} & Jurkat  & 12041   &  22173  \\
-                      & Mouse    & 14547   &   18487 \\
+\begin{tabular}{lll}
+                             & Original G-PTM-D & Enhanced G-PTM-D\\
+PSMs &  210566   &  220574  \\
+Unique Peptides &  85542   &  101980  \\
+Proteins &  9254   &  9587  \\
+PTM containing peptides &  12041   &  22173  \\
 \end{tabular}

### 2017-11-28 FigureGeneration/fig2-10ppmSearchCalib.py

@@ -6,7 +6,7 @@ maxi = 35000
 
 
 height = 6
-width = 7 
+width =  7 
 
 f, ((ax1), (ax3)) = plt.subplots(2, 1, sharex='col', sharey='row', figsize=(width,height))
 

### 2017-11-28 FigureGeneration/fig3-searchTypes.py

@@ -3,7 +3,7 @@ import matplotlib.pyplot as plt
 import matplotlib.patches as patches
 
 
-height = 2
+height = 2.1
 width = 7 
 fig = plt.figure(figsize=(width, height))
 
@@ -120,7 +120,7 @@ ax.get_yaxis().set_visible(False)
 
 ################
 
-fig.text(0.45, 0.00, 'mass differences between observed experimental mass and theoretical candidate mass', ha='center')
+fig.text(0.45, 0.05, 'mass differences between observed experimental mass and theoretical candidate mass', ha='center')
 
 #add red blue legend to plot
 #global x axis label instead of the three mass diff da 
@@ -137,7 +137,7 @@ plt.legend(handles=[blue_patch, red_patch],bbox_to_anchor=(1.05, 0.8), loc=2, bo
 
 plt.tight_layout()
 
-plt.subplots_adjust(right=0.83)
+plt.subplots_adjust(right=0.83, bottom = 0.3)
 
 plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig3-searchTypes.png', format='png', dpi=600)
 

### 2017-11-28 FigureGeneration/fig7-phosphoSulfoUncalibCalib.py

@@ -47,7 +47,7 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, ma
 plt.annotate('?', xy=(79.962, 22), xytext=(79.955,30), arrowprops=dict(arrowstyle="->"))
 plt.annotate('?', xy=(79.971, 20), xytext=(79.975,30), arrowprops=dict(arrowstyle="->"))
 
-plt.legend()
+plt.legend(loc=2)
 
 plt.ylabel("Count")
 plt.title("Uncalibrated")
@@ -92,7 +92,7 @@ plt.hist(massDiffsHighConf1decoy, bins  =int((max1-min1)/0.001), range=[min1, ma
 plt.annotate('sulfo', xy=(79.956, 23), xytext=(79.951,30), arrowprops=dict(arrowstyle="->"))
 plt.annotate('phospho', xy=(79.967, 28), xytext=(79.971,30), arrowprops=dict(arrowstyle="->"))
 
-plt.legend()
+plt.legend(loc=2)
 
 plt.ylabel("Count")
 plt.title("Calibrated")

### 2017-11-28 FigureGeneration/fig2-10ppmSearchCalib.py
djfkdj

@@ -5,10 +5,10 @@ import pandas
 maxi = 35000 
 
 
-height = 6
+height = 3
 width =  7 
 
-f, ((ax1), (ax3)) = plt.subplots(2, 1, sharex='col', sharey='row', figsize=(width,height))
+f, ((ax1), (ax3)) = plt.subplots(1, 2, figsize=(width,height))
 
 ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\1mm10ppmSearch\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')  
 
@@ -30,11 +30,8 @@ ax1.hist(A[(C<0.01) & B], bins  = 101, range=[-10.1,10.1])
 ax1.set_ylim([1,maxi])
 ax1.vlines(0,0,maxi,zorder = 10)
 ax1.set_ylabel('PSM count')
-ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
-                size='large', ha='right', va='center', xytext=(0,0.5))
-
-
-
+ax1.set_xlabel('Mass Error (ppm)')
+ax1.set_title('Uncalibrated')
 
 ok = pandas.read_csv(r"C:\Users\stepa\Data\PaperData\JurkatTrypsin\Calibrated\1mm10ppm\Task1-SearchTask\aggregatePSMs_1mm.psmtsv", sep='\t')  
 
@@ -57,15 +54,10 @@ ax3.set_ylim([1,maxi])
 ax3.vlines(0,0,maxi,zorder = 10)
 ax3.set_ylabel('PSM count')
 ax3.set_xlabel('Mass Error (ppm)')
-
-ax3.annotate(r"Calibrated", xy=(0, 0.5), xycoords=ax3.yaxis.label, textcoords='offset points',
-                size='large', ha='right', va='center', xytext=(0,0.5))
+ax3.set_title('Calibrated')
 
 
 plt.tight_layout()
-
-plt.subplots_adjust(left=0.3)
-
 
 plt.savefig(r'C:\Users\stepa\Source\StefanPaper\fig2-10ppmSearchCalib.png', format='png', dpi=600)
 

### 2017-11-28 paper.tex
djfkdj
dfjk

@@
 \begin{tabular}{lll}
-
                                                                                       & Orig        & Calib        \\
@@
 \end{figure}
-
-\newpage
 
@@
 \includegraphics{fig3-searchTypes.png}
-\caption{The three searches illustrated. Blue regions signify the allowed mass differences, while the pink ones show the mass differences that are not considered.}
+\caption{Schematic illustration of the three search types.
+Certain mass differences between the observed experimental mass and theoretical candidate masses are allowed (blue regions) or not considered (pink regions).}
 \label{fig:fig3-searchTypes}

### 2017-11-28 FigureGeneration/fig3-searchTypes.py

@@ -110,7 +110,7 @@ for p in [
 ax.set_xlim([-110,110])
 ax.set_ylim([-2, 9])
 
-plt.annotate('Only specific mass\ndifferences allowed', xy=(0,1), xytext=(-95,4), arrowprops=dict(arrowstyle="->"))
+plt.annotate('Multiple specific mass\ndifferences allowed', xy=(0,1), xytext=(-103,4), arrowprops=dict(arrowstyle="->"))
 ax.annotate("", xy=(42, 1), xytext=(10, 3.3), arrowprops=dict(arrowstyle="->"))
 ax.annotate("", xy=(80, 1), xytext=(20, 3.3), arrowprops=dict(arrowstyle="->"))
 
@@ -120,7 +120,7 @@ ax.get_yaxis().set_visible(False)
 
 ################
 
-fig.text(0.45, 0.05, 'mass differences between observed experimental mass and theoretical candidate mass', ha='center')
+fig.text(0.45, 0.05, 'Mass difference (Da)', ha='center')
 
 #add red blue legend to plot
 #global x axis label instead of the three mass diff da 

### 2017-11-29 paper.tex
lloyd

@@
 G-PTM-D is a recently developed technique for global identification and localization of PTMs.
-Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search shown here significantly increase the accuracy, speed and scope of PTM identification.
+Spectral file calibration prior to applying G-PTM-D, and algorithmic enhancements in the peptide database search significantly increase the accuracy, speed and scope of PTM identification.
 We enhance G-PTM-D by using multi-notch searches, and demonstrate its effectiveness in identification of numerous types of PTMs, including high-mass modifications such as glycosylations.
@@
 Spectral calibration is essential for distinguishing modifications with similar mass, and so we recommend running the G-PTM-D procedure on calibrated spectra files: for this purpose we implemented a post-acquisition spectral calibration procedure.
-To address the long search times of the open search, we replace the search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
+To address the long search times of the open search, we replaced the search in the first stage by a specialized search called a multi-notch search that only allows specific mass differences which correspond to pre-selected masses (e.g.
 PTMs and amino acids).
@@
 We developed a modified version of the Morpheus software for bottom-up spectral database searching\citep{Wenger_2013} that we call MetaMorpheus, which integrates the database search procedure with spectral calibration and the enhanced G-PTM-D workflow.
-MetaMorpheus is freely available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}.
+MetaMorpheus is open-source and freely available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}.
 MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
@@
 This tryptically digested cell lysate corresponds to 28 fractions from human Jurkat cells.
-The dataset is available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215}, and are described in detail elsewhere~\citep{Sheynkman2013, Shortreed_2015, Cesnik_2016}.
+The dataset is available at \url{https://db.systemsbiology.net/sbeams/cgi/PeptideAtlas/PASS_View?identifier=PASS00215}, and is described in detail elsewhere~\citep{Sheynkman2013, Shortreed_2015, Cesnik_2016}.
 
@@
 
-The data analysis was performed using MetaMorpheus Version: 0.0.201.0.
-The following search settings were used: protease = trypsin; maximum missed cleavages = 2; minimum peptide length = 5; maximum peptide length = unspecified; initiator methionine behavior = Variable; fixed modifications = Carbamidomethyl of C; variable modifications = Oxidation of M; max modification isoforms = 4096.
+The data analysis was performed using MetaMorpheus version 0.0.201.
+The following search settings were used: protease = trypsin; maximum missed cleavages = 2; minimum peptide length = 5; maximum peptide length = unspecified; initiator methionine behavior = variable; fixed modifications = carbamidomethyl of C; variable modifications = oxidation of M; max modification isoforms = 4096.
 Data files were processed on a Dell Inc. Precision Tower 7910 computer using Windows version 6.2.9200.0 with a 64-Bit processor and 24 cores operating at 2.993GHz and 128GB installed RAM.
@@
  \includegraphics{fig1-calibErrors.png}
- \caption{Errors in m/z value of select peaks before (top panels) and after (bottom panels) calibration. These errors are plotted vs two variables: m/z value (left) and the retention time (right).}
+ \caption{Errors in m/z values of select peaks before (top panels) and after (bottom panels) calibration. These errors are plotted vs two variables: m/z value (left) and the retention time (right).}
  \label{fgr:fig1-calibErrors}
@@
 \caption{Histograms of PSM mass differences found in various Dalton ranges.
-MM, Missed Monoisotopic peak; Asp, Aspartic Acid; Asn, Asparagine; Glu, Glutamic Acid; Gln, Glutamine; Xle, Leucine/Isoleucine; Val, Valine; Thr, Threonine; Pro, Proline.}
+Mm, Missed Monoisotopic peak; Asp, Aspartic Acid; Asn, Asparagine; Glu, Glutamic Acid; Gln, Glutamine; Xle, Leucine/Isoleucine; Val, Valine; Thr, Threonine; Pro, Proline.}
 \label{fig:fig4jurkat-1012}
@@
 Another example of this relates to peptide modifications.
-The phosphopeptide \texttt{PEP[Phospho]SK} could be reported with a mass error of -79.96 Da when the better identification would have been \texttt{PEPSK}.
+The phosphopeptide \texttt{PEPS[Phospho]K} could be reported with a mass error of -79.96 Da when the better identification would have been \texttt{PEPSK}.
 These simple errors are relatively easy to spot, but not when there are missed cleavages with multiple amino acids and/or multiple PTMs.
 
-The alternative we recommend, is the use of a multi-notch search, which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the fewest false positive PSM identifications.
+The alternative we recommend is the use of a multi-notch search, which offers a means of reducing the search space and limiting the number of theoretical candidates to those plausible for the scan, increasing the search speed and most importantly, producing the most correct PSMs with the fewest false positive PSM identifications.
 We compared searches of all Jurkat fractions using a $\pm 500$ Da wide mass search to a multi-notch search with notches at 0 and 79.961 with 0.02 Dalton tolerance.
@@
 These enhancements demonstrate that the enhanced G-PTM-D approach provides great value to bottom-up proteomic data analysis.
-It is fast enough included as part of every regular work-flow, and will yield significantly increased numbers of correctly identified PTM peptides.
+It is fast enough to be included as part of every regular work-flow, and will yield significantly increased numbers of correctly identified PTM peptides.
 

### 2017-11-29 FigureGeneration/absGraphic.py
jdfkj

@@ -1,0 +1,88 @@
+import numpy as np
+import matplotlib.pyplot as plt
+import matplotlib.patches as patches
+
+from PIL import Image
+from io import BytesIO
+
+
+height = 1.75 
+width = 3.25 
+fig = plt.figure(figsize=(width, height))
+
+#######
+
+ax = fig.add_subplot(111)
+for p in [
+    patches.Rectangle(
+        (-110, 0), 106, 1,
+        alpha=0.4,
+        facecolor="red"
+    ),
+    patches.Rectangle(
+        (-4, 0), 8, 1,
+    ),
+    patches.Rectangle(
+        (4, 0), 34, 1,
+        alpha=0.4,
+        facecolor="red"
+    ),
+    patches.Rectangle(
+        (38, 0), 8, 1,
+    ),
+    patches.Rectangle(
+        (46, 0), 30, 1,
+        alpha=0.4,
+        facecolor="red"
+    ),
+    patches.Rectangle(
+        (76, 0), 8, 1,
+    ),
+    patches.Rectangle(
+        (84, 0), 26, 1,
+        alpha=0.4,
+        facecolor="red"
+    )
+]:
+    ax.add_patch(p)
+
+
+ax.set_xlim([-110,110])
+ax.set_ylim([-2, 9])
+
+plt.annotate('Multiple specific mass\ndifferences allowed', xy=(0,1), xytext=(-103,4), arrowprops=dict(arrowstyle="->"))
+ax.annotate("", xy=(42, 1), xytext=(10, 3.3), arrowprops=dict(arrowstyle="->"))
+ax.annotate("", xy=(80, 1), xytext=(20, 3.3), arrowprops=dict(arrowstyle="->"))
+
+plt.title('Multi-Notch Search')
+
+ax.get_yaxis().set_visible(False)
+
+################
+
+fig.text(0.45, 0.05, 'Mass difference (Da)', ha='center')
+
+#######
+
+blue_patch = patches.Patch(label='Allowed')
+red_patch = patches.Patch(color='red', label='Not\nAllowed', alpha=0.4)
+
+
+plt.tight_layout()
+
+# save figure
+# (1) save the image in memory in PNG format
+
+png1 = BytesIO()
+plt.savefig(png1, format='png', dpi=300)
+
+# (2) load this image into PIL
+png2 = Image.open(png1)
+print(png2.info['dpi'])
+
+# (3) save as TIFF
[…]

### 2017-11-29 FigureGeneration/absGraphic.py
graphics

@@ -88,7 +88,7 @@ ax.set_ylim([-15, 12])
 
 #ax.axis('off')
 
-ax.annotate('Multi-Notch Sofware Filtering', xytext=(1.5, -6.5), xy=(0,0), fontsize=8)
+ax.annotate('Multi-Notch Sofware Filtering', xytext=(1.7, -6.6), xy=(0,0), fontsize=8)
 
 
 #plt.tight_layout()
@@ -103,6 +103,11 @@ ax.spines["bottom"].set_visible(False)
 ax.get_yaxis().set_visible(False)
 
 ax.set_xticks([0,1,16])
+ax.set_xticklabels(['0', '1', '79.966'])
+
+
+plt.tick_params(axis='both', which='major', labelsize=8)
+plt.tick_params(axis='both', which='minor', labelsize=8)
 
 ax.arrow(6,20,0,-10, head_width=1, head_length=5, facecolor="red")
 ax.arrow(10,20,0,-10, head_width=1, head_length=5, facecolor="red")
@@ -117,7 +122,7 @@ ax.arrow(1,0,0,-8, head_width=1, head_length=5)
 ax.arrow(16,0,0,-8, head_width=1, head_length=5)
 
 
-ax.annotate('Accepted Mass Diffs (Da)', xytext=(2.5, -27), xy=(0,0), fontsize=8)
+ax.annotate('Accepted Mass Diffs (Da)', xytext=(2.4, -27), xy=(0,0), fontsize=8)
 
 plt.subplots_adjust(bottom=0.15)
 

### 2017-11-29 paper.tex
dfjkdjf

@@
 
-\title[Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus]
-  {Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus}
+\title{Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus}
+
+\keywords{G-PTM-D, post-translational modification discovery, proteomics, database search, calibration, co-isolation, post-translational modification}
 

### 2017-11-29 paper.tex
djfkdjf

@@
 
-The database augmentation step allowed 209 modification types, including PTMs, adducts, and chemical modifications. See Supplementary Table 1 for a complete listing.
+The database augmentation step allowed 208 modification types, including PTMs, adducts, and chemical modifications.
+See Supplementary Table 1 for a complete listing.
 
@@
 
-%\begin{suppinfo}
+\begin{suppinfo}
 
-%The following files are available free of charge.
-%\begin{itemize}
-%  \item Filename: brief description
-%  \item Filename: brief description
-%\end{itemize}
+The following files are available free of charge.
+\begin{itemize}
+  \item Supp1.xlsx: The complete list of modifications used for enhanced G-PTM-D
+  \item Supp2Calibration.docx: brief description
+\end{itemize}
 

### 2017-11-29 FigureGeneration/absGraphic.py
djfk

@@ -141,7 +141,7 @@ png2 = Image.open(png1)
 print(png2.info['dpi'])
 
 # (3) save as TIFF
-png2.save(r'C:\Users\stepa\Source\StefanPaper\absGraphic.tiff',dpi=[300,300])
+png2.save(r'C:\Users\stepa\Source\StefanPaper\absGraphic.tif',dpi=[300,300])
 
 png1.close()
 

### 2017-12-04 paper.tex
jdfkdjf

@@
 Multi-notch searches are also shown to be useful in contexts other than G-PTM-D, producing superior results when used instead of standard narrow-window and open database searches.
+
 \end{abstract}
@@
 
-%\begin{acknowledgement}
+\begin{acknowledgement}
 
-%The authors thank \ldots
-%\end{acknowledgement}
+We thank the members of the Smith lab for testing G-PTM-D and other MetaMorpheus functionality (Leah V.
+Schaffer, Anthony J.
+Cesnik, Rob Millikin, Lei Lu, Zach Rolfs, Sean Dai, Rachel Miller, Rachel Knoener), who contributed input to the improvement of mzLib, and MetaMorpheus.
+We thank the MetaMorpheus users outside of this lab who submitted useful comments on GitHub.
+This work was supported by the grant DBI-1458524 from the National Science Foundation.
+
+\end{acknowledgement}
 

### 2017-12-05 paper.tex
djfkdjf

@@
 The database augmentation step allowed 208 modification types, including PTMs, adducts, and chemical modifications.
-See Supplementary Table 1 for a complete listing.
+See Table S-1 for a complete listing.
 
@@
 Step 2: The identifications from Step 1 provide a theoretical list of m/z peaks that should match closely to observed m/z peaks in the spectra file, and the discrepancies between the theoretical and observed peaks are used to inform a peak shifting procedure.
-The calibration algorithm is described in detail in Supplementary Material 2.
+The calibration algorithm is described in detail in Document S-2.
 
@@
 In a conventional search, only the isolation m/z value of 726.65 would be used to select a range of theoretical peptides for comparison.
-Assuming correct deisotoping and charge state calculation, the mass 1447.74 Da would be searched.
+Assuming correct deisotoping and charge state calculation, the mass 2175.96 Da would be searched.
 However, all peptides in the range 725.4 to 727.9 Th are co-fragmented and a better solution would be to use all co-isolated precursor masses.
@@
 Each deconvoluted mass is used to select a range of theoreticals to use for identification.
-The result of this is that often, more than one peptide can be identified from a single MS2 scan.
+The result of this is that often, more than one peptide can be identified from a single MS2 scan, as shown in Figure~\ref{fig:fig5-coIsolationSpectrum}.
 This is extremely beneficial in both open mass and multi-notch searches because it reduces the number of the more difficult to interpret mass differences and substantially improves G-PTM-D (see below).
@@
 
-The next step, database augmentation, outputs significantly different augmented databases because of the difference between accepted mass differences in the open and the multi-notch searches.
-At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly, from 22.1 hours to 2.9 hours.
+Further improvements arise from using a multi-notch search to create the augmented database.
+The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see Document S-2).
+The resulting augmented database is greatly improved compared to using an open search for this step.
+At this point, another advantage of the enhanced approach is clear; the overall search time dropped significantly from 22.1 hours to 2.9 hours, primarily due to the narrow tolerance multi-notch search versus the open search.
 
@@
 
-We thank the members of the Smith lab for testing G-PTM-D and other MetaMorpheus functionality (Leah V.
+We thank the members of the Smith lab (Leah V.
 Schaffer, Anthony J.
-Cesnik, Rob Millikin, Lei Lu, Zach Rolfs, Sean Dai, Rachel Miller, Rachel Knoener), who contributed input to the improvement of mzLib, and MetaMorpheus.
+Cesnik, Rob Millikin, Lei Lu, Zach Rolfs, Sean Dai, Rachel Miller, and Rachel Knoener) for testing G-PTM-D and contributing to the improvement of mzLib and MetaMorpheus.
 We thank the MetaMorpheus users outside of this lab who submitted useful comments on GitHub.

### 2018-02-15 paper.tex
djfkf
review
okay

@@
 \usepackage{booktabs}
-\usepackage{hyperref}
 
@@
 \email{smith@chem.wisc.edu}
-\affiliation[UwMadison]
-{University of Wisconsin-Madison}
+\altaffiliation{Genome Center of Wisconsin, University of Wisconsin-Madison, Madison, WI, USA}
+\affiliation[UW Madison]
+{Department of Chemistry, University of Wisconsin-Madison, Madison, WI, USA}
 
@@
 \begin{figure}[H]
-  \includegraphics[scale=0.5]{paperDiagram.png}
-  \caption{The three steps of the G-PTM-D workflow are indicated in the light beige color. The dark orange color corresponds to the inputs.}
+  \includegraphics[scale=0.75]{paperDiagram.png}
+  \caption{Comparing the original G-PTM-D workflow with the proposed enhanced workflow. Major changes are highlighted in pink.}
   \label{fgr:diagram}
@@
 The database augmentation step allowed 208 modification types, including PTMs, adducts, and chemical modifications.
-See Table S-1 for a complete listing.
+See supplement Supp1 for a complete listing.
 
@@
 
-We propose a limited multi-notch search as a remedy to the trade-offs caused by choosing a narrow (few ppm) or wide (few Da) precursor mass tolerance during a traditional narrow-window search.
+A multi-notch search is a remedy to the trade-offs caused by choosing a narrow (few ppm) or wide (few Da) precursor mass tolerance during a traditional narrow-window search.
+It is not novel, and many useful search tools such as Tandem, MSGF+, Coment, MSFragger, Sequest have this search functionality under names such as "isotope error option".
 This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (e.g.
@@
 
+The results of a multi-notch search are assessed using notch-specific FDRs, computed individually for each notch, similarly to MsFragger.
+
 \subsection{Co-isolation}
@@
 Further improvements arise from using a multi-notch search to create the augmented database.
-The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see Document S-2).
+The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see supplement Supp2Calibration).
 The resulting augmented database is greatly improved compared to using an open search for this step.
@@
 
-%\end{suppinfo}
+\end{suppinfo}
 

### 2018-02-15 FigureGeneration/fig1-calibErrors.py

@@ -39,7 +39,7 @@ ax1.plot(nice['mz'],nice['LabelTh'], 'bo',markersize = .3)
 ax1.hlines(0, min(nice['mz']), max(nice['mz']), zorder = 10)
 ax1.set_ylim([minY,maxY])
 
-ax1.set_ylabel('m/z error (Th)')
+ax1.set_ylabel('m/z error')
 ax1.annotate(r"Uncalibrated", xy=(0, 0.5), xycoords=ax1.yaxis.label, textcoords='offset points',
                 size='large', ha='right', va='center', xytext=(0,0.5))
 
@@ -50,8 +50,8 @@ ax2.set_ylim([minY,maxY])
 ax3.plot(nice2['mz'],nice2['LabelTh'], 'bo',markersize = .3)
 ax3.hlines(0, min(nice2['mz']), max(nice2['mz']), zorder = 10)
 
-ax3.set_ylabel('m/z error (Th)')
-ax3.set_xlabel('m/z (Th)')
+ax3.set_ylabel('m/z error')
+ax3.set_xlabel('m/z')
 
 ax3.annotate(r"Calibrated", xy=(0, 0.5), xycoords=ax3.yaxis.label, textcoords='offset points',
                 size='large', ha='right', va='center', xytext=(0,0.5))

### 2018-02-16 paper.tex
oka
yea

@@
 A multi-notch search is a remedy to the trade-offs caused by choosing a narrow (few ppm) or wide (few Da) precursor mass tolerance during a traditional narrow-window search.
-It is not novel, and many useful search tools such as Tandem, MSGF+, Coment, MSFragger, Sequest have this search functionality under names such as "isotope error option".
+It is not novel, and many useful search tools such as Tandem, MSGF+, Comet, MSFragger, Sequest have this search functionality under names such as "isotope error option".
 This limited multi-notch search allows mass differences between precursor and theoretical peptides of 0, 1.0029 and 2.0054 Da, but it limits the width around these three mass differences to a few ppm (e.g.
@@
 
-The results of a multi-notch search are assessed using notch-specific FDRs, computed individually for each notch, similarly to MsFragger.
+The results of a multi-notch search are assessed using notch-specific FDRs, computed individually for each notch, similarly to MSFragger.
+It is important to note that there exist post-processing tools that would push identifications with unknown mass shifts and large numbers of missed cleavages down the list, such as the PeptideProphet extended mass model.
 

### 2018-02-16 paper.tex
jkjf

@@
 The database augmentation step allowed 208 modification types, including PTMs, adducts, and chemical modifications.
-See supplement Supp1 for a complete listing.
+See Table S1 for a complete listing.
 
@@
 Step 2: The identifications from Step 1 provide a theoretical list of m/z peaks that should match closely to observed m/z peaks in the spectra file, and the discrepancies between the theoretical and observed peaks are used to inform a peak shifting procedure.
-The calibration algorithm is described in detail in Document S-2.
+The calibration algorithm is described in detail in Document S1.
 
@@
 
-The following files are available free of charge.
+The following supplementary information is available:
 \begin{itemize}
-  \item Supp1.xlsx: The complete list of modifications used for enhanced G-PTM-D
-  \item Supp2Calibration.docx: brief description
+  \item Table S1: The complete list of modifications used for enhanced G-PTM-D
+  \item Document S1: A brief description of the calibration methodology
+  \item Document S2: A 
 \end{itemize}

### 2018-02-16 supportingInfo.tex
jkjf

\documentclass{article}

\usepackage[letterpaper, margin=1.3in]{geometry}
\usepackage[version=3]{mhchem} % Formula subscripts using \ce{}
\usepackage[T1]{fontenc}       % Use modern font encodings
\usepackage{multirow}
\usepackage{subcaption}
\usepackage{booktabs}
\usepackage{fancyhdr}

\pagestyle{fancy}
\fancyhf{}
\lhead{Supoprting Information}
\fancyfoot[C]{S--\thepage}
\begin{document}

{\Huge Enhanced Global PTM Discovery (G-PTM-D) with MetaMorpheus}

{\Large Stefan K. Solntsev\textsuperscript{1}, Michael R. Shortreed\textsuperscript{1}, Brian L. Frey\textsuperscript{1}, Lloyd M. Smith\textsuperscript{1,2}}\\
{\footnotesize 1 Department of Chemistry, University of Wisconsin-Madison, Madison, WI, USA}\\
{\footnotesize 2 Genome Center of Wisconsin, University of Wisconsin-Madison, Madison, WI, USA}\\

\section{Table of Contents}

\begin{enumerate}
\item \textbf{Table S1} The complete list of modifications used for enhanced G-PTM-D\\
      See Excel file
\item \textbf{Document S1}   A brief description of the calibration methodology      \hfill      Page S-1
\item \textbf{Document S2}         \hfill      Page S-2
\end{enumerate}

\newpage

{\huge \textbf{Document S1:} A brief description of the calibration methodology}

The calibration procedure is motivated by the software lock mass concept recently reported by the Mann group. 
In MetaMorpheus it is comprised of a nested cycle of a database searches, data point acquisition procedures, calibration curve fitting, and peak alignment procedures.

\begin{figure}[ht]
\includegraphics{supFig1.png}
\end{figure}

The narrow-window database search is run with precursor and parent mass tolerances that are appropriate for the uncalibrated spectra file. The target PSMs within 1\% false discovery rate are good identifications to be used in the data point acquisition step.

The data point acquisition considers every identification in the confident PSMs, and searches the corresponding fragmentation spectra for peaks matching the fragments, considering different charge states and isotopic peaks for each fragment. If there is ambiguity as to which peak match a certain fragment, the corresponding match is discarded, because of the desire to use the most confident peaks for calibration, and an abundance of unambiguous peaks. If a sufficient number of such peaks is found, the identification is confident enough for calibration. Peaks corresponding to the unfragmented peptide in neighboring MS spectra are then gathered. Every accepted peak is a data point for the subsequent calibration curve fitting: every data point is comprised of the following numerical values:

\begin{itemize}
\item \textbf{Features:} m/z value,Retention time,Total ion count,Injection time
\item \textbf{Label:} m/z error
\end{itemize}

The inputs are the values that are available for every peak in the uncalibrated spectra, and would be used as inputs to the chosen calibration function. In addition to those, the confident identification, by means of a theoretical isotopic distribution, provides a Label for each data point, which is the difference between the observed and the expected m/z value of the peak. The calibration procedure relies on the assumption that peaks with similar inputs would have similar labels in the m/z value of the error.

Once the peaks are gathered, the data point set is split into a training and testing sets, where the training set is used to create several calibration curves, and a testing set is used to select the best curve based on the Mean Squared Error.

The contribution of every input variable to the error is modeled either as a linear function (chosen for simplicity and to account for instrument miscalibration) or as a random forest. The random forest model allows for non-linear relationship between an input variable and the m/z error, such as random ambient time-resolved fluctuations. The trained model is used to calibrate all the peaks in the spectra, by shifting the m/z value of the peak by an appropriate amount.

This procedure is repeated on a single collection of data points, and on multiple rounds of database searching.

If a pre-calibrated file is given as an input, no changes are made. This is due to the fact that the calibration procedure only accepts changes if a test database search results show improvement.

Resulting calibrated spectra are written to an mzML file that may be used in subsequent MetaMorpheus tasks, or in any external software that may benefit from increased mass accuracy in the spectra.

\newpage

{\huge \textbf{Document S2:} }

\end{document}

### 2018-03-18 paper.tex
n
from michael

@@
 MetaMorpheus is open-source and freely available at \url{https://github.com/smith-chem-wisc/MetaMorpheus}.
+It has attractive properties like Morpheus, such as the ability of running large-scale search tasks on memory-restricted machines.
 MetaMorpheus heavily utilizes an auxiliary library called mzLib, which is a vast library of software tools for analyzing and processing mass spectral data.
@@
 The results of a multi-notch search are assessed using notch-specific FDRs, computed individually for each notch, similarly to MSFragger.
-It is important to note that there exist post-processing tools that would push identifications with unknown mass shifts and large numbers of missed cleavages down the list, such as the PeptideProphet extended mass model.
+MetaMorpheus users interested in further downstream analysis using Percolator can use the pre-formatted Percolator output, which is a tab-delimited file with the file extension \texttt{"forPercolator.tsv"}.
+It is important to note that there exist post-processing tools that would push identifications with unknown mass shifts and large numbers of missed cleavages down the list, such as the PeptideProphet~\cite{Ma2012} extended mass model.
+
+Protein-level inference in MetaMorpheus uses all available observed peptides, including PTM-containing peptides. A minimal list of proteins to explain the observed set of peptides is constructed.
+Briefly, peptides that could only be a digestion product of one protein in the database ("unique peptides") are used to construct an initial protein set, and any peptide that could be a digestion product of this initial set is associated with those protein(s) and removed from the pool of unassigned peptides.
+The remaining unassigned peptides are assigned to proteins using the greedy algorithm.
+Proteins with identical peptide sets are grouped. Protein group FDR is estimated with the top-picked method.
+In this approach, each protein is paired with its reversed decoy. The PSM with the higher notch q-value (or lower score, in the case of a tie) is retrieved for each protein in the target-decoy pair.
+The protein associated with this PSM is considered the lower-quality hit of the pair.
+The protein group containing this protein is removed from the protein group set.
+Protein groups are then ranked by the notch q-value of their best PSM, and protein q-value is calculated as the number of decoys divided by the number of targets in the ranked list.
 
@@
   \item Document S1: A brief description of the calibration methodology
-  \item Document S2: A 
 \end{itemize}

### 2018-03-18 supportingInfo.tex
from michael

@@
 \item \textbf{Document S1}   A brief description of the calibration methodology      \hfill      Page S-1
-\item \textbf{Document S2}         \hfill      Page S-2
 \end{enumerate}
@@
 
-If a pre-calibrated file is given as an input, no changes are made. This is due to the fact that the calibration procedure only accepts changes if a test database search results show improvement.
+If a well-calibrated file is given as an input, no changes are made. This is because the calibration procedure only accepts changes if a test database search results in improvement.
 

### 2018-03-18 paper.tex
mkay

@@
 Further improvements arise from using a multi-notch search to create the augmented database.
-The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see supplement Supp2Calibration).
+The enhanced G-PTM-D procedure employed a narrow 5 ppm tolerance around 208 mass difference notches (see Table S1).
+The list of modifications added by G-PTM-D was generated based on a histogram of mass shifts from an open search.
+Any mass shift that was confidently identified 10 or more times, was considered as a possible modification.
+If a match was found in Unimod, the modification was added to our list.
 The resulting augmented database is greatly improved compared to using an open search for this step.
